---
title: "Benchmark rsofun tag v2.1"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(rsofun)
library(dplyr)
library(readr)
library(ingestr)
library(ggplot2)
library(knitr)
```

## Description

Benchmarking rsofun v2.1.

## Load data

Load drivers data frame (created by `prepare_inputs_FLUXNET2015_ensemble.Rmd`).
```{r}
load("~/data/rsofun_benchmarking/df_drivers_fluxnet2015.Rdata")
```

Change fluxes to units per second.
```{r}
secs_per_tstep <- 60 * 60 * 24
df_drivers_fluxnet2015 <- df_drivers_fluxnet2015 %>% 
  dplyr::select(sitename, forcing) %>% 
  unnest(forcing) %>% 
  mutate(prec = prec / secs_per_tstep, ppfd = ppfd / secs_per_tstep) %>% 
  group_by(sitename) %>% 
  nest() %>% 
  right_join(
    df_drivers_fluxnet2015 %>% 
      dplyr::select(-forcing),
    by = "sitename"
  ) %>% 
  dplyr::select(sitename, forcing = data, params_siml, siteinfo, df_soiltexture) %>% 
  ungroup()
```


## Calibrate model

Define calibration sites.
```{r}
flue_sites <- readr::read_csv( "~/data/flue/flue_stocker18nphyt.csv" ) %>%
              dplyr::filter( !is.na(cluster) ) %>% 
              distinct(site) %>% 
              pull(site)

calibsites <- siteinfo_fluxnet2015 %>% 
  dplyr::filter(!(sitename %in% c("DE-Akm", "IT-Ro1"))) %>%  # excluded because fapar data could not be downloaded (WEIRD)
  # dplyr::filter(!(sitename %in% c("AU-Wom"))) %>%  # excluded because no GPP data was found in FLUXNET file
  dplyr::filter(sitename != "FI-Sod") %>%  # excluded because some temperature data is missing
  dplyr::filter( c4 %in% c(FALSE, NA) & classid != "CRO" & classid != "WET" ) %>%
  dplyr::filter( sitename %in% flue_sites ) %>%
  pull(sitename)
```

Define calibration settings.
```{r}
settings_calib <- list(
  method              = "gensa",
  targetvars          = c("gpp"),
  timescale           = list( gpp = "d" ),
  maxit               = 5,
  sitenames           = calibsites,
  metric              = "rmse",
  dir_results         = "./",
  name                = "FULL",
  par                 = list( kphio       = list( lower=0.03, upper=0.1, init= 0.05 ),
                              soilm_par_a = list( lower=0.0,  upper=1.0, init=0.0 ),
                              soilm_par_b = list( lower=0.0,  upper=1.5, init=0.6 ) )
 )
```

Use the [ingestr](https://github.com/stineb/ingestr) package once again, now for collecting calibration target data. I.e., GPP based on the nighttime flux decomposition method.
```{r warning=FALSE, message=FALSE}
filn <- "~/data/rsofun_benchmarking/ddf_fluxnet_gpp.Rdata"
load(filn)
```

Calibrate the model.
```{r warning=FALSE, message=FALSE}
set.seed(1982)
settings_calib <- calib_sofun(
  df_drivers = dplyr::filter(df_drivers_fluxnet2015, sitename %in% calibsites),  # use only one site
  ddf_obs = ddf_fluxnet_gpp,
  settings = settings_calib
  )
```

The calibrated parameters are returned by `calib_sofun()` as part of the list:
```{r}
print(settings_calib$par_opt)
save(settings_calib, file = "./settings_calib.Rdata")
```

Update model parameters.
```{r}
params_modl <- list(
	kphio           = 0.05,
	soilm_par_a     = 1.0,
	soilm_par_b     = 0.0,
	vpdstress_par_a = 9999,
	vpdstress_par_b = 9999,
	vpdstress_par_m = 9999
	)
params_modl <- update_params(params_modl, settings_calib)
```


## Run model

```{r warning=FALSE, message=FALSE}
df_output <- runread_pmodel_f(
     df_drivers_fluxnet2015,
     params_modl = params_modl, 
     makecheck = TRUE,
     parallel = FALSE
     )
```

## Run evaluation

Do evaluation only for sites where simulation was run.
```{r}
evalsites <- df_output %>% 
  mutate(ntsteps = purrr::map_dbl(data, ~nrow(.))) %>% 
  dplyr::filter(ntsteps > 0) %>% 
  pull(sitename)
```

Load standard benchmarking file with observational data for evaluation.
```{r}
load("~/data/rsofun_benchmarking/obs_eval_fluxnet2015.Rdata")
```


Define evaluation settings.
```{r}
settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = evalsites,
  agg       = 8  # An integer specifying the number of days used to define the width of bins for daily data aggregated to several days
  )
```

And finally run the evaluation.
```{r warning=FALSE, message=FALSE, error=FALSE}
out_eval <- eval_sofun( 
  df_output, 
  settings_eval, 
  settings_sims, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

## Evaluation results

### Metrics table

```{r}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable
```

### Visualisations

```{r message=FALSE, warning=FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```


## Appendix

### Site list

```{r}
siteinfo_fluxnet2015 %>% 
  dplyr::filter(sitename %in% evalsites) %>% 
  kable()
```