---
title: "ConstantSelfThinning"
output:
  html_document: 
    fig_width: 10
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(dplyr)
library(tibble)
library(rsofun)
library(ggplot2)
```

## Example

### Simulation settings
### Simulation settings

Manually select some sites from which we're going to use the data for evaluation and calibration.
```{r}
mysites <- c("ORNL")
```

Create a site meta info table that contains all the site-specific information that is used to force site-simulations (e.g. starting year, number of simulations years, elevation, etc.). For FLUXNET2015 data, required meta info is provided by the `rsofun` package (data frame `rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015`).
```{r}
siteinfo <- data.frame(sitename="ORNL", lon = 0.0, lat = 0.0, elv = 0.0)
```

Now specify the simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
      spinup                = TRUE,
      spinupyears           = 1800, #1793,
      recycle               = 1, # 9 or 11 changed to 1 when aggregating forcing into 1 year
      firstyeartrend        = 1998,
      nyeartrend            = 1, #9 or 11 (longer transient years)
      outputhourly          = TRUE,
      outputdaily           = TRUE,
      do_U_shaped_mortality = FALSE,
      update_annualLAImax   = FALSE,
      do_closedN_run        = TRUE,
      method_photosynth     = "pmodel", # gs_leuning or pmodel
      method_mortality      = "const_selfthin" # dbh or cstarvation or growthrate or const_selfthin
      )
```

### Define model parameters

#### Tile-level parameters

```{r}
params_tile <- list(
  soiltype     = 3,     # Sand = 1, LoamySand = 2, SandyLoam = 3, SiltLoam = 4, FrittedClay = 5, Loam = 6, Clay = 7
  FLDCAP       = 0.4,   # soil property: field capacity 
  WILTPT       = 0.05,  # soil property: wilting point
  K1           = 2.0,   # turnover rate of fast SOM per year
  K2           = 0.05,  # turnover rate of slow SOM per year
  K_nitrogen   = 8.0,   # mineral Nitrogen turnover rate
  MLmixRatio   = 0.8,   # the ratio of C and N returned to litters from microbes
  etaN         = 0.025, # loss rate with runoff
  LMAmin       = 0.02,  # minimum LMA, boundary condition
  fsc_fine     = 1.0,   # fraction of fast turnover carbon in fine biomass
  fsc_wood     = 0.0,   # fraction of fast turnover carbon in wood biomass
  GR_factor    = 0.33,  # growth respiration factor
  l_fract      = 0.0,   # fraction of the carbon retained after leaf drop
  retransN     = 0.0,   # retranslocation coefficient of Nitrogen
  f_initialBSW = 0.2,
  f_N_add      = 0.02   # re-fill of N for sapwood
  )

```

#### Species-level parameters

```{r}
params_species <- tibble(
  lifeform      = rep(1,16),                 # 0 for grasses; 1 for trees
  phenotype     = rep(0,16),                 # 0 for Deciduous; 1 for Evergreen
  pt            = rep(0,16),                 # 0 for C3; 1 for C4
  seedlingsize  = rep(0.05,16),              # initial size of seedlings
  LMA           = rep(0.035,16),             # Leaf mass per unit area: 0.035, 0.085, 0.135
  phiRL         = rep(3.5,16),               # Root:Shoot ratio:  4.0, 6.0, 8.0
  LNbase        = rep(0.8E-3,16),            # kgN m-2 leaf, Vmax = 0.03125*LNbase
  laimax        = rep(3.5,16),               # maximum crown LAI
  LAI_light     = rep(4.0,16),               # Light-limited crown LAI
  Nfixrate0     = rep(0,16),                 # 0.03 kgN kgRootC-1 yr-1
  NfixCost0     = rep(12,16),                # 12, 24 gC/gN
  phiCSA        = rep(0.25E-4,16),           # ratio of sapwood area to leaf area
  mortrate_d_c  = rep(0.01,16),              # canopy tree mortality rate, year-1
  mortrate_d_u  = rep(0.075,16),             # understory tree mortality rate, year-1
  maturalage    = rep(5,16),                 # the age that can reproduce
  fNSNmax       = rep(5,16)                  # multiplier for NSNmax as sum of potential bl and br
  ) 
```

#### Soil parameters

By layers.
```{r}
# adopted from datatypes.mod.f90 l.538
params_soil <- tibble(
  type              = c("Coarse","Medium","Fine","CM","CF","MF","CMF","Peat","MCM"),
  GMD               = c(0.7, 0.4, 0.3, 0.1, 0.1, 0.07, 0.007, 0.3, 0.3),
  GSD               = c(5.0, 5.3, 7.4, 6.1, 6.1, 14.0, 15.0, 7.4, 7.4),
  vwc_sat           = c(0.380, 0.445, 0.448, 0.412, 0.414, 0.446, 0.424, 0.445, 0.445),
  chb               = c(3.5,6.4,11.0,4.8,6.3,8.4,6.3,6.4,6.4),
  psi_sat_ref       = c(-600, -790, -910, -1580, -1680, -1880, -5980, -790, -790), # Pa
  k_sat_ref         = c(130.8, 75.1, 53.2, 12.1, 11.1, 12.7, 1.69, 53.2, 53.2), # mol/(s MPa m)
  alphaSoil         = rep(1, 9),
  heat_capacity_dry = c(1.2e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.4e6, 1.0)
  )
```

#### Initial cohort specification

```{r}
init_cohort <- tibble(
 init_cohort_species = rep(2, 10),
 init_cohort_nindivs = rep(1,10),    # initial individual density, individual/m2
 init_cohort_bsw     = rep(0.05,10), # initial biomass of sapwood, kg C/individual
 init_cohort_bHW     = rep(0.0, 10), # initial biomass of heartwood, kg C/tree
 init_cohort_nsc     = rep(0.05,10)  # initial non-structural biomass
)
```

#### Initial soil pools

high N input --> Deciduous, low  N input --> Evergreen
```{r}
init_soil <- list(
 init_fast_soil_C    = 0.0,    # initial fast soil C, kg C/m2
 init_slow_soil_C    = 0.0,    # initial slow soil C, kg C/m2
 init_Nmineral       = 0.015,  # Mineral nitrogen pool, (kg N/m2)
 N_input             = 0.0008  # annual N input to soil N pool, kgN m-2 yr-1
)
```

### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Get input

```{r}
forcing <- read_delim( paste0(path.package("rsofun"), "/extdata/ORNL_forcing.txt"), col_names = TRUE, delim = "\t") %>% 
  mutate(date = lubridate::ymd_hm( paste0( as.character(YEAR), "-01-01 00:00" ) ) + days(DOY-1) + hours(HOUR) ) %>% 
  mutate(sitename = "ORNL") %>% 
  dplyr::filter(!(mday(date)==29 & month(date)==2))

## convert rain to units per seconds
dt_secs <- lubridate::interval(forcing$date[1], forcing$date[2]) %>% 
  time_length("seconds")
forcing <- forcing %>% 
  mutate(RAIN = RAIN / dt_secs)

if (params_siml$method_photosynth == "gs_leuning"){
  forcing <- forcing %>% 
    dplyr::group_by(lubridate::month(date),lubridate::day(date),lubridate::hour(date)) %>% 
    summarise_at(vars(1:13), list(~mean(., na.rm = TRUE)))
  forcing <- forcing[,-c(1:3)]
}

```

For P-model runs, aggregate to daily forcing.
```{r}
if (params_siml$method_photosynth == "pmodel" && dt_secs != (60*60*24)){
#  forcing <- forcing %>% 
#    dplyr::group_by(lubridate::date(date)) %>% 
#    summarise_at(vars(1:13), list(~mean(., na.rm = TRUE))) %>% 
#    dplyr::select(-1)
  forcing <- forcing %>% 
    dplyr::group_by(lubridate::month(date),lubridate::day(date)) %>% 
    summarise_at(vars(1:13), list(~mean(., na.rm = TRUE)))
  forcing <- forcing[,-c(1:2)]
}

# Changing values of PAR for running simulations
#forcing <- forcing %>% mutate(PAR = PAR*1.30) # levels = *1, *1.15 and *1.30

# Replicating forcing data to run longer transient years
#forcing <- bind_rows(replicate(10, forcing, simplify = FALSE))

```

### Run the model for different cases

Run the model for all environmental conditions, species and mortality formulations.
```{r}
out <- run_lm3ppa_f_bysite( "ORNL", 
                            params_siml, 
                            siteinfo, 
                            forcing, # ddf_input
                            params_tile, 
                            params_species, 
                            params_soil, 
                            init_cohort, 
                            init_soil, 
                            makecheck = TRUE)

out$output_annual_tile %>% 
  ggplot() +
  geom_line(aes(x = year, y = GPP)) +
  theme_classic()+labs(x = "Year", y = "GPP")

out$output_annual_tile %>% 
  ggplot() +
  geom_line(aes(x = year, y = plantC)) +
  theme_classic()+labs(x = "Year", y = "plantC")

out$output_annual_tile %>% 
  ggplot() + geom_line(aes(x = year, y = CAI)) +
  theme_classic()+labs(x = "Year", y = "CAI")

out$output_annual_tile %>% 
  ggplot() + geom_point(aes(x = log(QMD), y = log(Density12))) +
  theme_classic()+labs(x = "log(QMD)", y = "log(Density12)") + scale_x_continuous(limits = c(-3,-0.5))

```

### Write model outputs

```{r}
setwd("/Users/lmarques/rsofun/Simulations_pmodel")
write.csv(ea1sa1md_out_annual_tile,"ea1sa1md_out_annual_tile.csv") # for kphio = 0.05
#write.csv(ea2sa1md_out_annual_tile,"ea2sa1md_out_annual_tile.csv") # for kphio = 0.0575
#write.csv(ea3sa1md_out_annual_tile,"ea3sa1md_out_annual_tile.csv") # for kphio = 0.065
```

### Read model outputs
```{r}
setwd("/Users/lmarques/rsofun/Simulations_pmodel")
ea1sa1md_out_annual_tile <- read.csv("ea1sa1md_out_annual_tile.csv")
ea2sa1md_out_annual_tile <- read.csv("ea2sa1md_out_annual_tile.csv")
ea3sa1md_out_annual_tile <- read.csv("ea3sa1md_out_annual_tile.csv")
```

##### Stand development
```{r}
# Changes in LUE with WD level 1
# Mortality Constant Self-thinning
ggplot() + 
  geom_line(data=out$output_annual_tile, aes(x=year, y=plantC, color='Control')) + 
  scale_color_manual("Levels of LUE", breaks = c("Control", "+15%", "+30%"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") +
  ggtitle("Mortality formulation Constant Self-thinning")
```

#### Self-thinning relathionships

From new annual_tile_output

```{r}
# Changes in LUE with WD level 1
# Mortality Constant Self-thinning
ggplot() + 
  geom_point(data=subset(out$output_annual_tile,year>=100), aes(x=log(QMD), y=log(Density12), color='Control')) +
  scale_color_manual(labels = c("Control","+15%", "+30%"),values = c("#F8766D", "#00BA38", "#619CFF"))+
  labs(col = "Level of LUE", x = "Ln(QMD)", y = "Ln(N)") + 
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") +
  ggtitle("Mortality formulation Constant Self-thinning")
```

