---
title: "Example: Calibrating the P-model for C4 photosynthesis"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(rsofun)
library(dplyr)
library(readr)
library(ingestr)
library(ggplot2)
```

## Prepare drivers

### Get FLUXNET 2015 data

This uses the R package [*ingestr*](https://stineb.github.io/ingestr/).
```{r}
siteinfo <- ingestr::siteinfo_fluxnet2015 %>%
  dplyr::filter(c4)
```

### Get meteo data

The following ingests meteorological data from the FLUXNET 2015 files for variables daytime temperature, precipitation, daytime VPD, shortwave incoming radiation, net radiation, and atmospheric pressure. Arguments that are specific for this data source are provided in the `settings` list. Unfortunately, FLUXNET 2015 doesn't provide daytime VPD. But we can derive it using the ingestr R package as described [here](https://stineb.github.io/ingestr/articles/calc_daytime_vpd.html) and done below. This writes files with daytime VPD into the directory specified by `settings_fluxnet$dir_hh`. The data object `ddf_fluxnet` is organised as a nested table with rows for sites and time series nested inside the column `data`. See [here](https://tidyr.tidyverse.org/reference/nest.html) for how to handle nested dataframes. 
```{r message=FALSE, warning=FALSE}
ddf_fluxnet <- ingestr::ingest(
  siteinfo  = siteinfo,
  source    = "fluxnet",
  getvars   = list(temp = "TA_F_DAY", prec = "P_F", vpd  = "VPD_F_DAY", ppfd =  "SW_IN_F", patm = "PA_F"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  settings  = list(dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", getswc = FALSE),
  timescale = "d"
  )
```

Some meteo data is not available from FLUXNET. Extract it from CRU global climate files instead.
```{r}
ddf_cru <- ingestr::ingest(
  siteinfo  = siteinfo,
  source    = "cru",
  getvars   = list(ccov = "cld"),
  dir       = "~/data/cru/ts_4.01/"
  )
```

Combine the two meteo data frames into one, containing `ccov` (cloud cover) from CRU and all other variables from FLUXNET.
```{r}
ddf_meteo <- ddf_fluxnet %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru %>% 
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>% 
  group_by(sitename) %>% 
  tidyr::nest()
```









Specify additional simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = FALSE,
  tempstress         = FALSE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = FALSE,
  lgn3               = FALSE,
  lgr4               = TRUE
	)
```

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters), global simulation parameters are wrapped inside an additional column `params_siml`, added to the site meta info dataframe.
```{r}
siteinfo <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```

### Define model parameters

First, let's do it by hand (calibration of parameters is shown later).
```{r}
params_modl <- list(
	kphio           = 0.05,
	soilm_par_a     = 1.0,
	soilm_par_b     = 0.0,
	vpdstress_par_a = 0.2,
	vpdstress_par_b = 0.2,
	vpdstress_par_m = 5
	)
```

### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific. Values entered here take no effect.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Get input

Now, meteorological forcing data and fAPAR data for each time step have to be collected. These steps are described in the vignette `vignettes/prepare_inputs_rsofun.Rmd` for FLUXNET 2015 simulations. All forcing data, combined with model and simulations parameters are collected into one big nested data frame `df_drivers`, which stores all required information by rows for each site. An example of how this looks like is provided as part of the rsofun package.

```{r}
rsofun::df_drivers
```

```{r}
df_drivers$params_siml[[1]]$lgr3 <- FALSE
df_drivers$params_siml[[1]]$lgr4 <- TRUE
```

### Run the model

Run the model for all the sites specified in the first step.
```{r}
## run for a single site
mod <- run_pmodel_f_bysite( 
  df_drivers$sitename[1], 
  df_drivers$params_siml[[1]], 
  df_drivers$siteinfo[[1]], 
  df_drivers$forcing[[1]], 
  df_drivers$df_soiltexture[[1]], 
  params_modl = params_modl, 
  makecheck = TRUE 
  )
```

```{r}
mod_c3 <- mod
mod_c4 <- mod

mod_c3 %>% 
  dplyr::select(date, gpp_c3 = gpp) %>% 
  left_join(
    mod_c4 %>% 
      dplyr::select(date, gpp_c4 = gpp),
    by = "date"
  ) %>% 
  ggplot(aes(gpp_c3, gpp_c4)) +
  geom_point()
```


Run for the full set of sites
```{r}
ptm <- proc.time()
df_output <- runread_pmodel_f(
     df_drivers, 
     params_modl = params_modl, 
     makecheck = TRUE,
     parallel = FALSE
     )
print(ptm)

# microbenchmark::microbenchmark(
#   runread_pmodel_f(
#     df_drivers, 
#     params_modl = params_modl, 
#     makecheck = TRUE,
#     parallel = TRUE,
#     ncores = 4
#     ),
#   runread_pmodel_f(
#     df_drivers, 
#     params_modl = params_modl, 
#     makecheck = TRUE,
#     parallel = FALSE
#     ),
#   times = 5,
#   units = 's'
# )
```

```{r}
ggplot() +
  geom_line(data = df_output$data[[1]], aes(x=date, y=gpp), color = 'black') + 
  labs(title = df_output$sitename[[1]], subtitle = "SOFUN output")
```

### Calibrate

Define calibration settings.
```{r}
settings_calib <- list(
  method              = "gensa",
  targetvars          = c("gpp"),
  timescale           = list( gpp = "d" ),
  maxit               = 5, # (5 for gensa) (30 for optimr)    #
  sitenames           = "FR-Pue",
  metric              = "rmse",
  dir_results         = "./",
  name                = "ORG",
  par                 = list( kphio = list( lower=0.02, upper=0.07, init=0.0496 ) )
 )
```

Use the [ingestr](https://github.com/stineb/ingestr) package once again, now for collecting calibration target data. I.e., GPP based on the nighttime flux decomposition method.
```{r warning=FALSE, message=FALSE}
settings_ingestr_fluxnet <- list(
  dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", 
  getswc = FALSE,
  filter_ntdt = TRUE,
  threshold_GPP = 0.8,
  remove_neg = FALSE
  )

ddf_fluxnet_gpp <- ingestr::ingest(
  siteinfo = dplyr::filter(siteinfo, sitename == "FR-Pue"),
  source    = "fluxnet",
  getvars = list(gpp = "GPP_NT_VUT_REF",
                 gpp_unc = "GPP_NT_VUT_SE"),
  dir = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  settings = settings_ingestr_fluxnet,
  timescale = "d"
  )
```

Calibrate the model.
```{r warning=FALSE, message=FALSE}
set.seed(1982)
settings_calib <- calib_sofun(
  df_drivers = dplyr::filter(df_drivers, sitename %in% settings_calib$sitenames),  # use only one site
  ddf_obs = ddf_fluxnet_gpp,
  settings = settings_calib
  )
```

The calibrated parameters are returned by `calib_sofun()` as part of the list:
```{r}
print(settings_calib$par_opt)
```


### Evaluate

Run the model once again with these parameters and evaluate results.
```{r warning=FALSE, message=FALSE}
params_modl$kphio <- settings_calib$par_opt["kphio"]

df_output <- runread_pmodel_f(
     df_drivers, 
     params_modl = params_modl, 
     makecheck = TRUE,
     parallel = FALSE
     )
```


Get evaluation data (benchmarking data).
```{r message=FALSE, warning=FALSE}
## get data for idfferent time scales separately
settings_fluxnet <- list(
  getswc = FALSE,
  filter_ntdt = TRUE,
  threshold_GPP = 0.8,
  remove_neg = FALSE
  )

ddf_eval <- ingestr::ingest(
  siteinfo  = siteinfo %>% dplyr::filter(sitename %in% mysites),
  source    = "fluxnet",
  getvars   = list(gpp = "GPP_NT_VUT_REF",
                   gpp_unc = "GPP_NT_VUT_SE"),
  dir       = paste0(path.package("ingestr"), "/extdata/"),
  settings  = settings_fluxnet,
  timescale = "d"
  )

mdf_eval <- ingestr::ingest(
  siteinfo  = siteinfo %>% dplyr::filter(sitename %in% mysites),
  source    = "fluxnet",
  getvars   = list(gpp = "GPP_NT_VUT_REF",
                   gpp_unc = "GPP_NT_VUT_SE"),
  dir       = paste0(path.package("ingestr"), "/extdata/"),
  settings  = settings_fluxnet,
  timescale = "m"
  )

adf_eval <- ingestr::ingest(
  siteinfo  = siteinfo %>% dplyr::filter(sitename %in% mysites),
  source    = "fluxnet",
  getvars   = list(gpp = "GPP_NT_VUT_REF",
                   gpp_unc = "GPP_NT_VUT_SE"),
  dir       = paste0(path.package("ingestr"), "/extdata/"),
  settings  = settings_fluxnet,
  timescale = "y"
  )
```

Use rsofun to create a standardised object used for benchmarking the model output.
```{r}
settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = mysites,
  agg       = 8  # An integer specifying the number of days used to define the width of bins for daily data aggregated to several days
  )
obs_eval <- collect_obs_eval( 
  siteinfo = siteinfo %>% dplyr::filter(sitename %in% mysites),
  settings = settings_eval, 
  adf = adf_eval, 
  mdf = mdf_eval, 
  ddf = ddf_eval 
  )
```

`obs_eval` is now a list of data frames for different temporal resolutions. The data frames have rows for sites and time series for each site nested inside the column `data`. 

`df_output` is the model output, also a data frame with rows for sites and time series for each site nested inside a column named `data`.

And finally do the evaluation.
```{r warning=FALSE, message=FALSE}
out_eval <- eval_sofun( 
  df_output, 
  settings_eval, 
  settings_sims, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

Print some results.
```{r}
out_eval$gpp$fluxnet$metrics$xdaily_pooled
```

Get the [rbeni](https://github.com/stineb/rbeni) R package for nice plotting functions that can be used with the output of `eval_sofun()`.
```{r}
library(rbeni)
gg <- out_eval$gpp$fluxnet$data$xdf %>% 
  analyse_modobs2("mod", "obs", type = "heat")
gg$gg +
  labs(title = "FR-Pue: modelled vs. observed GPP", 
       x = expression(paste("Modelled GPP (gC m"^{-2}, "d"^{-1}, ")")), 
       y = expression(paste("Observed GPP (gC m"^{-2}, "d"^{-1}, ")")))
```
