---
title: "Hypothesis testing"
output:
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(dplyr)
library(rsofun)
library(ggplot2)
```

## Example

### Simulation settings

Manually select some sites from which we're going to use the data for evaluation and calibration.
```{r}
mysites <- c("ORNL")
```

Create a site meta info table that contains all the site-specific information that is used to force site-simulations (e.g. starting year, number of simulations years, elevation, etc.). For FLUXNET2015 data, required meta info is provided by the `rsofun` package (data frame `rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015`).
```{r}
siteinfo <- data.frame(sitename="ORNL", lon = 0.0, lat = 0.0, elv = 0.0)
```

Now specify the simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
      spinup                = TRUE,
      spinupyears           = 1793,
      recycle               = 11, # 9 or 11
      firstyeartrend        = 1998,
      nyeartrend            = 11, #9 or 11
      outputhourly          = FALSE,
      outputdaily           = TRUE,
      do_U_shaped_mortality = FALSE,
      update_annualLAImax   = FALSE,
      do_closedN_run        = TRUE,
      method_photosynth     = "pmodel" # gs_leuning or pmodel
      )
```

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters).
```{r}
#settings_sims <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```

### Define model parameters

#### Tile-level parameters

```{r}
params_tile <- list(
  soiltype     = 3,
  FLDCAP       = 0.4,
  WILTPT       = 0.05,
  K1           = 2.0,  # turnover rate of fast SOM per year
  K2           = 0.1,   # 0.12 # 0.2, # 0.075 # turnover rate of slow SOM per year
  K_nitrogen   = 0.0, # 1.2 # 2.5 # rate of a year, 2.5
  etaN         = 0.0, # 0.025 # loss rate with runoff
  MLmixRatio   = 0.6, # 0.8 # fraction of microbes to litter
  l_fract      = 0.0,
  retransN     = 0.0,
  f_N_add      = 0.02, # 0.005, # 0.02
  f_initialBSW = 0.005
  # alpha_FR   =  1.2; gamma_FR = 12.0; dSlowSOM = 2.5 * iSOM + 1.5; rho_N_up0 = 0.1
  )

```

#### Species-level parameters

```{r}
params_species <- tibble(
  lifeform      = c(0,       1,        1,        1,        1,      1,        1,        1,      1,        1,        1),    # 0: grass; 1 Woody
  phenotype     = c(0,       1,        1,        1,        1,      1,        1,        1,      1,        1,        0),    # 0: Deciduous; 1 Evergreen
  pt            = c(1,       0,        0,        0,        0,      0,        0,        0,      0,        0,        0),    # 0: C3; 1: C4
  seedlingsize  = c(0.02,    0.05,     0.05,     0.05,     0.05,   0.05,     0.05,     0.05,   0.05,     0.05,     0.05),  # initial size of seedlings
  LMA           = c(0.035,   0.14,     0.14,     0.14,     0.14,   0.14,    0.14,   0.14,     0.14,     0.14,     0.035),# Leaf mass per unit area
  phiRL         = c(4.0,     1.0,      2.0,      3.0,      4.0,     5.0,     6.0,      7.0,     8.0,    9.0,      3.0),
  LNbase        = c(0.8E-3,  0.8E-3,   0.8E-3,   0.8E-3,   0.8E-3, 0.8E-3,   0.8E-3,   0.8E-3, 0.8E-3,   0.8E-3,   0.8E-3),   # kgN m-2 leaf, Vmax = 0.03125*LNbase
  laimax        = c(3.0,     3.5,      3.5,      3.5,      3.5,    3.5,      3.5,      3.5,    3.5,      3.5,      3.8),  # maximum crown LAI
  LAI_light     = c(3.0,     3.5,      3.5,      3.5,      3.5,    3.5,      3.5,      3.5,    3.5,      3.5,      3.8),  # Light-limited crown LAI
  Nfixrate0     = c(0.0,     0.0,      0.0,      0.0,      0.0,    0.0,      0.0,      0.0,    0.0,      0.0,      0.0),  # 0.03 kgN kgRootC-1 yr-1
  NfixCost0     = c(12.,     12.,      12.,      12.,      12.,    12.,      12.,      12.,    12.,      12.,      12.),  # 12, 24 gC/gN
  phiCSA        = c(1.25E-4, 0.25E-4,  0.25E-4,  0.25E-4,  0.25E-4,0.25E-4,  0.25E-4,  0.25E-4,0.25E-4,  0.25E-4,  0.25E-4),
  mortrate_d_c  = c(0.02,    0.02,     0.02,     0.02,     0.02,   0.02,     0.02,     0.02,   0.02,     0.02,     0.02),    # canopy tree mortality rate, year-1
  mortrate_d_u  = c(0.2,     0.08,     0.08,     0.08,     0.08,   0.08,     0.08,     0.08,   0.08,     0.08,     0.08),    # understory tree mortality rate, year-1
  maturalage    = c(0.5,     5,        5,        5,        5,      5,        5,        5,      5,        5,        5),
  fNSNmax       = c(5,       5,        5,        5,        5,      5,        5,        5,      5,        5,        5)
  ) %>% 
  bind_rows(., slice(., 1:5))
```

#### Soil parameters

By layers.
```{r}
# adopted from datatypes.mod.f90 l.538
params_soil <- tibble(
  type              = c("Coarse",  "Medium",   "Fine",    "CM",     "CF",     "MF",    "CMF",    "Peat",    "MCM"),
  GMD               = c(0.7, 0.4, 0.3, 0.1, 0.1, 0.07, 0.007, 0.3, 0.3),
  GSD               = c(5.0, 5.3, 7.4, 6.1, 6.1, 14.0, 15.0, 7.4, 7.4),
  vwc_sat           = c(0.380, 0.445, 0.448, 0.412, 0.414, 0.446, 0.424, 0.445, 0.445),
  chb               = c( 3.5,   6.4,  11.0,   4.8,   6.3,   8.4,   6.3,   6.4,   6.4),
  psi_sat_ref       = c(-600., -790., -910., -1580., -1680., -1880., -5980., -790., -790.), # Pa
  k_sat_ref         = c(130.8, 75.1, 53.2, 12.1, 11.1, 12.7, 1.69, 53.2, 53.2),      # mol/(s MPa m)
  alphaSoil         = rep(1.0, 9),
  heat_capacity_dry = c(1.2e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.4e6,   1.0)
  )
```

#### Initial cohort specification

```{r}
init_cohort <- tibble(
 init_cohort_species = c(1,      2,        3,        4,      5,        6,        7,      8,        9,       10),
 init_cohort_nindivs = c(0.02,   0.02,     0.02,     0.02,   0.02,     0.02,     0.02,   0.02,     0.02,    0.02),
 init_cohort_bsw     = c(0.2,    0.2,      0.2,      0.2,    0.2,      0.2,      0.2,    0.2,      0.2,     0.2),
 init_cohort_bHW     = c(0.0,    0.0,      0.0,      0.0,    0.0,      0.0,      0.0,    0.0,      0.0,     0.0),
 init_cohort_nsc     = c(0.5,    0.5,      0.5,      0.5,    0.5,      0.5,      0.5,    0.5,      0.5,     0.5)
)

```

#### Initial soil pools

```{r}
# high N input --> Deciduous
# low  N input --> Evergreen
init_soil <- list(
 init_fast_soil_C    = 0.15,   # kg C m-2    # C/N = 15, 0.75 kg SOM ~ 50 gN
 init_slow_soil_C    = 16.0,   # kg C m-2    # C/N = 40
 init_Nmineral       = 2.5E-3, # kg N m-2
 N_input             = 0.0E-3  # 2.4E-3, # kg N m-2 yr-1, N deposit: 0.8 g N m-2 yr-1
)

```

### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Get input

Get the input from the nice weird looking file with half-hourly data.

The file has the following variables (units):

 Year  DOY  Hour  T_air          Q_air          Wind_speed      Precip          Pressure        R_global_in      R_longwave_in   CO2
                      K          kg/kg          m/s     kg_H2O/m2/s               pa               W/m2            W/m2            ppm     

```{r}
forcing <- read_delim( paste0(path.package("rsofun"), "/extdata/ORNL_forcing.txt"), col_names = TRUE, delim = "\t") %>% 
  mutate(date = lubridate::ymd_hm( paste0( as.character(YEAR), "-01-01 00:00" ) ) + days(DOY-1) + hours(HOUR) ) %>% 
  mutate(sitename = "ORNL") %>% 
  dplyr::filter(!(mday(date)==29 & month(date)==2))

## convert rain to units per seconds
dt_secs <- lubridate::interval(forcing$date[1], forcing$date[2]) %>% 
  time_length("seconds")
forcing <- forcing %>% 
  mutate(RAIN = RAIN / dt_secs)

# Changing values of PAR for running simulations
#forcing <- forcing %>% mutate(Swdown = Swdown*1.15) # levels = *1, *1.15 and *1.30

```

For P-model runs, aggregate to daily forcing.
```{r}
if (params_siml$method_photosynth == "pmodel" && dt_secs != (60*60*24)){
  forcing <- forcing %>% 
    dplyr::group_by(lubridate::date(date)) %>% 
    summarise_at(vars(1:13), list(~mean(., na.rm = TRUE))) %>% 
    dplyr::select(-1)
}
```

### Run the model

Run the model for all the sites specified in the first step.
```{r}
out <- run_lm3ppa_f_bysite( "ORNL", 
                            params_siml, 
                            siteinfo, 
                            forcing, # ddf_input
                            params_tile, 
                            params_species, 
                            params_soil, 
                            init_cohort, 
                            init_soil, 
                            makecheck = TRUE)

#save(out, file = "~/rsofun/output/out_lm3ppa_pmodel.RData")
rsofun_out_hourly_tile <- out$output_hourly_tile
rsofun_out_daily_tile <- out$output_daily_tile
rsofun_out_daily_cohorts <- out$output_daily_cohorts
rsofun_out_annual_tile <- out$output_annual_tile
rsofun_out_annual_cohorts <- out$output_annual_cohorts

rsofun_out_annual_tile %>% 
  ggplot() +
  geom_line(aes(x = year, y = GPP)) +
  labs(title = "LM3-PPA with gs leuning photosynthesis at ORNL", subtitle = "rsofun output")

rsofun_out_annual_tile %>% 
  ggplot() +
  geom_line(aes(x = year, y = Rauto)) +
  labs(title = "LM3-PPA with gs leuning photosynthesis at ORNL", subtitle = "rsofun output")

```

Run the model for with the p-model photosynthesis.
```{r}
out <- run_lm3ppa_f_bysite( "ORNL", 
                            params_siml, 
                            siteinfo, 
                            forcing, # ddf_input
                            params_tile, 
                            params_species, 
                            params_soil, 
                            init_cohort, 
                            init_soil, 
                            makecheck = TRUE)

#save(out, file = "~/rsofun/output/out_lm3ppa_pmodel.RData")
pmodel_out_hourly_tile <- out$output_hourly_tile
pmodel_out_daily_tile <- out$output_daily_tile
pmodel_out_daily_cohorts <- out$output_daily_cohorts
pmodel_out_annual_tile <- out$output_annual_tile
pmodel_out_annual_cohorts <- out$output_annual_cohorts

pmodel_out_annual_tile %>% 
  ggplot() +
  geom_line(aes(x = year, y = GPP)) +
  labs(title = "LM3-PPA with p-model photosynthesis at ORNL", subtitle = "rsofun output")

pmodel_out_annual_tile %>% 
  ggplot() +
  geom_line(aes(x = year, y = Rauto)) +
  labs(title = "LM3-PPA with p-model photosynthesis at ORNL", subtitle = "rsofun output")

```

### Estimating a mean value for LUE

LUE = Annual sum GPP / Annual sum PAR
```{r}

Annual_sum_GPP <- rsofun_out_annual_tile  %>% filter (year >= 1794)  %>% summarise_at(c("GPP"), sum, na.rm = TRUE)
Annual_sum_PAR <- forcing %>% summarise_at(c("PAR"), sum, na.rm = TRUE)

LUE <- Annual_sum_GPP/Annual_sum_PAR

```

See that this LUE is a lower value and produces much higher biomass than previous. We consider 1e-08 as LUE level 1.

### Run the model

Run the model for all environmental conditions, species and mortality formulations.
```{r}
out <- run_lm3ppa_f_bysite( "ORNL", 
                            params_siml, 
                            siteinfo, 
                            forcing, # ddf_input
                            params_tile, 
                            params_species, 
                            params_soil, 
                            init_cohort, 
                            init_soil, 
                            makecheck = TRUE)

eb3sb3ma_out_hourly_tile <- out$output_hourly_tile
eb3sb3ma_out_daily_tile <- out$output_daily_tile
eb3sb3ma_out_daily_cohorts <- out$output_daily_cohorts
eb3sb3ma_out_annual_tile <- out$output_annual_tile
eb3sb3ma_out_annual_cohorts <- out$output_annual_cohorts

out$output_annual_tile %>% 
  ggplot() +
  geom_line(aes(x = year, y = GPP)) +
  labs(title = "LM3-PPA with gs leuning photosynthesis at ORNL", subtitle = "rsofun output")

out$output_daily_tile %>% 
  mutate(date = lubridate::ymd(paste0(year, "-01-01")) + lubridate::days(doy-1)) %>% 
  ggplot(aes(x=date, y=GPP)) +
  geom_line() + 
  labs(title = "LM3-PPA with gs leuning photosynthesis at ORNL", subtitle = "rsofun output")

```

Same models for conditions by default
```{r}

ea1sb1ma_out_hourly_tile <- ea1sa1ma_out_hourly_tile
ea1sb1ma_out_daily_tile <- ea1sa1ma_out_daily_tile
ea1sb1ma_out_daily_cohorts <- ea1sa1ma_out_daily_cohorts
ea1sb1ma_out_annual_tile <- ea1sa1ma_out_annual_tile
ea1sb1ma_out_annual_cohorts <- ea1sa1ma_out_annual_cohorts

ea2sb1ma_out_hourly_tile <- ea2sa1ma_out_hourly_tile
ea2sb1ma_out_daily_tile <- ea2sa1ma_out_daily_tile
ea2sb1ma_out_daily_cohorts <- ea2sa1ma_out_daily_cohorts
ea2sb1ma_out_annual_tile <- ea2sa1ma_out_annual_tile
ea2sb1ma_out_annual_cohorts <- ea2sa1ma_out_annual_cohorts

ea3sb1ma_out_hourly_tile <- ea3sa1ma_out_hourly_tile
ea3sb1ma_out_daily_tile <- ea3sa1ma_out_daily_tile
ea3sb1ma_out_daily_cohorts <- ea3sa1ma_out_daily_cohorts
ea3sb1ma_out_annual_tile <- ea3sa1ma_out_annual_tile
ea3sb1ma_out_annual_cohorts <- ea3sa1ma_out_annual_cohorts

eb1sa1ma_out_hourly_tile <- ea1sa1ma_out_hourly_tile
eb1sa1ma_out_daily_tile <- ea1sa1ma_out_daily_tile
eb1sa1ma_out_daily_cohorts <- ea1sa1ma_out_daily_cohorts
eb1sa1ma_out_annual_tile <- ea1sa1ma_out_annual_tile
eb1sa1ma_out_annual_cohorts <- ea1sa1ma_out_annual_cohorts

eb1sb1ma_out_hourly_tile <- ea1sa1ma_out_hourly_tile
eb1sb1ma_out_daily_tile <- ea1sa1ma_out_daily_tile
eb1sb1ma_out_daily_cohorts <- ea1sa1ma_out_daily_cohorts
eb1sb1ma_out_annual_tile <- ea1sa1ma_out_annual_tile
eb1sb1ma_out_annual_cohorts <- ea1sa1ma_out_annual_cohorts

eb1sa2ma_out_hourly_tile <- ea1sa2ma_out_hourly_tile
eb1sa2ma_out_daily_tile <- ea1sa2ma_out_daily_tile
eb1sa2ma_out_daily_cohorts <- ea1sa2ma_out_daily_cohorts
eb1sa2ma_out_annual_tile <- ea1sa2ma_out_annual_tile
eb1sa2ma_out_annual_cohorts <- ea1sa2ma_out_annual_cohorts

eb1sa3ma_out_hourly_tile <- ea1sa3ma_out_hourly_tile
eb1sa3ma_out_daily_tile <- ea1sa3ma_out_daily_tile
eb1sa3ma_out_daily_cohorts <- ea1sa3ma_out_daily_cohorts
eb1sa3ma_out_annual_tile <- ea1sa3ma_out_annual_tile
eb1sa3ma_out_annual_cohorts <- ea1sa3ma_out_annual_cohorts

eb1sb2ma_out_hourly_tile <- ea1sb2ma_out_hourly_tile
eb1sb2ma_out_daily_tile <- ea1sb2ma_out_daily_tile
eb1sb2ma_out_daily_cohorts <- ea1sb2ma_out_daily_cohorts
eb1sb2ma_out_annual_tile <- ea1sb2ma_out_annual_tile
eb1sb2ma_out_annual_cohorts <- ea1sb2ma_out_annual_cohorts

eb1sb3ma_out_hourly_tile <- ea1sb3ma_out_hourly_tile
eb1sb3ma_out_daily_tile <- ea1sb3ma_out_daily_tile
eb1sb3ma_out_daily_cohorts <- ea1sb3ma_out_daily_cohorts
eb1sb3ma_out_annual_tile <- ea1sb3ma_out_annual_tile
eb1sb3ma_out_annual_cohorts <- ea1sb3ma_out_annual_cohorts

eb2sb1ma_out_hourly_tile <- eb2sa1ma_out_hourly_tile
eb2sb1ma_out_daily_tile <- eb2sa1ma_out_daily_tile
eb2sb1ma_out_daily_cohorts <- eb2sa1ma_out_daily_cohorts
eb2sb1ma_out_annual_tile <- eb2sa1ma_out_annual_tile
eb2sb1ma_out_annual_cohorts <- eb2sa1ma_out_annual_cohorts

eb3sb1ma_out_hourly_tile <- eb3sa1ma_out_hourly_tile
eb3sb1ma_out_daily_tile <- eb3sa1ma_out_daily_tile
eb3sb1ma_out_daily_cohorts <- eb3sa1ma_out_daily_cohorts
eb3sb1ma_out_annual_tile <- eb3sa1ma_out_annual_tile
eb3sb1ma_out_annual_cohorts <- eb3sa1ma_out_annual_cohorts

```


### Aggregating cohorts across years and cohorts

```{r}

rsofun_out_annual_cohorts <- rsofun_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))

ea1sa1ma_out_annual_cohorts <- ea1sa1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea1sa2ma_out_annual_cohorts <- ea1sa2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea1sa3ma_out_annual_cohorts <- ea1sa3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea2sa1ma_out_annual_cohorts <- ea2sa1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea2sa2ma_out_annual_cohorts <- ea2sa2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea2sa3ma_out_annual_cohorts <- ea2sa3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea3sa1ma_out_annual_cohorts <- ea3sa1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea3sa2ma_out_annual_cohorts <- ea3sa2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea3sa3ma_out_annual_cohorts <- ea3sa3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))

ea1sb1ma_out_annual_cohorts <- ea1sb1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea1sb2ma_out_annual_cohorts <- ea1sb2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea1sb3ma_out_annual_cohorts <- ea1sb3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea2sb1ma_out_annual_cohorts <- ea2sb1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea2sb2ma_out_annual_cohorts <- ea2sb2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea2sb3ma_out_annual_cohorts <- ea2sb3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea3sb1ma_out_annual_cohorts <- ea3sb1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea3sb2ma_out_annual_cohorts <- ea3sb2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
ea3sb3ma_out_annual_cohorts <- ea3sb3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))

eb1sa1ma_out_annual_cohorts <- eb1sa1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb1sa2ma_out_annual_cohorts <- eb1sa2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb1sa3ma_out_annual_cohorts <- eb1sa3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb2sa1ma_out_annual_cohorts <- eb2sa1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb2sa2ma_out_annual_cohorts <- eb2sa2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb2sa3ma_out_annual_cohorts <- eb2sa3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb3sa1ma_out_annual_cohorts <- eb3sa1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb3sa2ma_out_annual_cohorts <- eb3sa2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb3sa3ma_out_annual_cohorts <- eb3sa3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))

eb1sb1ma_out_annual_cohorts <- eb1sb1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density)%>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb1sb2ma_out_annual_cohorts <- eb1sb2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb1sb3ma_out_annual_cohorts <- eb1sb3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb2sb1ma_out_annual_cohorts <- eb2sb1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb2sb2ma_out_annual_cohorts <- eb2sb2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb2sb3ma_out_annual_cohorts <- eb2sb3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb3sb1ma_out_annual_cohorts <- eb3sb1ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb3sb2ma_out_annual_cohorts <- eb3sb2ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))
eb3sb3ma_out_annual_cohorts <- eb3sb3ma_out_annual_cohorts %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) %>% mutate(basalArea = pi*(dbh/2)*density) %>% group_by(year) %>% mutate(highBA= map_dbl(basalArea,~sum(basalArea[basalArea>=.x]))-basalArea) %>%
  ungroup %>% mutate(highBAclass = cut(highBA ,breaks = seq(0,600,100)))

rsofun_out_annual_cohortsMax <- rsofun_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)

ea1sa1ma_out_annual_cohortsMax <- ea1sa1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge) 
ea1sa2ma_out_annual_cohortsMax <- ea1sa2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea1sa3ma_out_annual_cohortsMax <- ea1sa3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea2sa1ma_out_annual_cohortsMax <- ea2sa1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea2sa2ma_out_annual_cohortsMax <- ea2sa2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea2sa3ma_out_annual_cohortsMax <- ea2sa3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea3sa1ma_out_annual_cohortsMax <- ea3sa1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea3sa2ma_out_annual_cohortsMax <- ea3sa2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea3sa3ma_out_annual_cohortsMax <- ea3sa3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)

ea1sb1ma_out_annual_cohortsMax <- ea1sb1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea1sb2ma_out_annual_cohortsMax <- ea1sb2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea1sb3ma_out_annual_cohortsMax <- ea1sb3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea2sb1ma_out_annual_cohortsMax <- ea2sb1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea2sb2ma_out_annual_cohortsMax <- ea2sb2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea2sb3ma_out_annual_cohortsMax <- ea2sb3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea3sb1ma_out_annual_cohortsMax <- ea3sb1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea3sb2ma_out_annual_cohortsMax <- ea3sb2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
ea3sb3ma_out_annual_cohortsMax <- ea3sb3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)

eb1sa1ma_out_annual_cohortsMax <- eb1sa1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb1sa2ma_out_annual_cohortsMax <- eb1sa2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb1sa3ma_out_annual_cohortsMax <- eb1sa3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb2sa1ma_out_annual_cohortsMax <- eb2sa1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb2sa2ma_out_annual_cohortsMax <- eb2sa2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb2sa3ma_out_annual_cohortsMax <- eb2sa3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb3sa1ma_out_annual_cohortsMax <- eb3sa1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb3sa2ma_out_annual_cohortsMax <- eb3sa2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb3sa3ma_out_annual_cohortsMax <- eb3sa3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)

eb1sb1ma_out_annual_cohortsMax <- eb1sb1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb1sb2ma_out_annual_cohortsMax <- eb1sb2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb1sb3ma_out_annual_cohortsMax <- eb1sb3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb2sb1ma_out_annual_cohortsMax <- eb2sb1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb2sb2ma_out_annual_cohortsMax <- eb2sb2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb2sb3ma_out_annual_cohortsMax <- eb2sb3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb3sb1ma_out_annual_cohortsMax <- eb3sb1ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb3sb2ma_out_annual_cohortsMax <- eb3sb2ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)
eb3sb3ma_out_annual_cohortsMax <- eb3sb3ma_out_annual_cohorts %>% group_by(year) %>% summarize(maxAge = max(age), DBHmaxAge = dbh[which.max(age)]) %>% mutate(GrowthRate = DBHmaxAge/maxAge)

rsofun_out_annual_cohortsMaxAgg <- tibble(rsofun_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,rsofun_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))

ea1sa1ma_out_annual_cohortsMaxAgg <- tibble(ea1sa1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea1sa1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea1sa2ma_out_annual_cohortsMaxAgg <- tibble(ea1sa2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea1sa2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea1sa3ma_out_annual_cohortsMaxAgg <- tibble(ea1sa3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea1sa3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea2sa1ma_out_annual_cohortsMaxAgg <- tibble(ea2sa1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea2sa1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea2sa2ma_out_annual_cohortsMaxAgg <- tibble(ea2sa2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea2sa2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea2sa3ma_out_annual_cohortsMaxAgg <- tibble(ea2sa3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea2sa3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea3sa1ma_out_annual_cohortsMaxAgg <- tibble(ea3sa1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea3sa1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea3sa2ma_out_annual_cohortsMaxAgg <- tibble(ea3sa2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea3sa2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea3sa3ma_out_annual_cohortsMaxAgg <- tibble(ea3sa3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea3sa3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))

ea1sb1ma_out_annual_cohortsMaxAgg <- tibble(ea1sb1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea1sb1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea1sb2ma_out_annual_cohortsMaxAgg <- tibble(ea1sb2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea1sb2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea1sb3ma_out_annual_cohortsMaxAgg <- tibble(ea1sb3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea1sb3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea2sb1ma_out_annual_cohortsMaxAgg <- tibble(ea2sb1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea2sb1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea2sb2ma_out_annual_cohortsMaxAgg <- tibble(ea2sb2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea2sb2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea2sb3ma_out_annual_cohortsMaxAgg <- tibble(ea2sb3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea2sb3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea3sb1ma_out_annual_cohortsMaxAgg <- tibble(ea3sb1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea3sb1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea3sb2ma_out_annual_cohortsMaxAgg <- tibble(ea3sb2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea3sb2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
ea3sb3ma_out_annual_cohortsMaxAgg <- tibble(ea3sb3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,ea3sb3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))

eb1sa1ma_out_annual_cohortsMaxAgg <- tibble(eb1sa1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb1sa1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb1sa2ma_out_annual_cohortsMaxAgg <- tibble(eb1sa2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb1sa2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb1sa3ma_out_annual_cohortsMaxAgg <- tibble(eb1sa3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb1sa3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb2sa1ma_out_annual_cohortsMaxAgg <- tibble(eb2sa1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb2sa1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb2sa2ma_out_annual_cohortsMaxAgg <- tibble(eb2sa2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb2sa2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb2sa3ma_out_annual_cohortsMaxAgg <- tibble(eb2sa3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb2sa3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb3sa1ma_out_annual_cohortsMaxAgg <- tibble(eb3sa1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb3sa1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb3sa2ma_out_annual_cohortsMaxAgg <- tibble(eb3sa2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb3sa2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb3sa3ma_out_annual_cohortsMaxAgg <- tibble(eb3sa3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb3sa3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))

eb1sb1ma_out_annual_cohortsMaxAgg <- tibble(eb1sb1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb1sb1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb1sb2ma_out_annual_cohortsMaxAgg <- tibble(eb1sb2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb1sb2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb1sb3ma_out_annual_cohortsMaxAgg <- tibble(eb1sb3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb1sb3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb2sb1ma_out_annual_cohortsMaxAgg <- tibble(eb2sb1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb2sb1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb2sb2ma_out_annual_cohortsMaxAgg <- tibble(eb2sb2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb2sb2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb2sb3ma_out_annual_cohortsMaxAgg <- tibble(eb2sb3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb2sb3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb3sb1ma_out_annual_cohortsMaxAgg <- tibble(eb3sb1ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb3sb1ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb3sb2ma_out_annual_cohortsMaxAgg <- tibble(eb3sb2ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb3sb2ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))
eb3sb3ma_out_annual_cohortsMaxAgg <- tibble(eb3sb3ma_out_annual_cohortsMax %>% summarise(maxAge_mean=mean(maxAge,na.rm=T), sdmaxAge=sd(maxAge,na.rm=T)) ,eb3sb3ma_out_annual_cohortsMax %>% summarise(meanGR=mean(GrowthRate,na.rm=T), sdGR=sd(GrowthRate,na.rm=T)))

#rsofun_out_annual_cohorts <- rsofun_out_annual_cohorts %>% group_by(cohort) %>% mutate(DeltaWood = ifelse(year>1794, wood-lag(wood), NA)) %>% mutate(Deltadbh = ifelse(year>1794, dbh-lag(dbh), NA)) %>% mutate(Mortality = ifelse(year>1794, lag(density)-density, NA)) %>% mutate(Mortality = ifelse(Mortality<0,0,Mortality)) %>% mutate(Longevity=1/Mortality) %>% mutate(Longevity = ifelse(Longevity=="Inf",NA,Longevity)) %>% mutate(dbhclass = cut(dbh ,breaks = seq(0,3,0.5))) 

#rsofun_out_annual_cohortsSum <- rsofun_out_annual_cohorts[c(2,6:29)] %>% group_by(year) %>% summarise_all(funs(mean, sum),na.rm=T) 

```

### Nesting data from simulations

Collect data into a tidy data frame with simulations in rows, "treatment"  and wrapped data (time series of variables) in columns.
```{r}

simul_data <- tibble(
  type = c(rep("annual_tile",10),rep("annual_cohorts",10),rep("annual_cohorts_Max",10),rep("annual_cohorts_MaxAgg",10),
           rep("annual_tile",9),rep("annual_cohorts",9),rep("annual_cohorts_Max",9),rep("annual_cohorts_MaxAgg",9),
           rep("annual_tile",9),rep("annual_cohorts",9),rep("annual_cohorts_Max",9),rep("annual_cohorts_MaxAgg",9),
           rep("annual_tile",9),rep("annual_cohorts",9),rep("annual_cohorts_Max",9),rep("annual_cohorts_MaxAgg",9)),
  LUE = c(0,1,1,1,2,2,2,3,3,3,0,1,1,1,2,2,2,3,3,3,0,1,1,1,2,2,2,3,3,3,0,1,1,1,2,2,2,3,3,3,
          1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,rep(1,72)),
  GS = c(rep(1,148)),
  PAR = c(rep(1,76),
          1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,
          1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3,1,1,1,2,2,2,3,3,3),
  rho_wood = c(1,1,2,3,1,2,3,1,2,3,1,1,2,3,1,2,3,1,2,3,1,1,2,3,1,2,3,1,2,3,1,1,2,3,1,2,3,1,2,3,rep(1,36),
               1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3, rep(1,36)),
  phiRL = c(rep(1,40),1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,
            rep(1,36),1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3),
  Mortality = c(rep("A",148)),
  simuls = list(
    rsofun_out_annual_tile,
    ea1sa1ma_out_annual_tile,
    ea1sa2ma_out_annual_tile,
    ea1sa3ma_out_annual_tile,
    ea2sa1ma_out_annual_tile,
    ea2sa2ma_out_annual_tile,
    ea2sa3ma_out_annual_tile,
    ea3sa1ma_out_annual_tile,
    ea3sa2ma_out_annual_tile,
    ea3sa3ma_out_annual_tile,
    rsofun_out_annual_cohorts,
    ea1sa1ma_out_annual_cohorts,
    ea1sa2ma_out_annual_cohorts,
    ea1sa3ma_out_annual_cohorts,
    ea2sa1ma_out_annual_cohorts,
    ea2sa2ma_out_annual_cohorts,
    ea2sa3ma_out_annual_cohorts,
    ea3sa1ma_out_annual_cohorts,
    ea3sa2ma_out_annual_cohorts,
    ea3sa3ma_out_annual_cohorts,
    rsofun_out_annual_cohortsMax,
    ea1sa1ma_out_annual_cohortsMax,
    ea1sa2ma_out_annual_cohortsMax,
    ea1sa3ma_out_annual_cohortsMax,
    ea2sa1ma_out_annual_cohortsMax,
    ea2sa2ma_out_annual_cohortsMax,
    ea2sa3ma_out_annual_cohortsMax,
    ea3sa1ma_out_annual_cohortsMax,
    ea3sa2ma_out_annual_cohortsMax,
    ea3sa3ma_out_annual_cohortsMax,
    rsofun_out_annual_cohortsMaxAgg,
    ea1sa1ma_out_annual_cohortsMaxAgg,
    ea1sa2ma_out_annual_cohortsMaxAgg,
    ea1sa3ma_out_annual_cohortsMaxAgg,
    ea2sa1ma_out_annual_cohortsMaxAgg,
    ea2sa2ma_out_annual_cohortsMaxAgg,
    ea2sa3ma_out_annual_cohortsMaxAgg,
    ea3sa1ma_out_annual_cohortsMaxAgg,
    ea3sa2ma_out_annual_cohortsMaxAgg,
    ea3sa3ma_out_annual_cohortsMaxAgg,

    ea1sb1ma_out_annual_tile,
    ea1sb2ma_out_annual_tile,
    ea1sb3ma_out_annual_tile,
    ea2sb1ma_out_annual_tile,
    ea2sb2ma_out_annual_tile,
    ea2sb3ma_out_annual_tile,
    ea3sb1ma_out_annual_tile,
    ea3sb2ma_out_annual_tile,
    ea3sb3ma_out_annual_tile,
    ea1sb1ma_out_annual_cohorts,
    ea1sb2ma_out_annual_cohorts,
    ea1sb3ma_out_annual_cohorts,
    ea2sb1ma_out_annual_cohorts,
    ea2sb2ma_out_annual_cohorts,
    ea2sb3ma_out_annual_cohorts,
    ea3sb1ma_out_annual_cohorts,
    ea3sb2ma_out_annual_cohorts,
    ea3sb3ma_out_annual_cohorts,
    ea1sb1ma_out_annual_cohortsMax,
    ea1sb2ma_out_annual_cohortsMax,
    ea1sb3ma_out_annual_cohortsMax,
    ea2sb1ma_out_annual_cohortsMax,
    ea2sb2ma_out_annual_cohortsMax,
    ea2sb3ma_out_annual_cohortsMax,
    ea3sb1ma_out_annual_cohortsMax,
    ea3sb2ma_out_annual_cohortsMax,
    ea3sb3ma_out_annual_cohortsMax,
    ea1sb1ma_out_annual_cohortsMaxAgg,
    ea1sb2ma_out_annual_cohortsMaxAgg,
    ea1sb3ma_out_annual_cohortsMaxAgg,
    ea2sb1ma_out_annual_cohortsMaxAgg,
    ea2sb2ma_out_annual_cohortsMaxAgg,
    ea2sb3ma_out_annual_cohortsMaxAgg,
    ea3sb1ma_out_annual_cohortsMaxAgg,
    ea3sb2ma_out_annual_cohortsMaxAgg,
    ea3sb3ma_out_annual_cohortsMaxAgg,
    
    eb1sa1ma_out_annual_tile,
    eb1sa2ma_out_annual_tile,
    eb1sa3ma_out_annual_tile,
    eb2sa1ma_out_annual_tile,
    eb2sa2ma_out_annual_tile,
    eb2sa3ma_out_annual_tile,
    eb3sa1ma_out_annual_tile,
    eb3sa2ma_out_annual_tile,
    eb3sa3ma_out_annual_tile,
    eb1sa1ma_out_annual_cohorts,
    eb1sa2ma_out_annual_cohorts,
    eb1sa3ma_out_annual_cohorts,
    eb2sa1ma_out_annual_cohorts,
    eb2sa2ma_out_annual_cohorts,
    eb2sa3ma_out_annual_cohorts,
    eb3sa1ma_out_annual_cohorts,
    eb3sa2ma_out_annual_cohorts,
    eb3sa3ma_out_annual_cohorts,
    eb1sa1ma_out_annual_cohortsMax,
    eb1sa2ma_out_annual_cohortsMax,
    eb1sa3ma_out_annual_cohortsMax,
    eb2sa1ma_out_annual_cohortsMax,
    eb2sa2ma_out_annual_cohortsMax,
    eb2sa3ma_out_annual_cohortsMax,
    eb3sa1ma_out_annual_cohortsMax,
    eb3sa2ma_out_annual_cohortsMax,
    eb3sa3ma_out_annual_cohortsMax,
    eb1sa1ma_out_annual_cohortsMaxAgg,
    eb1sa2ma_out_annual_cohortsMaxAgg,
    eb1sa3ma_out_annual_cohortsMaxAgg,
    eb2sa1ma_out_annual_cohortsMaxAgg,
    eb2sa2ma_out_annual_cohortsMaxAgg,
    eb2sa3ma_out_annual_cohortsMaxAgg,
    eb3sa1ma_out_annual_cohortsMaxAgg,
    eb3sa2ma_out_annual_cohortsMaxAgg,
    eb3sa3ma_out_annual_cohortsMaxAgg,
    
    eb1sb1ma_out_annual_tile,
    eb1sb2ma_out_annual_tile,
    eb1sb3ma_out_annual_tile,
    eb2sb1ma_out_annual_tile,
    eb2sb2ma_out_annual_tile,
    eb2sb3ma_out_annual_tile,
    eb3sb1ma_out_annual_tile,
    eb3sb2ma_out_annual_tile,
    eb3sb3ma_out_annual_tile,
    eb1sb1ma_out_annual_cohorts,
    eb1sb2ma_out_annual_cohorts,
    eb1sb3ma_out_annual_cohorts,
    eb2sb1ma_out_annual_cohorts,
    eb2sb2ma_out_annual_cohorts,
    eb2sb3ma_out_annual_cohorts,
    eb3sb1ma_out_annual_cohorts,
    eb3sb2ma_out_annual_cohorts,
    eb3sb3ma_out_annual_cohorts,
    eb1sb1ma_out_annual_cohortsMax,
    eb1sb2ma_out_annual_cohortsMax,
    eb1sb3ma_out_annual_cohortsMax,
    eb2sb1ma_out_annual_cohortsMax,
    eb2sb2ma_out_annual_cohortsMax,
    eb2sb3ma_out_annual_cohortsMax,
    eb3sb1ma_out_annual_cohortsMax,
    eb3sb2ma_out_annual_cohortsMax,
    eb3sb3ma_out_annual_cohortsMax,
    eb1sb1ma_out_annual_cohortsMaxAgg,
    eb1sb2ma_out_annual_cohortsMaxAgg,
    eb1sb3ma_out_annual_cohortsMaxAgg,
    eb2sb1ma_out_annual_cohortsMaxAgg,
    eb2sb2ma_out_annual_cohortsMaxAgg,
    eb2sb3ma_out_annual_cohortsMaxAgg,
    eb3sb1ma_out_annual_cohortsMaxAgg,
    eb3sb2ma_out_annual_cohortsMaxAgg,
    eb3sb3ma_out_annual_cohortsMaxAgg
  )
)

```

### Exploring relationships and plotting

#### Stand develop: Stand biomass vs. time

From annual_tile_output: PlanC (Biomass)

##### LUE (mol/s/m2)
```{r}

library(scales)
hue_pal()(3)
show_col(hue_pal()(3))

# Changes in LUE with WD level 1
ggLUE1sa1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[2]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[5]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[8]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in LUE with WD level 2
ggLUE1sa2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[3]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[6]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[9]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE at WD 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in LUE with WD level 3
ggLUE1sa3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[4]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[7]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[10]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE at WD 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in LUE with root:shoot level 1
ggLUE1sb1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[41]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[44]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[47]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in LUE with root:shoot level 2
ggLUE1sb2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[42]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[45]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[48]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE at R:S 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in LUE with root:shoot level 3
ggLUE1sb3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[43]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[46]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[49]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE at R:S 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 


# Additional plot for this: B* vs. initial slope of (B vs. t) 
# Changes in LUE with WD level 1

ggplot() + 
  geom_line(data=simul_data$simuls[[2]][1:150,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[5]][1:150,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[8]][1:150,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

ggplot() + 
  geom_line(data=simul_data$simuls[[2]][1300:1804,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[5]][1300:1804,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[8]][1300:1804,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

slope1 <- summary(lm(plantC ~ year, simul_data$simuls[[2]][1:150,]))$coefficients[2,1]
sd1 <- summary(lm(plantC ~ year, simul_data$simuls[[2]][1:150,]))$coefficients[2,2]
slope2 <- summary(lm(plantC ~ year, simul_data$simuls[[5]][1:150,]))$coefficients[2,1]
sd2 <- summary(lm(plantC ~ year, simul_data$simuls[[5]][1:150,]))$coefficients[2,2]
slope3 <- summary(lm(plantC ~ year, simul_data$simuls[[8]][1:150,]))$coefficients[2,1]
sd3 <- summary(lm(plantC ~ year, simul_data$simuls[[8]][1:150,]))$coefficients[2,2]

mean1 <- simul_data$simuls[[2]][1300:1804,] %>% summarise(mean1=mean(plantC,na.rm=T), sd1=sd(plantC,na.rm=T))
mean2 <- simul_data$simuls[[5]][1300:1804,] %>% summarise(mean2=mean(plantC,na.rm=T), sd2=sd(plantC,na.rm=T))
mean3 <- simul_data$simuls[[8]][1300:1804,] %>% summarise(mean3=mean(plantC,na.rm=T), sd3=sd(plantC,na.rm=T))

slopes <- c(slope1,slope2,slope3)
sdSlopes <- c(sd1,sd2,sd3)
meanBstar <- c(mean1$mean1,mean2$mean2, mean3$mean3)
sdBstar <- c(mean1$sd1,mean2$sd2, mean3$sd3)
level <- as.factor(c("Low","Medium","High"))
dataSlope <- data.frame(level,slopes,sdSlopes,meanBstar,sdBstar)

ggplot(data=dataSlope) +
  geom_point(aes(x=slopes, y=meanBstar,color=level)) +
  geom_errorbar(aes(x = slopes, ymin=meanBstar-sdBstar, ymax=meanBstar+sdBstar, color=level),width=0) + 
  scale_color_manual("Levels of LUE at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Slope of B ~ t", y = expression(paste("Mean biomass at maturity (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Additional plot for this: B* vs. initial slope of (B vs. t) 
# Changes in LUE with R:S level 1

ggplot() + 
  geom_line(data=simul_data$simuls[[41]][1:150,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[44]][1:150,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[47]][1:150,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

ggplot() + 
  geom_line(data=simul_data$simuls[[41]][1300:1804,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[44]][1300:1804,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[47]][1300:1804,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

slope1 <- summary(lm(plantC ~ year, simul_data$simuls[[41]][1:150,]))$coefficients[2,1]
sd1 <- summary(lm(plantC ~ year, simul_data$simuls[[41]][1:150,]))$coefficients[2,2]
slope2 <- summary(lm(plantC ~ year, simul_data$simuls[[44]][1:150,]))$coefficients[2,1]
sd2 <- summary(lm(plantC ~ year, simul_data$simuls[[44]][1:150,]))$coefficients[2,2]
slope3 <- summary(lm(plantC ~ year, simul_data$simuls[[47]][1:150,]))$coefficients[2,1]
sd3 <- summary(lm(plantC ~ year, simul_data$simuls[[47]][1:150,]))$coefficients[2,2]

mean1 <- simul_data$simuls[[41]][1300:1804,] %>% summarise(mean1=mean(plantC,na.rm=T), sd1=sd(plantC,na.rm=T))
mean2 <- simul_data$simuls[[44]][1300:1804,] %>% summarise(mean2=mean(plantC,na.rm=T), sd2=sd(plantC,na.rm=T))
mean3 <- simul_data$simuls[[47]][1300:1804,] %>% summarise(mean3=mean(plantC,na.rm=T), sd3=sd(plantC,na.rm=T))

slopes <- c(slope1,slope2,slope3)
sdSlopes <- c(sd1,sd2,sd3)
meanBstar <- c(mean1$mean1,mean2$mean2, mean3$mean3)
sdBstar <- c(mean1$sd1,mean2$sd2, mean3$sd3)
level <- as.factor(c("Low","Medium","High"))
dataSlope <- data.frame(level,slopes,sdSlopes,meanBstar,sdBstar)

ggplot(data=dataSlope) +
  geom_point(aes(x=slopes, y=meanBstar,color=level)) +
  geom_errorbar(aes(x = slopes, ymin=meanBstar-sdBstar, ymax=meanBstar+sdBstar, color=level),width=0) + 
  scale_color_manual("Levels of LUE at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Slope of B ~ t", y = expression(paste("Mean biomass at maturity (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 
```

##### GS
```{r}

```

##### Radiation
```{r}

# Changes in Radiation with WD level 1
ggRad1sa1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[77]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[80]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[83]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Radiation at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in Radiation with WD level 2
ggRad1sa2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[78]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[81]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[84]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Radiation at WD 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in Radiation with WD level 3
ggRad1sa3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[79]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[82]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[85]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Radiation at WD 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in Radiation with root:shoot level 1
ggRad1sb1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[113]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[116]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[119]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Radiation at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in Radiation with root:shoot level 2
ggLRad1sb2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[114]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[117]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[120]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Radiation at R:S 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in Radiation with root:shoot level 3
ggLUE1sb3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[115]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[118]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[121]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Radiation at R:S 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 


# Additional plot for this: B* vs. initial slope of (B vs. t) 
# Changes in Radiation with WD level 1

ggplot() + 
  geom_line(data=simul_data$simuls[[77]][1:150,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[80]][1:150,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[83]][1:150,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

ggplot() + 
  geom_line(data=simul_data$simuls[[77]][1300:1804,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[80]][1300:1804,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[83]][1300:1804,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

slope1 <- summary(lm(plantC ~ year, simul_data$simuls[[77]][1:150,]))$coefficients[2,1]
sd1 <- summary(lm(plantC ~ year, simul_data$simuls[[77]][1:150,]))$coefficients[2,2]
slope2 <- summary(lm(plantC ~ year, simul_data$simuls[[80]][1:150,]))$coefficients[2,1]
sd2 <- summary(lm(plantC ~ year, simul_data$simuls[[80]][1:150,]))$coefficients[2,2]
slope3 <- summary(lm(plantC ~ year, simul_data$simuls[[83]][1:150,]))$coefficients[2,1]
sd3 <- summary(lm(plantC ~ year, simul_data$simuls[[83]][1:150,]))$coefficients[2,2]

mean1 <- simul_data$simuls[[77]][1300:1804,] %>% summarise(mean1=mean(plantC,na.rm=T), sd1=sd(plantC,na.rm=T))
mean2 <- simul_data$simuls[[80]][1300:1804,] %>% summarise(mean2=mean(plantC,na.rm=T), sd2=sd(plantC,na.rm=T))
mean3 <- simul_data$simuls[[83]][1300:1804,] %>% summarise(mean3=mean(plantC,na.rm=T), sd3=sd(plantC,na.rm=T))

slopes <- c(slope1,slope2,slope3)
sdSlopes <- c(sd1,sd2,sd3)
meanBstar <- c(mean1$mean1,mean2$mean2, mean3$mean3)
sdBstar <- c(mean1$sd1,mean2$sd2, mean3$sd3)
level <- as.factor(c("Low","Medium","High"))
dataSlope <- data.frame(level,slopes,sdSlopes,meanBstar,sdBstar)

ggplot(data=dataSlope) +
  geom_point(aes(x=slopes, y=meanBstar,color=level)) +
  geom_errorbar(aes(x = slopes, ymin=meanBstar-sdBstar, ymax=meanBstar+sdBstar, color=level),width=0) + 
  scale_color_manual("Levels of Radiation at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Slope of B ~ t", y = expression(paste("Mean biomass at maturity (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Additional plot for this: B* vs. initial slope of (B vs. t) 
# Changes in Radiation with R:S level 1

ggplot() + 
  geom_line(data=simul_data$simuls[[113]][1:150,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[116]][1:150,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[119]][1:150,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

ggplot() + 
  geom_line(data=simul_data$simuls[[113]][1300:1804,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[116]][1300:1804,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[119]][1300:1804,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

slope1 <- summary(lm(plantC ~ year, simul_data$simuls[[113]][1:150,]))$coefficients[2,1]
sd1 <- summary(lm(plantC ~ year, simul_data$simuls[[113]][1:150,]))$coefficients[2,2]
slope2 <- summary(lm(plantC ~ year, simul_data$simuls[[116]][1:150,]))$coefficients[2,1]
sd2 <- summary(lm(plantC ~ year, simul_data$simuls[[116]][1:150,]))$coefficients[2,2]
slope3 <- summary(lm(plantC ~ year, simul_data$simuls[[119]][1:150,]))$coefficients[2,1]
sd3 <- summary(lm(plantC ~ year, simul_data$simuls[[119]][1:150,]))$coefficients[2,2]

mean1 <- simul_data$simuls[[113]][1300:1804,] %>% summarise(mean1=mean(plantC,na.rm=T), sd1=sd(plantC,na.rm=T))
mean2 <- simul_data$simuls[[116]][1300:1804,] %>% summarise(mean2=mean(plantC,na.rm=T), sd2=sd(plantC,na.rm=T))
mean3 <- simul_data$simuls[[119]][1300:1804,] %>% summarise(mean3=mean(plantC,na.rm=T), sd3=sd(plantC,na.rm=T))

slopes <- c(slope1,slope2,slope3)
sdSlopes <- c(sd1,sd2,sd3)
meanBstar <- c(mean1$mean1,mean2$mean2, mean3$mean3)
sdBstar <- c(mean1$sd1,mean2$sd2, mean3$sd3)
level <- as.factor(c("Low","Medium","High"))
dataSlope <- data.frame(level,slopes,sdSlopes,meanBstar,sdBstar)

ggplot(data=dataSlope) +
  geom_point(aes(x=slopes, y=meanBstar,color=level)) +
  geom_errorbar(aes(x = slopes, ymin=meanBstar-sdBstar, ymax=meanBstar+sdBstar, color=level),width=0) + 
  scale_color_manual("Levels of Radiation at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Slope of B ~ t", y = expression(paste("Mean biomass at maturity (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

```

##### Wood density
```{r}

# Changes in WD with LUE level 1
ggWD1ea1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[2]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[3]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[4]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Wood density at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in WD with LUE level 2
ggWD1ea2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[5]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[6]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[7]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Wood density at LUE 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in WD with LUE level 3
ggWD1ea3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[8]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[9]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[10]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Wood density at LUE 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in WD with Rad level 1
ggWD1eb1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[77]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[78]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[79]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Wood density at Rad 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in WD with Rad level 2
ggWD1eb2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[80]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[81]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[82]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Wood density at Rad 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in WD with Rad level 3
ggWD1eb3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[83]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[84]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[85]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Wood density at Rad 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 


# Additional plot for this: B* vs. initial slope of (B vs. t) 

ggplot() + 
  geom_line(data=simul_data$simuls[[2]][1:150,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[3]][1:150,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[4]][1:150,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

ggplot() + 
  geom_line(data=simul_data$simuls[[2]][1300:1804,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[3]][1300:1804,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[4]][1300:1804,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

slope1 <- summary(lm(plantC ~ year, simul_data$simuls[[2]][1:150,]))$coefficients[2,1]
sd1 <- summary(lm(plantC ~ year, simul_data$simuls[[2]][1:150,]))$coefficients[2,2]
slope2 <- summary(lm(plantC ~ year, simul_data$simuls[[3]][1:150,]))$coefficients[2,1]
sd2 <- summary(lm(plantC ~ year, simul_data$simuls[[3]][1:150,]))$coefficients[2,2]
slope3 <- summary(lm(plantC ~ year, simul_data$simuls[[4]][1:150,]))$coefficients[2,1]
sd3 <- summary(lm(plantC ~ year, simul_data$simuls[[4]][1:150,]))$coefficients[2,2]

mean1 <- simul_data$simuls[[2]][1300:1804,] %>% summarise(mean1=mean(plantC,na.rm=T), sd1=sd(plantC,na.rm=T))
mean2 <- simul_data$simuls[[3]][1300:1804,] %>% summarise(mean2=mean(plantC,na.rm=T), sd2=sd(plantC,na.rm=T))
mean3 <- simul_data$simuls[[4]][1300:1804,] %>% summarise(mean3=mean(plantC,na.rm=T), sd3=sd(plantC,na.rm=T))

slopes <- c(slope1,slope2,slope3)
sdSlopes <- c(sd1,sd2,sd3)
meanBstar <- c(mean1$mean1,mean2$mean2, mean3$mean3)
sdBstar <- c(mean1$sd1,mean2$sd2, mean3$sd3)
level <- as.factor(c("Low","Medium","High"))
dataSlope <- data.frame(level,slopes,sdSlopes,meanBstar,sdBstar)

ggplot(data=dataSlope) +
  geom_point(aes(x=slopes, y=meanBstar,color=level)) +
  geom_errorbar(aes(x = slopes, ymin=meanBstar-sdBstar, ymax=meanBstar+sdBstar, color=level),width=0) + 
  scale_color_manual("Levels of Wood density at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Slope of B ~ t", y = expression(paste("Mean biomass at maturity (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

```

##### Root:Shoot ratio
```{r}

# Changes in R:S with LUE level 1
ggRS1ea1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[41]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[42]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[43]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in R:S with LUE level 2
ggRS1ea2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[44]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[45]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[46]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at LUE 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in R:S with LUE level 3
ggRS1ea3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[47]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[48]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[49]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at LUE 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in R:S with Rad level 1
ggRS1eb1 <- ggplot() + 
  geom_line(data=simul_data$simuls[[113]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[114]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[115]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at Rad 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in R:S with Rad level 2
ggRS1eb2 <- ggplot() + 
  geom_line(data=simul_data$simuls[[116]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[117]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[118]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at Rad 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

# Changes in R:S with Rad level 3
ggRS1eb3 <- ggplot() + 
  geom_line(data=simul_data$simuls[[119]], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[120]], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[121]], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at Rad 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 


# Additional plot for this: B* vs. initial slope of (B vs. t) 

ggplot() + 
  geom_line(data=simul_data$simuls[[41]][1:150,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[42]][1:150,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[43]][1:150,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

ggplot() + 
  geom_line(data=simul_data$simuls[[41]][1300:1804,], aes(x=year, y=plantC, color='Low')) + 
  geom_line(data=simul_data$simuls[[42]][1300:1804,], aes(x=year, y=plantC, color='Medium')) +
  geom_line(data=simul_data$simuls[[43]][1300:1804,], aes(x=year, y=plantC, color='High')) +
  scale_color_manual("Levels of Root:Shoot ratio at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Year", y = expression(paste("Biomass (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

slope1 <- summary(lm(plantC ~ year, simul_data$simuls[[41]][1:150,]))$coefficients[2,1]
sd1 <- summary(lm(plantC ~ year, simul_data$simuls[[41]][1:150,]))$coefficients[2,2]
slope2 <- summary(lm(plantC ~ year, simul_data$simuls[[42]][1:150,]))$coefficients[2,1]
sd2 <- summary(lm(plantC ~ year, simul_data$simuls[[42]][1:150,]))$coefficients[2,2]
slope3 <- summary(lm(plantC ~ year, simul_data$simuls[[43]][1:150,]))$coefficients[2,1]
sd3 <- summary(lm(plantC ~ year, simul_data$simuls[[43]][1:150,]))$coefficients[2,2]

mean1 <- simul_data$simuls[[41]][1300:1804,] %>% summarise(mean1=mean(plantC,na.rm=T), sd1=sd(plantC,na.rm=T))
mean2 <- simul_data$simuls[[42]][1300:1804,] %>% summarise(mean2=mean(plantC,na.rm=T), sd2=sd(plantC,na.rm=T))
mean3 <- simul_data$simuls[[43]][1300:1804,] %>% summarise(mean3=mean(plantC,na.rm=T), sd3=sd(plantC,na.rm=T))

slopes <- c(slope1,slope2,slope3)
sdSlopes <- c(sd1,sd2,sd3)
meanBstar <- c(mean1$mean1,mean2$mean2, mean3$mean3)
sdBstar <- c(mean1$sd1,mean2$sd2, mean3$sd3)
level <- as.factor(c("Low","Medium","High"))
dataSlope <- data.frame(level,slopes,sdSlopes,meanBstar,sdBstar)

ggplot(data=dataSlope) +
  geom_point(aes(x=slopes, y=meanBstar,color=level)) +
  geom_errorbar(aes(x = slopes, ymin=meanBstar-sdBstar, ymax=meanBstar+sdBstar, color=level),width=0) + 
  scale_color_manual("Levels of Root:Shoot ratio at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Slope of B ~ t", y = expression(paste("Mean biomass at maturity (Kg C ", m^-2, " ", yr^-1, ") "))) + 
  theme_classic() 

```


##### Stand biomass Plots
```{r}

library(ggpubr)
ggarrange(ggLUE1sa1,ggLUE1sa2,ggLUE1sa3, ncol = 1, nrow = 3, common.legend = T)

```

#### Self-thinning relathionships

From new annual_tile_output

##### LUE
```{r}

# Changes in LUE with WD level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[2]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[5]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[8]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of LUE at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[2]][1000:1804,], aes(x=QMD, y=Density12, color='Low')) + 
  geom_point(data=simul_data$simuls[[5]][1000:1804,], aes(x=QMD, y=Density12, color='Medium')) +
  geom_point(data=simul_data$simuls[[8]][1000:1804,], aes(x=QMD, y=Density12, color='High')) +
  scale_color_manual("Levels of LUE", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Quadratic Mean Diameter (m)", y = expression(paste("Density (trees ", ha^-1, ") "))) + 
  theme_classic() 

# Changes in LUE with WD level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[3]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[6]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[9]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of LUE at WD 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in LUE with WD level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[4]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[7]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[10]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of LUE at WD 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in LUE with R:S level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[41]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[44]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[47]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of LUE at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[41]][1000:1804,], aes(x=QMD, y=Density12, color='Low')) + 
  geom_point(data=simul_data$simuls[[44]][1000:1804,], aes(x=QMD, y=Density12, color='Medium')) +
  geom_point(data=simul_data$simuls[[47]][1000:1804,], aes(x=QMD, y=Density12, color='High')) +
  scale_color_manual("Levels of LUE at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Quadratic Mean Diameter (m)", y = expression(paste("Density (trees ", ha^-1, ") "))) + 
  theme_classic() 

# Changes in LUE with WD level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[42]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[45]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[48]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of LUE at R:S 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in LUE with WD level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[43]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[46]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[49]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of LUE at R:S 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

```

##### GS
```{r}



```

##### PAR
```{r}

# Changes in Radiation with WD level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[77]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[80]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[83]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of Radiation at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[77]][1000:1804,], aes(x=QMD, y=Density12, color='Low')) + 
  geom_point(data=simul_data$simuls[[80]][1000:1804,], aes(x=QMD, y=Density12, color='Medium')) +
  geom_point(data=simul_data$simuls[[83]][1000:1804,], aes(x=QMD, y=Density12, color='High')) +
  scale_color_manual("Levels of Radiation at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Quadratic Mean Diameter (m)", y = expression(paste("Density (trees ", ha^-1, ") "))) + 
  theme_classic() 

# Changes in Radiation with WD level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[78]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[81]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[84]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of Radiation at WD 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in Radiation with WD level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[79]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[82]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[85]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of Radiation at WD 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in Radiation with R:S level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[113]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[116]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[119]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of Radiation at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[113]][1000:1804,], aes(x=QMD, y=Density12, color='Low')) + 
  geom_point(data=simul_data$simuls[[116]][1000:1804,], aes(x=QMD, y=Density12, color='Medium')) +
  geom_point(data=simul_data$simuls[[119]][1000:1804,], aes(x=QMD, y=Density12, color='High')) +
  scale_color_manual("Levels of Radiation at R:S 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Quadratic Mean Diameter (m)", y = expression(paste("Density (trees ", ha^-1, ") "))) + 
  theme_classic() 

# Changes in Radiation with WD level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[114]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[117]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[120]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of Radiation at R:S 2", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in Radiation with WD level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[115]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[118]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[121]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of Radiation at R:S 3", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

```

##### WD
```{r}

# Changes in WD with LUE level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[2]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[3]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[4]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of WD at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[2]][1000:1804,], aes(x=QMD, y=Density12, color='Low')) + 
  geom_point(data=simul_data$simuls[[3]][1000:1804,], aes(x=QMD, y=Density12, color='Medium')) +
  geom_point(data=simul_data$simuls[[4]][1000:1804,], aes(x=QMD, y=Density12, color='High')) +
  scale_color_manual("Levels of WD", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Quadratic Mean Diameter (m)", y = expression(paste("Density (trees ", ha^-1, ") "))) + 
  theme_classic() 

# Changes in WD with LUE level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[5]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[6]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[7]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of WD at LUE 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in WD with LUE level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[8]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[9]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[10]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of WD at LUE 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in WD with Rad level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[77]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[78]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[79]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of WD at Rad 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in WD with Rad level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[80]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[81]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[82]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of WD at Rad 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in WD with Rad level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[83]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[84]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[85]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of WD at Rad 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 
```

##### Root:Shoot ratio
```{r}
# Changes in R:S with LUE level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[41]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[42]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[43]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of R:S at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[41]][1000:1804,], aes(x=QMD, y=Density12, color='Low')) + 
  geom_point(data=simul_data$simuls[[42]][1000:1804,], aes(x=QMD, y=Density12, color='Medium')) +
  geom_point(data=simul_data$simuls[[43]][1000:1804,], aes(x=QMD, y=Density12, color='High')) +
  scale_color_manual("Levels of R:S", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Quadratic Mean Diameter (m)", y = expression(paste("Density (trees ", ha^-1, ") "))) + 
  theme_classic() 

# Changes in R:S with LUE level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[44]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[45]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[46]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of R:S at LUE 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in R:S with LUE level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[47]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[48]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[49]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of R:S at LUE 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in R:S with Rad level 1
ggplot() + 
  geom_point(data=simul_data$simuls[[113]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[114]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[115]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of R:S at Rad 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in R:S with Rad level 2
ggplot() + 
  geom_point(data=simul_data$simuls[[116]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[117]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[118]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of R:S at Rad 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

# Changes in R:S with Rad level 3
ggplot() + 
  geom_point(data=simul_data$simuls[[119]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Low')) + 
  geom_point(data=simul_data$simuls[[120]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='Medium')) +
  geom_point(data=simul_data$simuls[[121]][1000:1804,], aes(x=log(QMD), y=log(Density12), color='High')) +
  scale_color_manual("Levels of R:S at Rad 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "Ln(QMD)", y = "Ln(N)") + 
  theme_classic() 

```

##### Self-thinning Plots
```{r}

library(ggpubr)
ggarrange(ggLUE2,ggGS2,ggPAR2, ncol = 3, nrow = 1, common.legend = T)
ggarrange(ggWD2,ggRSR2,ggVCMAX2, ncol = 3, nrow = 1, common.legend = T)

```


#### Size bin (x, biomass, volume) vs. Growth (y, volume or mass increment per year); data = individual-level

From annual_cohort_output: dbh class NPP_yr

##### LUE
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[12]], aes(x=dbhclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[15]], aes(x=dbhclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[18]], aes(x=dbhclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of LUE at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "DBH class (m)", y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 


# Plot relative enhancement: NPP(level 2) / NPP(level 1) [and NPP(level 3) / NPP(level 1)] per size bin

RelEnan2_1 <- simul_data$simuls[[15]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr
RelEnan3_1 <- simul_data$simuls[[18]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr

```

##### GS
```{r}


```

##### PAR
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[86]], aes(x=dbhclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[89]], aes(x=dbhclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[92]], aes(x=dbhclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of Radiation at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = "DBH class (m)", y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 


# Plot relative enhancement: NPP(level 2) / NPP(level 1) [and NPP(level 3) / NPP(level 1)] per size bin

RelEnan2_1 <- simul_data$simuls[[50]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr
RelEnan3_1 <- simul_data$simuls[[53]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr
```

##### WD
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[12]], aes(x=dbhclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[13]], aes(x=dbhclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[14]], aes(x=dbhclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of WD at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "DBH class (m)", y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 


# Plot relative enhancement: NPP(level 2) / NPP(level 1) [and NPP(level 3) / NPP(level 1)] per size bin

RelEnan2_1 <- simul_data$simuls[[15]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr
RelEnan3_1 <- simul_data$simuls[[18]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr

```

##### R:S Ratio
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[50]], aes(x=dbhclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[51]], aes(x=dbhclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[52]], aes(x=dbhclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of Rhoot:Shoot ratio at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = "DBH class (m)", y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 

# Plot relative enhancement: NPP(level 2) / NPP(level 1) [and NPP(level 3) / NPP(level 1)] per size bin

RelEnan2_1 <- simul_data$simuls[[50]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr
RelEnan3_1 <- simul_data$simuls[[51]]$NPP_yr /simul_data$simuls[[12]]$NPP_yr

```


##### Size bin plots
```{r}

library(ggpubr)
ggarrange(ggLUE3,ggGS3,ggPAR3, ncol = 3, nrow = 1, common.legend = T)
ggarrange(ggWD3,ggRSR3,ggVCMAX3, ncol = 3, nrow = 1, common.legend = T)

```

#### Light bin (x, sum of basal area of trees with higher basal area than current one) vs. Growth (y, volume or mass increment per year); data = individual-level

From annual_cohort_output: dbh and NPP_yr

##### LUE
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[12]], aes(x=highBAclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[15]], aes(x=highBAclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[18]], aes(x=highBAclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of LUE at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = expression(paste("Sum BA of higher trees (", m^2, " " ,ha^-1, ") ")), y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 

```

##### PAR
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[86]], aes(x=highBAclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[89]], aes(x=highBAclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[92]], aes(x=highBAclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of Radiation at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = expression(paste("Sum BA of higher trees (", m^2, " " ,ha^-1, ") ")), y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 

```

##### WD
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[12]], aes(x=highBAclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[13]], aes(x=highBAclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[14]], aes(x=highBAclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of WD at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Sum BA of higher trees (", m^2, " " ,ha^-1, ") ")), y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 

```

##### R:S
```{r}

ggplot() + 
  geom_boxplot(data=simul_data$simuls[[50]], aes(x=highBAclass, y=NPP_yr, color='Low')) + 
  geom_boxplot(data=simul_data$simuls[[51]], aes(x=highBAclass, y=NPP_yr, color='Medium')) +
  geom_boxplot(data=simul_data$simuls[[52]], aes(x=highBAclass, y=NPP_yr, color='High')) +
  scale_color_manual("Levels of Rhoot:Shoot ratio at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Sum BA of higher trees (", m^2, " " ,ha^-1, ") ")), y = expression(paste("NPP (Kg C ", m^-2, " " ,yr^-1, ") "))) + 
  theme_classic() 

```
#### Longevity vs. growth rates

Hypothesis: The negative relationship between longevity vs. growth rates arises from variations between species, but disappears when looking at variations between sites.

Two plots:

- Different species at a given site: x = growth rate, y = longevity (or 1/mortality). Each point is one species (differing by wood density, root:shoot ratio). Add info either by using color ~site (given LUE, growing season length, PAR); or by combining data from multiple sites and normalising growth rates and longevity by site-level mean.
- Same species at different sites (differing by LUE, growing season length, PAR): x = growth rate, y = longevity. Each point is one site (show mean and standard deviations with x and y errorbars)


##### LUE

Todo:

- longevity is maximum age at tile-level, plot mean across multiple years after spinup
- plot change in DBH instead of NPP. Take the oldest cohort, and calculate DBH/age

##### LUE with Age max and dbh for oldest cohort (age max)
##### Same species at different sites (differing by LUE, growing season length, PAR)

```{r}

ggplot() + 
  geom_point(data=simul_data$simuls[[32]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[32]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[32]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[35]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[35]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[35]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[38]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[38]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[38]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of LUE at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ea1sa1ma_out_annual_cohortsMaxAgg
ea2sa1ma_out_annual_cohortsMaxAgg
ea3sa1ma_out_annual_cohortsMaxAgg

```

##### Radiation with Age max
##### Same species at different sites (differing by LUE, growing season length, PAR)

```{r}

ggplot() + 
  geom_point(data=simul_data$simuls[[104]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[104]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[104]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[107]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[107]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[107]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[110]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[110]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[110]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of Radiation at WD 1", breaks = c("Low", "Medium", "High"), values = c("#F8766D" ,"#00BA38", "#619CFF"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

eb1sa1ma_out_annual_cohortsMaxAgg
eb2sa1ma_out_annual_cohortsMaxAgg
eb3sa1ma_out_annual_cohortsMaxAgg
```

##### Wood density with Age max 
##### Different species at a given site: Each point is one species (differing by wood density, root:shoot ratio). Add info either by using color ~site (given LUE, growing season length, PAR).

```{r}

ggplot() + 
  geom_point(data=simul_data$simuls[[32]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[32]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[32]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[33]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[33]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[33]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[34]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[34]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[34]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[35]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[35]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[35]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[36]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[36]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[36]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[37]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[37]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[37]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at LUE 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[38]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[38]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[38]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[39]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[39]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[39]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[40]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[40]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[40]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at LUE 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[104]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[104]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[104]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[105]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[105]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[105]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[106]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[106]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[106]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at Rad 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[107]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[107]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[107]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[108]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[108]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[108]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[109]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[109]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[109]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at Rad 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[110]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[110]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[110]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[111]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[111]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[111]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[112]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[112]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[112]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at Rad 3", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

eb1sa1ma_out_annual_cohortsMaxAgg
eb1sa2ma_out_annual_cohortsMaxAgg
ea2sa3ma_out_annual_cohortsMaxAgg

```

##### Rhoot:Shoot ratio with Age max 
##### Different species at a given site: Each point is one species (differing by wood density, root:shoot ratio). Add info either by using color ~site (given LUE, growing season length, PAR).

```{r}

ggplot() + 
  geom_point(data=simul_data$simuls[[68]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[68]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[68]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[69]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[69]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[69]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[70]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[70]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[70]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at LUE 1", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[71]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[71]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[71]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[72]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[72]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[72]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[73]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[73]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[73]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at LUE 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ggplot() + 
  geom_point(data=simul_data$simuls[[74]], aes(x=meanGR, y=maxAge_mean, color='Low')) + 
  geom_errorbar(data=simul_data$simuls[[74]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Low'),width=0)+
    geom_errorbar(data=simul_data$simuls[[74]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Low'),width=0)+
  geom_point(data=simul_data$simuls[[75]], aes(x=meanGR, y=maxAge_mean, color='Medium')) + 
  geom_errorbar(data=simul_data$simuls[[75]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax= maxAge_mean+sdmaxAge,color='Medium'),width=0) +
  geom_errorbar(data=simul_data$simuls[[75]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='Medium'),width=0)+
geom_point(data=simul_data$simuls[[76]], aes(x=meanGR, y=maxAge_mean, color='High')) + 
  geom_errorbar(data=simul_data$simuls[[76]], aes(x=meanGR, ymin=maxAge_mean-sdmaxAge,ymax=maxAge_mean+sdmaxAge,color='High'),width=0) +
  geom_errorbar(data=simul_data$simuls[[76]], aes(xmin=meanGR-sdGR,xmax= meanGR+sdGR , y=maxAge_mean,color='High'),width=0)+
  scale_color_manual("Levels of WD at LUE 2", breaks = c("Low", "Medium", "High"), values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x = expression(paste("Growth rate (DBH ",age^-1,  ") ")), y = "Longevity") +  
  theme_classic() 

ea1sb1ma_out_annual_cohortsMaxAgg
ea3sb2ma_out_annual_cohortsMaxAgg
ea3sb3ma_out_annual_cohortsMaxAgg

```
