---
title: "calibrate_corrections"
author: "Beni Stocker"
date: "`r Sys.Date()`"
# output:
#   html_document:
#     toc: true
#     toc_float: true
#     toc_depth: 4
#     number_sections: true
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
header-includes:
   - \usepackage{amsmath}
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(captioner)
tab_nums <- captioner( prefix = "Table S", auto_space=FALSE, style = "i" )
fig_nums <- captioner( prefix = "Figure S", auto_space=FALSE, style = "i" )
library(dplyr)
library(readr)
load("siteinfo.Rdata")
load("settings_sims.Rdata")
load("settings_eval.Rdata")
source("prepare_setup_sofun.R")
source("prepare_input_sofun.R")
source("calib_sofun.R")
source("runread_sofun.R")
source("update_params.R")
source("eval_sofun.R")
setup_sofun <- list( 
  model      = "pmodel",
  dir        = "~/sofun/trunk",
  do_compile = FALSE,
  simsuite   = FALSE
  )
```

In the first step, the calibration of the quantum yield efficiency was done excluding data where air temperature was low ($<15^{\circ}$C) and where soil moisture was low ($<0.6$ relative soil water content). This left apparent model-data mismatch, of which two are particularly apparent:

- Positive bias in the early and late season at temperate and boreal sites. A hypothesis for explaining this is an underestimation of negative effects on photosynthesis by low temperature.
- Positive bias during the dry season and during drought periods in general at relatively dry sites. As we showed in previous work, this is due to an underestimation of the soil moisture effect.

These observations motivate the formulation of additional temperature and soil moisture stress functions. We follow an empirical approach here and use two sets of sites to investigate these two apparent biases. First, we determine empirically the functional relationships from the data, and then define and calibrate empirical stress functions which are to be applied as multipliers to simulated GPP by the P-model.

<!-- # Correcting the temperature bias

Calibration is done for a subset of sites, selected "by hand".
```{r}
calibsites <- c( "BE-Lon", "CA-Man", "CA-NS1", "CA-NS3", "CA-NS6", "CA-Qfo", 
                 "CH-Lae", "DE-Geb", "DE-Gri", "DE-Hai", "DK-Sor", "FI-Hyy", 
                 "FR-Fon", "IT-Col", "NL-Hor", "NL-Loo", "RU-Fyo", "US-Ha1", 
                 "US-MMS", "US-UMB", "US-WCr" )
```

```{r, message=FALSE, include=FALSE}
## determined by hand for temperature ramp calibration
calibsites <- c( "BE-Lon", "CA-Man", "CA-NS1", "CA-NS3", "CA-NS6", "CA-Qfo", "CH-Lae", "DE-Geb", "DE-Gri", "DE-Hai", 
                 "DK-Sor", "FI-Hyy", "FR-Fon", "IT-Col", "NL-Hor", "NL-Loo", "RU-Fyo", "US-Ha1", "US-MMS", "US-UMB", "US-WCr" )

settings_calib <- list(
  name             = "calib_temp",
  par              = list( kphio          = list( lower=0.01, upper=0.12, init=0.05 ),
                           temp_ramp_edge = list( lower=0.01, upper=20, init=5 ) ),
  method           = "optimr",
  targetvars       = c("gpp"),
  datasource       = list( gpp = "fluxnet2015" ),
  timescale        = list( gpp = "d" ),
  path_fluxnet2015 = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_gepisat     = "~/data/gepisat/v3_fluxnet2015/daily_gpp/",
  maxit            = 30,
  sitenames        = calibsites,
  filter_temp_min  = NA,
  filter_temp_max  = NA,
  filter_soilm_min = NA,
  filter_drought   = FALSE 
 )

settings_sims <- prepare_setup_sofun( 
  settings = settings_sims, 
  settings_calib = settings_calib, 
  write_paramfils = FALSE 
  )

## overwrite
settings_sims$sitenames <- settings_calib$sitenames
settings_sims$path_output_nc <- "~/sofun/output_nc_fluxnet2015_sofun/s22/"

## calibrate parameters
source("calib_sofun.R")
settings_calib <- calib_sofun(
  setup = setup_sofun,
  settings_calib = settings_calib,
  settings_sims = settings_sims,
  overwrite = TRUE
  )

## overwrite
params_opt <- read_csv( paste0("params_opt_", settings_calib$name,".csv") )
nothing <- update_params( params_opt, settings = settings_sims )
rm("mod")
mod <- runread_sofun( 
  settings = settings_sims, 
  setup = setup_sofun 
  )

## Evaluate outputs
settings_eval$sitenames <- settings_calib$sitenames
settings_eval$sitenames_siteplots <- settings_calib$sitenames

# Generate data used for plotting and get performance metrics
filn <- "out_eval.Rdata"
overwrite <- TRUE
if (!file.exists(filn)|overwrite){
  out_eval <- eval_sofun( mod, settings_eval, settings_sims, siteinfo, doplot=FALSE, overwrite = TRUE )
  save( out_eval, file = filn )
} else {
  load( filn )
}

```


```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
plot_by_doy_allsites( out_eval$data$meandoydf_stats, makepdf=FALSE )
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
plot_by_doy_allzones( out_eval$data$meandoydf_byclim_stats, makepdf=FALSE )
```
 -->

# Correcting the soil moisture bias

Calibration is done for the subset of sites where Stocker et al. (2018) identified clear drought impacts.

```{r, message=FALSE, include=FALSE}
## Use only sites for calibration for which ANN method by Stocker et al. (2018) worked fine,
## and exclude sites where C4 vegetation is present.
calibsites <- read_csv( "~/data/flue/flue_stocker18nphyt.csv" ) %>%
              dplyr::filter( cluster %in% c("cGR", "cDD") ) %>%
              dplyr::select( site ) %>%
              distinct() %>%

              ## additionally exclude C4
              dplyr::filter( !(site %in% dplyr::filter(siteinfo$light, c4 )$mysitename) ) %>%

              dplyr::select( site ) %>%
              unlist()


settings_calib <- list(
  name             = "calib_soilm",
  par              = list( kphio       = list( lower=0.01, upper=0.12, init=0.05 ),
                           soilm_par_a = list( lower=0.0, upper=0.5, init=0.2 ),
                           soilm_par_b = list( lower=0.0, upper=2.0, init=0.5 ) ),
  method           = "optimr",
  targetvars       = c("gpp"),
  datasource       = list( gpp = "fluxnet2015" ),
  timescale        = list( gpp = "d" ),
  path_fluxnet2015 = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_gepisat     = "~/data/gepisat/v3_fluxnet2015/daily_gpp/",
  maxit            = 30,
  sitenames        = calibsites,
  filter_temp_min  = NA,
  filter_temp_max  = NA,
  filter_soilm_min = NA,
  filter_drought   = TRUE 
 )

## complement simulation settings
settings_sims <- prepare_setup_sofun( 
  settings = settings_sims, 
  settings_calib = settings_calib, 
  write_paramfils = FALSE 
  )

## overwrite set of sites for reduced selection
settings_sims$sitenames <- settings_calib$sitenames

## write output to a separate folder
settings_sims$path_output_nc <- "~/sofun/output_nc_fluxnet2015_sofun/s23/"

## calibrate parameters
settings_calib <- calib_sofun(
  setup = setup_sofun,
  settings_calib = settings_calib,
  settings_sims = settings_sims,
  overwrite = TRUE
  )

## update parameters for site-level simulations, run and read sofun
params_opt <- read_csv( paste0("params_opt_", settings_calib$name,".csv") )
nothing <- update_params( params_opt, settings = settings_sims )
rm("mod")
mod <- runread_sofun( 
  settings = settings_sims, 
  setup = setup_sofun 
  )

## Evaluate outputs
settings_eval$sitenames <- settings_calib$sitenames
settings_eval$sitenames_siteplots <- settings_calib$sitenames
filn <- "out_eval.Rdata"
overwrite <- TRUE
if (!file.exists(filn)|overwrite){
  out_eval <- eval_sofun( mod, settings_eval, settings_sims, siteinfo, doplot=FALSE, overwrite = TRUE )
  save( out_eval, file = filn )
} else {
  load( filn )
}

```


```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
plot_by_doy_allsites( out_eval$data$meandoydf_stats, makepdf=FALSE )
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
plot_by_doy_allzones( out_eval$data$meandoydf_byclim_stats, makepdf=FALSE )
```
