---
title: "eval_sofun"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: true
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
#     number_sections: true
header-includes:
   - \usepackage{amsmath}
bibliography: bibliography.bib
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(captioner)
tab_nums <- captioner( prefix = "Table S", auto_space=FALSE, style = "i" )
fig_nums <- captioner( prefix = "Figure S", auto_space=FALSE, style = "i" )
```

# Evaluation of GPP against FLUXNET data

## Sites selection

Sites used for the model evaluation are shown on the map below. An overview table is in the Appendix (last section of this document).

```{r siteoverview_fig, echo=FALSE, warning=FALSE, message=FALSE}
require(ncdf4, quietly = TRUE)
ncfiln <- "../data/greve/ep_over_p_cru_ncep.nc"
if (!file.exists(ncfiln)) {
  epop <- array( 1, dim=c(720,360) )
} else {
  nc <- nc_open( ncfiln )
  epop <- ncvar_get( nc, varid="EP_OVER_P_CRU_NCEP" )
}
source("plot_map_siteoverview.R")
suppressMessages( plot_map_siteoverview( siteinfo, 1/epop ) ) # , plotfiln="fig/map_sites.pdf"
cap_siteoverview_fig <- fig_nums( "siteoverview_fig", caption=" Geographical distribution of sites selected for the bias evaluation. Sites listed in Table S1 as group 1 are in green, sites of group 2 are in black. The color of land area represents aridity, quantified as the ratio of potential evapotranspiration over precipitation from [@greve14]" )
```

`r fig_nums("siteoverview_fig")`

## Data processing

Daily data are used from the FLUXNET 2015 Tier 1 dataset, downloaded on 13. November, 2016. We use GPP as the mean of values based on the Nighttime Partitioning and the Daytime Partitioning method, both based on the Variable U-Star Threshold method, named `GPP_NT_VUT_REF` and `GPP_DT_VUT_REF`. In the FLUXNET 2015 dataset, daily values are sums over half-hourly data. We use only daily values where less than 50% of respective half-hourly data is gap-filled. We further removed data points where the daytime and nighttime methods (`GPP_DT_VUT_REF` and `GPP_NT_VUT_REF`, resp.) are inconsistent. I.e., the upper and lower 2.5% quantile of the difference between each method's GPP quantification. Finally, we removed all negative daily GPP values.

This is implemented by the following function:
```{r eval=FALSE}
clean_fluxnet_gpp <- function( gpp_nt, gpp_dt, qflag_nt, qflag_dt, cutoff=0.80 ){
  ##--------------------------------------------------------------------
  ## Cleans daily data.
  ## gpp_nt: based on nighttime flux decomposition ("NT")
  ## gpp_dt: based on daytime flux decomposition ("DT")
  ##--------------------------------------------------------------------

  ## Remove data points that are based on too much gap-filled data in the underlying half-hourly data
  gpp_nt[ which(qflag_nt < cutoff) ] <- NA  ## based on fraction of data based on gap-filled half-hourly
  gpp_dt[ which(qflag_dt < cutoff) ] <- NA  ## based on fraction of data based on gap-filled half-hourly

  ## Remove data points where the two flux decompositions are inconsistent,
  ## i.e. where the residual of their regression is above the 97.5% or below the 2.5% quantile.
  res  <- gpp_nt - gpp_dt
  q025 <- quantile( res, probs = 0.025, na.rm=TRUE )
  q975 <- quantile( res, probs = 0.975, na.rm=TRUE )

  gpp_nt[ res > q975 | res < q025  ] <- NA
  gpp_dt[ res > q975 | res < q025  ] <- NA

  ## remove negative GPP
  gpp_nt[ which(gpp_nt<0) ] <- NA
  gpp_dt[ which(gpp_dt<0) ] <- NA

  return( list( gpp_nt=gpp_nt, gpp_dt=gpp_dt ) )
}
```



## Metrics

Several performance metrics are calculated for different features of GPP variability. The performance metrics are:

- R$^2$
- RMSE
- slope (of regression observed over modelled)
- bias

The features of variability in GPP, for which model-observation agreement is calculated, are:

- mean annual values (giving "spatial" correlation)
- annual anomalies from mean across years
- daily values, absolute
- mean across X-day periods, absolute
- mean seasonal cycle (mean by day of year)
- daily anomalies from mean seasonal cycle

## Results of evaluation


```{r, echo=FALSE, message=FALSE}
# Generate data used for plotting and get performance metrics
source("eval_sofun.R")
filn <- "eval_sofun.Rdata"
if (!file.exists(filn)){
  out_eval <- eval_sofun( mod, settings_eval, settings_sims, siteinfo, doplot=FALSE )
  save( out_eval, file = filn )
} else {
  load( filn )
}
```

### Annual values

Annual values in modelled and simulated GPP can be decomposed into the mean annual GPP per site $\bar{X}_i$ and its anomaly from the multi-annual mean $X'_{i,t}$:
$$
X_{i,t} = \bar{X}_i + X'_{i,t}
$$

#### Spatial correlation

Comparing the multi-annual mean per site in simulated and observed GPP ($\bar{X}_i$) yields a "spatial correlation".

```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
plot_modobs_spatial( out_eval$data$meandf, makepdf=FALSE )
```

#### Annual GPP anomalies

Comparing annual anomalies ($X'_{i,t}$) yields insight into whether the model accurately simulates interannual variability.

```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
plot_modobs_anomalies_annual( out_eval$data$iavdf, out_eval$data$iavdf_stats, makepdf=FALSE )
```

### Combined spatial/annual

The two above can be combined into a single plot. This shows the same as in the SI of our submitted manuscript: Figures S14 and S15 [here](http://rpubs.com/stineb/si_soilm_global).

```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
plot_modobs_spatial_annual( out_eval$data$meandf, out_eval$data$linmod_meandf, out_eval$data$adf_stats, makepdf=FALSE )
```

```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE}
successcodes <- read_csv( "successcodes.csv" )
do.sites <- dplyr::filter( successcodes, successcode==1 | successcode==2 )$mysitename
plot_modobs_spatial_annual( 
  filter( out_eval$data$meandf, sitename %in% do.sites), 
  lm( gpp_obs ~ gpp_mod, data = filter( out_eval$data$meandf, sitename %in% do.sites) ), 
  filter( out_eval$data$adf_stats, sitename %in% do.sites), 
  makepdf=FALSE )
```


<!-- ### Daily values -->

<!-- Ignoring that we don't expect a LUE-type linear relationship between absorbed light $I_{\text{abs}}$ and GPP, we can still evaluate the daily GPP estimated by the P-model: -->
<!-- $$ -->
<!-- \text{GPP}(d) = \text{LUE}(m|d)\; \times \; I_{\text{abs}}(d) -->
<!-- $$ -->
<!-- Here, $LUE(m|d)$ is the monthly varying light use efficiency simulated by the P-model using forcing data averaged to monthly means. $m|d$ refers to the month of a given day. -->

<!-- The correlation is still quite good... -->

<!-- ```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE} -->
<!-- modobs_ddf <- plot_modobs_daily( out_eval$data$ddf, makepdf=FALSE, xlim=c(0,25), ylim=c(0,25) ) -->
<!-- ``` -->

<!-- ### Values aggregated to X days -->

<!-- Aggregated to longer periods, the performance slightly improves. Here, I aggregated modeled and observational data to 5-days periods. -->

<!-- ```{r, fig.width=7.6, fig.height=7, echo=FALSE, message=FALSE} -->
<!-- modobs_ddf <- plot_modobs_xdaily( out_eval$data$xdf, makepdf=FALSE, xlim=c(0,20), ylim=c(0,20) ) -->
<!-- ``` -->

# Appendix

```{r siteoverview, echo=FALSE, warning=FALSE, message=FALSE}
require(readr, quietly = TRUE)
require(dplyr, quietly = TRUE)
siteinfo <- read_csv("siteinfo_eval.csv") %>%
            mutate( Reference = paste0("[@", sitename ,"]") ) %>%
            mutate( Period = paste0(as.character(year_start), "-", as.character(year_end)) ) %>% 
            select( -year_start, -year_end )
siteinfo %>%
  dplyr::rename( Site=sitename, Lon.=lon, Lat.=lat, Elevation=elv, Veg.=classid, Clim.=koeppen_code, N = ndailygpp ) %>%
  select( Site, Lon., Lat., Period, Veg., Clim., N, Reference ) %>% 
  knitr::kable( caption = "Sites used for evaluation. Lon. is longitude, negative values indicate west longitude; Lat. is latitude, positive values indicate north latitude; Veg. is vegetation type: deciduous broadleaf forest (DBF); evergreen broadleaf forest (EBF); evergreen needleleaf forest (ENF); grassland (GRA); mixed deciduous and evergreen needleleaf forest (MF); savanna ecosystem (SAV); shrub ecosystem (SHR); wetland (WET)." )
```

