---
title: "Vignette for package rsofun"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

Load the package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output
```{r, eval=FALSE}
library(rsofun)
```

# Setup
Define data sources used for SOFUN input. This is based on keywords passed on as strings. E.g. `"fluxnet2015"` triggers specific functions that read input data from specific files, here all the site-level meteo data from FLUXNET 2015. Define the input settings as a list.
```{r}
settings_input <-  list( 
  temperature             = "fluxnet2015",
  precipitation           = "fluxnet2015",
  vpd                     = "fluxnet2015",
  ppfd                    = "fluxnet2015",
  netrad                  = NA,
  fapar                   = "MODIS_FPAR_MCD15A3H",
  dir_fluxnet2015         = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  dir_MODIS_FPAR_MCD15A3H = "/alphadata01/bstocker/data/fapar_MODIS_FPAR_MCD15A3H_fluxnet2015_gee_subset/"
  )
```

Define the simulation settings as a list.
```{r}
settings_sims <- list( 
  ensemble      = TRUE,
  simsuite      = "fluxnet2015",
  path_siteinfo = "/alphadata01/bstocker/sofun/input_fluxnet2015_sofun/siteinfo_fluxnet2015_sofun+whc.csv",
  lonlat        = FALSE,
  path_output   = "/alphadata01/bstocker/sofun/output_fluxnet2015_sofun/sXXX/"
  )
```

- `path_siteinfo`: Must contain site names in first column.

Define model setup as a list.
```{r}
setup_sofun <- list( 
  model      = "pmodel_simsuite", 
  do_compile = FALSE,
  dir        = "/alphadata01/bstocker/sofun/trunk/"
  )
```

- `model`: compilation option and name of executable 
- `dir`: directory of the model

Define model calibration settings as a list.
```{r, eval=FALSE}
settings_calib <- list( 
  par        = list( k_decay = c(1,20) ),
  method     = "gensa",
  targetvars = c("gpp_nt", "gpp_dt"),
  datasource = list(  gpp_nt = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
                      gpp_dt = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/" ),
  varnames   = list( gpp_nt = "GPP_NT_VUT_REF", gpp_dt = "GPP_DT_VUT_REF" ),
  datenames  = list( gpp_nt = "TIMESTAMP", gpp_dt = "TIMESTAMP" )
  # targetvars = c("fapar"),
  # datasource = list( fapar = "/alphadata01/bstocker/data/fapar_MODIS_FPAR_MCD15A3H_fluxnet2015_gee_subset/" ),
  # varnames   = list( fapar = "modisvar" ),
  # datenames  = list( fapar = "date" )
 )
```

- `par`: list of parameters to calibrate with their lower and upper boundaries
- `targetvars`: name of variable in SOFUN output
- `datasource`: directory where site-specific files are located, a list of paths with names corresponding to 'targetvars'. The paths are pointing to directories where site-specific files are located. The first found file is read where the site name is contained in the file name (using `grepl()`).
- `varnames`: name of column that holds the data for the target variable in the original file.
- `datenames`: name of column that holds the date information in the original file.


# Execute

Create run directory with all the simulation parameter files (defining simulation years, length, etc.). This returs the `settings_sims` list, complemented by additional information.
```{r, eval=FALSE}
settings_sims <- prepare_setup_sofun( settings = settings_sims )
```

Prepare SOFUN input (climate input, CO2, etc.), returns the path where inputs are located
```{r, eval=FALSE}
path_input <- prepare_input_sofun( settings = settings_input )
```

Run SOFUN with standard parameters
```{r, eval=FALSE}
error <- run_sofun( path_forcing, setup = setup_sofun, par = list( k_decay = 4.0 ) )
```

Calibrate SOFUN, returns calibrated parameters. Set `r setup_sofun$is_calib <- TRUE` to prevent SOFUN output to be written, calibration output written instead
```{r, eval=FALSE}
setup_sofun$is_calib <- TRUE
params_opt <- calib_sofun( setup = setup_sofun, settings_calib = settings_calib, settings_sims = settings_sims )
```

Run SOFUN with calibrated parameters. Set `r setup_sofun$is_calib <- FALSE` to write normal SOFUN output.
```{r, eval=FALSE}
setup_sofun$is_calib <- FALSE
error <- run_sofun( path_forcing, setup = setup_sofun, par = params_opt )
```

Read SOFUN output
```{r, eval=FALSE}
out <- read_output_sofun( settings = settings_sims )
```

