---
title: "Vignette for package rsofun"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

# Environment

Load the package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r, eval=FALSE}
library(sofunr)  ## Not yet implemented
```

For the time being, we just go with a set of wrapper functions `prepare_setup_sofun()`, `prepare_input_sofun()`, and `runread_sofun()`. Load the respective functions.
```{r}
source("prepare_metainfo_fluxnet2015.R")
source("prepare_setup_sofun.R")
source("prepare_input_sofun.R")
source("runread_sofun.R")
```

Some other crap.
```{r}
systr <- "''"    # for Mac
```

# Setup
## Simulation settings
Define the simulation settings as a list. Components of the simulation settings should specify only variables that are identical across simulations within an ensemble. Other variables that differ between simulation within an ensemble (e.g. start and end year of the simulation, vegetation type C3 or C4, longitude and latitude of the site, soil type, etc.) are to be specified in the meta info file `path_siteinfo` (CSV file created "by hand" or some other more clever method for large ensembles - up to you.). The function `prepare_setup_sofun()` complements the simulation settings list by these additional components as lists with elements for each site, nested within the simulation settings list.
```{r}
settings_sims <- list( 
  path_siteinfo   = "./metainfo_fluxnet2015",
  ensemble        = TRUE,
  lonlat          = FALSE,
  name            = "fluxnet2015",
  dir_sofun       = "~/sofun/trunk/",
  path_output     = "~/sofun/output_fluxnet2015_sofun_TEST/",
  path_output_nc  = "~/sofun/output_nc_fluxnet2015_sofun_TEST/",
  path_input      = "~/sofun/input_fluxnet2015_sofun_TEST/",
  grid            = NA,
  implementation  = "fortran",
  in_ppfd         = TRUE,
  recycle         = 1,
  spinupyears     = 10,
  loutplant       = FALSE,
  loutgpp         = FALSE,
  loutwaterbal    = FALSE,
  loutdtemp_soil  = FALSE,
  loutdgpp        = FALSE,
  loutdrd         = FALSE,
  loutdtransp     = FALSE,
  lncoutdtemp     = FALSE,
  lncoutdfapar    = FALSE,
  lncoutdgpp      = TRUE, 
  lncoutdwaterbal = TRUE
  )
```

- `path_siteinfo`: Path (character string) of a CSV file that contains all the information needed for defining simulations. One row for each simulation in an ensemble, typically sites. Columns are as follows:
    - site name, must be: column number 1 
    - longitude of site, column must be named 'lon'
    - latitude of site, column must be named 'lat'
    - elevation of site, column must be named 'elv'
    - years for which simulation is to be done (corresponding to data availability from site), 
      requires two columns named 'year_start' and 'year_end'.
- `ensemble`: TRUE if an ensemble of site-level simulations are to be run (This may become obsolte, this information is given already if the number of rows in the CSV file `path_siteinfo` is larger than one.)
- `lonlat`: TRUE if simulation(s) are on a spatial longitude/latitude grid (multiple gridcells in one simulation). For point-scale (site-scale) simulations, `lonlat` is FALSE.
- `name`: a character string specifying the name of the ensemble (e.g. 'fluxnet2015') or of a single simulation.
- `dir_sofun`: Path (character string) where the model sits (corresponding to the parent directory of the respective git repository).
- `path_output`: Path (character string) where model output (ascii text files) is written to.
- `path_output_nc`: Path (character string) where NetCDF model output is written to.
- `path_input`: Path (character string) where model input is located. 
- `grid`: Character string defining the type of grid used, e.g. `halfdeg` for half-degree resolution in lon. and lat. (only used if `lonlat=TRUE`).
- `implementation`: Character string specifying whether Fortran (`implementation= "fortran"`) or Python (`implementation= "python"`) version is to be used.
- `in_ppfd`: Switch (`TRUE` of `FALSE`) whether PPFD should be read from data (prescribed) or simulated online using SPLASH and prescribed fractional cloud cover data.
- `recycle`: Periodicity (integer, number of years) of repeating forcing years during spinup (e.g. if the simulation start year is 1982 and `recycle=3`, then forcing years 1979-1981 are repeated during the duration of the model spinup, so that the last year of the spinup is forcing year 1981. For `recycle=1` and simulation start year 1982, forcing year 1981 is used for all years of the model spinup. Here, 'forcing year' refers to the year AD in the climate, CO2, fAPAR, etc. data used as model forcing.
- `spinupyears`: Integer, number of model spinup years before the transient simulation. Use `spinupyears > 0` if the model contains pool variables that that are simulated by dynamics that depend on their current state (typically soil water storage, or plant and soil carbon pools). Typically `spinupyears = 10` is sufficient to bring soil water pools to equilibrium across the globe (unless you chose a large soil water holding capacity).
- `lout<var>`: Switch (`TRUE` of `FALSE`) whether variable `<var>` is to be written to output (ascii text file for time series). To be anbandoned, only NetCDF output should be maintained.
- `lncout<var>`: Switch (`TRUE` of `FALSE`) whether variable `<var>` is to be written to output (NetCDF file).

## Input settings
Define data sources used for SOFUN input. This is based on keywords passed on as strings. E.g. `"fluxnet2015"` triggers specific functions that read input data from specific files, here all the site-level meteo data from FLUXNET 2015. Define the input settings as a list.
```{r}
settings_input <-  list( 
  temperature              = "fluxnet2015",
  precipitation            = "fluxnet2015",
  vpd                      = "fluxnet2015",
  ppfd                     = "fluxnet2015",
  netrad                   = NA,
  cloudcover               = "cru_ts4_01",
  fapar                    = "MODIS_FPAR_MCD15A3H",
  co2                      = "cmip",
  path_cx1data             = "~/data/",
  path_fluxnet2015         = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_watch_wfdei         = "~/data/watch_wfdei/",
  path_cru_ts4_01          = "~/data/cru/ts_4.01/",
  path_MODIS_FPAR_MCD15A3H = "~/data/fapar_MODIS_FPAR_MCD15A3H_fluxnet2015_gee_subset/"
  )
```

- `temperature`: A character string specifying the source for temperature data. Any of `"fluxnet2015"`, `"watch_wfdei"`, and/or `"cru"`. This can also be a vector of strings (e.g. `c("fluxnet2015", "watch_wfdei")`, to specify priorities: first take FLUXNET 2015 data for periods where data is available. For remaining years (given by `settings_sims$date_start` and `settings_sims$date_end`), use WATCH-WFDEI data. If `"fluxnet2015"` is specified for any of `temperature`, `precipitation`, `vpd`, `ppfd`, or `netrad`, then `path_fluxnet2015` must be specified as well in the settings.
- `precipitation`: See `temperature`.
- `vpd`: See `temperature`.
- `ppfd`: See `temperature`.
- `netrad`: See `temperature`.
- `fapar`: A character string specifying the type of fAPAR data used as input (e.g. `"MODIS_FPAR_MCD15A3H"`), or `NA` in case no fAPAR data is used as forcing (internally simulated fAPAR).
- `co2`: A character string specifying which CO$_2$ file should be used (globally uniform and identical for each site in an ensemble). All available CO$_2$ forcing files are located on CX1 at `/work/bstocker/labprentice/data/co2/`. `co2="cmip"` specifies that the CMIP-standard CO2 file `cCO2_rcp85_const850-1765.dat` should be used.
- `path_cx1data`: the local directory that mirrors the data directory on CX1 (very handy!).

## Model setup
Define model setup as a list.
```{r}
setup_sofun <- list( 
  model      = "pmodel",
  dir        = "~/sofun/trunk",
  do_compile = FALSE,
  simsuite   = FALSE
  )
```

- `model`: For Fortran version: A character string specifying the compilation option. The name of the executable is derived from this as `"run<setup_sofun$model>"`.
- `dir`: A path (character) specifying the directory of where the executables are located (corresponds to the parent directory of the model git repository).
- `do_compile`: If `TRUE`, the model code is compiled as `make <setup_sofun$model>`. If `FALSE`, compiled executables are used (compiled with gfortran on a Mac 64-bit).
- `simsuite`: If `TRUE`, the SOFUN option for running an entire simulation suite (ensemble of simulations) with a single executable. (In the Fortran implementation, this this requires the main program file `sofun_simsuite.f90` to be compiled instead of `sofun.f90`). This is to be preferred over a set of individual runs submitted individually when doing calibration runs because the cost across the entire ensemble (mod-obs error) can thus be calculated online.


## Calibration settings
Define model calibration settings as a list.
```{r, eval=FALSE}
settings_calib <- list( 
  par        = list( k_decay = c(1,20) ),
  method     = "gensa",
  targetvars = c("gpp_nt", "gpp_dt"),
  datasource = list(  gpp_nt = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
                      gpp_dt = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/" ),
  varnames   = list( gpp_nt = "GPP_NT_VUT_REF", gpp_dt = "GPP_DT_VUT_REF" ),
  datenames  = list( gpp_nt = "TIMESTAMP", gpp_dt = "TIMESTAMP" )
  # targetvars = c("fapar"),
  # datasource = list( fapar = "~/data/fapar_MODIS_FPAR_MCD15A3H_fluxnet2015_gee_subset/" ),
  # varnames   = list( fapar = "modisvar" ),
  # datenames  = list( fapar = "date" )
 )
```

- `par`: list of parameters to calibrate with their lower and upper boundaries
- `targetvars`: name of variable in SOFUN output
- `datasource`: directory where site-specific files are located, a list of paths with names corresponding to 'targetvars'. The paths are pointing to directories where site-specific files are located. The first found file is read where the site name is contained in the file name (using `grepl()`).
- `varnames`: name of column that holds the data for the target variable in the original file.
- `datenames`: name of column that holds the date information in the original file.


# Workflow

## Prepare meta info file
This step is either done by hand so that the file has a certain structure (for column names etc see above: `path_siteinfo`). Below is an example for creating such a meta info file (info for each site in an ensemble) for the FLUXNET 2015 Tier 1 ensemble.
```{r, eval=FALSE}
siteinfo <- prepare_metainfo_fluxnet2015( settings_sims = settings_sims, settings_input = settings_input, overwrite = FALSE, filn_elv_watch = paste0( settings_input$path_cx1data, "watch_wfdei/WFDEI-elevation.nc" ) )
```

- `overwrite`: Set to `TRUE` if meta info file in wide format is to be overwritten (original is in long format, a wide-format version is used by `prepare_metainfo_fluxnet2015()`).
- `filn_elv_watch`: File name (full path, character string) of the WATCH-WFDEI elevation file (NetCDF format) used to complement missing elevation information.


## Prepare simulation setup
Create a run directory with all the simulation parameter files (defining simulation years, length, etc.). This returs the `settings_sims` list, complemented by additional information.
```{r, eval=FALSE}
settings_sims <- prepare_setup_sofun( settings = settings_sims, write_paramfils = FALSE )
```

## Prepare inputs
Prepare SOFUN input (climate input, CO2, etc.). Complements `settings_input`. This will require inputs from the user through the prompt, entered in the console to specify whether data files should be downloaded from Imperial CX1. In case you chose to download, you must have access to CX1 and be connected to the Imperial VPN. Once asked (see console!), enter your user name on CX1. This also requires that no additional entering of the password is required. In order to set this up, you need to generate an SSH key pair beforehand (see (here)[https://www.digitalocean.com/community/tutorials/how-to-set-up-ssh-keys--2]). 
```{r, eval=FALSE}
inputdata <- prepare_input_sofun( settings_input = settings_input, settings_sims = settings_sims, return_data = TRUE )
```

## Run the model
Run SOFUN with standard parameters and read output in once.
```{r, eval=FALSE}
out <- runread_sofun( settings = settings_sims, setup = setup_sofun )  
# to be done: pass calibratable parameters as additional argument , par = list( k_decay = 4.0 )
```

# BELOW NOT YET IMPLEMENTED

Calibrate SOFUN, returns calibrated parameters. Set `r setup_sofun$is_calib <- TRUE` to prevent SOFUN output to be written, calibration output written instead
```{r, eval=FALSE}
setup_sofun$is_calib <- TRUE
params_opt <- calib_sofun( setup = setup_sofun, settings_calib = settings_calib, settings_sims = settings_sims )
```

Run SOFUN with calibrated parameters. Set `r setup_sofun$is_calib <- FALSE` to write normal SOFUN output.
```{r, eval=FALSE}
setup_sofun$is_calib <- FALSE
out <- runread_sofun( path_forcing, setup = setup_sofun, par = params_opt )
```

