---
title: "Vignette for package rsofun"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

Load the package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r, eval=FALSE}
library(rsofun)  ## Not yet implemented
```

For the time being, we just go with a set of wrapper functions `prepare_setup_sofun()`, `prepare_input_sofun()`, and `runread_sofun()`. Load the respective functions.
```{r}
source("prepare_setup_sofun.R")
source("prepare_input_sofun.R")
source("runread_sofun.R")
```

Some other crap.
```{r}
systr <- "''"    # for Mac
```

# Setup
Define the simulation settings as a list.
```{r}
settings_sims <- list( 
  path_siteinfo = "/alphadata01/bstocker/sofun/input_fluxnet2015_sofun/siteinfo_fluxnet2015_sofun+whc.csv",
  ensemble      = TRUE,
  lonlat        = FALSE,
  name          = "fluxnet2015",
  dir_sofun     = "/alphadata01/bstocker/sofun/trunk/",
  path_output   = "/alphadata01/bstocker/sofun/output_fluxnet2015_sofun_TEST/",
  path_input    = "/alphadata01/bstocker/sofun/input_fluxnet2015_sofun_TEST/",
  grid          = NA,
  implementation= "fortran"
  )
```

- `path_siteinfo`: Path (character string) of a CSV file that contains all the information needed for defining simulations. One row for each simulation in an ensemble, typically sites. Columns are as follows:
    - site name, must be: column number 1 
    - longitude of site, column must be named 'lon'
    - latitude of site, column must be named 'lat'
    - elevation of site, column must be named 'elv'
    - years for which simulation is to be done (corresponding to data availability from site), 
      requires two columns named 'year_start' and 'year_end'.
- `ensemble`: TRUE if an ensemble of site-level simulations are to be run (This may become obsolte, this information is given already if the number of rows in the CSV file `path_siteinfo` is larger than one.)
- `lonlat`: TRUE if simulation(s) are on a spatial longitude/latitude grid (multiple gridcells in one simulation). For point-scale (site-scale) simulations, `lonlat` is FALSE.
- `name`: a character string specifying the name of the ensemble (e.g. 'fluxnet2015') or of a single simulation.
- `dir_sofun`: Path (character string) where the model sits (corresponding to the parent directory of the respective git repository).
- `path_output`: Path (character string) where model output is written to.
- `path_input`: Path (character string) where model input is located. 
- `grid`: Character string defining the type of grid used, e.g. `halfdeg` for half-degree resolution in lon. and lat. (only used if `lonlat=TRUE`).
- `implementation`: Character string specifying whether Fortran (`implementation= "fortran"`) or Python (`implementation= "python"`) version is to be used.


Define data sources used for SOFUN input. This is based on keywords passed on as strings. E.g. `"fluxnet2015"` triggers specific functions that read input data from specific files, here all the site-level meteo data from FLUXNET 2015. Define the input settings as a list.
```{r}
settings_input <-  list( 
  temperature              = "fluxnet2015",
  precipitation            = "fluxnet2015",
  vpd                      = "fluxnet2015",
  ppfd                     = "fluxnet2015",
  netrad                   = NA,
  fapar                    = "MODIS_FPAR_MCD15A3H",
  path_cx1data             = "/alphadata01/bstocker/data/",
  path_fluxnet2015         = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_MODIS_FPAR_MCD15A3H = "/alphadata01/bstocker/data/fapar_MODIS_FPAR_MCD15A3H_fluxnet2015_gee_subset/"
  )
```

Define model setup as a list.
```{r}
setup_sofun <- list( 
  model      = "pmodel_simsuite",
  do_compile = FALSE
  )
```

- `model`: compilation option and name of executable 
- `dir`: directory of the model

Define model calibration settings as a list.
```{r, eval=FALSE}
settings_calib <- list( 
  par        = list( k_decay = c(1,20) ),
  method     = "gensa",
  targetvars = c("gpp_nt", "gpp_dt"),
  datasource = list(  gpp_nt = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
                      gpp_dt = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/" ),
  varnames   = list( gpp_nt = "GPP_NT_VUT_REF", gpp_dt = "GPP_DT_VUT_REF" ),
  datenames  = list( gpp_nt = "TIMESTAMP", gpp_dt = "TIMESTAMP" )
  # targetvars = c("fapar"),
  # datasource = list( fapar = "/alphadata01/bstocker/data/fapar_MODIS_FPAR_MCD15A3H_fluxnet2015_gee_subset/" ),
  # varnames   = list( fapar = "modisvar" ),
  # datenames  = list( fapar = "date" )
 )
```

- `par`: list of parameters to calibrate with their lower and upper boundaries
- `targetvars`: name of variable in SOFUN output
- `datasource`: directory where site-specific files are located, a list of paths with names corresponding to 'targetvars'. The paths are pointing to directories where site-specific files are located. The first found file is read where the site name is contained in the file name (using `grepl()`).
- `varnames`: name of column that holds the data for the target variable in the original file.
- `datenames`: name of column that holds the date information in the original file.


# Workflow

Create run directory with all the simulation parameter files (defining simulation years, length, etc.). This returs the `settings_sims` list, complemented by additional information.
```{r, eval=FALSE}
settings_sims <- prepare_setup_sofun( settings = settings_sims )
```

Prepare SOFUN input (climate input, CO2, etc.). Complements `settings_input`.
```{r, eval=FALSE}
settings_input <- prepare_input_sofun( settings_input = settings_input, settings_sims = settings_sims )
```

Run SOFUN with standard parameters and read output in once.
```{r, eval=FALSE}
out <- runread_sofun( settings = settings_sims, setup = setup_sofun )  
# to be done: pass calibratable parameters as additional argument , par = list( k_decay = 4.0 )
```

Calibrate SOFUN, returns calibrated parameters. Set `r setup_sofun$is_calib <- TRUE` to prevent SOFUN output to be written, calibration output written instead
```{r, eval=FALSE}
setup_sofun$is_calib <- TRUE
params_opt <- calib_sofun( setup = setup_sofun, settings_calib = settings_calib, settings_sims = settings_sims )
```

Run SOFUN with calibrated parameters. Set `r setup_sofun$is_calib <- FALSE` to write normal SOFUN output.
```{r, eval=FALSE}
setup_sofun$is_calib <- FALSE
out <- runread_sofun( path_forcing, setup = setup_sofun, par = params_opt )
```

