---
title: "Calibrating multiple C-model parameters with GPP, Reco, and LAI features"
author: "Benjamin Stocker"
output: rmarkdown::html_vignette
code-fold: true
vignette: >
  %\VignetteIndexEntry{Calibrating multiple C-model parameters with GPP, Reco, and LAI features}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.5
)
```

## Aim

This vignette extends the feature-oriented likelihood developed in
`vignette("calibration_cmodel_ch_oe1")` to a joint calibration of several
C-model parameters at the temperate grassland site CH-Oe1. GPP, ecosystem
respiration (Reco), and leaf area index (LAI) are separate calibration targets.

Every target is reduced to annually replicated time-series features. GPP and
Reco contribute annual totals, while LAI contributes an annual mean. All three
also contribute peak magnitude, peak timing, growing-season onset and
termination, and quantile-based seasonal amplitude. A separate residual
variance is assigned to every target-feature combination and integrated out
under an inverse-gamma prior. Consequently, the Latin hypercube contains only
biological model parameters, not weakly identified error parameters.

The sample size used below is suitable for demonstrating the workflow, not for
a definitive eight-dimensional calibration. Increase `n_samples` and inspect
posterior stability before interpreting parameter estimates.

```{r libraries}
library(rsofun)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

if (!exists(
  ".calib_lhs_worker_stop_recovery",
  envir = asNamespace("rsofun"),
  inherits = FALSE
)) {
  stop(
    "The installed rsofun is stale. Reinstall this working tree, or run ",
    "devtools::load_all() before rendering the vignette."
  )
}
```

## Drivers and three observation targets

The FluxDataKit forcing supplies GPP and NEE. Under its sign convention,
observed Reco is `GPP + NEE`. MODIS LAI is taken from the quality-controlled
data prepared in `vignette("example_cmodel")`.

```{r data, eval=TRUE}
fdk_path <- "~/data_2/FluxDataKit/v3.4/zenodo_upload"

df_drivers <- readRDS(file.path(
  fdk_path,
  "rsofun_driver_data_v3.4.2.rds"
)) |>
  dplyr::filter(sitename == "CH-Oe1")

use_cseed <- 5
use_nseed <- use_cseed / 20
df_drivers <- df_drivers |>
  dplyr::mutate(
    forcing = purrr::map(forcing, ~ dplyr::mutate(
      .,
      fharv = 0,
      dno3 = 1.5 / (2 * 365),
      dnh4 = 1.5 / (2 * 365),
      cseed = ifelse(lubridate::month(date) == 2 &
                       lubridate::mday(date) == 15, use_cseed, 0),
      nseed = ifelse(lubridate::month(date) == 2 &
                       lubridate::mday(date) == 15, use_nseed, 0)
    ))
  )
df_drivers$params_siml[[1]]$spinupyears <- 2002
df_drivers$params_siml[[1]]$recycle <- 5
df_drivers$params_siml[[1]]$c_only <- TRUE

flux_observations <- df_drivers |>
  dplyr::select(sitename, forcing) |>
  tidyr::unnest(forcing) |>
  dplyr::transmute(
    sitename,
    date,
    gpp = dplyr::if_else(gpp_qc >= 0.8, gpp, NA_real_),
    reco = dplyr::if_else(
      gpp_qc >= 0.8 & nee_qc >= 0.8, gpp + nee, NA_real_
    )
  )

data("modis_lai_ch_oe1", package = "rsofun")
model_period <- range(flux_observations$date)
lai_observations <- modis_lai_ch_oe1 |>
  dplyr::filter(dplyr::between(date, model_period[1], model_period[2])) |>
  dplyr::transmute(sitename = "CH-Oe1", date, lai)
```

## Baseline parameters

The P-model values below reproduce the MAP estimates in the last column of
Table 2 (setup S6) of Arán Paredes et al. (2025): `kphio = 0.08`,
`kphio_par_a = -0.001`, `kphio_par_b = 27.5`, `soilm_thetastar = 26.7`,
`beta_unitcostratio = 187.5`, `tau_acclim = 1.0`, and `kc_jmax = 0.58`.
Parameters that are not estimated in setup S6 retain their existing model
defaults. Note that the S6 value of `beta_unitcostratio` lies at the lower bound
of its truncated normal prior; it should not be replaced by the posterior mean
or by the rounded value discussed elsewhere in the article.

The three quantum-yield parameters have a distinct and coupled meaning:

```text
phi(T) = kphio * max(0, min(1, 1 + kphio_par_a * (T - kphio_par_b)^2))
```

Thus `kphio` is the maximum quantum yield, `kphio_par_a` controls the curvature,
and `kphio_par_b` locates the temperature optimum. The shape and location
parameters remain fixed at their S6 estimates, as does `kphio`. The calibrated
subset below instead targets C-model phenology, allocation, turnover, growth,
and respiration parameters.

The turnover constants are annual values. Baseline values of `1.9 yr-1` for
leaves and `0.7 yr-1` for roots correspond to expected tissue lifetimes of
about 0.53 and 1.43 years. They define the prior centres for the turnover
calibration. Allocation to wood remains fixed at zero.

Leaf structure and nitrogen partitioning are also fixed. In particular,
`r_ctostructn_leaf = 30.9` is the previously selected value for a temperate
grass; it is held fixed during the multi-target calibration.

```{r parameters}
pars <- list(
  # P-model: informed by Arán Paredes et al. (2025), setup S6
  kphio = 0.08,
  kphio_par_a = -0.001,
  kphio_par_b = 27.5,
  # C-model phenology; unrelated to the P-model quantum-yield parameters above
  dehardening_gdd_midpoint = 150.0,
  dehardening_gdd_slope = 0.05,
  dehardening_gdd_base = 5.0,
  soilm_thetastar = 26.7,
  soilm_betao = 0.0,
  beta_unitcostratio = 187.5,
  rd_to_vcmax = 0.014,
  tau_acclim = 1.0,
  kc_jmax = 0.58,

  # Plant C and N
  f_nretain = 0.5,
  fpc_tree_max = 0.95,
  growtheff = 0.6,
  r_root = 2 * 0.913,
  r_sapw = 2 * 0.044,
  exurate = 0.003,
  k_decay_leaf = 1.9,
  k_decay_root = 0.7,
  k_decay_labl = 1.9,
  k_decay_sapw = 1.9,
  r_cton_root = 37.0,
  r_cton_wood = 100.0,
  r_cton_seed = 15.0,
  nv_vcmax25 = 0.02 * 13681.77,
  ncw_min = 0.08 * 1.116222,
  r_n_cw_v = 0.0,
  r_ctostructn_leaf = 30.9,
  kbeer = 0.5,

  # Phenology
  gddbase = 5.0,
  ramp = 0.0,
  phentype = 2.0,

  # Soil physics
  perc_k1 = 5.0,
  thdiff_wp = 0.2,
  thdiff_whc15 = 0.8,
  thdiff_fc = 0.4,
  forg = 0.01,
  wbwp = 0.029,
  por = 0.421,
  fsand = 0.82,
  fclay = 0.06,
  fsilt = 0.12,

  # Water and energy balance
  kA = 107.0,
  kalb_sw = 0.17,
  kalb_vis = 0.03,
  kb = 0.20,
  kc = 0.25,
  kCw = 1.05,
  kd = 0.50,
  ke = 0.0167,
  keps = 23.44,
  kWm = 220.0,
  kw = 0.26,
  komega = 283.0,
  maxmeltrate = 3.0,

  # Soil biogeochemistry
  klitt_af10 = 1.2,
  klitt_as10 = 0.35,
  klitt_bg10 = 0.35,
  kexu10 = 50.0,
  ksoil_fs10 = 0.021,
  ksoil_sl10 = 7.0e-4,
  ntoc_crit1 = 0.45,
  ntoc_crit2 = 0.76,
  cton_microb = 10.0,
  cton_soil = 9.77,
  fastfrac = 0.985,

  # N uptake and transformations
  eff_nup = 1.0e-4,
  minimumcostfix = 1.0,
  fixoptimum = 25.15,
  a_param_fix = -3.62,
  b_param_fix = 0.27,
  maxnitr = 5.0e-5,
  non = 0.01,
  n2on = 5.0e-4,
  kn = 83.0,
  kdoc = 17.0,
  docmax = 1.0,
  dnitr2n2o = 0.01,

  # Prescribed allocation in the C-only setup
  frac_leaf = 0.5,
  frac_wood = 0.0,
  frac_avl_labl = 0.1,
  tmppar = 9999.0,
  nuptake_kc = 600.0,
  nuptake_kv = 5.0,
  nuptake_vmax = 0.2
)

stopifnot(pars$frac_wood == 0)
```

## Annual time-series features

The helper below first interpolates internal gaps onto a daily grid and then
applies a centred 15-day moving average. Onset and termination are the first
and last crossings around the seasonal peak of a threshold located 15% of the
way from the 5th to the 95th smoothed quantile. Peak timing is represented by
week of year.

Fluxes use an annual total calculated over the common observation period. LAI
is a state variable rather than a flux, so its annual-level feature is the
annual mean. Features are only retained for years with sufficient paired data.

```{r feature-extraction}
annual_series_features <- function(
    data, value_name, annual_stat = c("total", "mean"),
    minimum_observations
) {
  annual_stat <- match.arg(annual_stat)
  series <- data |>
    dplyr::transmute(
      year = lubridate::year(date),
      day = lubridate::yday(date),
      value = .data[[value_name]]
    ) |>
    dplyr::filter(is.finite(value))
  series <- split(series, series$year)

  extract_year <- function(x) {
    if (nrow(x) < minimum_observations) return(NULL)
    daily <- tibble::tibble(
      day = seq_len(365),
      value = stats::approx(
        x$day, x$value, xout = seq_len(365), ties = mean
      )$y
    ) |>
      dplyr::mutate(
        smooth = as.numeric(stats::filter(
          value, rep(1 / 15, 15), sides = 2
        )),
        week = (day - 1L) %/% 7L + 1L
      )
    usable <- dplyr::filter(daily, is.finite(smooth))
    if (nrow(usable) < minimum_observations) return(NULL)

    weekly <- daily |>
      dplyr::filter(is.finite(value)) |>
      dplyr::group_by(week) |>
      dplyr::summarise(value = mean(value), .groups = "drop")
    peak <- weekly[which.max(weekly$value), ]
    lower <- stats::quantile(usable$smooth, 0.05, names = FALSE)
    upper <- stats::quantile(usable$smooth, 0.95, names = FALSE)
    threshold <- lower + 0.15 * (upper - lower)
    peak_day <- usable$day[which.max(usable$smooth)]
    active <- dplyr::filter(usable, smooth >= threshold)
    before_peak <- active$day[active$day <= peak_day]
    after_peak <- active$day[active$day >= peak_day]
    if (!length(before_peak) || !length(after_peak)) return(NULL)

    annual_value <- if (annual_stat == "total") {
      sum(daily$value, na.rm = TRUE)
    } else {
      mean(daily$value, na.rm = TRUE)
    }
    annual_name <- if (annual_stat == "total") {
      "annual_total"
    } else {
      "annual_mean"
    }

    tibble::tibble(
      year = x$year[[1]],
      annual_level = annual_value,
      peak_value = peak$value,
      peak_week = peak$week,
      onset = min(before_peak),
      termination = max(after_peak),
      amplitude = upper - lower
    ) |>
      tidyr::pivot_longer(
        -year, names_to = "feature", values_to = "value"
      ) |>
      dplyr::mutate(
        feature = dplyr::if_else(
          feature == "annual_level", annual_name, feature
        )
      )
  }

  dplyr::bind_rows(lapply(series, extract_year))
}

pair_features <- function(
    data, observed_name, modelled_name, target, annual_stat,
    minimum_observations, feature_function
) {
  # Both feature sets must use identical dates. Otherwise data gaps in the
  # observations would be mistaken for model-data differences.
  paired <- data |>
    dplyr::filter(
      is.finite(.data[[observed_name]]),
      is.finite(.data[[modelled_name]])
    )
  observed <- feature_function(
    paired, observed_name, annual_stat, minimum_observations
  ) |>
    dplyr::rename(observed = value)
  modelled <- feature_function(
    paired, modelled_name, annual_stat, minimum_observations
  ) |>
    dplyr::rename(modelled = value)

  dplyr::inner_join(observed, modelled, by = c("year", "feature")) |>
    dplyr::mutate(target = target)
}

build_feature_pairs <- function(
    simulated, flux_observations, lai_observations,
    feature_function, pair_function
) {
  matched_flux <- flux_observations |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(
      is.finite(gpp_modelled), is.finite(reco_modelled)
    )
  matched_lai <- lai_observations |>
    dplyr::inner_join(
      dplyr::select(simulated, sitename, date, lai_modelled),
      by = c("sitename", "date")
    ) |>
    dplyr::filter(is.finite(lai), is.finite(lai_modelled))

  dplyr::bind_rows(
    pair_function(
      matched_flux, "gpp", "gpp_modelled", "GPP", "total", 200,
      feature_function
    ),
    pair_function(
      matched_flux, "reco", "reco_modelled", "Reco", "total", 200,
      feature_function
    ),
    pair_function(
      matched_lai, "lai", "lai_modelled", "LAI", "mean", 12,
      feature_function
    )
  )
}
```

## Feature-specific variance priors

As explained in the companion vignette, if normal residual variance
$v=\sigma^2$ has an inverse-gamma prior, it can be integrated out exactly. The
prior is specified through `prior_rms = sqrt(E[v])`, in the native units of
each feature. `alpha = 3` is retained after the sensitivity analysis in the
single-parameter vignette.

The values below are explicit assumptions. Flux totals are in
g C m$^{-2}$ yr$^{-1}$, peak fluxes and amplitudes in
g C m$^{-2}$ d$^{-1}$, timing variables in weeks or days, and LAI magnitude
features in m$^2$ m$^{-2}$.

```{r variance-priors}
make_variance_prior <- function(prior_rms, alpha = 3) {
  list(
    alpha = alpha,
    beta = (alpha - 1) * prior_rms^2,
    prior_rms = prior_rms
  )
}

feature_variance_priors <- list(
  GPP__annual_total = make_variance_prior(150),
  GPP__peak_value = make_variance_prior(0.75),
  GPP__peak_week = make_variance_prior(2),
  GPP__onset = make_variance_prior(14),
  GPP__termination = make_variance_prior(14),
  GPP__amplitude = make_variance_prior(0.75),

  Reco__annual_total = make_variance_prior(150),
  Reco__peak_value = make_variance_prior(0.75),
  Reco__peak_week = make_variance_prior(2),
  Reco__onset = make_variance_prior(14),
  Reco__termination = make_variance_prior(14),
  Reco__amplitude = make_variance_prior(0.75),

  LAI__annual_mean = make_variance_prior(0.5),
  LAI__peak_value = make_variance_prior(0.75),
  LAI__peak_week = make_variance_prior(3),
  LAI__onset = make_variance_prior(21),
  LAI__termination = make_variance_prior(21),
  LAI__amplitude = make_variance_prior(0.5)
)

inverse_gamma_marginal_loglik <- function(residual, alpha, beta) {
  n <- length(residual)
  if (!n || any(!is.finite(residual))) return(-Inf)
  lgamma(alpha + n / 2) - lgamma(alpha) + alpha * log(beta) -
    (alpha + n / 2) * log(beta + sum(residual^2) / 2) -
    n * log(2 * pi) / 2
}
```

## Joint feature likelihood

The model is run once per parameter sample. GPP, Reco, and LAI features are
then evaluated separately, and their log marginal likelihoods are added. A
sample must produce every requested target-feature group; otherwise it receives
zero posterior weight.

```{r joint-feature-likelihood}
multitarget_feature_loglik <- function(
    par, model_error, drivers, flux_observations, lai_observations,
    feature_function, pair_function, pair_builder, variance_priors,
    marginal_function
) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(
      sitename,
      date,
      gpp_modelled = gpp,
      reco_modelled = rleaf + rroot + rwood + rhet + rgrow,
      lai_modelled = lai
    )
  pairs <- pair_builder(
    simulated = simulated,
    flux_observations = flux_observations,
    lai_observations = lai_observations,
    feature_function = feature_function,
    pair_function = pair_function
  ) |>
    dplyr::mutate(
      residual = observed - modelled,
      residual = dplyr::if_else(
        feature == "peak_week",
        pmin(abs(residual), 52 - abs(residual)),
        residual
      ),
      prior_name = paste(target, feature, sep = "__")
    )

  score <- vapply(names(variance_priors), function(prior_name) {
    prior <- variance_priors[[prior_name]]
    residual <- pairs$residual[pairs$prior_name == prior_name]
    marginal_function(residual, prior$alpha, prior$beta)
  }, numeric(1))
  if (any(!is.finite(score))) return(-Inf)
  sum(score)
}
```

## Parameters and parallel LHS calibration

The quantum-yield shape and location parameters remain fixed at the S6 values,
and wood allocation remains zero. The calibrated subset controls temperature
recovery after hardening, growth efficiency, leaf allocation, leaf and root
turnover, and root respiration.

```{r calibration-parameters}
calibration_parameters <- list(
  dehardening_gdd_midpoint = list(
    mean = pars$dehardening_gdd_midpoint, sd = 30,
    distribution = "normal", lower = 30, upper = 300
  ),
  dehardening_gdd_slope = list(
    mean = pars$dehardening_gdd_slope, sd = 0.015,
    distribution = "lognormal", lower = 0.005, upper = 0.15
  ),
  dehardening_gdd_base = list(
    mean = pars$dehardening_gdd_base, sd = 1.5,
    distribution = "normal", lower = 0, upper = 12
  ),
  growtheff = list(
    mean = pars$growtheff, sd = 0.06,
    distribution = "normal", lower = 0.35, upper = 0.85
  ),
  frac_leaf = list(
    mean = pars$frac_leaf, sd = 0.08,
    distribution = "normal", lower = 0.1, upper = 0.9
  ),
  k_decay_leaf = list(
    mean = 1.9, sd = 0.6,
    distribution = "lognormal", lower = 0.7, upper = 4
  ),
  k_decay_root = list(
    mean = 0.7, sd = 0.25,
    distribution = "lognormal", lower = 0.1, upper = 1.5
  ),
  r_root = list(
    mean = pars$r_root, sd = 0.45,
    distribution = "lognormal", lower = 0.5, upper = 3.5
  )
)

feature_preflight <- multitarget_feature_loglik(
  par = pars,
  model_error = list(),
  drivers = df_drivers,
  flux_observations = flux_observations,
  lai_observations = lai_observations,
  feature_function = annual_series_features,
  pair_function = pair_features,
  pair_builder = build_feature_pairs,
  variance_priors = feature_variance_priors,
  marginal_function = inverse_gamma_marginal_loglik
)
if (!is.finite(feature_preflight)) {
  stop("The multi-target feature likelihood failed its serial preflight check.")
}

n_samples <- 250
multitarget_calibration <- calib_sofun_lhs(
  pars = pars,
  par_calib = calibration_parameters,
  model_error = list(),
  likelihood = multitarget_feature_loglik,
  n_samples = n_samples,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 46,
  progress = TRUE,
  drivers = df_drivers,
  flux_observations = flux_observations,
  lai_observations = lai_observations,
  feature_function = annual_series_features,
  pair_function = pair_features,
  pair_builder = build_feature_pairs,
  variance_priors = feature_variance_priors,
  marginal_function = inverse_gamma_marginal_loglik
)
```

## Posterior-weight diagnostics

The effective sample size and ranked weights reveal whether the finite LHS
posterior is supported by several parameter combinations or collapses onto one.
Because eight parameters are varied, 250 samples are only a computationally
light demonstration.

```{r posterior-diagnostics}
posterior_diagnostics <- multitarget_calibration$samples |>
  dplyr::arrange(dplyr::desc(weight)) |>
  dplyr::mutate(cumulative_weight = cumsum(weight))

tibble::tibble(
  effective_sample_size = multitarget_calibration$effective_sample_size,
  failed_evaluations = sum(!multitarget_calibration$samples$evaluation_ok),
  maximum_weight = max(multitarget_calibration$samples$weight),
  samples_to_80_percent = which(
    posterior_diagnostics$cumulative_weight >= 0.8
  )[1]
)

ggplot(
  dplyr::mutate(
    posterior_diagnostics,
    rank = dplyr::row_number(),
    weight_for_plot = pmax(weight, 1e-16)
  ),
  aes(rank, weight_for_plot)
) +
  geom_hline(
    yintercept = 1 / nrow(posterior_diagnostics),
    linetype = "dashed", colour = "grey50"
  ) +
  geom_line(colour = "#0072B2") +
  geom_point(colour = "#0072B2", size = 0.8) +
  scale_y_log10() +
  labs(
    x = "Sample rank (highest posterior weight first)",
    y = "Posterior weight (log scale)"
  ) +
  theme_classic()
```

## Re-run the MAP model

The returned MAP parameter list includes fixed and calibrated values and can be
passed directly back to the C-model.

```{r map-run}
pars_map <- multitarget_calibration$par
stopifnot(pars_map$frac_wood == 0)

map_output <- rsofun::runread_cnmodel_f(
  drivers = df_drivers,
  par = pars_map
)
map_daily <- map_output |>
  dplyr::select(sitename, data) |>
  tidyr::unnest(data) |>
  dplyr::transmute(
    sitename,
    date,
    gpp_modelled = gpp,
    reco_modelled = rleaf + rroot + rwood + rhet + rgrow,
    lai_modelled = lai,
    lma,
    cleaf,
    croot,
    cwood
  )
```

## Target-feature diagnostics at the MAP

The table exposes every annual observed-modelled comparison rather than only
the summed likelihood. The following table gives the corresponding conditional
posterior error distributions. These errors were integrated out during
calibration and are recovered here only for interpretation.

```{r map-feature-diagnostics}
map_feature_pairs <- build_feature_pairs(
  simulated = map_daily,
  flux_observations = flux_observations,
  lai_observations = lai_observations,
  feature_function = annual_series_features,
  pair_function = pair_features
)
map_feature_pairs

feature_error_posteriors <- map_feature_pairs |>
  dplyr::mutate(
    residual = observed - modelled,
    residual = dplyr::if_else(
      feature == "peak_week",
      pmin(abs(residual), 52 - abs(residual)), residual
    ),
    prior_name = paste(target, feature, sep = "__")
  ) |>
  dplyr::group_by(target, feature, prior_name) |>
  dplyr::summarise(
    n_years = dplyr::n(),
    sum_squared_residual = sum(residual^2),
    .groups = "drop"
  ) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    prior_rms = feature_variance_priors[[prior_name]]$prior_rms,
    alpha_posterior = feature_variance_priors[[prior_name]]$alpha +
      n_years / 2,
    beta_posterior = feature_variance_priors[[prior_name]]$beta +
      sum_squared_residual / 2,
    posterior_rms = sqrt(beta_posterior / (alpha_posterior - 1)),
    posterior_median = sqrt(1 / stats::qgamma(
      0.5, shape = alpha_posterior, rate = beta_posterior
    )),
    posterior_lower = sqrt(1 / stats::qgamma(
      0.975, shape = alpha_posterior, rate = beta_posterior
    )),
    posterior_upper = sqrt(1 / stats::qgamma(
      0.025, shape = alpha_posterior, rate = beta_posterior
    ))
  ) |>
  dplyr::ungroup() |>
  dplyr::select(
    target, feature, n_years, prior_rms, posterior_rms,
    posterior_median, posterior_lower, posterior_upper
  )
feature_error_posteriors
```

## Observed and MAP seasonal cycles

For fluxes, modelled values are averaged only over days with a corresponding
quality-controlled observation. LAI observations retain the MODIS acquisition
dates. The shaded envelope spans the minimum to maximum observed value for each
day of year across years; the black and orange lines show the observed and MAP-model
means. Faceting with free vertical scales keeps all three targets legible.

```{r map-seasonal-cycles, fig.width=7.5, fig.height=8}
seasonal_cycle <- function(data, observed_name, modelled_name, target_name) {
  data |>
    dplyr::filter(
      is.finite(.data[[observed_name]]),
      is.finite(.data[[modelled_name]]),
      !(lubridate::month(date) == 2 & lubridate::mday(date) == 29)
    ) |>
    dplyr::mutate(
      day = lubridate::yday(date) - dplyr::if_else(
        lubridate::leap_year(date) & lubridate::month(date) > 2,
        1L, 0L
      )
    ) |>
    dplyr::group_by(day) |>
    dplyr::summarise(
      observed_min = min(.data[[observed_name]]),
      observed_mean = mean(.data[[observed_name]]),
      observed_max = max(.data[[observed_name]]),
      modelled_mean = mean(.data[[modelled_name]]),
      .groups = "drop"
    ) |>
    dplyr::mutate(target = target_name)
}

matched_flux_cycles <- flux_observations |>
  dplyr::inner_join(map_daily, by = c("sitename", "date"))

flux_cycles <- dplyr::bind_rows(
  seasonal_cycle(matched_flux_cycles, "gpp", "gpp_modelled", "GPP"),
  seasonal_cycle(matched_flux_cycles, "reco", "reco_modelled", "Reco")
)

lai_cycles <- lai_observations |>
  dplyr::inner_join(
    dplyr::select(map_daily, sitename, date, lai_modelled),
    by = c("sitename", "date")
  ) |>
  seasonal_cycle("lai", "lai_modelled", "LAI")

dplyr::bind_rows(flux_cycles, lai_cycles) |>
  ggplot(aes(day)) +
  geom_ribbon(
    aes(
      ymin = observed_min,
      ymax = observed_max,
      fill = "Observed min–max"
    ),
    alpha = 0.25,
    colour = NA
  ) +
  geom_line(aes(y = observed_mean, colour = "Observed mean"), linewidth = 0.7) +
  geom_line(aes(y = modelled_mean, colour = "MAP model mean"), linewidth = 0.8) +
  facet_wrap(~ target, scales = "free_y", ncol = 1) +
  scale_x_continuous(
    breaks = c(15, 46, 74, 105, 135, 166, 196, 227, 258, 288, 319, 349),
    labels = month.abb
  ) +
  scale_colour_manual(values = c(
    "Observed mean" = "black",
    "MAP model mean" = "#D55E00"
  )) +
  scale_fill_manual(values = c("Observed min–max" = "grey55")) +
  labs(x = NULL, y = NULL, colour = NULL, fill = NULL) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

## MAP constraints and parameter estimates

```{r map-summary}
map_parameters <- tibble::tibble(
  parameter = names(calibration_parameters),
  prior_centre = vapply(
    calibration_parameters, function(x) x$mean, numeric(1)
  ),
  map = unlist(pars_map[names(calibration_parameters)])
)
map_parameters

tibble::tibble(
  quantity = c(
    "Peak-season median LMA (g C m-2 leaf)",
    "Wood allocation",
    "Mean standing wood C"
  ),
  value = c(
    median(map_daily$lma[map_daily$lma > 0], na.rm = TRUE),
    pars_map$frac_wood,
    mean(map_daily$cwood, na.rm = TRUE)
  )
)
```

## References

Arán Paredes, J., Bernhard, F., Hufkens, K., Marcadella, M., and Stocker,
B. D. (2025). rsofun v5.1: a model-data integration framework for simulating
ecosystem processes. *Geoscientific Model Development*, 18, 9855–9878.
doi:10.5194/gmd-18-9855-2025.

Leifeld, J., Meyer, S., Budge, K., Sebastia, M.-T., Zimmermann, M., and Fuhrer,
J. (2015). Turnover of grassland roots in mountain ecosystems revealed by
radiocarbon analysis: role of temperature and management. *PLOS ONE*, 10,
e0119184. doi:10.1371/journal.pone.0119184.
