---
title: "Calibrating the C-model at CH-Oe1"
author: "Benjamin Stocker"
output: rmarkdown::html_vignette
code-fold: true
vignette: >
  %\VignetteIndexEntry{Calibrating the C-model at CH-Oe1}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.5
)
```

## Aim and calibration sequence

This vignette contains the complete calibration workflow for the carbon-only
C-model at the temperate grassland site CH-Oe1. The calibration is deliberately
staged so that the ecosystem flux data do not compensate for poorly constrained
photosynthesis, leaf structure, turnover, or allocation parameters.

1. Initialise the P-model parameters at the posterior MAP estimates from setup
   S6 of Arán Paredes et al. (2025), Table 2.
2. Estimate the structural leaf parameter against a representative temperate-
   grass leaf mass per area (LMA).
3. Estimate leaf and root turnover using biologically constrained priors for a
   temperate grass.
4. Fix allocation to wood to zero.
5. Diagnose posterior-weight collapse by varying only `kphio` against GPP while
   holding every other model parameter constant, comparing daily and
   mean-annual likelihoods.
6. Restore the multiple-parameter calibration against monthly GPP, monthly
   ecosystem respiration (Reco), and LAI.

Monthly aggregation reduces the effect of strongly autocorrelated daily
residuals. It does not assert that monthly residuals are necessarily independent;
that assumption should still be checked after calibration.

The completed LMA calibration is shown with `eval=FALSE` and its cached result
is used below. The flux calibration remains executable and starts with a
reduced LHS design.

Functions evaluated on parallel workers use explicit namespace qualifiers such
as `dplyr::filter()` and `tidyr::unnest()`. Worker processes do not inherit the
packages attached in the interactive R session, so unqualified package
functions should be avoided inside a likelihood.

```{r libraries}
library(rsofun)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

# Rendering with library(rsofun) uses the installed package, not uninstalled
# edits under R/. Fail before starting expensive simulations if that installed
# namespace does not yet contain Fortran-STOP worker recovery.
if (!exists(
  ".calib_lhs_worker_stop_recovery",
  envir = asNamespace("rsofun"),
  inherits = FALSE
)) {
  stop(
    "The installed rsofun is stale. Reinstall this working tree, or run ",
    "devtools::load_all() before rendering the vignette."
  )
}
```

## Drivers and observations

This example uses the FluxDataKit v3.4.2 driver object. Adjust `fdk_path` to the
local download location. The forcing contains daily GPP and NEE observations;
observed Reco is calculated from the ecosystem carbon balance as
`Reco = NEE + GPP` under the FLUXNET sign convention used here.

```{r data, eval=TRUE}
fdk_path <- "~/data_2/FluxDataKit/v3.4/zenodo_upload"

df_drivers <- readRDS(file.path(
  fdk_path,
  "rsofun_driver_data_v3.4.2.rds"
)) |>
  filter(sitename == "CH-Oe1")

use_cseed <- 5
use_nseed <- use_cseed / 20

df_drivers <- df_drivers |>
  mutate(
    forcing = purrr::map(forcing, ~ mutate(
      .,
      fharv = 0,
      dno3 = 1.5 / (2 * 365),
      dnh4 = 1.5 / (2 * 365),
      cseed = ifelse(month(date) == 2 & mday(date) == 15, use_cseed, 0),
      nseed = ifelse(month(date) == 2 & mday(date) == 15, use_nseed, 0)
    ))
  )

df_drivers$params_siml[[1]]$spinupyears <- 2002
df_drivers$params_siml[[1]]$recycle <- 5
df_drivers$params_siml[[1]]$c_only <- TRUE

daily_observations <- df_drivers |>
  select(sitename, forcing) |>
  unnest(forcing) |>
  transmute(
    sitename,
    date,
    gpp = if_else(gpp_qc >= 0.8, gpp, NA_real_),
    reco = if_else(gpp_qc >= 0.8 & nee_qc >= 0.8, gpp + nee, NA_real_)
  )

# Quality-controlled MODIS MOD15A2H LAI; see example_cmodel.Rmd for QC details.
data("modis_lai_ch_oe1", package = "rsofun")
model_period <- range(daily_observations$date)
lai_observations <- modis_lai_ch_oe1 |>
  filter(between(date, model_period[1], model_period[2])) |>
  select(date, lai)
```

## Baseline parameters

The P-model values below reproduce the MAP estimates in the last column of
Table 2 (setup S6) of Arán Paredes et al. (2025): `kphio = 0.08`,
`kphio_par_a = -0.001`, `kphio_par_b = 27.5`, `soilm_thetastar = 26.7`,
`beta_unitcostratio = 187.5`, `tau_acclim = 1.0`, and `kc_jmax = 0.58`.
Parameters that are not estimated in setup S6 retain their existing model
defaults. Note that the S6 value of `beta_unitcostratio` lies at the lower bound
of its truncated normal prior; it should not be replaced by the posterior mean
or by the rounded value discussed elsewhere in the article.

The three quantum-yield parameters have a distinct and coupled meaning:

```text
phi(T) = kphio * max(0, min(1, 1 + kphio_par_a * (T - kphio_par_b)^2))
```

Thus `kphio` is the maximum quantum yield, `kphio_par_a` controls the curvature,
and `kphio_par_b` locates the temperature optimum. The shape and location
parameters remain fixed at their S6 estimates. For the deliberately simplified
diagnostic calibration below, only the maximum, `kphio`, is allowed to depart
from its S6 estimate. All C-model parameters remain fixed.

The turnover constants are annual values. We use baseline values of `1.9
yr-1` for leaves and `0.7 yr-1` for roots, corresponding to expected tissue
lifetimes of about 0.53 and 1.43 years, respectively. These values also define
the prior centres below, but both rates are estimated during calibration. Wood
allocation is fixed to zero throughout.

The model reports LMA in grams of carbon per square metre of leaf. Define the
target in the same units. The value below is an explicit working assumption of
`25 g C m-2`, approximately `50 g dry mass m-2` for a carbon fraction of 0.5.
Replace it with a site- or trait-database-derived value and uncertainty when
available.

LMA depends directly on `r_ctostructn_leaf` and also on the partitioning of leaf
nitrogen into structural and metabolic components. To avoid an underidentified
trait-only calibration, keep `r_ctostructn_leaf`, `ncw_min`, `r_n_cw_v`, and `nv_vcmax25`
fixed. Since `r_n_cw_v = 0` here, LMA depends only on `r_ctostructn_leaf` and is set to a value of 30.9.

```{r parameters}
pars <- list(
  # P-model: informed by Arán Paredes et al. (2025), setup S6
  kphio = 0.08,
  kphio_par_a = -0.001,
  kphio_par_b = 27.5,
  # C-model phenology; unrelated to the P-model quantum-yield parameters above
  dehardening_gdd_midpoint = 150.0,
  dehardening_gdd_slope = 0.05,
  dehardening_gdd_base = 5.0,
  soilm_thetastar = 26.7,
  soilm_betao = 0.0,
  beta_unitcostratio = 187.5,
  rd_to_vcmax = 0.014,
  tau_acclim = 1.0,
  kc_jmax = 0.58,

  # Plant C and N
  f_nretain = 0.5,
  fpc_tree_max = 0.95,
  growtheff = 0.6,
  r_root = 2 * 0.913,
  r_sapw = 2 * 0.044,
  exurate = 0.003,
  k_decay_leaf = 1.9,
  k_decay_root = 0.7,
  k_decay_labl = 1.9,
  k_decay_sapw = 1.9,
  r_cton_root = 37.0,
  r_cton_wood = 100.0,
  r_cton_seed = 15.0,
  nv_vcmax25 = 0.02 * 13681.77,
  ncw_min = 0.08 * 1.116222,
  r_n_cw_v = 0.0,
  r_ctostructn_leaf = 30.9, # from LMA calibration below
  kbeer = 0.5,

  # Phenology
  gddbase = 5.0,
  ramp = 0.0,
  phentype = 2.0,

  # Soil physics
  perc_k1 = 5.0,
  thdiff_wp = 0.2,
  thdiff_whc15 = 0.8,
  thdiff_fc = 0.4,
  forg = 0.01,
  wbwp = 0.029,
  por = 0.421,
  fsand = 0.82,
  fclay = 0.06,
  fsilt = 0.12,

  # Water and energy balance
  kA = 107.0,
  kalb_sw = 0.17,
  kalb_vis = 0.03,
  kb = 0.20,
  kc = 0.25,
  kCw = 1.05,
  kd = 0.50,
  ke = 0.0167,
  keps = 23.44,
  kWm = 220.0,
  kw = 0.26,
  komega = 283.0,
  maxmeltrate = 3.0,

  # Soil biogeochemistry
  klitt_af10 = 1.2,
  klitt_as10 = 0.35,
  klitt_bg10 = 0.35,
  kexu10 = 50.0,
  ksoil_fs10 = 0.021,
  ksoil_sl10 = 7.0e-4,
  ntoc_crit1 = 0.45,
  ntoc_crit2 = 0.76,
  cton_microb = 10.0,
  cton_soil = 9.77,
  fastfrac = 0.985,

  # N uptake and transformations
  eff_nup = 1.0e-4,
  minimumcostfix = 1.0,
  fixoptimum = 25.15,
  a_param_fix = -3.62,
  b_param_fix = 0.27,
  maxnitr = 5.0e-5,
  non = 0.01,
  n2on = 5.0e-4,
  kn = 83.0,
  kdoc = 17.0,
  docmax = 1.0,
  dnitr2n2o = 0.01,

  # Prescribed allocation in the C-only setup
  frac_leaf = 0.5,
  frac_wood = 0.0,
  frac_avl_labl = 0.1,
  tmppar = 9999.0,
  nuptake_kc = 600.0,
  nuptake_kv = 5.0,
  nuptake_vmax = 0.2
)

stopifnot(pars$frac_wood == 0)

# LMA is already fixed in `pars`; retain this alias for the broader calibration
# setup below.
pars_lma <- pars
```

## Calibration setup 1: `kphio` and GPP model error

This diagnostic setup varies only `kphio` and one GPP residual standard
deviation. The `kphio` prior is normal with mean `0.08` and standard deviation
`0.04`, truncated to `0.03--0.12`. All other model parameters remain fixed.
Two likelihoods use the same simulations and priors but retain very different
amounts of temporal information:

- the daily likelihood treats every quality-controlled daily GPP residual as
  an observation;
- the mean-annual likelihood reduces the record to one value: the mean of the
  paired annual GPP totals.

The comparison reveals how repeated, autocorrelated observations can make the
likelihood much sharper than a single aggregated target.
The daily error is expressed in daily GPP units; the annual error is expressed
in annual-total GPP units, so their numerical MAP values are not directly
comparable.

```{r gpp-likelihoods}
daily_gpp_loglik <- function(par, model_error, drivers, observations) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp)

  matched <- observations |>
    dplyr::select(sitename, date, gpp) |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(is.finite(gpp), is.finite(gpp_modelled))

  if (nrow(matched) == 0) return(-Inf)

  sum(stats::dnorm(
    matched$gpp,
    mean = matched$gpp_modelled,
    sd = model_error$sigma_gpp_daily,
    log = TRUE
  ))
}

mean_annual_gpp_loglik <- function(par, model_error, drivers, observations) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp)

  annual_gpp <- observations |>
    dplyr::select(sitename, date, gpp) |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(is.finite(gpp), is.finite(gpp_modelled)) |>
    dplyr::mutate(year = lubridate::year(date)) |>
    dplyr::group_by(sitename, year) |>
    dplyr::summarise(
      observed = sum(gpp),
      modelled = sum(gpp_modelled),
      .groups = "drop"
    )

  if (nrow(annual_gpp) == 0) return(-Inf)

  mean_annual <- annual_gpp |>
    dplyr::summarise(
      observed = mean(observed),
      modelled = mean(modelled)
    )

  stats::dnorm(
    mean_annual$observed,
    mean = mean_annual$modelled,
    sd = model_error$sigma_gpp_annual,
    log = TRUE
  )
}
```

```{r simple-gpp-calibrations, eval=TRUE}
kphio_prior <- list(
  kphio = list(
    mean = 0.08, sd = 0.04,
    distribution = "normal", lower = 0.03, upper = 0.12
  )
)
stopifnot(identical(names(kphio_prior), "kphio"))

daily_gpp_calibration <- calib_sofun_lhs(
  pars = pars,
  par_calib = kphio_prior,
  model_error = list(
    sigma_gpp_daily = list(
      mean = 0.8, sd = 0.3,
      distribution = "lognormal", lower = 0.05, upper = 4
    )
  ),
  likelihood = daily_gpp_loglik,
  n_samples = 250,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 42,
  progress = TRUE,
  drivers = df_drivers,
  observations = daily_observations
)

annual_gpp_calibration <- calib_sofun_lhs(
  pars = pars,
  par_calib = kphio_prior,
  model_error = list(
    sigma_gpp_annual = list(
      mean = 100, sd = 50,
      distribution = "lognormal", lower = 10, upper = 500
    )
  ),
  likelihood = mean_annual_gpp_loglik,
  n_samples = 250,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 43,
  progress = TRUE,
  drivers = df_drivers,
  observations = daily_observations
)
```

### Prior and posterior distributions of `kphio`

The unweighted LHS draws approximate the prior; the normalized importance
weights applied to those same draws approximate the posterior.

```{r simple-kphio-prior-posterior, eval=TRUE}
distribution_data <- function(calibration, target_label) {
  dplyr::bind_rows(
    calibration$samples |>
      dplyr::transmute(
        kphio, target = target_label, distribution = "Prior",
        density_weight = 1 / dplyr::n()
      ),
    calibration$samples |>
      dplyr::transmute(
        kphio, target = target_label, distribution = "Posterior",
        density_weight = weight
      )
  )
}

kphio_distributions <- dplyr::bind_rows(
  distribution_data(daily_gpp_calibration, "Daily GPP"),
  distribution_data(annual_gpp_calibration, "One mean annual GPP value")
)

ggplot(
  kphio_distributions,
  aes(kphio, weight = density_weight, colour = distribution, fill = distribution)
) +
  geom_density(alpha = 0.2) +
  geom_vline(xintercept = pars$kphio, linetype = "dashed") +
  facet_wrap(~ target, ncol = 1) +
  scale_colour_manual(values = c(Prior = "grey40", Posterior = "red")) +
  scale_fill_manual(values = c(Prior = "grey70", Posterior = "red")) +
  labs(x = "Maximum quantum yield (kphio)", y = "Density", colour = NULL,
       fill = NULL) +
  theme_classic()
```

```{r simple-calibration-diagnostics, eval=TRUE}
tibble(
  target = c("Daily GPP", "One mean annual GPP value"),
  effective_sample_size = c(
    daily_gpp_calibration$effective_sample_size,
    annual_gpp_calibration$effective_sample_size
  ),
  maximum_weight = c(
    max(daily_gpp_calibration$samples$weight),
    max(annual_gpp_calibration$samples$weight)
  ),
  map_kphio = c(
    daily_gpp_calibration$par$kphio,
    annual_gpp_calibration$par$kphio
  ),
  map_gpp_error = c(
    daily_gpp_calibration$model_error$sigma_gpp_daily,
    annual_gpp_calibration$model_error$sigma_gpp_annual
  )
)
```

### Mean annual cycle of daily GPP

Re-run the C-model with the MAP `kphio` from each calibration. To make the
seasonal cycles directly comparable, calculate modelled means only for dates
that have a quality-controlled GPP observation.

```{r simple-gpp-annual-cycle, eval=TRUE}
map_gpp_cycle <- function(calibration, source_label) {
  rsofun::runread_cnmodel_f(
    drivers = df_drivers,
    par = calibration$par
  ) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp) |>
    dplyr::inner_join(
      dplyr::filter(daily_observations, is.finite(gpp)),
      by = c("sitename", "date")
    ) |>
    dplyr::filter(is.finite(gpp_modelled)) |>
    dplyr::mutate(day_of_year = lubridate::yday(date)) |>
    dplyr::group_by(day_of_year) |>
    dplyr::summarise(gpp = mean(gpp_modelled), .groups = "drop") |>
    dplyr::mutate(source = source_label)
}

observed_gpp_cycle <- daily_observations |>
  dplyr::filter(is.finite(gpp)) |>
  dplyr::mutate(day_of_year = lubridate::yday(date)) |>
  dplyr::group_by(day_of_year) |>
  dplyr::summarise(gpp = mean(gpp), .groups = "drop") |>
  dplyr::mutate(source = "Observed")

gpp_annual_cycles <- dplyr::bind_rows(
  observed_gpp_cycle,
  map_gpp_cycle(daily_gpp_calibration, "MAP from daily GPP"),
  map_gpp_cycle(annual_gpp_calibration, "MAP from mean annual GPP")
)

ggplot(gpp_annual_cycles, aes(day_of_year, gpp, colour = source)) +
  geom_line() +
  scale_x_continuous(
    breaks = c(15, 46, 74, 105, 135, 166, 196, 227, 258, 288, 319, 349),
    labels = month.abb
  ) +
  scale_colour_manual(values = c(
    "Observed" = "black",
    "MAP from daily GPP" = "tomato",
    "MAP from mean annual GPP" = "royalblue"
  )) +
  labs(
    x = NULL,
    y = expression("Mean daily GPP (g C " * m^-2 * d^-1 * ")"),
    colour = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom"
  )
```

<!--
## Calibration setup 2: multiple parameters and observations

This restores the broader calibration against monthly GPP, monthly Reco, and
MODIS LAI. The modelled respiratory components are summed as
`Reco = rleaf + rroot + rwood + rhet + rgrow`. Months require at least 20 valid
daily observations. LAI is evaluated on the MODIS acquisition dates.

```{r monthly-likelihood, eval=FALSE}
monthly_flux_loglik <- function(
    par, model_error, drivers, observations, lai_observations,
    minimum_days = 20
) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(
      sitename, date, gpp_modelled = gpp,
      reco_modelled = rleaf + rroot + rwood + rhet + rgrow,
      lai_modelled = lai
    )

  matched <- observations |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::mutate(month = lubridate::floor_date(date, "month"))

  aggregate_monthly <- function(data, observed, modelled) {
    data |>
      dplyr::filter(is.finite({{ observed }}), is.finite({{ modelled }})) |>
      dplyr::group_by(sitename, month) |>
      dplyr::summarise(
        n_days = dplyr::n(), observed = mean({{ observed }}),
        modelled = mean({{ modelled }}), .groups = "drop"
      ) |>
      dplyr::filter(n_days >= minimum_days)
  }

  monthly_gpp <- aggregate_monthly(matched, gpp, gpp_modelled)
  monthly_reco <- aggregate_monthly(matched, reco, reco_modelled)
  matched_lai <- lai_observations |>
    dplyr::inner_join(dplyr::select(simulated, date, lai_modelled), by = "date") |>
    dplyr::filter(is.finite(lai), is.finite(lai_modelled))

  if (!nrow(monthly_gpp) || !nrow(monthly_reco) || !nrow(matched_lai)) return(-Inf)

  sum(stats::dnorm(monthly_gpp$observed, monthly_gpp$modelled,
                   model_error$sigma_gpp_monthly, log = TRUE)) +
    sum(stats::dnorm(monthly_reco$observed, monthly_reco$modelled,
                     model_error$sigma_reco_monthly, log = TRUE)) +
    sum(stats::dnorm(matched_lai$lai, matched_lai$lai_modelled,
                     model_error$sigma_lai, log = TRUE))
}
```

Some samples can violate C--N pool constraints. A failed run is retained with
`log_likelihood = -Inf`, `weight = 0`, and `evaluation_ok = FALSE`.

```{r monthly-calibration, eval=FALSE}
flux_parameters <- list(
  dehardening_gdd_midpoint = list(
    mean = pars_lma$dehardening_gdd_midpoint, sd = 30,
    distribution = "normal", lower = 30, upper = 300
  ),
  dehardening_gdd_slope = list(
    mean = pars_lma$dehardening_gdd_slope, sd = 0.015,
    distribution = "lognormal", lower = 0.005, upper = 0.15
  ),
  dehardening_gdd_base = list(
    mean = pars_lma$dehardening_gdd_base, sd = 1.5,
    distribution = "normal", lower = 0, upper = 12
  ),
  growtheff = list(
    mean = pars_lma$growtheff, sd = 0.06,
    distribution = "normal", lower = 0.35, upper = 0.85
  ),
  frac_leaf = list(
    mean = pars_lma$frac_leaf, sd = 0.08,
    distribution = "normal", lower = 0.1, upper = 0.9
  ),
  k_decay_leaf = list(
    mean = 1.9, sd = 0.6,
    distribution = "lognormal", lower = 0.7, upper = 4.0
  ),
  k_decay_root = list(
    mean = 0.7, sd = 0.25,
    distribution = "lognormal", lower = 0.1, upper = 1.5
  ),
  r_root = list(
    mean = pars_lma$r_root, sd = 0.45,
    distribution = "lognormal", lower = 0.5, upper = 3.5
  )
)

flux_errors <- list(
  sigma_gpp_monthly = list(
    mean = 0.8, sd = 0.3,
    distribution = "lognormal", lower = 0.05, upper = 4
  ),
  sigma_reco_monthly = list(
    mean = 0.8, sd = 0.3,
    distribution = "lognormal", lower = 0.05, upper = 4
  ),
  sigma_lai = list(
    mean = 0.5, sd = 0.2,
    distribution = "lognormal", lower = 0.1, upper = 2
  )
)

flux_calibration <- calib_sofun_lhs(
  pars = pars_lma,
  par_calib = flux_parameters,
  model_error = flux_errors,
  likelihood = monthly_flux_loglik,
  n_samples = 25,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 44,
  progress = TRUE,
  drivers = df_drivers,
  observations = daily_observations,
  lai_observations = lai_observations
)
```

### Multiple-target posterior diagnostics

```{r posterior-diagnostics, eval=FALSE}
posterior_diagnostics <- flux_calibration$samples |>
  arrange(desc(weight)) |>
  mutate(cumulative_weight = cumsum(weight))

n_to_80_percent <- which(posterior_diagnostics$cumulative_weight >= 0.8)[1]

list(
  effective_sample_size = flux_calibration$effective_sample_size,
  failed_evaluations = sum(!flux_calibration$samples$evaluation_ok),
  n_to_80_percent = n_to_80_percent,
  top_samples = slice_head(posterior_diagnostics, n = 10)
)
```

## Re-run and evaluate the MAP simulation

The returned MAP list contains all fixed and calibrated parameters, including
the zero wood-allocation constraint. Re-run the model once with this list before
evaluating either target.

```{r map-run, eval=FALSE}
pars_map <- flux_calibration$par
stopifnot(pars_map$frac_wood == 0)

output_map <- runread_cnmodel_f(
  drivers = df_drivers,
  par = pars_map
)

map_daily <- output_map |>
  select(sitename, data) |>
  unnest(data) |>
  transmute(
    sitename,
    date,
    gpp_modelled = gpp,
    reco_modelled = rleaf + rroot + rwood + rhet + rgrow,
    lai_modelled = lai,
    lma,
    cleaf,
    croot,
    cwood
  )

map_monthly <- daily_observations |>
  inner_join(map_daily, by = c("sitename", "date")) |>
  mutate(month = floor_date(date, "month")) |>
  group_by(sitename, month) |>
  summarise(
    gpp_observed = mean(gpp, na.rm = TRUE),
    gpp_modelled = mean(gpp_modelled[is.finite(gpp)], na.rm = TRUE),
    reco_observed = mean(reco, na.rm = TRUE),
    reco_modelled = mean(reco_modelled[is.finite(reco)], na.rm = TRUE),
    .groups = "drop"
  )
```

```{r map-plots, eval=FALSE}
map_monthly |>
  select(sitename, month, starts_with("gpp"), starts_with("reco")) |>
  pivot_longer(
    -c(sitename, month),
    names_to = c("variable", "source"),
    names_sep = "_",
    values_to = "value"
  ) |>
  ggplot(aes(month, value, colour = source)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y", ncol = 1) +
  scale_colour_manual(values = c(observed = "black", modelled = "red")) +
  labs(x = NULL, y = expression("Monthly mean flux (g C " * m^-2 * d^-1 * ")")) +
  theme_classic()
```

Evaluate LAI on the MODIS acquisition dates rather than interpolating the
observations to every model day:

```{r map-lai-plot, eval=FALSE}
map_lai <- lai_observations |>
  inner_join(select(map_daily, date, lai_modelled), by = "date")

ggplot(map_lai, aes(date)) +
  geom_line(aes(y = lai_modelled, colour = "Modelled"), linewidth = 0.4) +
  geom_point(aes(y = lai, colour = "MODIS"), size = 1.2, alpha = 0.8) +
  scale_colour_manual(values = c(Modelled = "red", MODIS = "black")) +
  labs(x = NULL, y = expression("LAI (m"^2 * " m"^-2 * ")"), colour = NULL) +
  theme_classic()
```

Finally, confirm that the staged constraints remain satisfied in the MAP run:

```{r map-constraints, eval=FALSE}
map_checks <- tibble(
  quantity = c(
    "Peak-season median LMA (g C m-2 leaf)",
    "Mean wood allocation",
    "Mean standing wood C"
  ),
  value = c(
    median(map_daily$lma[map_daily$lma > 0], na.rm = TRUE),
    pars_map$frac_wood,
    mean(map_daily$cwood, na.rm = TRUE)
  )
)

map_checks
```

-->

## Reference

Arán Paredes, J., Bernhard, F., Hufkens, K., Marcadella, M., and Stocker,
B. D. (2025). rsofun v5.1: a model-data integration framework for simulating
ecosystem processes. *Geoscientific Model Development*, 18, 9855–9878.
doi:10.5194/gmd-18-9855-2025.

Leifeld, J., Meyer, S., Budge, K., Sebastia, M.-T., Zimmermann, M., and Fuhrer,
J. (2015). Turnover of grassland roots in mountain ecosystems revealed by
radiocarbon analysis: role of temperature and management. *PLOS ONE*, 10,
e0119184. doi:10.1371/journal.pone.0119184.
