---
title: "Calibrating kphio with GPP features"
author: "Benjamin Stocker"
output: rmarkdown::html_vignette
code-fold: true
vignette: >
  %\VignetteIndexEntry{Calibrating kphio with GPP features}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.5
)
```

## Aim

This vignette isolates one question: how does the temporal representation of
GPP observations affect a Bayesian calibration of the maximum quantum yield
`kphio`? All other C-model parameters are held fixed for the temperate
grassland site CH-Oe1. This deliberately simple setup makes posterior-weight
collapse visible without confounding it with parameter interactions.

Four likelihood constructions are compared: daily GPP, monthly mean GPP, one
mean annual GPP value, and annually replicated seasonal GPP features. The first
three diagnose how repeated autocorrelated residuals affect the finite LHS
posterior. The feature likelihood provides the proposed solution: retain
seasonal timing, magnitude, and interannual information while integrating
feature-specific error variances under transparent inverse-gamma priors.

The companion vignette
`vignette("calibration_cmodel_multitarget_ch_oe1")` extends this approach to a
joint calibration of multiple C-model parameters against GPP, ecosystem
respiration (Reco), and LAI.

Functions evaluated on parallel workers use explicit namespace qualifiers such
as `dplyr::filter()` and `tidyr::unnest()`. Worker processes do not inherit the
packages attached in the interactive R session, so unqualified package
functions should be avoided inside a likelihood.

```{r libraries}
library(rsofun)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

# Rendering with library(rsofun) uses the installed package, not uninstalled
# edits under R/. Fail before starting expensive simulations if that installed
# namespace does not yet contain Fortran-STOP worker recovery.
if (!exists(
  ".calib_lhs_worker_stop_recovery",
  envir = asNamespace("rsofun"),
  inherits = FALSE
)) {
  stop(
    "The installed rsofun is stale. Reinstall this working tree, or run ",
    "devtools::load_all() before rendering the vignette."
  )
}
```

## Drivers and observations

This example uses the FluxDataKit v3.4.2 driver object. Adjust `fdk_path` to the
local download location. Only quality-controlled daily GPP is used as a target
in this vignette.

```{r data, eval=TRUE}
fdk_path <- "~/data_2/FluxDataKit/v3.4/zenodo_upload"

df_drivers <- readRDS(file.path(
  fdk_path,
  "rsofun_driver_data_v3.4.2.rds"
)) |>
  filter(sitename == "CH-Oe1")

use_cseed <- 5
use_nseed <- use_cseed / 20

df_drivers <- df_drivers |>
  mutate(
    forcing = purrr::map(forcing, ~ mutate(
      .,
      fharv = 0,
      dno3 = 1.5 / (2 * 365),
      dnh4 = 1.5 / (2 * 365),
      cseed = ifelse(month(date) == 2 & mday(date) == 15, use_cseed, 0),
      nseed = ifelse(month(date) == 2 & mday(date) == 15, use_nseed, 0)
    ))
  )

df_drivers$params_siml[[1]]$spinupyears <- 2002
df_drivers$params_siml[[1]]$recycle <- 5
df_drivers$params_siml[[1]]$c_only <- TRUE

daily_observations <- df_drivers |>
  select(sitename, forcing) |>
  unnest(forcing) |>
  transmute(
    sitename,
    date,
    gpp = if_else(gpp_qc >= 0.8, gpp, NA_real_)
  )
```

## Baseline parameters

The P-model values below reproduce the MAP estimates in the last column of
Table 2 (setup S6) of Arán Paredes et al. (2025): `kphio = 0.08`,
`kphio_par_a = -0.001`, `kphio_par_b = 27.5`, `soilm_thetastar = 26.7`,
`beta_unitcostratio = 187.5`, `tau_acclim = 1.0`, and `kc_jmax = 0.58`.
Parameters that are not estimated in setup S6 retain their existing model
defaults. Note that the S6 value of `beta_unitcostratio` lies at the lower bound
of its truncated normal prior; it should not be replaced by the posterior mean
or by the rounded value discussed elsewhere in the article.

The three quantum-yield parameters have a distinct and coupled meaning:

```text
phi(T) = kphio * max(0, min(1, 1 + kphio_par_a * (T - kphio_par_b)^2))
```

Thus `kphio` is the maximum quantum yield, `kphio_par_a` controls the curvature,
and `kphio_par_b` locates the temperature optimum. The shape and location
parameters remain fixed at their S6 estimates. For the deliberately simplified
diagnostic calibration below, only the maximum, `kphio`, is allowed to depart
from its S6 estimate. All C-model parameters remain fixed.

The turnover constants are annual values. Baseline values of `1.9 yr-1` for
leaves and `0.7 yr-1` for roots correspond to expected tissue lifetimes of
about 0.53 and 1.43 years. They remain fixed here, as does zero allocation to
wood.

Leaf structure and nitrogen partitioning are also fixed. In particular,
`r_ctostructn_leaf = 30.9` is the previously selected value for a temperate
grass; it is not inferred from GPP in this diagnostic vignette.

```{r parameters}
pars <- list(
  # P-model: informed by Arán Paredes et al. (2025), setup S6
  kphio = 0.08,
  kphio_par_a = -0.001,
  kphio_par_b = 27.5,
  # C-model phenology; unrelated to the P-model quantum-yield parameters above
  dehardening_gdd_midpoint = 150.0,
  dehardening_gdd_slope = 0.05,
  dehardening_gdd_base = 5.0,
  soilm_thetastar = 26.7,
  soilm_betao = 0.0,
  beta_unitcostratio = 187.5,
  rd_to_vcmax = 0.014,
  tau_acclim = 1.0,
  kc_jmax = 0.58,

  # Plant C and N
  f_nretain = 0.5,
  fpc_tree_max = 0.95,
  growtheff = 0.6,
  r_root = 2 * 0.913,
  r_sapw = 2 * 0.044,
  exurate = 0.003,
  k_decay_leaf = 1.9,
  k_decay_root = 0.7,
  k_decay_labl = 1.9,
  k_decay_sapw = 1.9,
  r_cton_root = 37.0,
  r_cton_wood = 100.0,
  r_cton_seed = 15.0,
  nv_vcmax25 = 0.02 * 13681.77,
  ncw_min = 0.08 * 1.116222,
  r_n_cw_v = 0.0,
  r_ctostructn_leaf = 30.9,
  kbeer = 0.5,

  # Phenology
  gddbase = 5.0,
  ramp = 0.0,
  phentype = 2.0,

  # Soil physics
  perc_k1 = 5.0,
  thdiff_wp = 0.2,
  thdiff_whc15 = 0.8,
  thdiff_fc = 0.4,
  forg = 0.01,
  wbwp = 0.029,
  por = 0.421,
  fsand = 0.82,
  fclay = 0.06,
  fsilt = 0.12,

  # Water and energy balance
  kA = 107.0,
  kalb_sw = 0.17,
  kalb_vis = 0.03,
  kb = 0.20,
  kc = 0.25,
  kCw = 1.05,
  kd = 0.50,
  ke = 0.0167,
  keps = 23.44,
  kWm = 220.0,
  kw = 0.26,
  komega = 283.0,
  maxmeltrate = 3.0,

  # Soil biogeochemistry
  klitt_af10 = 1.2,
  klitt_as10 = 0.35,
  klitt_bg10 = 0.35,
  kexu10 = 50.0,
  ksoil_fs10 = 0.021,
  ksoil_sl10 = 7.0e-4,
  ntoc_crit1 = 0.45,
  ntoc_crit2 = 0.76,
  cton_microb = 10.0,
  cton_soil = 9.77,
  fastfrac = 0.985,

  # N uptake and transformations
  eff_nup = 1.0e-4,
  minimumcostfix = 1.0,
  fixoptimum = 25.15,
  a_param_fix = -3.62,
  b_param_fix = 0.27,
  maxnitr = 5.0e-5,
  non = 0.01,
  n2on = 5.0e-4,
  kn = 83.0,
  kdoc = 17.0,
  docmax = 1.0,
  dnitr2n2o = 0.01,

  # Prescribed allocation in the C-only setup
  frac_leaf = 0.5,
  frac_wood = 0.0,
  frac_avl_labl = 0.1,
  tmppar = 9999.0,
  nuptake_kc = 600.0,
  nuptake_kv = 5.0,
  nuptake_vmax = 0.2
)

stopifnot(pars$frac_wood == 0)

```

## Comparing four GPP likelihoods

This diagnostic setup varies only `kphio` and one GPP residual standard
deviation. The `kphio` prior is normal with mean `0.08` and standard deviation
`0.04`, truncated to `0.03--0.12`. All other model parameters remain fixed.
Four likelihoods use the same model-parameter prior but retain very different
amounts of temporal information:

- the daily likelihood treats every quality-controlled daily GPP residual as
  an observation;
- the monthly likelihood uses monthly mean GPP when at least 20 paired daily
  observations are available;
- the mean-annual likelihood reduces the record to one value: the mean of the
  paired annual GPP totals;
- the feature likelihood uses annual GPP totals, peak-week GPP, timing of peak
  GPP, growing-season onset and termination, and seasonal amplitude, each
  calculated separately for individual years. It omits their redundant
  mean-annual aggregate.

The comparison reveals how repeated, autocorrelated observations can make the
likelihood much sharper than a single aggregated target.
The daily and monthly-mean errors are expressed in daily GPP units, whereas the
mean-annual error is expressed in annual-total GPP units. The annual error's
numerical MAP value is therefore not directly comparable with the other two.

```{r gpp-likelihoods}
daily_gpp_loglik <- function(par, model_error, drivers, observations) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp)

  matched <- observations |>
    dplyr::select(sitename, date, gpp) |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(is.finite(gpp), is.finite(gpp_modelled))

  if (nrow(matched) == 0) return(-Inf)

  sum(stats::dnorm(
    matched$gpp,
    mean = matched$gpp_modelled,
    sd = model_error$sigma_gpp_daily,
    log = TRUE
  ))
}

monthly_gpp_loglik <- function(
    par, model_error, drivers, observations, minimum_days = 20
) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp)

  monthly_gpp <- observations |>
    dplyr::select(sitename, date, gpp) |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(is.finite(gpp), is.finite(gpp_modelled)) |>
    dplyr::mutate(month = lubridate::floor_date(date, "month")) |>
    dplyr::group_by(sitename, month) |>
    dplyr::summarise(
      n_days = dplyr::n(),
      observed = mean(gpp),
      modelled = mean(gpp_modelled),
      .groups = "drop"
    ) |>
    dplyr::filter(n_days >= minimum_days)

  if (nrow(monthly_gpp) == 0) return(-Inf)

  sum(stats::dnorm(
    monthly_gpp$observed,
    mean = monthly_gpp$modelled,
    sd = model_error$sigma_gpp_monthly,
    log = TRUE
  ))
}

mean_annual_gpp_loglik <- function(par, model_error, drivers, observations) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp)

  annual_gpp <- observations |>
    dplyr::select(sitename, date, gpp) |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(is.finite(gpp), is.finite(gpp_modelled)) |>
    dplyr::mutate(year = lubridate::year(date)) |>
    dplyr::group_by(sitename, year) |>
    dplyr::summarise(
      observed = sum(gpp),
      modelled = sum(gpp_modelled),
      .groups = "drop"
    )

  if (nrow(annual_gpp) == 0) return(-Inf)

  mean_annual <- annual_gpp |>
    dplyr::summarise(
      observed = mean(observed),
      modelled = mean(modelled)
    )

  stats::dnorm(
    mean_annual$observed,
    mean = mean_annual$modelled,
    sd = model_error$sigma_gpp_annual,
    log = TRUE
  )
}

# Extract one set of features for every year. Annual replication is essential:
# it supplies information about each feature's residual variance. The 15-day
# centred moving average is used only to locate the 15%-of-amplitude crossings.
gpp_features_by_year <- function(data, value, minimum_days = 200) {
  value <- rlang::ensym(value)
  series <- data |>
    dplyr::transmute(
      date, year = lubridate::year(date), day = lubridate::yday(date),
      gpp = !!value
    ) |>
    dplyr::filter(is.finite(gpp))
  series <- split(series, series$year)

  extract_year <- function(x) {
    if (nrow(x) < minimum_days) return(NULL)
    # Interpolate missing observation days before applying the moving average.
    daily <- tibble::tibble(
      day = seq_len(365),
      gpp = stats::approx(x$day, x$gpp, xout = seq_len(365),
                          ties = mean)$y
    ) |>
      dplyr::mutate(
        smooth = as.numeric(stats::filter(gpp, rep(1 / 15, 15), sides = 2)),
        week = (day - 1L) %/% 7L + 1L
      )
    usable <- dplyr::filter(daily, is.finite(smooth))
    if (nrow(usable) < minimum_days) return(NULL)

    weekly <- daily |>
      dplyr::filter(is.finite(gpp)) |>
      dplyr::group_by(week) |>
      dplyr::summarise(gpp = mean(gpp), .groups = "drop")
    peak <- weekly[which.max(weekly$gpp), ]
    lower <- stats::quantile(usable$smooth, 0.05, names = FALSE)
    upper <- stats::quantile(usable$smooth, 0.95, names = FALSE)
    threshold <- lower + 0.15 * (upper - lower)
    peak_day <- usable$day[which.max(usable$smooth)]
    active <- dplyr::filter(usable, smooth >= threshold)
    before_peak <- active$day[active$day <= peak_day]
    after_peak <- active$day[active$day >= peak_day]
    if (!length(before_peak) || !length(after_peak)) return(NULL)

    tibble::tibble(
      year = x$year[[1]],
      annual_gpp = sum(x$gpp),
      peak_gpp = peak$gpp,
      peak_week = peak$week,
      onset = min(before_peak),
      termination = max(after_peak),
      amplitude = upper - lower
    )
  }

  dplyr::bind_rows(lapply(series, extract_year)) |>
    tidyr::pivot_longer(-year, names_to = "feature", values_to = "value")
}

# If residuals are normal with variance v and v ~ inverse-gamma(alpha, beta),
# this is log p(residuals) after integrating v from zero to infinity.
inverse_gamma_marginal_loglik <- function(residual, alpha, beta) {
  n <- length(residual)
  if (!n || any(!is.finite(residual))) return(-Inf)
  lgamma(alpha + n / 2) - lgamma(alpha) + alpha * log(beta) -
    (alpha + n / 2) * log(beta + sum(residual^2) / 2) -
    n * log(2 * pi) / 2
}

feature_gpp_loglik <- function(
    par, model_error, drivers, observations, feature_function,
    variance_priors, marginal_function
) {
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = par) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp)
  matched <- observations |>
    dplyr::select(sitename, date, gpp) |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(is.finite(gpp), is.finite(gpp_modelled))
  if (nrow(matched) == 0) return(-Inf)

  # `feature_function` is passed explicitly through calib_sofun_lhs(). This is
  # important for PSOCK parallelism: unlike an object in the likelihood's
  # global environment, an explicit argument is serialized to every worker.
  observed <- feature_function(matched, gpp) |>
    dplyr::rename(observed = value)
  modelled <- feature_function(matched, gpp_modelled) |>
    dplyr::rename(modelled = value)
  features <- dplyr::inner_join(observed, modelled,
    by = c("year", "feature")) |>
    dplyr::mutate(residual = observed - modelled)
  if (!nrow(features)) return(-Inf)

  # Week 1 and week 52 are neighbours, so use circular rather than linear
  # distance for this feature.
  is_peak_week <- features$feature == "peak_week"
  features$residual[is_peak_week] <- pmin(
    abs(features$residual[is_peak_week]),
    52 - abs(features$residual[is_peak_week])
  )

  score <- vapply(names(variance_priors), function(feature_name) {
    prior <- variance_priors[[feature_name]]
    residual <- features$residual[features$feature == feature_name]
    marginal_function(
      residual, alpha = prior$alpha, beta = prior$beta
    )
  }, numeric(1))
  if (any(!is.finite(score))) return(-Inf)
  sum(score)
}
```

```{r simple-gpp-calibrations, eval=TRUE}
kphio_prior <- list(
  kphio = list(
    mean = 0.08, sd = 0.04,
    distribution = "normal", lower = 0.03, upper = 0.12
  )
)
stopifnot(identical(names(kphio_prior), "kphio"))

daily_gpp_calibration <- calib_sofun_lhs(
  pars = pars,
  par_calib = kphio_prior,
  model_error = list(
    sigma_gpp_daily = list(
      mean = 0.8, sd = 0.3,
      distribution = "lognormal", lower = 0.05, upper = 4
    )
  ),
  likelihood = daily_gpp_loglik,
  n_samples = 500,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 42,
  progress = TRUE,
  drivers = df_drivers,
  observations = daily_observations
)

monthly_gpp_calibration <- calib_sofun_lhs(
  pars = pars,
  par_calib = kphio_prior,
  model_error = list(
    sigma_gpp_monthly = list(
      mean = 0.8, sd = 0.3,
      distribution = "lognormal", lower = 0.05, upper = 4
    )
  ),
  likelihood = monthly_gpp_loglik,
  n_samples = 500,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 44,
  progress = TRUE,
  drivers = df_drivers,
  observations = daily_observations
)

annual_gpp_calibration <- calib_sofun_lhs(
  pars = pars,
  par_calib = kphio_prior,
  model_error = list(
    sigma_gpp_annual = list(
      mean = 100, sd = 50,
      distribution = "lognormal", lower = 10, upper = 500
    )
  ),
  likelihood = mean_annual_gpp_loglik,
  n_samples = 500,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 43,
  progress = TRUE,
  drivers = df_drivers,
  observations = daily_observations
)
```

### Feature-oriented calibration with integrated error variances

Mean annual GPP is deliberately omitted from this likelihood because it is
calculated from the same annual totals that are already included. Using both
would count essentially the same information twice. The retained features are
calculated separately for each year, producing a set of annual residuals for
each feature.

#### Interpreting the variance priors

The mismatch between model and observations is unlikely to have the same size
for all features. An error of two weeks may be plausible for peak timing, for
example, whereas an error of two units has a completely different meaning for
daily GPP. Each feature therefore gets its own unknown residual variance
$v=\sigma^2$, expressed in the square of that feature's native units.

Instead of fixing $\sigma$, assign its variance the prior

$$
v \sim \mathrm{InvGamma}(\alpha,\beta).
$$

The code specifies this prior through `prior_rms`, defined as
$\sqrt{E[v]}$. This is easier to interpret than the inverse-gamma scale
parameter. For example, `prior_rms = 2` for `peak_week` says that, before
seeing the CH-Oe1 residuals, a residual scale of roughly two weeks is plausible.
It does **not** fix the error at two weeks: the inverse-gamma distribution still
allows both smaller and larger values.

For an inverse-gamma distribution, $E[v]=\beta/(\alpha-1)$. Consequently,
the conversion

```r
beta = (alpha - 1) * prior_rms^2
```

makes the value entered as `prior_rms` equal to $\sqrt{E[v]}$. Here
`alpha = 3` gives every variance a proper but fairly broad prior. Increasing
`alpha` while preserving the same `prior_rms` would express greater confidence
in the chosen error scale; decreasing it towards 2 would make the prior wider.

#### Why the variances need not be sampled

For a proposed `kphio`, the model produces annual residuals
$r_1,\ldots,r_n$ for each feature. Conditional on a variance, these are
assumed normal. The normal likelihood and inverse-gamma variance prior are
conjugate, so the variance can be averaged over exactly:

$$
p(r \mid \mathrm{kphio}) =
\int_0^\infty p(r \mid \mathrm{kphio},v)\,p(v)\,\mathrm{d}v.
$$

This integral is the marginal likelihood. Intuitively, it asks: *Across all
plausible error variances, weighted by their prior plausibility, how well does
this `kphio` explain the annual feature values?* A poor fit cannot escape its
penalty merely by selecting an arbitrarily large error because large variances
are also penalized by the normalized likelihood and by the prior.

The resulting log marginal likelihood for one feature is

$$
\log p(r) = \log\Gamma(\alpha+n/2)-\log\Gamma(\alpha)
+\alpha\log\beta-(\alpha+n/2)\log\left(\beta+\tfrac12\sum r_i^2\right)
-\tfrac{n}{2}\log(2\pi).
$$

The two important updates in this expression are intuitive:

- $\alpha+n/2$ increases the information about the variance as more annual
  residuals become available;
- $\beta+\tfrac12\sum r_i^2$ increases the inferred error scale when the
  model has larger squared residuals.

The symbols used above mean:

- $\mathrm{kphio}$ is the model parameter being calibrated: the maximum
  quantum-yield efficiency;
- $r=(r_1,\ldots,r_n)$ is the vector of annual residuals for one feature, with
  $r_i$ equal to observed minus modelled feature value in year $i$;
- $n$ is the number of years with a valid observed and modelled feature value;
- $\sigma$ is the residual standard deviation and $v=\sigma^2$ is its variance;
- $\mathrm{InvGamma}(\alpha,\beta)$ is the inverse-gamma distribution used for
  $v$, with shape $\alpha$ and scale $\beta$;
- $E[v]$ is the prior expected value of the variance;
- $p(r\mid\mathrm{kphio},v)$ is the normal likelihood conditional on one
  variance, while $p(v)$ is the inverse-gamma prior density;
- $\int_0^\infty \cdot\,\mathrm{d}v$ means that the conditional likelihood is
  averaged over every positive value of $v$;
- $\log$ is the natural logarithm, $\Gamma$ is the gamma function,
  $\sum r_i^2$ is the sum of squared annual residuals, and $\pi$ is the circle
  constant.

The six feature contributions are added to obtain the likelihood for a
`kphio` sample. No variance is added to the Latin hypercube, so the feature
calibration samples only `kphio` rather than `kphio` plus six weakly identified
error parameters. The error variances have not been ignored: they have been
integrated over. Their conditional posterior distributions can still be
recovered after calibration, as demonstrated below.

```{r feature-gpp-calibration, eval=TRUE}

# Each inverse-gamma prior is defined through E[sigma^2]^(1/2), an interpretable
# prior RMS error in the feature's native units. alpha = 3 leaves substantial
# uncertainty while giving the variance a finite prior mean.
make_variance_prior <- function(prior_rms, alpha = 3) {
  list(
    alpha = alpha,
    beta = (alpha - 1) * prior_rms^2,
    prior_rms = prior_rms
  )
}
feature_variance_priors <- list(
  annual_gpp = make_variance_prior(150), # g C m-2 yr-1
  peak_gpp = make_variance_prior(0.75),  # g C m-2 d-1
  peak_week = make_variance_prior(2),    # weeks
  onset = make_variance_prior(14),       # days of year
  termination = make_variance_prior(14), # days of year
  amplitude = make_variance_prior(0.75)  # g C m-2 d-1
)

# Check the feature pipeline once in the main R process. This gives a local,
# informative failure before launching hundreds of worker evaluations.
feature_preflight <- feature_gpp_loglik(
  par = pars,
  model_error = list(),
  drivers = df_drivers,
  observations = daily_observations,
  feature_function = gpp_features_by_year,
  variance_priors = feature_variance_priors,
  marginal_function = inverse_gamma_marginal_loglik
)
if (!is.finite(feature_preflight)) {
  stop("The seasonal-feature likelihood failed its serial preflight check.")
}

feature_gpp_calibration <- calib_sofun_lhs(
  pars = pars,
  par_calib = kphio_prior,
  model_error = list(),
  likelihood = feature_gpp_loglik,
  n_samples = 500,
  n_cores = max(1, parallel::detectCores() - 1),
  seed = 45,
  progress = TRUE,
  drivers = df_drivers,
  observations = daily_observations,
  feature_function = gpp_features_by_year,
  variance_priors = feature_variance_priors,
  marginal_function = inverse_gamma_marginal_loglik
)
```

```{r feature-calibration-components, eval=TRUE}
feature_map_simulation <- rsofun::runread_cnmodel_f(
  drivers = df_drivers,
  par = feature_gpp_calibration$par
) |>
  dplyr::select(sitename, data) |>
  tidyr::unnest(data) |>
  dplyr::transmute(sitename, date, gpp_modelled = gpp)

feature_map_matched <- daily_observations |>
  dplyr::select(sitename, date, gpp) |>
  dplyr::inner_join(feature_map_simulation, by = c("sitename", "date")) |>
  dplyr::filter(is.finite(gpp), is.finite(gpp_modelled))
observed_features <- gpp_features_by_year(feature_map_matched, gpp) |>
  dplyr::rename(observed = value)
map_features <- gpp_features_by_year(feature_map_matched, gpp_modelled) |>
  dplyr::rename(map_model = value)
feature_comparison <- dplyr::inner_join(
  observed_features, map_features, by = c("year", "feature")
)
feature_comparison
```

The conjugate update also provides the conditional posterior of every error
standard deviation at the MAP `kphio`. These are diagnostics, not additional
LHS dimensions.

```{r feature-error-posteriors, eval=TRUE}
feature_error_posteriors <- feature_comparison |>
  dplyr::mutate(
    residual = observed - map_model,
    residual = dplyr::if_else(
      feature == "peak_week",
      pmin(abs(residual), 52 - abs(residual)), residual
    )
  ) |>
  dplyr::group_by(feature) |>
  dplyr::summarise(
    n_years = dplyr::n(),
    sum_squared_residual = sum(residual^2),
    .groups = "drop"
  ) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    prior_rms = feature_variance_priors[[feature]]$prior_rms,
    alpha_posterior = feature_variance_priors[[feature]]$alpha + n_years / 2,
    beta_posterior = feature_variance_priors[[feature]]$beta +
      sum_squared_residual / 2,
    posterior_rms = sqrt(beta_posterior / (alpha_posterior - 1)),
    posterior_median = sqrt(1 / stats::qgamma(
      0.5, shape = alpha_posterior, rate = beta_posterior
    )),
    posterior_lower = sqrt(1 / stats::qgamma(
      0.975, shape = alpha_posterior, rate = beta_posterior
    )),
    posterior_upper = sqrt(1 / stats::qgamma(
      0.025, shape = alpha_posterior, rate = beta_posterior
    ))
  ) |>
  dplyr::ungroup() |>
  dplyr::select(
    feature, n_years, prior_rms, posterior_rms, posterior_median,
    posterior_lower, posterior_upper
  )
feature_error_posteriors
```

### Sensitivity to the inverse-gamma shape parameter

The choice `alpha = 3` should be checked rather than treated as universal. The
comparison below uses `alpha = 2.1, 3, 5, 10`, while preserving every feature's
`prior_rms`. Therefore, all alternatives have the same prior mean variance and
differ only in how tightly the variance is concentrated around that scale.

Running four complete calibrations would repeat the expensive C-model
simulation four times. Instead, evaluate every successful `kphio` sample from
the main calibration once, retain the number and sum of squared annual residuals
for each feature, and reuse those sufficient statistics for all four analytical
marginal likelihoods.

```{r feature-alpha-sensitivity-simulations, eval=TRUE}
feature_residual_summary <- function(
    kphio, pars, drivers, observations, feature_function
) {
  candidate <- pars
  candidate$kphio <- kphio
  simulated <- rsofun::runread_cnmodel_f(drivers = drivers, par = candidate) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp)
  matched <- observations |>
    dplyr::select(sitename, date, gpp) |>
    dplyr::inner_join(simulated, by = c("sitename", "date")) |>
    dplyr::filter(is.finite(gpp), is.finite(gpp_modelled))

  observed <- feature_function(matched, gpp) |>
    dplyr::rename(observed = value)
  modelled <- feature_function(matched, gpp_modelled) |>
    dplyr::rename(modelled = value)
  dplyr::inner_join(observed, modelled, by = c("year", "feature")) |>
    dplyr::mutate(
      residual = observed - modelled,
      residual = dplyr::if_else(
        feature == "peak_week",
        pmin(abs(residual), 52 - abs(residual)), residual
      )
    ) |>
    dplyr::group_by(feature) |>
    dplyr::summarise(
      n = dplyr::n(),
      sum_squared_residual = sum(residual^2),
      .groups = "drop"
    )
}

sensitivity_design <- feature_gpp_calibration$samples |>
  dplyr::filter(evaluation_ok) |>
  dplyr::select(sample_id, kphio, log_prior)

evaluate_sensitivity_sample <- function(i) {
  tryCatch(
    feature_residual_summary(
      kphio = sensitivity_design$kphio[[i]],
      pars = pars,
      drivers = df_drivers,
      observations = daily_observations,
      feature_function = gpp_features_by_year
    ) |>
      dplyr::mutate(sample_id = sensitivity_design$sample_id[[i]]),
    error = function(condition) NULL
  )
}

n_sensitivity_workers <- min(
  max(1, parallel::detectCores() - 1), nrow(sensitivity_design)
)
message(sprintf(
  "Evaluating %d shared sensitivity simulations on %d worker(s).",
  nrow(sensitivity_design), n_sensitivity_workers
))
if (n_sensitivity_workers == 1) {
  sensitivity_residuals <- lapply(
    seq_len(nrow(sensitivity_design)), evaluate_sensitivity_sample
  )
} else {
  sensitivity_cluster <- parallel::makeCluster(n_sensitivity_workers)
  sensitivity_residuals <- tryCatch({
    parallel::clusterExport(
      sensitivity_cluster,
      c(
        "evaluate_sensitivity_sample", "feature_residual_summary",
        "sensitivity_design", "pars", "df_drivers", "daily_observations",
        "gpp_features_by_year"
      ),
      envir = environment()
    )
    parallel::parLapplyLB(
      sensitivity_cluster,
      seq_len(nrow(sensitivity_design)),
      function(i) evaluate_sensitivity_sample(i)
    )
  }, finally = {
    parallel::stopCluster(sensitivity_cluster)
  })
}
sensitivity_residuals <- dplyr::bind_rows(sensitivity_residuals)
message(sprintf(
  "Recovered feature residuals for %d/%d sensitivity samples.",
  dplyr::n_distinct(sensitivity_residuals$sample_id), nrow(sensitivity_design)
))
```

The likelihood calculation below is inexpensive: it contains no model call.
For each `alpha`, `beta` is recalculated to preserve `prior_rms`, and posterior
weights are normalized independently.

```{r feature-alpha-sensitivity-posteriors, eval=TRUE}
alpha_values <- c(2.1, 3, 5, 10)

alpha_posteriors <- tidyr::crossing(
  sensitivity_residuals,
  alpha = alpha_values
) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    prior_rms = feature_variance_priors[[feature]]$prior_rms,
    beta = (alpha - 1) * prior_rms^2,
    log_likelihood_feature =
      lgamma(alpha + n / 2) - lgamma(alpha) + alpha * log(beta) -
      (alpha + n / 2) * log(beta + sum_squared_residual / 2) -
      n * log(2 * pi) / 2
  ) |>
  dplyr::ungroup() |>
  dplyr::group_by(sample_id, alpha) |>
  dplyr::filter(dplyr::n() == length(feature_variance_priors)) |>
  dplyr::summarise(
    log_likelihood = sum(log_likelihood_feature),
    .groups = "drop"
  ) |>
  dplyr::left_join(sensitivity_design, by = "sample_id") |>
  dplyr::mutate(log_posterior = log_prior + log_likelihood) |>
  dplyr::group_by(alpha) |>
  dplyr::mutate(
    weight = exp(log_likelihood - max(log_likelihood)),
    weight = weight / sum(weight)
  ) |>
  dplyr::ungroup()

alpha_diagnostics <- alpha_posteriors |>
  dplyr::group_by(alpha) |>
  dplyr::summarise(
    map_kphio = kphio[which.max(log_posterior)],
    effective_sample_size = 1 / sum(weight^2),
    maximum_weight = max(weight),
    samples_to_80_percent = which(
      cumsum(sort(weight, decreasing = TRUE)) >= 0.8
    )[1],
    .groups = "drop"
  ) |>
  dplyr::mutate(
    map_difference_from_alpha_3 = map_kphio - map_kphio[alpha == 3],
    prior_cv_variance = 1 / sqrt(alpha - 2),
    prior_sigma_lower_over_rms = sqrt(
      (alpha - 1) / stats::qgamma(0.975, shape = alpha)
    ),
    prior_sigma_upper_over_rms = sqrt(
      (alpha - 1) / stats::qgamma(0.025, shape = alpha)
    )
  )
map_alpha_3 <- alpha_diagnostics$map_kphio[alpha_diagnostics$alpha == 3]
maximum_map_change <- max(abs(alpha_diagnostics$map_difference_from_alpha_3))
maximum_relative_map_change <- maximum_map_change / map_alpha_3
alpha_sensitivity_conclusion <- if (maximum_relative_map_change <= 0.05) {
  paste(
    "The maximum relative MAP change is no greater than 5%, supporting the",
    "conclusion that the MAP is insensitive over this range. Together with",
    "its variance coefficient of variation of one, this supports alpha = 3",
    "as a broad but finite-variance compromise."
  )
} else {
  paste(
    "The maximum relative MAP change exceeds 5%; the calibration should",
    "therefore be described as sensitive to alpha, and alpha = 3 cannot be",
    "justified from MAP stability alone."
  )
}
alpha_diagnostics
```

The last three columns quantify prior width independently of the calibration.
`prior_cv_variance` is the coefficient of variation of $v$; the two ratio
columns give the 95% prior interval for $\sigma$ relative to `prior_rms`. At
`alpha = 3`, the variance has a coefficient of variation of
`r signif(alpha_diagnostics$prior_cv_variance[alpha_diagnostics$alpha == 3], 3)`
and 95% of the prior standard-deviation distribution lies between
`r signif(alpha_diagnostics$prior_sigma_lower_over_rms[alpha_diagnostics$alpha == 3], 3)`
and
`r signif(alpha_diagnostics$prior_sigma_upper_over_rms[alpha_diagnostics$alpha == 3], 3)`
times `prior_rms`. This is the sense in which `alpha = 3` is informative enough
to exclude implausibly extreme errors, but still broad.

The ECDF compares the full `kphio` posteriors without kernel smoothing. Similar
curves and closely aligned MAP lines indicate that the inference is insensitive
to the reasonable range of `alpha` values considered here.

```{r feature-alpha-sensitivity-ecdf, eval=TRUE}
alpha_ecdf <- alpha_posteriors |>
  dplyr::arrange(alpha, kphio) |>
  dplyr::group_by(alpha) |>
  dplyr::mutate(cumulative_probability = cumsum(weight)) |>
  dplyr::ungroup() |>
  dplyr::mutate(alpha = factor(alpha, levels = alpha_values))

ggplot(alpha_ecdf, aes(kphio, cumulative_probability, colour = alpha)) +
  geom_step(linewidth = 0.8) +
  geom_vline(
    data = dplyr::mutate(
      alpha_diagnostics, alpha = factor(alpha, levels = alpha_values)
    ),
    aes(xintercept = map_kphio, colour = alpha),
    linetype = "dotted",
    show.legend = FALSE
  ) +
  labs(
    x = "Maximum quantum yield (kphio)",
    y = "Posterior cumulative probability",
    colour = expression(alpha)
  ) +
  theme_classic()
```

Finally, ranked weights show whether a more concentrated variance prior causes
importance-weight collapse. The dashed line is the equal-weight reference.

```{r feature-alpha-sensitivity-weights, eval=TRUE}
alpha_weight_ranks <- alpha_posteriors |>
  dplyr::group_by(alpha) |>
  dplyr::arrange(dplyr::desc(weight), .by_group = TRUE) |>
  dplyr::mutate(
    weight_rank = dplyr::row_number(),
    weight_for_plot = pmax(weight, 1e-16)
  ) |>
  dplyr::ungroup()

ggplot(alpha_weight_ranks, aes(weight_rank, weight_for_plot)) +
  geom_hline(
    yintercept = 1 / nrow(sensitivity_design),
    linetype = "dashed", colour = "grey50"
  ) +
  geom_line(colour = "#0072B2", linewidth = 0.6) +
  facet_wrap(~ alpha, labeller = ggplot2::label_bquote(alpha == .(alpha))) +
  scale_y_log10() +
  labs(
    x = "Sample rank (highest posterior weight first)",
    y = "Posterior weight (log scale)"
  ) +
  theme_classic()
```

Across these alternatives, the MAP range is
`r signif(diff(range(alpha_diagnostics$map_kphio)), 3)` and the largest absolute
deviation from the `alpha = 3` MAP is
`r signif(maximum_map_change, 3)`
(`r signif(100 * maximum_relative_map_change, 3)`% relative to the
`alpha = 3` MAP). `r alpha_sensitivity_conclusion`

### Prior and posterior empirical distributions of `kphio`

The empirical cumulative distribution function (ECDF) avoids kernel smoothing.
The prior ECDF gives every LHS draw equal mass; the posterior ECDF uses the
normalized posterior weights. A collapsed posterior therefore appears as one
large vertical step at the dominant sample's `kphio` value.

```{r simple-kphio-prior-posterior, eval=TRUE}
distribution_data <- function(calibration, target_label) {
  dplyr::bind_rows(
    calibration$samples |>
      dplyr::arrange(kphio) |>
      dplyr::transmute(kphio, cumulative_probability = dplyr::row_number() /
                         dplyr::n(), target = target_label,
                       distribution = "Prior"),
    calibration$samples |>
      dplyr::arrange(kphio) |>
      dplyr::transmute(kphio, cumulative_probability = cumsum(weight),
                       target = target_label, distribution = "Posterior")
  )
}

kphio_distributions <- dplyr::bind_rows(
  distribution_data(daily_gpp_calibration, "Daily GPP"),
  distribution_data(monthly_gpp_calibration, "Monthly GPP"),
  distribution_data(annual_gpp_calibration, "One mean annual GPP value"),
  distribution_data(feature_gpp_calibration, "Seasonal GPP features")
)

ggplot(
  kphio_distributions,
  aes(kphio, cumulative_probability, colour = distribution)
) +
  geom_step(linewidth = 0.7) +
  geom_vline(xintercept = pars$kphio, linetype = "dashed") +
  facet_wrap(~ target, ncol = 1) +
  scale_colour_manual(values = c(Prior = "grey40", Posterior = "red")) +
  labs(x = "Maximum quantum yield (kphio)",
       y = "Cumulative probability", colour = NULL) +
  theme_classic()
```

```{r simple-calibration-diagnostics, eval=TRUE}
samples_to_80_percent <- function(weight) {
  which(cumsum(sort(weight, decreasing = TRUE)) >= 0.8)[1]
}

tibble(
  target = c(
    "Daily GPP", "Monthly GPP", "One mean annual GPP value",
    "Seasonal GPP features"
  ),
  effective_sample_size = c(
    daily_gpp_calibration$effective_sample_size,
    monthly_gpp_calibration$effective_sample_size,
    annual_gpp_calibration$effective_sample_size,
    feature_gpp_calibration$effective_sample_size
  ),
  maximum_weight = c(
    max(daily_gpp_calibration$samples$weight),
    max(monthly_gpp_calibration$samples$weight),
    max(annual_gpp_calibration$samples$weight),
    max(feature_gpp_calibration$samples$weight)
  ),
  samples_to_80_percent = c(
    samples_to_80_percent(daily_gpp_calibration$samples$weight),
    samples_to_80_percent(monthly_gpp_calibration$samples$weight),
    samples_to_80_percent(annual_gpp_calibration$samples$weight),
    samples_to_80_percent(feature_gpp_calibration$samples$weight)
  ),
  map_kphio = c(
    daily_gpp_calibration$par$kphio,
    monthly_gpp_calibration$par$kphio,
    annual_gpp_calibration$par$kphio,
    feature_gpp_calibration$par$kphio
  ),
  map_sampled_error = c(
    daily_gpp_calibration$model_error$sigma_gpp_daily,
    monthly_gpp_calibration$model_error$sigma_gpp_monthly,
    annual_gpp_calibration$model_error$sigma_gpp_annual,
    NA_real_ # feature-specific variances are integrated out, not sampled
  )
)
```

### Distribution of posterior weights

Ranked weights directly reveal importance-weight collapse. A well-supported
posterior distributes weight across multiple samples; a collapsed posterior has
one point near one and almost all remaining points near zero. The logarithmic
axis retains differences among very small weights. Exact zeros and weights below
`1e-16` are drawn at that plotting floor.

```{r simple-posterior-weight-distributions, eval=TRUE}
posterior_weights <- dplyr::bind_rows(
  daily_gpp_calibration$samples |>
    dplyr::mutate(target = "Daily GPP"),
  monthly_gpp_calibration$samples |>
    dplyr::mutate(target = "Monthly GPP"),
  annual_gpp_calibration$samples |>
    dplyr::mutate(target = "One mean annual GPP value"),
  feature_gpp_calibration$samples |>
    dplyr::mutate(target = "Seasonal GPP features")
) |>
  dplyr::group_by(target) |>
  dplyr::arrange(dplyr::desc(weight), .by_group = TRUE) |>
  dplyr::mutate(
    weight_rank = dplyr::row_number(),
    weight_for_plot = pmax(weight, 1e-16)
  ) |>
  dplyr::ungroup()

ggplot(posterior_weights, aes(weight_rank, weight_for_plot)) +
  geom_hline(
    yintercept = 1 / nrow(daily_gpp_calibration$samples),
    linetype = "dashed",
    colour = "grey50"
  ) +
  geom_line(colour = "#0072B2", linewidth = 0.5) +
  geom_point(colour = "#0072B2", size = 0.8) +
  facet_wrap(~ target, ncol = 1) +
  scale_y_log10() +
  labs(
    x = "Sample rank (highest posterior weight first)",
    y = "Posterior weight (log scale)"
  ) +
  theme_classic()
```

### Mean annual cycle of daily GPP

Re-run the C-model with the MAP `kphio` from each calibration. To make the
seasonal cycles directly comparable, calculate modelled means only for dates
that have a quality-controlled GPP observation.

```{r simple-gpp-annual-cycle, eval=TRUE}
map_gpp_cycle <- function(calibration, source_label) {
  rsofun::runread_cnmodel_f(
    drivers = df_drivers,
    par = calibration$par
  ) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::transmute(sitename, date, gpp_modelled = gpp) |>
    dplyr::inner_join(
      dplyr::filter(daily_observations, is.finite(gpp)),
      by = c("sitename", "date")
    ) |>
    dplyr::filter(is.finite(gpp_modelled)) |>
    dplyr::mutate(day_of_year = lubridate::yday(date)) |>
    dplyr::group_by(day_of_year) |>
    dplyr::summarise(gpp = mean(gpp_modelled), .groups = "drop") |>
    dplyr::mutate(source = source_label)
}

observed_gpp_cycle <- daily_observations |>
  dplyr::filter(is.finite(gpp)) |>
  dplyr::mutate(day_of_year = lubridate::yday(date)) |>
  dplyr::group_by(day_of_year) |>
  dplyr::summarise(gpp = mean(gpp), .groups = "drop") |>
  dplyr::mutate(source = "Observed")

gpp_annual_cycles <- dplyr::bind_rows(
  observed_gpp_cycle,
  map_gpp_cycle(daily_gpp_calibration, "MAP from daily GPP"),
  map_gpp_cycle(monthly_gpp_calibration, "MAP from monthly GPP"),
  map_gpp_cycle(annual_gpp_calibration, "MAP from mean annual GPP"),
  map_gpp_cycle(feature_gpp_calibration, "MAP from GPP features")
)

ggplot(gpp_annual_cycles, aes(day_of_year, gpp, colour = source)) +
  geom_line() +
  scale_x_continuous(
    breaks = c(15, 46, 74, 105, 135, 166, 196, 227, 258, 288, 319, 349),
    labels = month.abb
  ) +
  scale_colour_manual(values = c(
    "Observed" = "black",
    "MAP from daily GPP" = "tomato",
    "MAP from monthly GPP" = "darkgreen",
    "MAP from mean annual GPP" = "royalblue",
    "MAP from GPP features" = "darkorange3"
  )) +
  labs(
    x = NULL,
    y = expression("Mean daily GPP (g C " * m^-2 * d^-1 * ")"),
    colour = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom"
  )
```


## References

Arán Paredes, J., Bernhard, F., Hufkens, K., Marcadella, M., and Stocker,
B. D. (2025). rsofun v5.1: a model-data integration framework for simulating
ecosystem processes. *Geoscientific Model Development*, 18, 9855–9878.
doi:10.5194/gmd-18-9855-2025.

Leifeld, J., Meyer, S., Budge, K., Sebastia, M.-T., Zimmermann, M., and Fuhrer,
J. (2015). Turnover of grassland roots in mountain ecosystems revealed by
radiocarbon analysis: role of temperature and management. *PLOS ONE*, 10,
e0119184. doi:10.1371/journal.pone.0119184.
