---
title: "Consistency check 1: BiomeE-Allocation vs. SOFUN"
output: html_document
---

Compare output files written by original BiomeE-Allocation (Enshneg Weng) with SOFUN in branch lm3ppa_BiomeE_test_rsofun. 

```{r setup, include=F,  eval=T}
library(dplyr)
library(rsofun)
load_dependencies_rsofun()
library(readr)
library(tidyr)
#if(!require(devtools)){install.packages(devtools)}
#devtools::install_github("stineb/rbeni")
#library(rbeni)
library(rlang)
```

Set local path of original BiomeE-Allocation repo and SOFUN.
```{r, include=F, eval=T}
path_orig_B  <- "/Users/benjaminstocker/cagibi/BiomeE-Allocation/model"
path_sofun_B <- "/Users/benjaminstocker/sofun"
path_orig_output_B  <- "/Users/benjaminstocker/cagibi/BiomeE-Allocation/model/output"
path_sofun_output_B <- "/Users/benjaminstocker/sofun/output"

path_orig_model_L  <- "/Users/lmarques/BiomeE-Allocation/model"
path_sofun_model_L <- "/Users/lmarques/sofun"
path_orig_output_L  <- "/Users/lmarques/BiomeE-Allocation/model/output"
path_sofun_output_L <- "/Users/lmarques/sofun/output"
```

Describe function from rbeni package
```{r, include=F, eval=T}
analyse_modobs2 <- function(
  df,
  mod,
  obs,
  type       = "points",
  filnam     = NA,
  xlim       = NULL,
  ylim       = NULL,
  use_factor = NULL,
  shortsubtitle = FALSE,
  ...
  ){

  require(ggplot2)
  require(dplyr)
  require(LSD)
  require(ggthemes)
  require(RColorBrewer)

  #if (identical(filnam, NA)) filnam <- "analyse_modobs.pdf"

  ## rename to 'mod' and 'obs' and remove rows with NA in mod or obs
  df <- df %>%
    as_tibble() %>%
    ungroup() %>%
    dplyr::select(mod=mod, obs=obs) %>%
    tidyr::drop_na(mod, obs)

  ## get linear regression (coefficients)
  linmod <- lm( obs ~ mod, data=df )

  ## construct metrics table using the 'yardstick' library
  df_metrics <- df %>%
    yardstick::metrics(obs, mod) %>%
    dplyr::bind_rows( tibble( .metric = "n",        .estimator = "standard", .estimate = summarise(df, numb=n()) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "slope",    .estimator = "standard", .estimate = coef(linmod)[2]) ) %>%
    # dplyr::bind_rows( tibble( .metric = "nse",      .estimator = "standard", .estimate = hydroGOF::NSE( obs, mod, na.rm=TRUE ) ) ) %>%
    dplyr::bind_rows( tibble( .metric = "mean_obs", .estimator = "standard", .estimate = summarise(df, mean=mean(obs, na.rm=TRUE)) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "prmse",    .estimator = "standard",
                       .estimate = filter(., .metric=="rmse") %>% dplyr::select(.estimate) %>% unlist() /
                         filter(., .metric=="mean_obs") %>% dplyr::select(.estimate) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "pmae",    .estimator = "standard",
                       .estimate = filter(., .metric=="mae") %>% dplyr::select(.estimate) %>% unlist() /
                         filter(., .metric=="mean_obs") %>% dplyr::select(.estimate) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "bias",        .estimator = "standard", .estimate = summarise(df, mean((mod-obs), na.rm=TRUE    )) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "pbias",       .estimator = "standard", .estimate = summarise(df, mean((mod-obs)/obs, na.rm=TRUE)) %>% unlist() ) )

  rsq_val <- df_metrics %>% filter(.metric=="rsq") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  rmse_val <- df_metrics %>% filter(.metric=="rmse") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  mae_val <- df_metrics %>% filter(.metric=="mae") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  bias_val <- df_metrics %>% filter(.metric=="bias") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  slope_val <- df_metrics %>% filter(.metric=="slope") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  n_val <- df_metrics %>% filter(.metric=="n") %>% dplyr::select(.estimate) %>% unlist() %>% unname()

  rsq_lab <- format( rsq_val, digits = 2 )
  rmse_lab <- format( rmse_val, digits = 3 )
  mae_lab <- format( mae_val, digits = 3 )
  bias_lab <- format( bias_val, digits = 3 )
  slope_lab <- format( slope_val, digits = 3 )
  n_lab <- format( n_val, digits = 3 )

  results <- tibble( rsq = rsq_val, rmse = rmse_val, mae = mae_val, bias = bias_val, slope = slope_val, n = n_val )

  if (shortsubtitle){
    subtitle <- bquote( italic(R)^2 == .(rsq_lab) ~~
                          RMSE == .(rmse_lab) )
  } else {
    subtitle <- bquote( italic(R)^2 == .(rsq_lab) ~~
                          RMSE == .(rmse_lab) ~~
                          bias == .(bias_lab) ~~
                          slope == .(slope_lab) ~~
                          italic(N) == .(n_lab) )
  }

  if (type=="heat"){

    # if (!identical(filnam, NA)) dev.off()
    source("~/LSD/R/LSD.heatscatter.R")
    gg <- heatscatter(
                  df$mod,
                  df$obs,
                  xlim=xlim,
                  ylim=ylim,
                  main="",
                  ggplot=TRUE )

    gg <- gg +
      geom_smooth(method='lm', color="red", size=0.5, se=FALSE) +
      geom_abline(intercept=0, slope=1, linetype="dotted") +
      theme_classic() +
      labs(
        subtitle = subtitle
        )

    if (!identical(filnam, NA)) {
      ggsave(filnam, width=5, height=5)
    }

  } else if (type=="hex"){

    ## ggplot hexbin
    gg <- df %>%
      ggplot2::ggplot(aes(x=mod, y=obs)) +
      geom_hex() +
      scale_fill_gradientn(
        colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5)) +
      geom_smooth(method='lm', color="red", size=0.5) +
      geom_abline(intercept=0, slope=1, linetype="dotted") +
      # coord_fixed() +
      # xlim(0,NA) +
      # ylim(0,NA) +
      theme_classic() +
      labs(
        subtitle = subtitle
        )

    if (!identical(filnam, NA)) {
      ggsave(filnam, width=5, height=5)
    }

  } else if (type=="points") {

    ## points
    gg <- df %>%
      ggplot(aes(x=mod, y=obs)) +
      geom_point() +
      geom_smooth(method='lm', color="red", size=0.5) +
      geom_abline(intercept=0, slope=1, linetype="dotted") +
      # coord_fixed() +
      # xlim(0,NA) +
      # ylim(0,NA) +
      theme_classic() +
      labs(
        # subtitle = expression( paste( italic(R)^2, "\n",
        #   "beni")),
        subtitle = subtitle
        )

    if (!identical(filnam, NA)) {
      ggsave(filnam, width=5, height=5)
    }

  }

  return(list(df_metrics=df_metrics, gg=gg, linmod=linmod, results = results))
}
```

Define plotting functions by variable.
```{r, include=F, eval=T}

plot_tseries_byvar <- function(var, df_biomeE, df_sofun){
  
  var_biomeE <- paste0(var, "_biomeE")
  var_sofun  <- paste0(var, "_sofun")
  
  df <- df_biomeE %>% 
    dplyr::select(year, {{var}}) %>% 
    dplyr::rename(biomeE = {{var}}) %>% 
    left_join(
      df_sofun %>% 
        dplyr::select(year,{{var}}) %>% 
        dplyr::rename(sofun = {{var}}),
      by = "year"
    ) %>% 
    pivot_longer(cols = c(biomeE, sofun), names_to = "source", values_to = "var")
  
  gg <- df %>% 
    ggplot(aes(x = year, y = var, color = source)) +
    geom_line() +
    labs(x = "Year", y = var)
  
  return(gg)
}

plot_modobs_byvar <- function(var, df_biomeE, df_sofun){
  
  var_biomeE <- paste0(var, "_biomeE")
  var_sofun  <- paste0(var, "_sofun")
  
  df <- df_biomeE %>% 
    dplyr::select(year, {{var}}) %>% 
    dplyr::rename(var_biomeE = {{var}}) %>% 
    left_join(
      df_sofun %>% 
        dplyr::select(year,{{var}}) %>% 
        dplyr::rename(var_sofun = {{var}}),
      by = "year"
    )
  modobs <- df %>% 
    analyse_modobs2("var_biomeE", "var_sofun", relative = TRUE)
  gg <- modobs$gg +
    labs(x = var_biomeE, y = var_sofun)
  
  return(gg)
}
```

## Running BiomeE-Allocation in terminal and getting outputs as csv files

```{r, include=F, eval=T}
setwd(path_orig_model_L)
system("gfortran src/main.f90 src/datatypes.f90 src/soil.f90 src/vegetation.f90 -o ess")
system("./ess")
```

#### Getting BiomeE outputs

```{r, include=F, eval=T}
setwd(path_orig_output_L)
#biomeE_out_hourly_tile <- read.csv("Hourly_tile_test.csv", sep=",")
#biomeE_out_daily_tile <- read.csv("Daily_tile_test.csv", sep=",")
#biomeE_out_daily_cohorts <- read.csv("Daily_cohorts_test.csv", sep=",")
biomeE_out_annual_tile <- read.csv("Annual_tile_test.csv", sep=",")
biomeE_out_annual_cohorts <- read.csv("Annual_cohorts_test.csv", sep=",")
```

## Running SOFUN in terminal and getting outputs as csv files

```{r, include=F, eval=T}
setwd(path_sofun_model_L)
system("git checkout lm3ppa_BiomeE_test_rsofun")
system("make clean")
system("make lm3ppa")
system("./runlm3ppa")
```

#### Getting SOFUN outputs

```{r, include=F, eval=T}
setwd(path_sofun_output_L)
#sofun_out_hourly_tile <- read.csv("Hourly_tile_test.csv", sep=",")
#sofun_out_daily_tile <- read.csv("Daily_tile_test.csv", sep=",")
#sofun_out_daily_cohorts <- read.csv("Daily_cohorts_test.csv", sep=",")
sofun_out_annual_tile <- read.csv("Annual_tile_test.csv", sep=",")
sofun_out_annual_cohorts <- read.csv("Annual_cohorts_test.csv", sep=",")
```

### Plots of hourly tile outputs

```{r, include=F, eval=F}
varnams_hourly_tile_biomeE <- names(biomeE_out_hourly_tile)[c(-1, -length(names(biomeE_out_hourly_tile)))]
varnams_hourly_tile_sofun <- names(sofun_out_hourly_tile)[c(-1,-length(names(biomeE_out_hourly_tile)))]
all.equal(varnams_hourly_tile_biomeE, varnams_hourly_tile_sofun)

for (varnam in varnams_hourly_tile_biomeE[1:15]){
  print(varnam)
  
  gg <- plot_tseries_byvar(varnam, biomeE_out_hourly_tile, sofun_out_hourly_tile)
  print(gg)

  gg <- plot_modobs_byvar(varnam, biomeE_out_hourly_tile, sofun_out_hourly_tile)
  print(gg)
}

out_modobs <- purrr::map(as.list(varnams_hourly_tile_biomeE),
                         ~plot_modobs_byvar(., biomeE_out_hourly_tile, sofun_out_hourly_tile))
```

```{r, include=T, eval=T}
# Other option for plots using ggplot2()
dim(biomeE_out_hourly_tile)
dim(sofun_out_hourly_tile)
biomeE_out_hourly_tile <- biomeE_out_hourly_tile[,-16]
sofun_out_hourly_tile <- sofun_out_hourly_tile[,-16]

plot_dfs <- lapply(names(biomeE_out_hourly_tile),function(nm)data.frame(col1 = biomeE_out_hourly_tile[,nm], col2 = sofun_out_hourly_tile[,nm]))

for (idx in seq_along(plot_dfs))
  print(ggplot()+geom_smooth(data = plot_dfs[[idx]], aes(x=col1, y=col2),color="blue") +
      ggtitle(names(biomeE_out_hourly_tile)[idx]))
```

### Plots of daily tile outputs

```{r, include=F, eval=F}
varnams_daily_tile_biomeE <- names(biomeE_out_daily_tile)[c(-1, -length(names(biomeE_out_daily_tile)))]
varnams_daily_tile_sofun <- names(sofun_out_daily_tile)[c(-1,-length(names(biomeE_out_daily_tile)))]
all.equal(varnams_daily_tile_biomeE, varnams_daily_tile_sofun)

for (varnam in varnams_daily_tile_biomeE[1:34]){
  print(varnam)
  
  gg <- plot_tseries_byvar(varnam, biomeE_out_daily_tile, sofun_out_daily_tile)
  print(gg)

  gg <- plot_modobs_byvar(varnam, biomeE_out_daily_tile, sofun_out_daily_tile)
  print(gg)
}

out_modobs <- purrr::map(as.list(varnams_daily_tile_biomeE),
                         ~plot_modobs_byvar(., biomeE_out_daily_tile, sofun_out_daily_tile))
```

```{r, include=T, eval=T}
# Other option for plots using ggplot2()
dim(biomeE_out_daily_tile)
dim(sofun_out_daily_tile)
biomeE_out_daily_tile <- biomeE_out_daily_tile[,-36]
sofun_out_daily_tile <- sofun_out_daily_tile[,-36]

plot_dfs <- lapply(names(biomeE_out_daily_tile),function(nm)data.frame(col1 = biomeE_out_daily_tile[,nm], col2 = sofun_out_daily_tile[,nm]))

for (idx in seq_along(plot_dfs))
  print(ggplot()+geom_smooth(data = plot_dfs[[idx]], aes(x=col1, y=col2),color="blue") +
      ggtitle(names(biomeE_out_daily_tile)[idx]))
```

### Plots of daily cohorts outputs

```{r, include=F, eval=F}
varnams_daily_cohorts_biomeE <- names(biomeE_out_daily_cohorts)[c(-1, -length(names(biomeE_out_daily_cohorts)))]
varnams_daily_cohorts_sofun <- names(sofun_out_daily_cohorts)[c(-1,-length(names(biomeE_out_daily_cohorts)))]
all.equal(varnams_daily_cohorts_biomeE, varnams_daily_cohorts_sofun)

for (varnam in varnams_daily_cohorts_biomeE[1:2]){
  print(varnam)
  
  gg <- plot_tseries_byvar(varnam, biomeE_out_daily_cohorts, sofun_out_daily_cohorts)
  print(gg)

  gg <- plot_modobs_byvar(varnam, biomeE_out_daily_cohorts, sofun_out_daily_cohorts)
  print(gg)
}

out_modobs <- purrr::map(as.list(varnams_daily_cohorts_biomeE),
                         ~plot_modobs_byvar(., biomeE_out_daily_cohorts, sofun_out_daily_cohorts))
```

```{r, include=T, eval=T}
# Other option for plots using ggplot2()
dim(biomeE_out_daily_cohorts)
dim(sofun_out_daily_cohorts)
biomeE_out_daily_cohorts <- biomeE_out_daily_cohorts[,-28]
sofun_out_daily_cohorts <- sofun_out_daily_cohorts[,-28]

biomeE_out_daily_cohorts_short <- biomeE_out_daily_cohorts[1:100000,]
sofun_out_daily_cohorts_short <- sofun_out_daily_cohorts[1:100000,]

plot_dfs <- lapply(names(biomeE_out_daily_cohorts_short),function(nm)data.frame(col1 = biomeE_out_daily_cohorts_short[,nm], col2 = sofun_out_daily_cohorts_short[,nm]))

for (idx in seq_along(plot_dfs))
  print(ggplot()+geom_smooth(data = plot_dfs[[idx]], aes(x=col1, y=col2),color="blue") +
      ggtitle(names(biomeE_out_daily_cohorts_short)[idx]))
```

### Plots of annual tile outputs

```{r, include=T, eval=T}
varnams_annual_tile_biomeE <- names(biomeE_out_annual_tile)[c(-1, -length(names(biomeE_out_annual_tile)))]
varnams_annual_tile_sofun <- names(sofun_out_annual_tile)[c(-1, -length(names(biomeE_out_annual_tile)))]
all.equal(varnams_annual_tile_biomeE, varnams_annual_tile_sofun)

for (varnam in varnams_annual_tile_biomeE[1:43]){
  print(varnam)
  
  gg <- plot_tseries_byvar(varnam, biomeE_out_annual_tile, sofun_out_annual_tile)
  print(gg)

  gg <- plot_modobs_byvar(varnam, biomeE_out_annual_tile, sofun_out_annual_tile)
  print(gg)
}

out_modobs <- purrr::map(as.list(varnams_annual_tile_biomeE),
                         ~plot_modobs_byvar(., biomeE_out_annual_tile, sofun_out_annual_tile))
```

```{r, include=T, eval=T}
# Other option for plots using ggplot2()
dim(biomeE_out_annual_tile)
dim(sofun_out_annual_tile)
biomeE_out_annual_tile <- biomeE_out_annual_tile[,-45]
sofun_out_annual_tile <- sofun_out_annual_tile[,-45]

plot_dfs <- lapply(names(biomeE_out_annual_tile),function(nm)data.frame(col1 = biomeE_out_annual_tile[,nm], col2 = sofun_out_annual_tile[,nm]))

for (idx in seq_along(plot_dfs))
  print(ggplot()+geom_smooth(data = plot_dfs[[idx]], aes(x=col1, y=col2),color="blue") +
      ggtitle(names(biomeE_out_annual_tile)[idx]))

```

See that N_fxd and N_loss plots are empty because these variables are equal to 0.

### Plots of annual cohorts outputs

```{r, include=F, eval=F}
varnams_annual_cohorts_biomeE <- names(biomeE_out_annual_cohorts)[c(-1, -length(names(biomeE_out_annual_cohorts)))]
varnams_annual_cohorts_sofun <- names(sofun_out_annual_cohorts)[c(-1, -length(names(sofun_out_annual_cohorts)))]
all.equal(varnams_annual_cohorts_biomeE, varnams_annual_cohorts_sofun)

for (varnam in varnams_annual_cohorts_biomeE[1:23]){
  print(varnam)
  
  gg <- plot_tseries_byvar(varnam, biomeE_out_annual_cohorts, sofun_out_annual_cohorts)
  print(gg)

  gg <- plot_modobs_byvar(varnam, biomeE_out_annual_cohorts, sofun_out_annual_cohorts)
  print(gg)
}

out_modobs <- purrr::map(as.list(varnams_annual_cohorts_biomeE),
                         ~plot_modobs_byvar(., biomeE_out_annual_cohorts, sofun_out_annual_cohorts))
```

```{r, include=T, eval=T}
# Other option for plots using ggplot2()
dim(biomeE_out_annual_cohorts)
dim(sofun_out_annual_cohorts)
sofun_out_annual_cohorts <- sofun_out_annual_cohorts[,-24]
biomeE_out_annual_cohorts <- biomeE_out_annual_cohorts[,-24]

plot_dfs <- lapply(names(biomeE_out_annual_cohorts),function(nm)data.frame(col1 = biomeE_out_annual_cohorts[,nm], col2 = sofun_out_annual_cohorts[,nm]))

for (idx in seq_along(plot_dfs))
  print(ggplot()+geom_smooth(data = plot_dfs[[idx]], aes(x=col1, y=col2),color="blue") +
      ggtitle(names(biomeE_out_annual_cohorts)[idx]))

```

Compared to BiomeE model outputs and found differences that are due to plant-soil feedbacks, i.e., when holding mineralN constant, the outputs are exactly identical. Slightly different dynamics are caused by mineralN. See file vegetation.mod.f90 in sofun and BiomE-Allocation (vegn%mineralN = 0.00025).



