---
title: "Example"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(rsofun)
load_dependencies_rsofun()
```

## Example

### Simulation settings

Manually select some sites from which we're going to use the data for evaluation and calibration.
```{r}
mysites <- c("BE-Vie", "DE-Tha", "DK-Sor", "FI-Hyy", "IT-Col", "NL-Loo", "US-MMS", "US-WCr", "US-UMB", "US-Syv", "DE-Hai", "IT-MBo", "US-GLE", "FR-Fon", "NL-Hor", "US-UMd", "AU-Dry", "DE-Obe", "IT-Tor", "US-Wi4")
```

Create a site meta info table that contains all the site-specific information that is used to force site-simulations (e.g. starting year, number of simulations years, elevation, etc.). For FLUXNET2015 data, required meta info is provided by the `rsofun` package (data frame `rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015`).
```{r}
path_siteinfo <- "~/.siteinfo_example_fortran.csv"
siteinfo <- rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015 %>% 
  dplyr::filter(sitename %in% mysites) %>%
  write_csv(path = path_siteinfo)
```

Now specify the simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = FALSE,
  tempstress         = FALSE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  # const_clim_year    = -9999,
  # const_lu_year      = -9999,
  # const_co2_year     = -9999,
  # const_ndep_year    = -9999,
  # const_nfert_year   = -9999,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)
```

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters).
```{r}
settings_sims <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```

### Define model parameters

First, let's do it by hand (calibration of parameters is shown later).
```{r}
params_modl <- list(
	kphio           = 0.04997714009213085,
	soilm_par_a     = 1.0,
	soilm_par_b     = 0.0,
	vpdstress_par_a = 0.2,
	vpdstress_par_b = 0.2,
	vpdstress_par_m = 5
	)
```

### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Get input

First, define input settings.
```{r}
settings_input <-  list(
    data                     = NA,
    temperature              = "fluxnet2015",
    precipitation            = "fluxnet2015",
    vpd                      = "fluxnet2015",
    ppfd                     = "fluxnet2015",
    netrad                   = "fluxnet2015",  #  c("fluxnet2015", "watch_wfdei"),
    patm                     = "fluxnet2015",
    netrad                   = NA,
    cloudcover               = "cru",
    path_input               = "~/sofun_inputs/example/",
    path_watch_wfdei         = "~/data/watch_wfdei/",
    path_cru                 = "~/data/cru/ts_4.01/",
    path_MODIS_FPAR_MCD15A3H = "~/data/fluxnet_subsets/fapar_MODIS_FPAR_MCD15A3H_gee_MCD15A3H_fluxnet2015_gee_subset/",
    path_MODIS_EVI_MOD13Q1   = "~/data/fluxnet_subsets/fapar_MODIS_EVI_MOD13Q1_gee_MOD13Q1_fluxnet2015_gee_subset/",
    path_co2                 = "~/data/co2/cCO2_rcp85_const850-1765.csv",
    path_fluxnet2015         = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
    path_fluxnet2015_hh      = "~/data/FLUXNET-2015_Tier1/20191024/HH/",
    get_from_remote          = FALSE,
    settings_gee             = get_settings_gee( 
      bundle      = "fpar", 
      python_path = "/Users/benjaminstocker/Library/Enthought/Canopy_64bit/User/bin/python",
      gee_path    = "~/gee_subset/gee_subset/"
      ),
  fapar = "MODIS_FPAR_MCD15A3H",
  splined_fapar = TRUE
  )
```

Then, get the input data.
```{r message=FALSE, warning=FALSE, echo=FALSE}
ddf_input <- prepare_input_sofun(
  settings_input             = settings_input,
  settings_sims              = settings_sims,
  overwrite_csv_climate_lev1 = FALSE,
  overwrite_csv_climate_lev2 = TRUE,
  overwrite_csv_climate_lev3 = TRUE,
  overwrite_rdata_climate    = TRUE,
  overwrite_csv_fapar        = FALSE,
  verbose                    = FALSE
  )
```

### Run the model

Run the model for all the sites specified in the first step.
```{r}
df_drivers <- collect_drivers_sofun( 
  settings       = settings_sims, 
  forcing        = ddf_input, 
  df_soiltexture = df_soiltexture
  )

## run for a single site
mod <- run_sofun_f_bysite( 
  df_drivers$sitename[1], 
  df_drivers$params_siml[[1]], 
  df_drivers$siteinfo[[1]], 
  df_drivers$forcing[[1]], 
  df_drivers$df_soiltexture[[1]], 
  params_modl = params_modl, 
  makecheck = TRUE 
  )

## Run for the full set of sites
ptm <- proc.time()
df_output <- runread_sofun_f(
     df_drivers, 
     params_modl = params_modl, 
     makecheck = TRUE,
     parallel = FALSE
     )
print(ptm)

# microbenchmark::microbenchmark(
#   runread_sofun_f(
#     df_drivers, 
#     params_modl = params_modl, 
#     makecheck = TRUE,
#     parallel = TRUE,
#     ncores = 4
#     ),
#   runread_sofun_f(
#     df_drivers, 
#     params_modl = params_modl, 
#     makecheck = TRUE,
#     parallel = FALSE
#     ),
#   times = 5,
#   units = 's'
# )

df_output$out_sofun[[1]] %>% 
  ggplot(aes(x=date, y=gpp)) +
  geom_line() + 
  labs(title = df_output$sitename[[1]], subtitle = "SOFUN output")
```

### Calibrate

Define calibration settings.
```{r}
## Define calibration settings common for all setups
settings_calib <- list(
  method              = "gensa",
  targetvars          = c("gpp"),
  timescale           = list( gpp = "d" ),
  path_fluxnet2015    = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  path_fluxnet2015_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/",
  threshold_GPP       = 0.5,
  path_gepisat        = "~/data/gepisat/v3_fluxnet2015/daily_gpp/",
  maxit               = 5, # (5 for gensa) (30 for optimr)    #
  sitenames           = mysites,
  filter_temp_max     = 35.0,
  filter_drought      = FALSE,
  metric              = "rmse",
  dir_results         = "./",
  name                = "ORG",
  par                 = list( kphio = list( lower=0.03, upper=0.07, init=0.0496 ) ),
  datasource          = list( gpp = "fluxnet2015_NT" ),
  filter_temp_min     = NA,
  filter_soilm_min    = NA
 )
```
 
Get calibration target data.
```{r warning=FALSE, message=FALSE}
ddf_obs_calib <- get_obs_calib( 
  settings_calib = settings_calib, 
  dplyr::select(df_drivers, sitename, siteinfo) %>% tidyr::unnest(siteinfo), 
  settings_input
  )
```

Calibrate the model.

```{r warning=FALSE, message=FALSE}
set.seed(1982)
settings_calib <- calib_sofun( 
  settings_calib, 
  df_drivers, 
  ddf_obs = ddf_obs_calib 
  )
```

The calibrated parameters are returned by `calib_sofun()` as part of the list:
```{r}
print(settings_calib$par_opt)
```


### Evaluate

Run the model once again with these parameters and evaluate results.
```{r warning=FALSE, message=FALSE}
mylist <- readr::read_csv("~/eval_pmodel/myselect_fluxnet2015.csv") %>% 
  dplyr::filter( use==1 ) %>% 
  dplyr::pull( Site )

settings_eval <- list(
  sitenames           = settings_sims$sitename,
  sitenames_siteplots = mylist,
  agg                 = 8,
  path_fluxnet2015_d  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_fluxnet2015_w  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_7d/original/unpacked/",
  path_fluxnet2015_m  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1m/original/unpacked/",
  path_fluxnet2015_y  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1y/original/unpacked/",
  path_gepisat_d      = "~/data/gepisat/v3_fluxnet2015/daily_gpp/",
  benchmark           = list( gpp = c("fluxnet2015_NT") ),
  remove_premodis     = TRUE
  )
```

Get evaluation data (benchmarking data).
```{r warning=FALSE, message=FALSE}
filn <- "./obs_eval.Rdata"
if (file.exists(filn)){
  load(filn)
} else {
  obs_eval  <- get_obs_eval( 
    settings_eval = settings_eval, 
    settings_sims = settings_sims, 
    overwrite     = TRUE, 
    light         = TRUE,
    add_forcing   = FALSE
  )
  save(obs_eval, file = filn)
} 
```

Now run the model with calibrated parameters.
```{r}
params_modl <- list(
  kphio           = settings_calib$par_opt[["kphio"]],
  soilm_par_a     = 1.0,
  soilm_par_b     = 0.0,
  vpdstress_par_a = 0.2,
  vpdstress_par_b = 0.2,
  vpdstress_par_m = 5
  )

mod <- runread_sofun_f(
  df_drivers, 
  params_modl = params_modl, 
  makecheck = TRUE
  ) %>% 
  rename(id = sitename) %>% 
  unnest(out_sofun)
```

And finally do the evaluation.
```{r warning=FALSE, message=FALSE}
out_eval <- eval_sofun( 
  mod, 
  settings_eval, 
  settings_sims, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

Print some results.
```{r}
out_eval$gpp$fluxnet2015$metrics$xdaily_pooled
```

```{r}
out_eval$gpp$fluxnet2015$data$xdf %>% rbeni::analyse_modobs2("mod", "obs", type = "heat")
```

```{r}
out_eval$gpp$fluxnet2015$data$ddf %>% 
  dplyr::filter(sitename=="BE-Vie" & year(date) < 2005) %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = obs), col="black") +
  geom_line(aes(y = mod), col="red") + 
  labs(title = "BE-Vie")

out_eval$gpp$fluxnet2015$data$ddf %>% 
  dplyr::filter(sitename=="AU-Dry" & year(date) > 2010) %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = obs), col="black") +
  geom_line(aes(y = mod), col="red") + 
  labs(title = "AU-Dry")
```
