---
title: "Example"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(rsofun)
load_dependencies_rsofun()
```

## Example

### Simulation settings

Manually select some sites from which we're going to use the data for evaluation and calibration.
```{r}
mysites <- c("BE-Vie", "DE-Tha", "DK-Sor", "FI-Hyy", "IT-Col", "NL-Loo", "US-MMS", "US-WCr", "US-UMB", "US-Syv", "DE-Hai", "IT-MBo", "US-GLE", "FR-Fon", "NL-Hor", "US-UMd", "AU-Dry", "DE-Obe", "IT-Tor", "US-Wi4")

mysites <- c("ORNL")
```

Create a site meta info table that contains all the site-specific information that is used to force site-simulations (e.g. starting year, number of simulations years, elevation, etc.). For FLUXNET2015 data, required meta info is provided by the `rsofun` package (data frame `rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015`).
```{r}
path_siteinfo <- "~/.siteinfo_example_fortran.csv"
siteinfo <- rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015 %>% 
  dplyr::filter(sitename %in% mysites) %>%
  write_csv(path = path_siteinfo)

siteinfo <- data.frame(sitename="ORNL", lon = 0.0, lat = 0.0, elv = 0.0)
```

Now specify the simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
      spinup                = TRUE,
      spinupyears           = 1800,
      recycle               = 11, # 9 or 11
      firstyeartrend        = 1998,
      nyeartrend            = 11, #9 or 11
      outputhourly          = FALSE,
      outputdaily           = TRUE,
      do_U_shaped_mortality = FALSE,
      update_annaulLAImax   = FALSE,
      do_closedN_run        = TRUE
      )
```

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters).
```{r}
#settings_sims <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```

### Define model parameters

#### Tile-level parameters

```{r}
params_tile <- list(
  soiltype     = 3,
  FLDCAP       = 0.4,
  WILTPT       = 0.05,
  K1           = 2.0,  # turnover rate of fast SOM per year
  K2           = 0.1,   # 0.12 # 0.2, # 0.075 # turnover rate of slow SOM per year
  K_nitrogen   = 0.0, # 1.2 # 2.5 # rate of a year, 2.5
  etaN         = 0.0, # 0.025 # loss rate with runoff
  MLmixRatio   = 0.6, # 0.8 # fraction of microbes to litter
  l_fract      = 0.0,
  retransN     = 0.0,
  f_N_add      = 0.02, # 0.005, # 0.02
  f_initialBSW = 0.005
  # alpha_FR   =  1.2; gamma_FR = 12.0; dSlowSOM = 2.5 * iSOM + 1.5; rho_N_up0 = 0.1
  )

```

#### Species-level parameters

```{r}
params_species <- tibble(
  lifeform      = c(0,       1,        1,        1,        1,      1,        1,        1,      1,        1,        1),    # 0: grass; 1 Woody
  phenotype     = c(0,       1,        1,        1,        1,      1,        1,        1,      1,        1,        0),    # 0: Deciduous; 1 Evergreen
  pt            = c(1,       0,        0,        0,        0,      0,        0,        0,      0,        0,        0),    # 0: C3; 1: C4
  seedlingsize  = c(0.02,    0.05,     0.05,     0.05,     0.05,   0.05,     0.05,     0.05,   0.05,     0.05,     0.05),  # initial size of seedlings
  LMA           = c(0.035,   0.14,     0.14,     0.14,     0.14,   0.14,    0.14,   0.14,     0.14,     0.14,     0.035),# Leaf mass per unit area
  phiRL         = c(4.0,     1.0,      2.0,      3.0,      4.0,     5.0,     6.0,      7.0,     8.0,    9.0,      3.0),
  LNbase        = c(0.8E-3,  0.8E-3,   0.8E-3,   0.8E-3,   0.8E-3, 0.8E-3,   0.8E-3,   0.8E-3, 0.8E-3,   0.8E-3,   0.8E-3),   # kgN m-2 leaf, Vmax = 0.03125*LNbase
  laimax        = c(3.0,     3.5,      3.5,      3.5,      3.5,    3.5,      3.5,      3.5,    3.5,      3.5,      3.8),  # maximum crown LAI
  LAI_light     = c(3.0,     3.5,      3.5,      3.5,      3.5,    3.5,      3.5,      3.5,    3.5,      3.5,      3.8),  # Light-limited crown LAI
  Nfixrate0     = c(0.0,     0.0,      0.0,      0.0,      0.0,    0.0,      0.0,      0.0,    0.0,      0.0,      0.0),  # 0.03 kgN kgRootC-1 yr-1
  NfixCost0     = c(12.,     12.,      12.,      12.,      12.,    12.,      12.,      12.,    12.,      12.,      12.),  # 12, 24 gC/gN
  phiCSA        = c(1.25E-4, 0.25E-4,  0.25E-4,  0.25E-4,  0.25E-4,0.25E-4,  0.25E-4,  0.25E-4,0.25E-4,  0.25E-4,  0.25E-4),
  mortrate_d_c  = c(0.02,    0.02,     0.02,     0.02,     0.02,   0.02,     0.02,     0.02,   0.02,     0.02,     0.02),    # canopy tree mortality rate, year-1
  mortrate_d_u  = c(0.2,     0.08,     0.08,     0.08,     0.08,   0.08,     0.08,     0.08,   0.08,     0.08,     0.08),    # understory tree mortality rate, year-1
  maturalage    = c(0.5,     5,        5,        5,        5,      5,        5,        5,      5,        5,        5),
  fNSNmax       = c(5,       5,        5,        5,        5,      5,        5,        5,      5,        5,        5)
  ) %>% 
  bind_rows(., slice(., 1:5))
```

#### Soil parameters

By layers.
```{r}
# adopted from datatypes.mod.f90 l.538
params_soil <- tibble(
  type              = c("Coarse",  "Medium",   "Fine",    "CM",     "CF",     "MF",    "CMF",    "Peat",    "MCM"),
  GMD               = c(0.7, 0.4, 0.3, 0.1, 0.1, 0.07, 0.007, 0.3, 0.3),
  GSD               = c(5.0, 5.3, 7.4, 6.1, 6.1, 14.0, 15.0, 7.4, 7.4),
  vwc_sat           = c(0.380, 0.445, 0.448, 0.412, 0.414, 0.446, 0.424, 0.445, 0.445),
  chb               = c( 3.5,   6.4,  11.0,   4.8,   6.3,   8.4,   6.3,   6.4,   6.4),
  psi_sat_ref       = c(-600., -790., -910., -1580., -1680., -1880., -5980., -790., -790.), # Pa
  k_sat_ref         = c(130.8, 75.1, 53.2, 12.1, 11.1, 12.7, 1.69, 53.2, 53.2),      # mol/(s MPa m)
  alphaSoil         = rep(1.0, 9),
  heat_capacity_dry = c(1.2e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.4e6,   1.0)
  )
```

#### Initial cohort specification

```{r}
init_cohort <- tibble(
 init_cohort_species = c(1,      2,        3,        4,      5,        6,        7,      8,        9,       10),
 init_cohort_nindivs = c(0.02,   0.02,     0.02,     0.02,   0.02,     0.02,     0.02,   0.02,     0.02,    0.02),
 init_cohort_bsw     = c(0.2,    0.2,      0.2,      0.2,    0.2,      0.2,      0.2,    0.2,      0.2,     0.2),
 init_cohort_bHW     = c(0.0,    0.0,      0.0,      0.0,    0.0,      0.0,      0.0,    0.0,      0.0,     0.0),
 init_cohort_nsc     = c(0.5,    0.5,      0.5,      0.5,    0.5,      0.5,      0.5,    0.5,      0.5,     0.5)
)

```

#### Initial soil pools

```{r}
# high N input --> Deciduous
# low  N input --> Evergreen
init_soil <- list(
 init_fast_soil_C    = 0.15,   # kg C m-2    # C/N = 15, 0.75 kg SOM ~ 50 gN
 init_slow_soil_C    = 16.0,   # kg C m-2    # C/N = 40
 init_Nmineral       = 2.5E-3, # kg N m-2
 N_input             = 0.0E-3  # 2.4E-3, # kg N m-2 yr-1, N deposit: 0.8 g N m-2 yr-1
)

```


### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Get input

Get the input from the nice weird looking file with half-hourly data.

The file has the following variables (units):

 Year  DOY  Hour  T_air          Q_air          Wind_speed      Precip          Pressure        R_global_in      R_longwave_in   CO2
                      K          kg/kg          m/s     kg_H2O/m2/s               pa               W/m2            W/m2            ppm     

```{r}
forcing <- read_delim("~/cagibi/BiomeE-Allocation/model/input/ORNL_forcing.txt", col_names = TRUE, delim = "\t") %>% 
  mutate(date = lubridate::ymd_hm( paste0( as.character(YEAR), "-01-01 00:00" ) ) + days(DOY-1) + hours(HOUR) ) %>% 
  mutate(sitename = "ORNL") %>% 
  filter(!(mday(date)==29 & month(date)==2)) %>% 
  dplyr::select(-date, -sitename)
#forcing <- subset(forcing, YEAR<=2006)
#ORNL_forcing.txt
#US-WCrforcing.txt
complete <- rep(seq(ymd("1998-01-01"), ymd("2008-12-31"), by = "days"), each = 24)

ddf_complete <- tibble(date = complete)

# forcing <- read_delim("~/BiomeE-Allocation/model/input/US-WCrforcing_clean.txt", col_names = TRUE, delim = "\t")

```


### Run the model

Run the model for all the sites specified in the first step.
```{r}
# siteinfo <- list(lon = 99, lat = 99, elv = 99)
mod <- run_lm3ppa_f_bysite( "ORNL", 
                            params_siml, 
                            siteinfo, 
                            forcing, # ddf_input
                            params_tile, 
                            params_species, 
                            params_soil, 
                            init_cohort, 
                            init_soil, 
                            makecheck = TRUE)

df_drivers <- collect_drivers_sofun( 
  settings       = settings_sims, 
  forcing        = ddf_input
  )

## run for a single site
mod <- run_lm3ppa_f_bysite( 
  df_drivers$sitename[1], 
  df_drivers$params_siml[[1]], 
  df_drivers$siteinfo[[1]], 
  df_drivers$forcing[[1]], 
  df_drivers$df_soiltexture[[1]], 
  params_modl = params_modl, 
  makecheck = TRUE,
  method = "lm3ppa"
  )

# ## Run for the full set of sites
# ptm <- proc.time()
# df_output <- runread_sofun_f(
#      df_drivers, 
#      params_modl = params_modl, 
#      makecheck = TRUE,
#      parallel = FALSE,
#      method = "lm3ppa"
#      )
# print(ptm)

# microbenchmark::microbenchmark(
#   runread_sofun_f(
#     df_drivers, 
#     params_modl = params_modl, 
#     makecheck = TRUE,
#     parallel = TRUE,
#     ncores = 4
#     ),
#   runread_sofun_f(
#     df_drivers, 
#     params_modl = params_modl, 
#     makecheck = TRUE,
#     parallel = FALSE
#     ),
#   times = 5,
#   units = 's'
# )

df_output$out_sofun[[1]] %>% 
  ggplot(aes(x=date, y=gpp)) +
  geom_line() + 
  labs(title = df_output$sitename[[1]], subtitle = "SOFUN output")
```

### Calibrate

Define calibration settings.
```{r}
## Define calibration settings common for all setups
settings_calib <- list(
  method              = "gensa",
  targetvars          = c("gpp"),
  timescale           = list( gpp = "d" ),
  path_fluxnet2015    = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  path_fluxnet2015_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/",
  threshold_GPP       = 0.5,
  path_gepisat        = "~/data/gepisat/v3_fluxnet2015/daily_gpp/",
  maxit               = 5, # (5 for gensa) (30 for optimr)    #
  sitenames           = mysites,
  filter_temp_max     = 35.0,
  filter_drought      = FALSE,
  metric              = "rmse",
  dir_results         = "./",
  name                = "ORG",
  par                 = list( kphio = list( lower=0.03, upper=0.07, init=0.0496 ) ),
  datasource          = list( gpp = "fluxnet2015_NT" ),
  filter_temp_min     = NA,
  filter_soilm_min    = NA
 )
```
 
Get calibration target data.
```{r warning=FALSE, message=FALSE}
ddf_obs_calib <- get_obs_calib( 
  settings_calib = settings_calib, 
  dplyr::select(df_drivers, sitename, siteinfo) %>% tidyr::unnest(siteinfo), 
  settings_input
  )
```

Calibrate the model.

```{r warning=FALSE, message=FALSE}
set.seed(1982)
settings_calib <- calib_sofun( 
  settings_calib, 
  df_drivers, 
  ddf_obs = ddf_obs_calib 
  )
```

The calibrated parameters are returned by `calib_sofun()` as part of the list:
```{r}
print(settings_calib$par_opt)
```


### Evaluate

Run the model once again with these parameters and evaluate results.
```{r warning=FALSE, message=FALSE}
mylist <- readr::read_csv("~/eval_pmodel/myselect_fluxnet2015.csv") %>% 
  dplyr::filter( use==1 ) %>% 
  dplyr::pull( Site )

settings_eval <- list(
  sitenames           = settings_sims$sitename,
  sitenames_siteplots = mylist,
  agg                 = 8,
  path_fluxnet2015_d  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_fluxnet2015_w  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_7d/original/unpacked/",
  path_fluxnet2015_m  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1m/original/unpacked/",
  path_fluxnet2015_y  = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1y/original/unpacked/",
  path_gepisat_d      = "~/data/gepisat/v3_fluxnet2015/daily_gpp/",
  benchmark           = list( gpp = c("fluxnet2015_NT") ),
  remove_premodis     = TRUE
  )
```

Get evaluation data (benchmarking data).
```{r warning=FALSE, message=FALSE}
filn <- "./obs_eval.Rdata"
if (file.exists(filn)){
  load(filn)
} else {
  obs_eval  <- get_obs_eval( 
    settings_eval = settings_eval, 
    settings_sims = settings_sims, 
    overwrite     = TRUE, 
    light         = TRUE,
    add_forcing   = FALSE
  )
  save(obs_eval, file = filn)
} 
```

Now run the model with calibrated parameters.
```{r}
params_modl <- list(
  kphio           = settings_calib$par_opt[["kphio"]],
  soilm_par_a     = 1.0,
  soilm_par_b     = 0.0,
  vpdstress_par_a = 0.2,
  vpdstress_par_b = 0.2,
  vpdstress_par_m = 5
  )

mod <- runread_sofun_f(
  df_drivers, 
  params_modl = params_modl, 
  makecheck = TRUE
  ) %>% 
  rename(id = sitename) %>% 
  unnest(out_sofun)
```

And finally do the evaluation.
```{r warning=FALSE, message=FALSE}
out_eval <- eval_sofun( 
  mod, 
  settings_eval, 
  settings_sims, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

Print some results.
```{r}
out_eval$gpp$fluxnet2015$metrics$xdaily_pooled
```

```{r}
out_eval$gpp$fluxnet2015$data$xdf %>% rbeni::analyse_modobs2("mod", "obs", type = "heat")
```

```{r}
out_eval$gpp$fluxnet2015$data$ddf %>% 
  dplyr::filter(sitename=="BE-Vie" & year(date) < 2005) %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = obs), col="black") +
  geom_line(aes(y = mod), col="red") + 
  labs(title = "BE-Vie")

out_eval$gpp$fluxnet2015$data$ddf %>% 
  dplyr::filter(sitename=="AU-Dry" & year(date) > 2010) %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = obs), col="black") +
  geom_line(aes(y = mod), col="red") + 
  labs(title = "AU-Dry")
```
