---
title: "Running the C-only model at CH-Oe1"
author: "Benjamin Stocker"
output: rmarkdown::html_vignette
code-fold: true
vignette: >
  %\VignetteIndexEntry{Running the C-only model at CH-Oe1}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.5
)
```

## Aim

This vignette demonstrates a site-level simulation with the carbon-only
configuration of the C-model. We use daily meteorological forcing from the
FLUXNET grassland site CH-Oe1. In the C-only configuration, carbon allocation
is prescribed and plant growth is not limited by nitrogen availability.

## Install rsofun

Install rsofun from `cnmodel` branch and install
```{r installation, eval=FALSE}
# # geco-bern packages not on CRAN
# if (!require(remotes)) {
#   install.packages(remotes)
# }

# # uncomment below if not already installed
# # remotes::install_github("geco-bern/rsofun", ref = "cnmodel")
# library(rsofun)

# When working from within the rsofun repo as the project locally
library(devtools)
devtools::document()
devtools::install()
```

Load other required packages.
```{r libraries}
library(rsofun)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(cowplot)
```

## Load data

The CH-Oe1 forcing and evaluation test data is distributed through FluxDataKit. Load the drivers for all sites and filter to retain only driver data for CH-Oe1 (any other site may be selected instead).

```{r data}
# Variant 1) load the manually downloaded file:
df_drivers <- readRDS(
  "~/data_2/FluxDataKit/v3.4/zenodo_upload/rsofun_driver_data_v3.4.2.rds"
) |> 
  filter(sitename == "CH-Oe1")

fdk_site_info <- tibble(read.csv(
  "~/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_info.csv"
))

fdk_site_fullyearsequence <- tibble(read.csv(
  "~/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_fullyearsequence.csv"
))

# # Variant 2) or automatically download and load the file:
# library(zen4R) # install.packages("zen4R")
# download_path <- tempdir()
# zen4R::download_zenodo(
#   path = download_path,
#   "10.5281/zenodo.12092555",
#   files = c(
#     "rsofun_driver_data_v3.2.rds",
#     "fdk_site_info.csv",
#     "fdk_site_fullyearsequence.csv"
#   )
# )

# driver <- readRDS(file.path(download_path, "rsofun_driver_data_v3.2.rds"))
# fdk_site_info <- tibble(read.csv(file.path(download_path, "fdk_site_info.csv")))
# fdk_site_fullyearsequence <- tibble(read.csv(file.path(
#   download_path,
#   "fdk_site_fullyearsequence.csv"
# )))
```

## Parameters

The parameter values follow `analysis/example_cnmodel.R`. Although nitrogen
parameters are supplied because the same Fortran interface serves both model
configurations, they do not constrain carbon allocation in this run.

```{r parameters}
pars <- list(
  # P-model
  kphio = 0.09,
  kphio_par_a = 0,
  kphio_par_b = 21.0,

  # Active recovery from cold hardening
  dehardening_gdd_midpoint = 90, # GDD midpoint of dehardening
  dehardening_gdd_slope = 0.1, # sigmoid slope (degree-day^-1)
  dehardening_gdd_base = 4.5, # base temperature for GDD accumulation (deg C)

  soilm_thetastar = 26.7,
  soilm_betao = 0.0,
  beta_unitcostratio = 187.5,
  rd_to_vcmax = 0.025,
  tau_acclim = 10.0,
  kc_jmax = 0.58,

  # Plant carbon and nitrogen
  f_nretain = 0.5,
  fpc_tree_max = 0.95,
  growtheff = 0.55,
  r_root = 2.5,
  r_sapw = 0,
  exurate = 0.003,
  k_decay_leaf = 5,
  k_decay_root = 5,
  k_decay_labl = 0,
  k_decay_sapw = 0.1,
  r_cton_root = 37.0,
  r_cton_wood = 100.0,
  r_cton_seed = 15.0,
  nv_vcmax25 = 0.02 * 13681.77,
  ncw_min = 0.08 * 1.116222,
  r_n_cw_v = 0.0,
  r_ctostructn_leaf = 30.86173,
  kbeer = 0.4,

  # Phenology
  gddbase = 5.0,
  ramp = 0.0,
  phentype = 2.0,

  # Soil physics
  perc_k1 = 5.0,
  thdiff_wp = 0.2,
  thdiff_whc15 = 0.8,
  thdiff_fc = 0.4,
  forg = 0.01,
  wbwp = 0.029,
  por = 0.421,
  fsand = 0.82,
  fclay = 0.06,
  fsilt = 0.12,

  # Water and energy balance
  kA = 107.0,
  kalb_sw = 0.17,
  kalb_vis = 0.03,
  kb = 0.20,
  kc = 0.25,
  kCw = 1.05,
  kd = 0.50,
  ke = 0.0167,
  keps = 23.44,
  kWm = 220.0,
  kw = 0.26,
  komega = 283.0,
  maxmeltrate = 3.0,

  # Soil biogeochemistry
  klitt_af10 = 1.2,
  klitt_as10 = 0.35,
  klitt_bg10 = 0.35,
  kexu10 = 50.0,
  ksoil_fs10 = 0.021,
  ksoil_sl10 = 7.0e-4,
  ntoc_crit1 = 0.45,
  ntoc_crit2 = 0.76,
  cton_microb = 10.0,
  cton_soil = 9.77,
  fastfrac = 0.985,

  # Nitrogen uptake and transformations
  eff_nup = 1.0e-4,
  minimumcostfix = 1.0,
  fixoptimum = 25.15,
  a_param_fix = -3.62,
  b_param_fix = 0.27,
  maxnitr = 5.0e-5,
  non = 0.01,
  n2on = 5.0e-4,
  kn = 83.0,
  kdoc = 17.0,
  docmax = 1.0,
  dnitr2n2o = 0.01,

  # Fixed allocation used by the C-only setup
  frac_leaf = 0.45,
  frac_wood = 0.0,
  frac_avl_labl = 0.3,
  tmppar = 9999.0,
  nuptake_kc = 600.0,
  nuptake_kv = 5.0,
  nuptake_vmax = 0.2
)
```

## Complement the driver data

The prepared driver object is a nested data frame containing data from multiple sites, including
metadata, simulation settings, and daily forcing. We add the management and
nitrogen-deposition columns expected by `runread_cnmodel_f()`. No harvest is
applied in this example, while 5 g C m^-2^ of seed carbon is added on 15
February of every year.

N-deposition is assumed to yield annual totals of 1.5 gN m^−2^ yr^−1^, evenly distributed over days and into NOy and NHx, respectively.

```{r drivers}
## add new required columns to forcing
use_cseed <- 5
cn_seed <- 20
use_nseed <- use_cseed / cn_seed

df_drivers <- df_drivers |>
  mutate(
    forcing = purrr::map(
      forcing,
      ~ mutate(
        .,
        fharv = 0.0,
        dno3 = 1.5 / (2 * 365),
        dnh4 = 1.5 / (2 * 365)
      )
    ),
    forcing = purrr::map(
      forcing,
      ~ mutate(
        .,
        fharv = ifelse(month(date) == 7 & mday(date) == 15, 0.0, 0.0),
        cseed = ifelse(month(date) == 2 & mday(date) == 15, use_cseed, 0.0),
        nseed = ifelse(month(date) == 2 & mday(date) == 15, use_nseed, 0.0)
      )
    )
  )

# Use a five-year recycling period for the spin-up.
df_drivers$params_siml[[1]]$spinupyears <- 2002
df_drivers$params_siml[[1]]$recycle <- 5

# This switch selects prescribed allocation and disables interactive C-N
# constraints. It is the essential difference from the CN run.
df_drivers$params_siml[[1]]$c_only <- TRUE
```

## Run the simulation

`runread_cnmodel_f()` runs the compiled model and returns one row per site, with
daily results in the nested `data` column.

```{r run-model}
output_uncalibrated <- runread_cnmodel_f(
  drivers = df_drivers,
  par = pars
)
```

ds
```{r postprocessing}
# Daily values
ddf <- output_uncalibrated |>
  select(sitename, data) |>
  unnest(data) |>
  mutate(
    nee = -(gpp - rleaf - rwood - rroot - rhet - rgrow)
  )

# Annual totals
adf <- ddf |>
  mutate(year = year(date)) |>
  group_by(year) |>
  summarise(
    lai = max(lai),
    gpp = sum(gpp),
    npp = sum(npp),
    npp_leaf = sum(npp_leaf),
    npp_root = sum(npp_root),
    npp_wood = sum(npp_wood),
    npp_seed = sum(npp_seed),
    cleaf = mean(cleaf),
    nleaf = mean(nleaf),
    clitt = mean(clitt),
    nlitt = mean(nlitt),
    croot = mean(croot),
    cwood = mean(cwood),
    cseed = mean(cseed),
    nup = sum(nup),
    netmin = sum(netmin),
    ninorg = mean(pno3 + pnh4),
    nloss = sum(nloss)
  )

# Mean annual totals
adf_mean <- adf |>
  filter(year %in% 2002:2008) |>
  summarise(across(where(is.numeric), mean)) |>
  dplyr::select(-year) |>
  mutate(cue = npp / gpp, cnleaf = cleaf / nleaf, cnlitt = clitt / nlitt) |>
  pivot_longer(cols = everything())

# Mean seasonal cycle
msc <- ddf |>
  mutate(
    month = month(date),
    doy = yday(date)
  ) |>
  group_by(doy) |>
  summarise(
    doy = mean(doy),
    gpp = mean(gpp),
    npp = mean(npp),
    npp_leaf = mean(npp_leaf),
    npp_root = mean(npp_root),
    npp_wood = mean(npp_wood),
    npp_seed = mean(npp_seed),
    rcex = mean(rcex),
    rleaf = mean(rleaf),
    rwood = mean(rwood),
    rroot = mean(rroot),
    rgrow = mean(rgrow),
    dclabl = mean(dclabl),
    nup = mean(nup),
    netmin = mean(netmin),
    ninorg = mean(pno3 + pnh4),
    nloss = mean(nloss),
    dnleaf = mean(dnleaf),
    dnroot = mean(dnroot),
    dnwood = mean(dnwood),
    dnseed = mean(dnseed),
    dnlabl = mean(dnlabl),
    nresorb = mean(nresorb)
  ) |>
  mutate(date = ymd("2023-01-01") + floor(doy) - 1)
```

## Summarise and visualise the output

### Some daily time series

```{r plot-output}
plot_data <- ddf |>
  select(date, gpp, npp, lai, cleaf, croot, nee) |>
  pivot_longer(-date, names_to = "variable", values_to = "value")

ggplot(plot_data, aes(date, value)) +
  geom_line(linewidth = 0.3) +
  facet_wrap(~ variable, scales = "free_y", ncol = 2) +
  labs(x = NULL, y = NULL) +
  theme_classic()
```

### Allocation

```{r allocationplot}
gg1 <- msc |>
  ggplot() +
  geom_ribbon(
    aes(
      date,
      ymin = 0,
      ymax = npp_leaf,
      fill = "Leaf production"
    )
  ) +
  geom_ribbon(
    aes(
      date,
      ymin = npp_leaf,
      ymax = npp_leaf + npp_seed + npp_root, # seed C is zero
      fill = "Root production"
    )
  ) +
  geom_ribbon(
    aes(
      date,
      ymin = npp_root + npp_leaf + npp_seed,
      ymax = npp_root + npp_leaf + npp_seed + rcex,
      fill = "Exudates"
    )
  ) +
  geom_ribbon(
    aes(
      date,
      ymin = npp_root + npp_leaf + npp_seed + rcex,
      ymax = npp_root + npp_leaf + npp_seed + rcex + dclabl,
      fill = "Labile turnover"
    )
  ) +
  geom_ribbon(
    aes(
      date,
      ymin = npp_root + npp_leaf + npp_seed + rcex + dclabl,
      ymax = npp_root +
        npp_leaf +
        npp_seed +
        rcex +
        dclabl +
        rleaf +
        rwood +
        rroot +
        rgrow,
      fill = "Respiration"
    )
  ) +
  geom_line(
    aes(
      date,
      gpp,
      color = "GPP"
    )
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b",
    expand = c(0, 0)
  ) +
  khroma::scale_color_okabeito(name = "") +
  scale_fill_manual(
    name = "",
    values = c(
      "Leaf production" = "#009E73",
      "Root production" = "#E69F00",
      "Exudates" = "#CC79A7",
      "Labile turnover" = "#0072B2",
      "Respiration" = "#56B4E9"
    )
  ) +
  theme_classic() +
  labs(
    x = "",
    y = expression(paste("C flux (gC m"^-2, " d"^-1, ")"))
  ) +
  scale_y_continuous(expand = c(0, 0))

msc_agg <- msc |>
  mutate(
    npp_leaf = npp_leaf + npp_seed,
    resp = rleaf + rwood + rroot + rgrow
  ) |>
  summarise(
    gpp = sum(gpp),
    `Leaf production` = sum(npp_leaf),
    `Root production` = sum(npp_root),
    Exudates = sum(rcex),
    `Labile turnover` = sum(dclabl),
    Respiration = sum(resp)
  )

tmp <- msc_agg |>
  dplyr::select(-gpp) |>
  pivot_longer(cols = everything()) |>
  mutate(category = "Sinks") |>
  bind_rows(
    tibble(
      name = "gpp",
      value = msc_agg$gpp,
      category = "GPP"
    )
  )

gg2 <- tmp |>
  mutate(
    name = factor(
      name,
      levels = rev(c(
        "Leaf production",
        "Root production",
        "Exudates",
        "Labile turnover",
        "Respiration"
      ))
    )
  ) |>
  ggplot() +
  geom_bar(
    aes(x = category, y = value, fill = name),
    stat = "identity",
    position = "stack"
  ) +
  theme_classic() +
  scale_fill_manual(
    name = "",
    values = c(
      "Leaf production" = "#009E73",
      "Root production" = "#E69F00",
      "Exudates" = "#CC79A7",
      "Labile turnover" = "#0072B2",
      "Respiration" = "#56B4E9"
    )
  ) +
  labs(
    x = "",
    y = expression(paste("Cumulative C flux (gC m"^-2, "yr"^-1, ")"))
  ) +
  theme(legend.position = "none")

plot_grid(
  gg1,
  gg2,
  rel_widths = c(1, 0.5),
  labels = c("a", "b")
)
```

## Evaluation

```{r}
ddf_modobs <- ddf |>
  transmute(
    date,
    nee_mod = nee,
    gpp_mod = gpp,
    reco_mod = rleaf + rwood + rroot + rhet
  ) |>
  left_join(
    df_drivers |>
      select(sitename, forcing) |>
      unnest(forcing) |>
      select(
        date,
        nee_obs = nee,
        gpp_obs = gpp,
        nee_qc,
        gpp_qc
      )
  ) |>
  mutate(reco_obs = nee_obs + gpp_obs)
```

### Leaf area index

We use the Terra MODIS MOD15A2H product, which provides 8-day LAI estimates at
500 m resolution. `Lai_500m` is stored with a scale factor of 0.1. The
`FparLai_QC` layer is an 8-bit quality flag. Following the
[MOD15 Collection 6.1 user guide](https://modis-land.gsfc.nasa.gov/pdf/MOD15_C61_UserGuide_April2020.pdf),
we retain observations for which:

- bit 0 (`MODLAND_QC`) reports good quality;
- bit 2 does not flag substantial dead-detector retrievals;
- bits 3--4 report no significant or mixed clouds; and
- bits 5--7 (`SCF_QC`) indicate the main radiative-transfer algorithm, either
  without saturation (0) or with usable saturation (1).

The filtering therefore removes cloud-affected pixels, empirical back-up
retrievals, failed retrievals, and other poor-quality observations. The
download and preparation are reproduced by
data-raw/generate_modis_lai_ch_oe1.R; the resulting quality-controlled time
series is distributed with rsofun as modis_lai_ch_oe1.

```{r modis-lai-data}
model_period <- range(ddf$date)

data("modis_lai_ch_oe1", package = "rsofun")

modis_lai <- modis_lai_ch_oe1 |>
  filter(between(date, model_period[1], model_period[2]))
```

```{r plot-lai}
ggplot() +
  geom_line(
    data = ddf,
    aes(date, lai, color = "Modelled"),
    linewidth = 0.4
  ) +
  geom_point(
    data = modis_lai,
    aes(date, lai, color = "MODIS"),
    size = 1.2,
    alpha = 0.8
  ) +
  scale_color_manual(
    name = "",
    values = c(Modelled = "red", MODIS = "black")
  ) +
  labs(
    x = NULL,
    y = expression(paste("LAI (m"^2, " m"^-2, ")"))
  ) +
  theme_classic()
```

### GPP

#### Time series

```{r}
ddf_modobs |>
  pivot_longer(
    c(gpp_mod, gpp_obs),
    names_to = "source",
    values_to = "gpp"
  ) |>
  ggplot(aes(date, gpp, color = source)) +
  geom_line() +
  scale_color_manual(
    name = "",
    values = c(
      gpp_obs = "black",
      gpp_mod = "red"
    ),
    labels = c(
      gpp_obs = "Observed",
      gpp_mod = "Modelled"
    )
  ) +
  labs(
    x = "",
    y = expression(paste("GPP (gC m"^-2, " d"^-1, ")"))
  ) +
  theme_bw()
```

#### Mean seasonal cycle

```{r}
msc_gpp <- ddf_modobs |>
  filter(!(month(date) == 2 & mday(date) == 29)) |>
  pivot_longer(
    c(gpp_mod, gpp_obs),
    names_to = "source",
    values_to = "gpp"
  ) |>
  mutate(
    month = month(date),
    day = mday(date)
  ) |>
  group_by(source, month, day) |>
  summarise(
    gpp_q25 = quantile(gpp, 0.25, na.rm = TRUE),
    gpp_q75 = quantile(gpp, 0.75, na.rm = TRUE),
    gpp = mean(gpp, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(if_all(c(gpp, gpp_q25, gpp_q75), is.finite)) |>
  mutate(date = make_date(2023, month, day))

msc_gpp |>
  ggplot(aes(date, gpp, color = source)) +
  geom_ribbon(
    aes(ymin = gpp_q25, ymax = gpp_q75, fill = source),
    alpha = 0.2,
    color = NA
  ) +
  geom_line() +
  scale_color_manual(
    name = "",
    values = c(
      gpp_obs = "black",
      gpp_mod = "red"
    ),
    labels = c(
      gpp_obs = "Observed",
      gpp_mod = "Modelled"
    )
  ) +
  scale_fill_manual(
    values = c(
      gpp_obs = "black",
      gpp_mod = "red"
    ),
    guide = "none"
  ) +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  labs(
    x = "",
    y = expression(paste("Mean GPP (gC m"^-2, " d"^-1, ")"))
  ) +
  theme_classic()
```

### Ecosystem respiration

Modelled ecosystem respiration is the sum of leaf, wood, root, and
heterotrophic respiration. Observed ecosystem respiration is derived from the
flux partitioning identity $\mathrm{NEE} = R_\mathrm{eco} - \mathrm{GPP}$, so
$R_\mathrm{eco} = \mathrm{NEE} + \mathrm{GPP}$.

#### Time series

```{r}
ddf_modobs |>
  pivot_longer(
    c(reco_mod, reco_obs),
    names_to = "source",
    values_to = "reco"
  ) |>
  ggplot(aes(date, reco, color = source)) +
  geom_line() +
  scale_color_manual(
    name = "",
    values = c(
      reco_obs = "black",
      reco_mod = "red"
    ),
    labels = c(
      reco_obs = "Observed",
      reco_mod = "Modelled"
    )
  ) +
  labs(
    x = "",
    y = expression(paste(R[eco], " (gC m"^-2, " d"^-1, ")"))
  ) +
  theme_bw()
```

#### Mean seasonal cycle

```{r}
msc_reco <- ddf_modobs |>
  filter(!(month(date) == 2 & mday(date) == 29)) |>
  pivot_longer(
    c(reco_mod, reco_obs),
    names_to = "source",
    values_to = "reco"
  ) |>
  mutate(
    month = month(date),
    day = mday(date)
  ) |>
  group_by(source, month, day) |>
  summarise(
    reco_q25 = quantile(reco, 0.25, na.rm = TRUE),
    reco_q75 = quantile(reco, 0.75, na.rm = TRUE),
    reco = mean(reco, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(if_all(c(reco, reco_q25, reco_q75), is.finite)) |>
  mutate(date = make_date(2023, month, day))

msc_reco |>
  ggplot(aes(date, reco, color = source)) +
  geom_ribbon(
    aes(ymin = reco_q25, ymax = reco_q75, fill = source),
    alpha = 0.2,
    color = NA
  ) +
  geom_line() +
  scale_color_manual(
    name = "",
    values = c(
      reco_obs = "black",
      reco_mod = "red"
    ),
    labels = c(
      reco_obs = "Observed",
      reco_mod = "Modelled"
    )
  ) +
  scale_fill_manual(
    values = c(
      reco_obs = "black",
      reco_mod = "red"
    ),
    guide = "none"
  ) +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  labs(
    x = "",
    y = expression(paste("Mean ", R[eco], " (gC m"^-2, " d"^-1, ")"))
  ) +
  theme_classic()
```

### NEE

#### Time series

```{r}
ddf_modobs |>
  pivot_longer(
    c(nee_mod, nee_obs),
    names_to = "source",
    values_to = "nee"
  ) |>
  ggplot(aes(date, nee, color = source)) +
  geom_line() +
  scale_color_manual(
    name = "",
    values = c(
      nee_obs = "black",
      nee_mod = "red"
    ),
    labels = c(
      nee_obs = "Observed",
      nee_mod = "Modelled"
    )
  ) +
  labs(
    x = "",
    y = expression(paste("NEE (gC m"^-2, " d"^-1, ")"))
  ) +
  theme_bw()
```

#### Mean seasonal cycle

```{r}
msc_nee <- ddf_modobs |>
  filter(!(month(date) == 2 & mday(date) == 29)) |>
  pivot_longer(
    c(nee_mod, nee_obs),
    names_to = "source",
    values_to = "nee"
  ) |>
  mutate(
    month = month(date),
    day = mday(date)
  ) |>
  group_by(source, month, day) |>
  summarise(
    nee_q25 = quantile(nee, 0.25, na.rm = TRUE),
    nee_q75 = quantile(nee, 0.75, na.rm = TRUE),
    nee = mean(nee, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(if_all(c(nee, nee_q25, nee_q75), is.finite)) |>
  mutate(date = make_date(2023, month, day))

msc_nee |>
  ggplot(aes(date, nee, color = source)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_ribbon(
    aes(ymin = nee_q25, ymax = nee_q75, fill = source),
    alpha = 0.2,
    color = NA
  ) +
  geom_line() +
  scale_color_manual(
    name = "",
    values = c(
      nee_obs = "black",
      nee_mod = "red"
    ),
    labels = c(
      nee_obs = "Observed",
      nee_mod = "Modelled"
    )
  ) +
  scale_fill_manual(
    values = c(
      nee_obs = "black",
      nee_mod = "red"
    ),
    guide = "none"
  ) +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  labs(
    x = "",
    y = expression(paste("Mean NEE (gC m"^-2, " d"^-1, ")"))
  ) +
  theme_classic()
```
