---
title: "calib_pmodel"
author: "Beni Stocker"
date: "12/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This document is compiled by "knitting" the Rmarkdown file `knit_calib_pmodel.Rmd`. All source code is available in the repository [calib_pmodel](https://bitbucket.org/labprentice/calib_pmodel) on bitbucket, group *labprentice* (login required). The workflow for model calibration is described below. 

## Collect data

Read forcing data and P-model output into a common dataframe. To collect all daily forcing data and flux measurement data from the FLUXNET 2015 dataset, fAPAR forcing data, and SOFUN output data into a single data frame, use `get_modobs.R`. 
Define ensemble name:
```{r}
simsuite <- "fluxnet2015"
outputset <- c( "s15" )
```

Run in R:
```{r get_modobs, eval=TRUE}
source("get_modobs.R")
get_modobs( simsuite = simsuite, outputset = outputset, outdir = "~/data/fluxnet_sofun/" )
```

This performs a data "cleaning" (see `clean_fluxnet.R`), calculates additional variables (e.g. soilm_mean, soilm_obs_mean) and writes all data as an R data frame (tibble) that we're using for the calibration into the file `df_modobs_<simsuite>_*.Rdata`.


## P-model calibration

After writing `df_modobs_<simsuite>_*.Rdata` by the previous step, we simply have to read in the data, remove cold and dry days and calibrate the apparent quantum yield efficiency parameter `kphio_app` using GPP from the FLUXNET 2015 data (`GPP_NT_VUT_REF`) and the linear regression function `lm()` using least-squares minimisation. After running SOFUN with an arbitrary parameter value `kphio_app`, the `calibrate_pmodel.R` calculates the optimised `kphio_app` to best match observations during relatively moist and warm days. SOFUN writes parameter values as meta information into NetCDF outputs. `calibrate_pmodel.R` reads this and and calculates by how much this (arbitrary) value has to be scaled to best match observations. These are the steps:

###  1. Load the calibration function

```{r}
source("calib_pmodel.R")
```

###  2. Load the data collected by 'get_modobs.R'

```{r}
load( paste0( "df_modobs_", simsuite, "_", outputset, "_with_SWC_v4.Rdata" ) )  # loads df_fluxnet
ndays_0 <- nrow( df_fluxnet )
print( paste( "total number of days, unfiltered: ", ndays_0 ) )
```

###  3. Aggregate to monthly data

```{r message=FALSE}
require(dplyr)
require(lubridate)
df_fluxnet <- df_fluxnet %>% group_by( mysitename, year(date), month(date) ) %>%
														 summarise_all( mean, na.rm=TRUE )
```

###  4. Read GPP NetCDF output to get parameter value of quantum yield efficiency that was used in simulations

```{r message=FALSE}
## a random site
runname <- "FR-Pue"

## output directory
dirn <- paste( myhome, "sofun/output_nc_", simsuite, "_sofun/", outputset, "/", sep="" )

## a random variable
filn <- paste0( runname, ".d.gpp.nc" )

## read meta information
require(ncdf4)
path         <- paste0( dirn, filn )
nc           <- nc_open( path )
kphio_used   <- ncatt_get( nc, varid=0, attname="param_kphio_GrC3" )$value %>% as.numeric()
fapar_source <- ncatt_get( nc, varid=0, attname="fapar_source" )$value
nc_close(nc)
```

The quantum yield efficiency parameter used in simulations was:
```{r}
print(kphio_used)
```

###  5. Evaluate optimised parameter for a range of soil moisture cutoffs

```{r message=FALSE}
soilm_cutoff_range <- seq(0.1, 0.9, 0.05)
list_sens_soilm <- lapply( soilm_cutoff_range, function(x) calib_pmodel( df_fluxnet, 5.0, x, kphio_used ) )
kphio_opt  <- lapply( list_sens_soilm, function(x) x$kphio_opt ) %>% unlist()
kphio_corr <- lapply( list_sens_soilm, function(x) x$kphio_corr ) %>% unlist()
df_sens_soilm_cutoff <- tibble( soilm_cutoff=soilm_cutoff_range, kphio_opt=kphio_opt, kphio_corr=kphio_corr )
with( df_sens_soilm_cutoff, plot( soilm_cutoff, kphio_opt, pch=16, main=paste0("Sensitivity to soil moisture cutoff (temp. cutoff = 5 deg C, fAPAR data: ", fapar_source, ")"), font.main=1, cex.main=0.9 ))
```

###  6. Evaluate optimised parameter for a range of temperature cutoffs

```{r message=FALSE}
temp_cutoff_range <- seq(0, 20, 1)
list_sens_temp <- lapply( temp_cutoff_range, function(x) calib_pmodel( df_fluxnet, x, 0.4, kphio_used ) )
kphio_opt  <- lapply( list_sens_temp, function(x) x$kphio_opt ) %>% unlist()
kphio_corr <- lapply( list_sens_temp, function(x) x$kphio_corr ) %>% unlist()
df_sens_temp_cutoff <- tibble( temp_cutoff=temp_cutoff_range, kphio_opt=kphio_opt, kphio_corr=kphio_corr )
with( df_sens_temp_cutoff, plot( temp_cutoff, kphio_opt, pch=16, main=paste0("Sensitivity to temperature cutoff (soilm. cutoff = 0.4, fAPAR data: ", fapar_source, ")"), font.main=1, cex.main=0.9 ))
```

Here, we chose to set the soil moisture cutoff to 0.4 and the temperature cutoff to 5$^\circ$C. The optimal quantum yield efficiency (`kphio`) is then:
```{r}
out_calib <- calib_pmodel( df_fluxnet, temp_cutoff=5.0, soilm_cutoff=0.4, kphio_used=kphio_used )
print( paste( "Optimal kphio:", format( out_calib$kphio_opt, digits=4 ) ) )
```
This requires a correction factor to be applied to the modelled values. The correction factor is:
```{r}
print( paste( "Correction factor:", format( out_calib$kphio_corr, digits=4 ) ) )
```

###  7. Plot modelled vs. observed for given soil moisture (0.4) and temperature (5 deg C) cutoff after applying the correction factor.

```{r message=FALSE}
source("analyse_modobs.R")
stats <- with( filter( df_fluxnet, temp>5 & soilm_mean>0.4), analyse_modobs( GPP_NT_VUT_REF, out_calib$kphio_corr * gpp, yintersect0=FALSE, xlab=expression( paste( "modelled GPP (gC m"^{-1}, " d"^{-1}, ")")), ylab=expression( paste( "observed GPP (gC m"^{-1}, " d"^{-1}, ")")), do.plot=TRUE ) )
```
  