---
title: "Cost functions for parameter calibration"
author: "Pepa Aran"
date: "2022-10-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cost functions for parameter calibration}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(rsofun)
library(dplyr)
library(ggplot2)
```

The `rsofun` package allows to calibrate parameters of the `pmodel` via the `calib_sofun()` function and various cost functions starting with `cost_*`. In this vignette, we go over some examples on how to use these cost functions for parameter calibration and how to write your own custom one.



A simple approach to parameter calibration is to fit the one that leads to the best GPP prediction performance, in terms of the RMSE (root mean squared error) . With `create_cost_function()`, we can create a cost function that corresponds to the calibration setups in Stocker et al., 2020 GMD. For example, following the `FULL` setup, we can calibrate parameters `kphio`, `soilm_par_a` and `soilm_par_b`. We must always specify which values the fixed parameters should take, i.e. the parameters that aren't calibrated.

```{r}
# Set parameter values
pars <- list(
    kphio           = 0.04607080,
    soilm_par_a     = 2.75687824,
    soilm_par_b     = 1.68140444,
    tau_acclim_tempstress = 7.35259044,
    par_shape_tempstress  = 0.09863961
  )

# Create the cost function
cost_rmse_full <- create_cost_function(
  params_modl = pars,
  setup = 'FULL'
)
```

Now we can run the calibration routine, as follows.
```{r}
# Define calibration settings and parameter ranges from previous work
settings <- list(
  method = 'BayesianTools',
  targetvars = c('gpp'),
  metric = cost_rmse_full, # our cost function
  dir_results = "./",
  control = list(
    sampler = 'DEzs',
    settings = list(
      burnin = 500,
      iterations = 1500
    )),
  par = list(
    kphio = list(lower=0.04, upper=0.2, init=0.05),
    soilm_par_a = list(lower=0.1, upper=5, init=2.4),
    soilm_par_b = list(lower=1, upper=2, init=1.5)
  )
)

# Calibrate the model and optimize the free parameters using
# demo datasets
pars_calib <- calib_sofun(
  drivers = p_model_drivers,
  obs = p_model_validation,
  settings = settings
)
```