---
title: "Overview for package rsofun"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
# output:
#   html_document:
#     toc: true
#     toc_float: true
#     toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
---

# Environment

Load the package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r message=FALSE}
library(rsofun)
library(dplyr)

## load all rsofun dependencies
load_dependencies_rsofun()

## other crap
systr <- "''"    # for Mac
knitr::opts_knit$set( root.dir = rprojroot::find_rstudio_root_file() ) # does not work properly
## for developer setup
if (!file.exists("bash"))    system("ln -s inst/bash bash")
if (!file.exists("extdata")) system("ln -s inst/extdata extdata")
```


## Simplest setup

This is the simplest setup, where the SOFUN is executed as a function for a single time step and location.

### Simulation settings

```{r}
settings_sims_simple <- list( 
  setup = "simple", 
  implementation  = "fortran", 
  dir_sofun = "~/sofun/trunk/" 
  )
```

### Model setup

Use the `demo_pmodel` compilation for the simple setup

```{r}
setup_sofun_simple <- list(
  model      = "demo_pmodel",
  dir        = "~/sofun/trunk",
  do_compile = FALSE,
  simsuite   = FALSE
  )
```


### Prepare setup

This the simple setup, this just checks whether the model code is in place.

```{r, message=FALSE, warning=FALSE}
settings_sims_simple <- prepare_setup_sofun( settings = settings_sims_simple, setup = setup_sofun_simple )
```

### Run the model

Run SOFUN in the simple setup.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
## First update calibratable parameters.
knitr::opts_knit$set( root.dir = rprojroot::find_rstudio_root_file() )
params_opt <- readr::read_csv( paste0( path.package("rsofun"), "/extdata/params_opt_kphio_soilm_global.csv" ) ) # as an example
nothing <- update_params( params_opt, settings_sims_simple$dir_sofun )

## Check if executable is available, otherwise copy it from the R package's external data
## This is not done inside the function 'pmodel()'. It would slow it down unnecessarily
if (!file.exists( paste0( settings_sims_simple$dir_sofun, "/rundemo_pmodel"))){
  system( paste0( "cp extdata/rundemo_pmodel ", settings_sims_simple$dir_sofun ) )
}

## Run for different temperatures and plot
ppfd <- 800
vpd  <- 1000
co2  <- 400
elv  <- 0
fapar <- 1

## Fortran
## update parameter
params_opt <- dplyr::tibble( kphio=0.05 ) # , temp_ramp_edge=NA, soilm_par_a=NA, soilm_par_b=NA
nothing <- update_params( params_opt, "~/sofun/trunk/" )

ptm <- proc.time()
pmodel_fortran <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~pmodel( temp = ., vpd = vpd, co2 = co2, ppfd = ppfd, fapar = fapar, elv = elv, implementation = "fortran", sofundir = settings_sims_simple$dir_sofun ) )
time_for <- proc.time() - ptm

## R
ptm <- proc.time()
pmodel_R <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~rpmodel( tc = ., vpd = vpd, co2 = co2, elv = elv, kphio = 0.05, fapar = fapar, ppfd = ppfd, method_optci="prentice14", method_jmaxlim = "wang17", do_ftemp_kphio = FALSE ) 
  )
time_R <- proc.time() - ptm

plot(  seq( 0, 35, length.out = 100 ), pmodel_fortran %>% purrr::map_dbl("gpp"), ylim = c(0,300), type = "l", xlab = "Temperature (deg C)", ylab = "GPP (gC/d/m2)" )
lines( seq( 0, 35, length.out = 100 ), pmodel_R %>% purrr::map_dbl("gpp"), col = "red" )
legend( "topright", c("Fortran", "R"), lty = 1, col = c("black", "red"), bty = "n")
```


## Site-scale simulations

The examples shown below are for site-scale simulations, i.e., where one model run covers a single point only. Note that calibration and evaluation steps (described further below) are currently available only for site-scale simulations.

### Simulation settings

The simulation settings define the model simulations for each site and the the user-defined local environment. Simulation settings are defined as a named list. The user specifies only variables that are identical across simulations within an ensemble. Other variables that differ between simulation within an ensemble (e.g. start and end year of the simulation, vegetation type C3 or C4, longitude and latitude of the site, soil type, etc.) are to be specified in the meta info file `path_siteinfo` (a CSV file created separately). The function `prepare_setup_sofun()` complements the simulation settings by these additional components as lists with elements for each site, nested within the simulation settings list.

In the example given here, we use a site meta information file for FLUXNET 2015 Tier 1 sites, provided along with the `rsofun` package. It's located at the path `/rsofun/extdata/siteinfo_example.csv`. For demonstration, we define a set of site-scale simulations of length one. That is, for only one site, here "FR-Pue". Note that FLUXNET 2015 data and MODIS FPAR data is provided along with the package `rsofun` for this one site for demonstration purposes only  and provides data only for a reduced number of years (2007-2014). The site meta information file looks like this:
```{r message=FALSE}
siteinfo <- readr::read_csv( paste0( path.package("rsofun"), "/extdata/siteinfo_example.csv" ) )
print( siteinfo )
```

To conduct P-model simulations with the Fortran implementation, respective model code needs to be downloaded separately from the Github [sofun](https://github.com/stineb/sofun) repository and placed in a separate directory (`dir_sofun` below). Builds of this Fortran code are available for public download at XXX in order to avoid having to compile from source.


```{r}
settings_sims_sitescale <- list(
  path_siteinfo   = paste0( path.package("rsofun"), "/extdata/siteinfo_example.csv"),
  ensemble        = TRUE,
  setup           = "site",
  name            = "example",
  dir_sofun       = "~/sofun/trunk/",
  path_output     = "~/sofun/output_example_sofun/",
  path_output_nc  = "~/sofun/output_nc_example_sofun/",
  path_input      = "~/sofun/input_example_sofun/",
  grid            = NA,
  implementation  = "fortran",
  in_ppfd         = TRUE,
  in_netrad       = FALSE,
  recycle         = 1,
  spinupyears     = 10,
  calibvars       = c("gpp"),  # needed later for calibration setup, any of "gpp", "fapar", and "transp"
  soilmstress     = TRUE,
  tempstress      = TRUE,
  loutdgpp        = TRUE,
  loutdwcont      = TRUE,
  loutdaet        = TRUE,
  loutdpet        = TRUE,
  loutdalpha      = TRUE
  )
```

- `path_siteinfo`: Path (character string) of a CSV file that contains all the information needed for defining simulations. One row for each simulation in an ensemble, typically sites. Specifying this element is required if `ensemble = TRUE`. Columns are as follows:
    - site name, must be: column number 1
    - longitude of site, column must be named 'lon'
    - latitude of site, column must be named 'lat'
    - elevation of site, column must be named 'elv'
    - years for which simulation is to be done (corresponding to data availability from site),
      requires two columns named 'year_start' and 'year_end'.
- `ensemble`: TRUE if an ensemble of site-level simulations are to be run (This may become obsolte, this information is given already if the number of rows in the CSV file `path_siteinfo` is larger than one.)
- `setup`: String. One of `"simple"` (SOFUN used as a function, single time step, single location), `"site"` (site-scale simulation), or `"lonlat"` (spatial simulation on a longitude-latitude grid).
- `name`: a character string specifying the name of the ensemble (e.g. 'fluxnet2015') or of a single simulation.
- `dir_sofun`: Path (character string) where the model sits (corresponding to the parent directory of the respective git repository).
- `path_output`: Path (character string) where model output (ascii text files) is written to.
- `path_output_nc`: Path (character string) where NetCDF model output is written to.
- `path_input`: Path (character string) where model input is located.
- `grid`: Character string defining the type of grid used, e.g. `halfdeg` for half-degree resolution in lon. and lat. (only used in `lonlat` setup).
- `implementation`: Character string specifying whether Fortran (`implementation= "fortran"`) or Python (`implementation= "python"`) version is to be used.
- `in_ppfd`: Switch (`TRUE` of `FALSE`) whether PPFD should be read from data (prescribed) or simulated online using SPLASH and prescribed fractional cloud cover data.
- `recycle`: Periodicity (integer, number of years) of repeating forcing years during spinup (e.g. if the simulation start year is 1982 and `recycle=3`, then forcing years 1979-1981 are repeated during the duration of the model spinup, so that the last year of the spinup is forcing year 1981. For `recycle=1` and simulation start year 1982, forcing year 1981 is used for all years of the model spinup. Here, 'forcing year' refers to the year AD in the climate, CO2, fAPAR, etc. data used as model forcing.
- `spinupyears`: Integer, number of model spinup years before the transient simulation. Use `spinupyears > 0` if the model contains pool variables that that are simulated by dynamics that depend on their current state (typically soil water storage, or plant and soil carbon pools). Typically `spinupyears = 10` is sufficient to bring soil water pools to equilibrium across the globe (unless you chose a large soil water holding capacity).
- `calibvars`: A vector of strings specifying which variables are to be calibrated. Use `c()` (used as default) for a standard model setup. See vignette `calib_sofun.pdf` for an example of a model calibration setup.
- `soilmstress`: Switch (`TRUE` of `FALSE`) defining whether soil moisture stress function is to be applied to GPP.
- `tempstress`: Switch (`TRUE` of `FALSE`) defining whether temperature stress function is to be applied to GPP.
- `lout<var>`: Switch (`TRUE` of `FALSE`) whether variable `<var>` is to be written to output (ascii text file for time series). To be anbandoned, only NetCDF output should be maintained.
- `lout<var>`: Switch (`TRUE` of `FALSE`) whether variable `<var>` is to be written to output (NetCDF file).


### Input settings

Input settings define data sources used for SOFUN input. Two modes of specifying inputs are available. Either providing the input data directly (element `data`, not available for `"lonlat"` simulations), or by specifying a keyword for the source dataset as strings. E.g. `"fluxnet2015"` triggers specific functions that read input data from specifically formatted files, here all the site-level meteo data from FLUXNET 2015. Define the input settings as a list.

```{r}
settings_input_sitescale <-  list(
  data                     = NA,
  temperature              = "fluxnet2015",
  precipitation            = "fluxnet2015",
  vpd                      = "fluxnet2015",
  ppfd                     = "fluxnet2015",
  netrad                   = "fluxnet2015",
  cloudcover               = NA,
  fapar                    = "MODIS_FPAR_MCD15A3H",
  splined_fapar            = TRUE,
  path_co2                 = paste0( path.package("rsofun"), "/extdata/cCO2_rcp85_const850-1765.dat" ),
  path_fluxnet2015         = paste0( path.package("rsofun"), "/extdata/" ),
  path_MODIS_FPAR_MCD15A3H = paste0( path.package("rsofun"), "/extdata/" ),

  ## below is information for downloading input data from remote server
  get_from_remote          = TRUE,
  path_remote_watch_wfdei  = "/work/bstocker/labprentice/data/watch_wfdei/",
  path_remote_co2          = "/work/bstocker/labprentice/data/co2/cCO2_rcp85_const850-1765.dat",
  path_remote_cru          = "/work/bstocker/labprentice/data/cru/ts_4.01/",
  path_remote_MODIS_FPAR_MCD15A3H  = "/work/bstocker/labprentice/data/fapar_MODIS_FPAR_MCD15A3H_fluxnet2015_gee_subset/",
  path_remote_MODIS_EVI_MOD13Q1 = "/work/bstocker/labprentice/data/evi_MODIS_EVI_MOD13Q1_fluxnet2015_gee_subset/",
  path_remote_fluxnet2015  = "/work/bstocker/labprentice/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_remote_watch_wfdei  = "/work/bstocker/labprentice/data/watch_wfdei/",
  uname_remote             = "bstocker",
  address_remote           = "login.cx1.hpc.ic.ac.uk"
  )
```

- `data`: A named list of data frames, each containing the input data for one site with a column `"date"` specifying the date, and columns `fapar`, `tempvarnam`, `precvarnam`, `vpdvarnam`, `ppfdvarnam`, and `netradvarnam` specifying the fraction of absorbed photosynthetically active radiation, temperature, precipitation, vapour pressure deficit, photosynthetic photon flux density, and net radiation, respictively. Elements of the list `data` must be named according to site names. If `data = NA`, data will be read from files. If `data` is provided, all other elements of the input settings will be ignored.
- `temperature`: A character string specifying the source for temperature data. Any of `"fluxnet2015"`, `"watch_wfdei"`, and/or `"cru"`. This can also be a vector of strings (e.g. `c("fluxnet2015", "watch_wfdei")`, to specify priorities: first take FLUXNET 2015 data for periods where data is available. For remaining years (given by `date_start` and `date_end`), use WATCH-WFDEI data. If `"fluxnet2015"` is specified for any of `temperature`, `precipitation`, `vpd`, `ppfd`, or `netrad`, then `path_fluxnet2015` must be specified as well in the settings.
- `precipitation`: See `temperature`.
- `vpd`: See `temperature`.
- `ppfd`: See `temperature`.
- `netrad`: See `temperature`.
- `fapar`: A character string specifying the type of fAPAR data used as input. Implemented for use of data from CX1 are `"MODIS_FPAR_MCD15A3H"` and `"MODIS_EVI_MOD13Q1"`. Use `NA` in case no fAPAR data is used as forcing (internally simulated fAPAR).
- `splined_fapar`: Logical defining whether splined fAPAR data is to be used. If `FALSE`, linearly interpolated fAPAR data is used.
- `co2`: A character string specifying which CO$_2$ file should be used (globally uniform and identical for each site in an ensemble). All available CO$_2$ forcing files are located on CX1 at `/work/bstocker/labprentice/data/co2/`. `co2="cmip"` specifies that the CMIP-standard CO2 file `cCO2_rcp85_const850-1765.dat` should be used.
- `path_fluxnet2015`: A character string specifying the path where standard FLUXNET 2015 CSV files are located.
- `path_watch_wfdei`: A character string specifying the path where standard WATCH-WFDEI NetCDF files are located.
- `path_cru_ts4_01`: A character string specifying the path where standard CRU NetCDF files are located.
- `path<fapar>`: A character string specifying the path where site-specific fapar files are located. This element is named according to the `fapar` setting (element `fapar`). E.g., an element named `path_MODIS_FPAR_MCD15A3H` is required if `fapar = MODIS_FPAR_MCD15A3H`.


### Model setup
Define model setup as a list.
```{r}
setup_sofun_sitescale <- list(
  model      = "pmodel",
  dir        = "~/sofun/trunk",
  do_compile = FALSE,
  simsuite   = FALSE
  )
```

- `model`: For Fortran version: A character string specifying the compilation option. The name of the executable is derived from this as `"run<setup_sofun$model>"`.
- `dir`: A path (character) specifying the directory of where the executables are located (corresponds to the parent directory of the model git repository).
- `do_compile`: If `TRUE`, the model code is compiled as `make <setup_sofun$model>`. If `FALSE`, compiled executables are used (compiled with gfortran on a Mac 64-bit).
- `simsuite`: If `TRUE`, the SOFUN option for running an entire simulation suite (ensemble of simulations) with a single executable. (In the Fortran implementation, this this requires the main program file `sofun_simsuite.f90` to be compiled instead of `sofun.f90`). This is to be preferred over a set of individual runs submitted individually when doing calibration runs because the cost across the entire ensemble (mod-obs error) can thus be calculated online.


<!-- ## Spatial simulations -->

<!-- The examples shown below are for spatial simulations, i.e., where each model run covers multiple points located on a cartesian grid in longitude and latitude. Calibration and evaluation steps are currently available only for site-scale simulations. -->

<!-- ### Simulation settings -->

<!-- In contrast to site-scale simulations, set `setup = "lonlat"`. Typically, we would run single simulations. Hence, set `ensemble = FALSE`. In this setup also the meta info file specified by `path_siteinfo` is obsolete (longitude, latitude and elevation are defined for each gridcell in respective input files). The simulation start year and length (number of years covered) are specified directly in the simulation settings (`firstyeartrend` and `nyeartrend`). For spatial simulations, we have to choice to run the model on different grids. Set `grid = "halfdeg"` to run on a $0.5^{\circ} \times 0.5^{\circ}$ grid. Alternatively, use `grid = "1x1deg"` to run on a $1^{\circ} \times 1^{\circ}$ grid (no other resolutions are implemented yet). All other elements of `settings_sims` are described in the tab for site-scale simulations and are used the same way here. -->

<!-- Set `c3 = TRUE` and `c4 = FALSE` here to use only one single generic PFT that follows the C3 photosynthesis pathway.  -->

<!-- ```{r} -->
<!-- settings_sims_lonlat <- list(  -->
<!--   path_siteinfo   = NA, -->
<!--   ensemble        = FALSE, -->
<!--   setup           = "lonlat", -->
<!--   name            = "global_TEST", -->
<!--   dir_sofun       = "~/sofun/trunk/", -->
<!--   path_output     = "~/sofun/output_global_TEST_sofun/", -->
<!--   path_output_nc  = "~/sofun/output_nc_global_TEST_sofun/", -->
<!--   path_input      = "~/sofun/input_global_TEST_sofun/", -->
<!--   grid            = "halfdeg", -->
<!--   implementation  = "fortran", -->
<!--   in_ppfd         = TRUE, -->
<!--   firstyeartrend  = 1982, -->
<!--   nyeartrend      = 35, -->
<!--   recycle         = 1, -->
<!--   spinupyears     = 10, -->
<!--   calibvars       = c(), -->
<!--   soilmstress     = FALSE, -->
<!--   tempstress      = FALSE, -->
<!--   c3              = TRUE, -->
<!--   c4              = FALSE, -->
<!--   loutplant       = FALSE, -->
<!--   loutgpp         = TRUE, -->
<!--   loutwaterbal    = TRUE, -->
<!--   loutdalpha      = TRUE, -->
<!--   loutdgpp        = TRUE, -->
<!--   loutdwcont      = TRUE -->
<!--   ) -->
<!-- ``` -->


<!-- ### Input settings -->

<!-- The general difference in driving spatial simulations as opposed to site-scale simulations is that input is read not from text files (separate files for each site containing time series for each year), but from NetCDF files containing input data for each gridcell. Since these files are large, the input preparation step creates links to original files and places links in the pre-defined location (`./input/global/`) instead of writing new standardised input files as is done for the site-scale simulations. The model implements different modules for reading from specific types of NetCDF input files. Specific functions are available for reading WATCH-WFDEI climate data, CRU climate data, MODIS FPAR fAPAR data, etc. Select then accordingly by specifying elements in the list `settings_input`.  -->

<!-- The `lonlat` setup requires additional paths to additional input datasets to be specified (see below), the information of which is provided through the meta info file for site-scale simulations. Other elements are described in tab for site-scale simulations. -->

<!-- ```{r} -->
<!-- settings_input_lonlat <-  list(  -->
<!--   temperature              = "watch_wfdei", -->
<!--   precipitation            = "watch_wfdei", -->
<!--   vpd                      = "watch_wfdei", -->
<!--   ppfd                     = "watch_wfdei", -->
<!--   netrad                   = NA, -->
<!--   cloudcover               = "cru", -->
<!--   fapar                    = "fAPAR3g", -->
<!--   path_cx1data             = "~/data/", -->
<!--   path_watch_wfdei         = "~/data/watch_wfdei/", -->
<!--   path_cru                 = "~/data/cru/ts_4.01/", -->
<!--   path_fAPAR3g             = "~/data/fAPAR/fAPAR3g_v2/fAPAR3g_v2_1982_2016_FILLED.nc", -->
<!--   path_soiltype            = "~/data/soil/hwsd/soil_type_hwsd_halfdeg.cdf", -->
<!--   path_landmask            = "~/data/landmasks/gicew_halfdeg.cdf", -->
<!--   path_elevation           = "~/data/watch_wfdei/WFDEI-elevation.nc", -->
<!--   path_whc                 = "~/data/soil/soilgrids/whc_soilgrids_halfdeg_FILLED.nc", -->
<!--   path_landcover           = "~/data/landcover/modis_landcover_halfdeg_2010_FILLED.nc", -->
<!--   path_co2                 = "~/data/co2/cCO2_rcp85_const850-1765.dat" -->
<!--   ) -->
<!-- ``` -->

<!-- - `path_soiltype`: Path (character string) specifying the location of the soil type input file. The current implementation of SOFUN works with a discrete number of soil types (nine different ones) corresponding to the classes of the Harmonized World Soil Database. The values of a set of physical soil parameters are given for each soil class within SOFUN (parameter text files under `sofun/params/params_soil_sc??.dat`). Soil parameters are used within SOFUN (so far) only for the calculation of soil temperature. -->
<!-- - `path_landmask`: Path (character string) specifying the location of land mask input file. This defines which gridcells are land as opposed to ice or water. SOFUN is executed only on land gridcells. SOFUN works with the land mask from [LUH](http://luh.umd.edu/data.shtml) which defines the *fraction* of each gridcell covered by ice or water.  -->
<!-- - `path_elevation`: Path (character string) specifying the location of elevation input file. This contains the elevation in meters for each gridcell. -->
<!-- - `path_whc`:  Path (character string) specifying the location of the input file used to define the soil water holding capacity. This is independent here from the information provided in the soil type file.  -->
<!-- - `path_landcover`: Path (character string) specifying the location of the land cover input file. This information is used in SOFUN to define the sensitivity of the water stress function, distinguishing between grasses and trees. -->
<!-- - `path_co2`: Path (character string) specifying the location of the co2 input file. This is a text file specifying the (globally uniform) atmospheric CO2 concentration for each year. -->

<!-- ### Model setup -->

<!-- In this case, set `model = "gpmodel"` in order to compile the model for "lonlat" simulations. Other elements are described in tab for site-scale simulations. -->

<!-- ```{r} -->
<!-- setup_sofun_lonlat <- list(  -->
<!--   model      = "gpmodel", -->
<!--   dir        = "~/sofun/trunk", -->
<!--   do_compile = TRUE, -->
<!--   simsuite   = FALSE -->
<!--   ) -->
<!-- ``` -->


# Workflow

The example shown below is for a set of site-scale simulations.

<!-- ## Prepare meta info file -->

<!-- This setp is required only for site-scale simulations. -->

<!-- This step is either done by hand so that the file has a certain structure (for column names etc see above: `path_siteinfo`). Below is an example for creating such a meta info file (info for each site in an ensemble) for the FLUXNET 2015 Tier 1 ensemble. -->

<!-- ```{r, eval=FALSE, message=FALSE, warning=FALSE} -->
<!-- siteinfo <- prepare_metainfo_fluxnet2015( -->
<!--   settings_sims = settings_sims_sitescale, -->
<!--   settings_input = settings_input_sitescale, -->
<!--   overwrite = FALSE, -->
<!--   filn_elv_watch = paste0( settings_input_sitescale$path_cx1data, "watch_wfdei/WFDEI-elevation.nc" ) -->
<!--   ) -->
<!-- save(siteinfo, file = "siteinfo.Rdata") -->
<!-- ``` -->

<!-- - `overwrite`: Set to `TRUE` if meta info file in wide format is to be overwritten (original is in long format, a wide-format version is used by `prepare_metainfo_fluxnet2015()`). -->
<!-- - `filn_elv_watch`: File name (full path, character string) of the WATCH-WFDEI elevation file (NetCDF format) used to complement missing elevation information. -->


## Prepare simulation setup

Create a run directory with all the simulation parameter files (defining simulation years, length, etc.). This returs the `settings_sims` list, complemented by additional information. Calibration settings are an optional argument. When passed on, simulation parameter files will contain information which target variables are to be written to special calibration output files (a single file written for an entire ensemble).

```{r, eval=TRUE, message=FALSE, warning=FALSE}
settings_sims_sitescale <- prepare_setup_sofun(
  settings = settings_sims_sitescale,
  setup = setup_sofun_sitescale,
  write_paramfils = TRUE
  )
```

## Prepare inputs
Prepare SOFUN input (climate input, CO2, etc.). Complements `settings_input`. This will require inputs from the user through the prompt, entered in the console to specify whether data files should be downloaded from Imperial CX1. In case you chose to download, you must have access to CX1 and be connected to the Imperial VPN. Once asked (see console!), enter your user name on CX1. This also requires that no additional entering of the password is required. In order to set this up, you need to generate an SSH key pair beforehand (see [here](https://www.digitalocean.com/community/tutorials/how-to-set-up-ssh-keys--2).
```{r, eval=FALSE, message=FALSE, warning=FALSE, results="hide"}
inputdata <- prepare_input_sofun(
  settings_input = settings_input_sitescale,
  settings_sims = settings_sims_sitescale,
  return_data = TRUE,
  overwrite_climate = TRUE,
  overwrite_fapar = TRUE,
  verbose = TRUE
  )
```

## Calibrate the model

Define the calibration settings as.
```{r}
## Define calibration settings common for all setups
settings_calib <- list(
  name             = "example",
  method           = "optimr",
  targetvars       = c("gpp"),
  timescale        = list( gpp = "d" ),
  par              = list( kphio = list( lower=0.01, upper=0.12, init=0.05 ) ),
  datasource       = list( gpp = "fluxnet2015_NT" ),
  path_fluxnet2015 = paste0( path.package("rsofun"), "/extdata/" ),
  path_gepisat     = paste0( path.package("rsofun"), "/extdata/" ),
  maxit            = 50,
  sitenames        = "FR-Pue",  # dplyr::filter(siteinfo$light, c4 %in% c(FALSE, NA) )$mysitename, # calibrate for non-C4 sites
  filter_temp_min  = 10.0,
  filter_temp_max  = 35.0,
  filter_soilm_min = NA,     # this is dangerous: removes entire sites where no soil moisture is measured
  filter_drought   = FALSE,
  metric           = "rmse",
  dir_results      = "/Users/benjaminstocker/rsofun/calib_results" # use absolute path!
 )
```

- `name`: A character string to define a name used for this calibration, used in file names containing calibration outputs.
- `par`: list of parameters to calibrate with their lower and upper boundaries. This is rigid. If you chose to use different parameters or in a different order, modify code in `calib_sofun.R` (function `cost_rmse()`) and in the model source code (`sofun/src/sofun.f90`).
- `targetvars`: Character string for name of variable in SOFUN output
- `datasource`: Named list of character strings with data source identifiers for each calibration target variable. The list is named corresponding to variable names defined by 'targetvars'. The identifier triggers certain functions to be used for reading and processing observational data. Use, e.g., `datasource = list( gpp = "fluxnet2015_NT" )` to specify that observational data for the target variable `"gpp"` comes from FLUXNET 2015 dataset with GPP data based on the night-time flux decomposition method (`"NT"`). Alternatively, use GPP data based on the daytime method (`"fluxnet2015_NT"`) or Tyler Davis' new method (unpublished) (`"fluxnet2015_Ty"`). If multiple data sources are selected (e.g., `datasource = list( gpp = c("fluxnet2015_NT", "fluxnet2015_DT") )`), their mean is used for calibration.
- `timescale`: Named list of characters specifying the time scale used for aggregating modelled and observational data in time series before calculating the cost function. For naming the list, see point above. Set, e.g., `timescale = list( gpp = "d" )` to calibrate to observational GPP aggregated to daily intervals (i.e. not aggregated). Use `"w"` for weekly, `"m"` for monthly, or `"y"` for annual (XXX NOT YET IMPLEMENTED).
- `path_fluxnet2015`: Path (character string) for where FLUXNET 2015 data is located (data files may be in subdirectories thereof). This settings list element needs to be specified if any of `datasource` is `"fluxnet2015"`.
- `path_gepisat`: Path (character string) for where GePiSaT GPP data is located (data files may be in subdirectories thereof). This settings list element needs to be specified if any of `datasource` is `"gepisat"`.
- `sitenames`: Vector of character strings for site names of sites' data to be used for calibration.
- `filter_temp_min`: Minimum temperature of data used for calibration. Data points for days where air temperature is below, are removed (replaced by NA) in the observational data.
- `filter_temp_max`: Maximum temperature of data used for calibration. Data points for days where air temperature is above, are removed (replaced by NA) in the observational data.
- `filter_soilm_min`: Minimum soil moisture (relative, normalised to maximum of mean across multiple depths at each site) of data used for calibration. Data points for days where soil moisture is below, are removed (replaced by NA) in the observational data.
- `metric`: A character string specifying the metric to be used for calulating the cost.

Do the calibration
```{r message=FALSE, warning=FALSE, include=FALSE}
settings_calib <- calib_sofun(
  setup          = setup_sofun_sitescale,
  settings_calib = settings_calib,
  settings_sims  = settings_sims_sitescale,
  settings_input = settings_input_sitescale
  )
```

## Run the model

Run SOFUN with calibrated parameters.
```{r, message=FALSE, warning=FALSE}
params_opt <- readr::read_csv( paste0( settings_calib$dir_results, "/params_opt_", settings_calib$name, ".csv" ) ) # as an example
nothing <- update_params( params_opt, settings_sims_sitescale$dir_sofun )
mod <- runread_sofun(
  settings = settings_sims_sitescale,
  setup = setup_sofun_sitescale
  )
```

## Evaluate the model

Define model evaluation settings as a list.
```{r}
settings_eval <- list(
  sitenames = "FR-Pue",
  benchmark = list( gpp = c("fluxnet2015_NT") ),
  sitenames_siteplots = "FR-Pue",
  agg = 5,
  path_fluxnet2015_d = paste0( path.package("rsofun"), "/extdata/" ),
  path_fluxnet2015_w = "",
  path_fluxnet2015_m = paste0( path.package("rsofun"), "/extdata/" ),
  path_fluxnet2015_y = paste0( path.package("rsofun"), "/extdata/" ),
  path_gepisat_d     = "",
  dir_figs           = "~/rsofun/fig/"
  )
```

- `benchmark`: Named list of character strings with data source identifiers for each calibration target variable. The list is named corresponding to variable names defined by 'targetvars'. The identifier triggers certain functions to be used for reading and processing observational data. Use, e.g., `datasource = list( gpp = "fluxnet2015_NT" )` to specify that observational data for the target variable `"gpp"` comes from FLUXNET 2015 dataset with GPP data based on the night-time flux decomposition method (`"NT"`). Alternatively, use GPP data based on the daytime method (`"fluxnet2015_NT"`) or Tyler Davis' new method (unpublished) (`"fluxnet2015_Ty"`). If multiple data sources are selected (e.g., `datasource = list( gpp = c("fluxnet2015_NT", "fluxnet2015_DT") )`), their mean is used for calibration.


```{r, eval=TRUE, message=FALSE, warning=FALSE}
## Get observational data for evaluation
obs_eval <- get_obs_eval( settings_eval = settings_eval, settings_sims = settings_sims_sitescale, overwrite = TRUE )
out_eval <- eval_sofun( mod, settings_eval, settings_sims_sitescale, obs_eval = obs_eval, overwrite = TRUE )
```

The object returned by `eval_sofun()` includes a set of standard functions that can be applied to the evaluation data itself. Here is one example for observed versus modelled values.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
modobs_daily   <- out_eval$plot$modobs_daily()
modobs_xdf     <- out_eval$plot$modobs_xdaily()
modobs_mdf     <- out_eval$plot$modobs_monthly()
modobs_meandoy <- out_eval$plot$modobs_meandoy()
modobs_meandoy <- out_eval$plot$modobs_annual()
out_eval$plot$by_doy_allsites()
out_eval$plot$by_xoy_allsites()
modobs_anomalies_daily  <- out_eval$plot$modobs_anomalies_daily()
modobs_anomalies_xdaily <- out_eval$plot$modobs_anomalies_xdaily()
modobs_anomalies_annual <- out_eval$plot$modobs_anomalies_annual()
modobs_meanxoy <- out_eval$plot$modobs_meanxoy()
out_eval$plot$by_doy_allzones()
#out_eval$plot$modobs_spatial_annual()
modobs_spatial <- out_eval$plot$modobs_spatial()
```



