---
title: "P-model usage"
author: "Koen Hufkens-->changed by YP"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{P-model usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

devtools::load_all("D:/Github/ingestr/")
devtools::load_all("D:/Github/rsofun/")
library(rsofun)
library(dplyr)
library(ggplot2)

pars <- list(par = c(
    kphio           = 0.04607080,
    soilm_par_a     = 2.75687824,
    soilm_par_b     = 1.68140444,
    tau_acclim_tempstress = 7.35259044,
    par_shape_tempstress  = 0.09863961
  ))
```

The `rsofun` package and framework includes two main models. The `pmodel` and `lm3-ppa` (which in part relies on pmodel compents). Here we give a short example on how to run the `pmodel` on the included demo datasets to familiarize yourself with both the data structure and the outputs.

## Demo data

The package includes two demo datasets to run and validate pmodel output. These files can be directly loaded into your workspace by typing:

## using multiple sites data to do this test-->add by YP 
These are real data from the fluxnet sites. We can use these data to run the model, together with observations of GPP we can also parameterize `pmodel` parameters.


```{r}
library(rsofun)
#koen put the prepared drivers in sofunCalVal package-->load it:
#devtools::load_all("D:/Github/sofunCalVal/")-->not run at this moment becasue cannot access the euler
#drivers of gpp simulation:
load(paste0("D:/Github/sofunCalVal/data/","p_model_fluxnet_drivers.RDA"))
#gpp based on the EC data:
#-->gpp and gpp_unc in this datasets are the same adopt from Beni et al., 2018 New Phyto.
#and sites number is not the same as the "p_model_fluxnet_drivers.RDA"
load(paste0("D:/Github/sofunCalVal/data/","p_model_fluxnet_calVal.RDA")) 
#compare the sites in two datasets:
pos<-match(p_model_fluxnet_calval$sitename,p_model_fluxnet_drivers$sitename)
#-->subset the drivers data from the sites have EC data:
p_model_fluxnet_drivers_new<-p_model_fluxnet_drivers[pos,]
```

## firstly to select 3 sites to run the first round of the processes
```{r}
p_model_drivers_subset<-p_model_fluxnet_drivers_new%>%
  dplyr::slice(1:3)
p_model_drivers_subset
```

## Running the model
With all data prepared we can run the model using `runread_pmodel_f()`. This function takes the nested data structure and runs the model site by site, returning nested model output results matching the input drivers.

```{r}
# optimized parameters from previous work
params_modl <- list(
    kphio           = 0.09423773,
    soilm_par_a     = 0.33349283,
    soilm_par_b     = 1.45602286,
    tau_acclim_tempstress = 10,
    par_shape_tempstress  = 0.0
  )

#run the model for these parameters-->new output for 3 sites-->add by YP
output_subset<-c()
##!!report to Koen-->the names of variables are not same as the demo data
names(p_model_drivers_subset)<-c("sitename","forcing","params_siml","site_info","params_soil")
#report error to Koen-->some sites do not have the fapar-->then model cannot run
p_model_drivers_subset<-p_model_drivers_subset %>%
  dplyr::filter(sitename != "AR-Vir")
# for(i in 1:nrow(p_model_drivers_new)){
for(i in 1:nrow(p_model_drivers_subset)){
  output <- rsofun::runread_pmodel_f(
  p_model_drivers_subset[i,],
  par = params_modl
  )
output_subset<-bind_rows(output_subset,output)
}
```

### plotting output
##
We can now visualize both the model output and the measured values together.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)


# the main model output-->for each site
for (i in 1:nrow(output_subset)){
  sitename_run<-output_subset$sitename[i]
  model_data <- output_subset %>%
    filter(sitename==sitename_run)%>%
  tidyr::unnest(data)

  validation_data <- p_model_fluxnet_calval %>%
  filter(sitename==sitename_run)%>%
  tidyr::unnest(data)

p_gpp<-ggplot() +
    geom_line(data = model_data,
    aes(date,gpp,col="modeled gpp")) +
    geom_line(data = validation_data,
    aes(date,gpp,col="EC gpp")) +
    scale_color_manual("",values = c("modeled gpp"="tomato",
    "EC gpp"="cyan2"))+
    labs(x = "Date",y = "GPP(unit)")
print(p_gpp)
}

```

## Calibrating model parameters-->for multiple sites

To optimize new parameters based upon driver data and a validation dataset we must first specify an optimization strategy and settings, as well as parameter ranges.

```{r}
 for (i in 1:nrow(output_subset)){
  #(1)prepare the data(model and EC based gpp)
  sitename_run<-output_subset$sitename[i]
  model_drivers <- p_model_fluxnet_drivers_new %>%
    filter(sitename==sitename_run)
  # model_data <- output_subset %>%
  #   filter(sitename==sitename_run)%>%
  # tidyr::unnest(data)

  validation_data <- p_model_fluxnet_calval %>%
  filter(sitename==sitename_run)
  
  #(2)specify the optimization strategy:
  settings <- list(
    method              = "bayesiantools",
    targetvars          = c("gpp"),
    timescale           = list(targets_obs = "y"),
    sitenames           = sitename_run,
    metric              = cost_rmse_kphio,
    dir_results         = "./",
    name                = "ORG",  #!!ask Koen if this corresponds to the Beni et al., 2020
    control = list(
      sampler = "DEzs",
      settings = list(
        burnin = 1000,
        iterations = 5000
      )
    ),
    par = list(
      kphio = list(lower=0.04, upper=0.1, init = 0.05),
      soilm_par_a = list(lower=0, upper=5, init = 3.5),
      soilm_par_b = list(lower=1, upper=5, init=3.5),
      tau_acclim_tempstress = list(lower=0, upper=15, init=1),
      par_shape_tempstress = list(lower=0, upper=0.1, init=0)
      )
  )

# By default `rsofun` supports both optimization using the `GenSA` and `BayesianTools` packages. The above statement provides settings for a `BayesianTools` approach using the DEza sampler. For this example the burnin and iterations are kept artificially low. In a real scenario you will have to increase these values orders of magnitude. Keep in mind that optimization routines rely on a cost function, which, depending on its structure influences parameter selection. A limited set of cost functions is provided all starting with `cost_*` but the model structure is transparent and custom cost functions can be easily written.
# 
# In addition starting values and ranges are provided for the free parameters in the model. Free parameters include, kphio, a, b, tau and shape. Be mindful that with newer version of rsofun additional parameters might be introduced, so re-check vignettes and model descriptions when updating existing code.
# 
# With all settings defined the optimization function `calib_sofun()` can be called with driver data and observations specified.

#(3) calibrate the model and optimize free parameters
pars <- calib_sofun(
    drivers = model_drivers,  
    obs = validation_data,
    settings = settings
)

}



```

```{r eval = FALSE}
```

When successful the optimized parameters can be used to run subsequent modelling efforts, in this case slightly improving the model fit over a more global parameter set.

```{r}
# reformatting the parameter list
params_modl <- list(
    kphio           = pars$par[1],
    soilm_par_a     = pars$par[2],
    soilm_par_b     = pars$par[3],
    tau_acclim_tempstress = pars$par[4],
    par_shape_tempstress  = pars$par[5]
  )

# run the model for these parameters
output_new <- rsofun::runread_pmodel_f(
  p_model_drivers,
  par = params_modl
  )

# we only have one site so we'll unnest
# the main model output
model_data_new <- output_new %>%
  filter(sitename == "FR-Pue") %>%
  tidyr::unnest(data)

  ggplot() +
    geom_line(
    data = model_data,
    aes(
      date,
      gpp
    ),
    colour = "blue",
    alpha = 0.3
  ) +
  geom_line(
    data = validation_data,
    aes(
      date,
      gpp
    )
  ) +
    geom_line(
    data = model_data_new,
    aes(
      date,
      gpp
    ),
    colour = "red"
  ) +
  labs(
    x = "Date",
    y = "GPP"
  )

```

For details on the optimization settings we refer to the manuals of [GenSA](https://cran.r-project.org/web/packages/GenSA/index.html) and [BayesianTools](https://github.com/florianhartig/BayesianTools).
