---
title: "P-model usage"
author: "Koen Hufkens, Josefa Arán, Benjamin Stocker"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{P-model usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 5
)

library(rsofun)
library(dplyr)
library(tidyr)
library(ggplot2)
library(khroma)
library(cowplot)

# define function for evaluation plots
plot_eval <- function(df, is_gpp = TRUE, return_gg1 = FALSE, return_gg2 = FALSE){
  
  use_col <- ifelse(is_gpp, "tomato", "royalblue")
  use_lab <- ifelse(
    is_gpp,
    expression(paste("GPP (g C m"^-2, "s"^-1, ")")), 
    expression(paste("LE (W m"^-2, ")"))
    )
  use_lab_obs <- ifelse(
    is_gpp,
    expression(paste("Observed GPP (g C m"^-2, "s"^-1, ")")), 
    expression(paste("Observed LE (W m"^-2, ")"))
    )
  use_lab_mod <- ifelse(
    is_gpp,
    expression(paste("Modeled GPP (g C m"^-2, "s"^-1, ")")), 
    expression(paste("Modeled LE (W m"^-2, ")"))
    )
  lims <- ifelse(
    is_gpp,
    c(-0.5, 10),
    c(-20, 175)
  )
   
  # Plot GPP observed and modelled quartiles by day-of-year
  gg1 <- df |> 
    tidyr::pivot_longer(c(mod, obs), names_to = "type", values_to = "var") |> 
    mutate(doy = lubridate::yday(date)) |> 
    group_by(doy, type) |> 
    summarise(
      var_q25 = quantile(var, probs = c(0.25), na.rm = TRUE),
      var_q75 = quantile(var, probs = c(0.75), na.rm = TRUE)
      ) |> 
    ggplot() +
    geom_ribbon(
      aes(
        x = doy,
        ymin = var_q25,
        ymax = var_q75,
        fill = type
        ),
      alpha = 0.75
    ) +
    scale_fill_manual(
      name = "",
      values = c(
        "mod" = use_col,
        "obs" = 'grey20'
        ),
      labels = c(
        "mod" = "Mod.",
        "obs" = "Obs."
      )) +
    theme_classic() +
    theme() +
    labs(
      x = 'Date',
      y = use_lab,
      colour = ""
    )
  
  # get fit metrics
  rsq_val <- yardstick::rsq(df, mod, obs) |> pull(.estimate)
  rmse_val <- yardstick::rmse(df, mod, obs) |> pull(.estimate)
  bias_val <- mean(df$mod - df$obs, na.rm = TRUE)
  linmod <- lm(obs ~ mod + 0, data = df)
  slope_val <- coef(linmod)
  n_val <- df |> tidyr::drop_na() |> nrow()
  
  rsq_lab <- format(rsq_val, digits = 3)
  rmse_lab <- format(rmse_val, digits = 3)
  bias_lab <- format(bias_val, digits = 3)
  slope_lab <- format(slope_val, digits = 3)
  n_lab <- n_val
  
  subtitle <- bquote(italic(R)^2 == .(rsq_lab) ~ ~
                    RMSE == .(rmse_lab) ~ ~
                    bias == .(bias_lab) ~ ~
                    slope == .(slope_lab) ~ ~
                    italic(N) == .(n_lab))
  
  # plot modelled vs. observed (actually the reverse)
  gg2 <- df |> 
    ggplot(aes(x = mod, y = obs)) +
    geom_hex(bins = 50, show.legend = FALSE) +
    theme_classic() +
    geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
    geom_hline(yintercept = 0, linetype = "dotted") +
    geom_vline(xintercept = 0, linetype = "dotted")+ 
    geom_smooth(method = "lm", formula = y ~ x + 0, color = "red", size = 0.5, se = FALSE) +
    xlim(lims[1], lims[2]) + 
    ylim(lims[1], lims[2]) +
    labs(
      x = use_lab_mod,
      y = use_lab_obs,
      subtitle = subtitle
    )
  
  if (is_gpp){
    gg2 <- gg2 +
      khroma::scale_fill_batlowW(trans = "log", reverse = TRUE)
  } else {
    gg2 <- gg2 +
      khroma::scale_fill_davos(trans = "log", reverse = TRUE)

  }
  
  if (return_gg1){
    return(gg1)
  } else if (return_gg2){
    return(gg2)
  } else {
    cow <- cowplot::plot_grid(gg1, gg2, nrow = 1)
    return(cow)
  }
}
```

The `rsofun` package and framework includes two main models. The `pmodel` and `biomee` (which in part relies on P-model components). Here we give a short example on how to run the `pmodel` on the included demo datasets to familiarize yourself with both the data structure and the outputs.

## Demo data

The package includes two demo datasets to run and validate pmodel output using GPP observations. These files can be directly loaded into your workspace by typing:

```{r}
p_model_drivers
p_model_validation
```

These are data from the French FR-Pue FLUXNET site. Information about data structure, variable names, and their meaning and units can be found in the reference pages of `p_model_drivers` and `p_model_validation`. We can use these data to run the model, together with observations of GPP we can also calibrate `pmodel` parameters.

Another two datasets are provided as an example to validate the model against leaf traits data, rather than fluxes. Measurements of Vcmax25 (aggregated over species) for a subset of 4 sites from the GlobResp database (Atkin et al., 2015) are given in `p_model_validation_vcmax25` and the corresponding forcing for the P-model is given in `p_model_drivers_vcmax25`. Since leaf traits are only measured once per site, the forcing used is a single year of average climate (the average measurements between 2001 and 2015 on each day of the year).

```{r}
p_model_drivers_vcmax25
p_model_validation_vcmax25
```

For the remainder of this vignette, we will use the GPP flux datasets. The workflow is exactly the same for leaf traits data.

The script `data-raw/generate_pmodel_driver_data.R` implements the subsetting of rsofun driver data to the demo data provided through this package.

## Running model

With all data prepared we can run the P-model using `runread_pmodel_f()`. This function takes the nested data structure and runs the model site by site, returning nested model output results matching the input drivers.

```{r}
# define model parameter values from previous work
# ------------------------------------------------------
# Note that in the phydro branch of rsofun,
#    whc must be included in params_modl, rather than in site_info
# ------------------------------------------------------
params_modl <- list(
  kphio              = 0.04998,    # setup ORG in Stocker et al. 2020 GMD
  kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
  kphio_par_b        = 1.0,
  soilm_thetastar    = 0.6 * p_model_drivers$site_info[[1]]$whc,
  beta_unitcostratio = 146.0,
  rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
  tau_acclim         = 30.0,
  kc_jmax            = 0.41
  )

# run the model for these parameters
output <- rsofun::runread_pmodel_f(
  p_model_drivers,
  par = params_modl
  )
```

### Plotting output

We can now visualize both the model output and the measured values together.

```{r}
# Create dataframe for plotting
df_plot <- output |>
  unnest(data) |>
  select(date, mod = gpp) |> 
  left_join(
    p_model_validation |>
      unnest(data) |> 
      select(date, obs = gpp),
    by = join_by(date)
  )

plot_eval(df_plot)
```

## Calibrating model parameters

To optimize new parameters based upon driver data and a validation dataset we must first specify an optimization strategy and settings, as well as a cost function and parameter ranges.

```{r}
settings <- list(
  method               = "GenSA",
  metric               = cost_rmse_pmodel,
  control = list(maxit = 3000),
  par = list(
    # all model parameters are calibratable
    kphio = list(lower = 0.03, upper = 0.99, init = 0.05),
    kphio_par_a = list(lower = -0.005, upper = 0, init = -0.0025),
    kphio_par_b = list(lower = 10, upper = 30, init = 20),
    soilm_thetastar = list(lower = 10, upper = 200, init = 100),
    
    # WHC is exceptional: provided as site info to enable specification 
    # by site in multi-site simulations, but also calibratable.
    whc = list(lower = 10, upper = 1000, init = 400)
  )
)
```

`rsofun` supports both optimization using the `GenSA` and `BayesianTools` packages. The above statement provides settings for a `GenSA` optimization approach. For this example the maximum number of iterations is kept artificially low. In a real scenario you will have to increase this value orders of magnitude. Keep in mind that optimization routines rely on a cost function, which, depending on its structure influences parameter selection. A limited set of cost functions is provided but the model structure is transparent and custom cost functions can be easily written. More details can be found in the "Parameter calibration and cost functions" vignette.

In addition starting values and ranges are provided for the free parameters in the model. Free parameters include: parameters for the quantum yield efficiency `kphio`, `kphio_par_a` and `kphio_par_b`, soil moisture stress parameter `soilm_thetastar`, and also `beta_unitcostratio`, `rd_to_vcmax`, `tau_acclim` and `kc_jmax` (see `?runread_pmodel_f`). Be mindful that with newer versions of `rsofun` additional parameters might be introduced, so re-check vignettes and function documentation when updating existing code.

With all settings defined the optimization function `calib_sofun()` can be called with driver data and observations specified. Extra arguments for the cost function (like what variable should be used as target to compute the root mean squared error (RMSE) and previous values for the parameters that aren't calibrated, which are needed to run the P-model).

```{r eval=FALSE, warning=FALSE}
# calibrate the model and optimize free parameters
pars <- calib_sofun(
  
  drivers = p_model_drivers,  
  obs = p_model_validation,
  settings = settings,
  
  # extra arguments passed to the cost function:
  targets = "gpp",             # define target variable GPP
  
  # fix non-calibrated parameters to previous 
  par_fixed = list(
    beta_unitcostratio = 146.0,
    rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
    tau_acclim         = 30.0,
    kc_jmax            = 0.41
  )
)
```

When successful the optimized parameters can be used to run subsequent modelling efforts, in this case improving the model fit over the initial parameter set. The graph shows observed and modelled quartiles of daily GPP by day-of-year.

```{r}
# Update the parameter list with calibrated value
params_modl <- list(
  kphio = 0.039889979, #pars$mod$par[["kphio"]],
  kphio_par_a = -0.002281718, # pars$mod$par[["kphio_par_a"]],
  kphio_par_b = 15.064324945, # pars$mod$par[["kphio_par_b"]],
  soilm_thetastar = 67.247956095, # pars$mod$par[["soilm_thetastar"]], 
  beta_unitcostratio = 146.0,
  rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
  tau_acclim         = 30.0,
  kc_jmax            = 0.41
  )

# root zone total water holding capacity (whc) is calibrated, but specified as
# a site parameter in the P-model call with function runread_pmodel_f()
p_model_drivers$site_info[[1]]$whc <- 512  # approximate value, should re-calibrate this

# Run the model with calibrated parameters
output_new <- rsofun::runread_pmodel_f(
  drivers = p_model_drivers,
  par = params_modl
  )

# Update dataframe for plotting
df_plot <- output_new |>
  unnest(data) |>
  select(date, mod = gpp) |> 
  left_join(
    p_model_validation |>
      unnest(data) |> 
      select(date, obs = gpp),
    by = join_by(date)
  )

plot_eval(df_plot)
```

For details on the optimization settings we refer to the manuals of [GenSA](https://cran.r-project.org/package=GenSA) and [BayesianTools](https://github.com/florianhartig/BayesianTools).

## Water-carbon coupled simulations

### SPLASH AET

In the standard-setup, SPLASH is used for the water balance and ET. 

Simulations shown above are using the standard-setup. It is selected by setting `use_gs = FALSE` in `params_siml`. 

Here are the comparisons of modelled and observed ET.
```{r}
# Create dataframe for plotting
df_plot <- output |>
  unnest(data) |>
  mutate(le =  le / (24 * 60 * 60)) |> 
  select(date, mod = le) |> 
  left_join(
    p_model_validation |>
      unnest(data) |> 
      select(date, obs = le),
    by = join_by(date)
  )

plot_eval(df_plot, is_gpp = FALSE)
```

### Gs-coupled diffusion ET

Set `use_gs = TRUE` in `params_siml` so that the internally predicted stomatal conductance ($G_s$) from P-model will be used in calculation of ET. ET is a weighted average of canopy transpiration ($T$) and soil evaporation. Canopy transpiration is calculated using the diffusion equation as:
$$
T = 1.6 \; G_s \; \text{VPD}
$$

Run the model.

```{r}
drivers_tmp <- p_model_drivers |> 
  mutate(
    params_siml = purrr::map(
      params_siml,
      ~mutate(
        ., 
        use_gs = TRUE,
        use_pml = FALSE,
        use_phydro = FALSE
        )
      ))

# run the model with the same model parameters
output <- rsofun::runread_pmodel_f(
  drivers_tmp,
  par = params_modl
)
```

Plot outputs for LE. 

```{r}
# Create dataframe for plotting
df_plot <- output |>
  unnest(data) |>
  mutate(le =  le / (24 * 60 * 60)) |> 
  select(date, mod = le) |> 
  left_join(
    p_model_validation |>
      unnest(data) |> 
      select(date, obs = le),
    by = join_by(date)
  )

plot_eval(df_plot, is_gpp = FALSE)
```

### Penman-Monteith ET

Set `use_gs` flag to TRUE in params_siml so that the internally predicted stomatal conductance ($G_s$) from P-model will be used in calculation of ET. ET is a weighted average of canopy transpiration ($T$) and soil evaporation. Canopy transpiration is calculated using the diffusion equation as:
$$
\lambda E = \frac{s(R_n - G) + \rho c_p D_a G_\mathrm{ah}}{s + \gamma (1 + G_\mathrm{ah}/G_\mathrm{sw})}
$$

Run the model.

```{r}
drivers_tmp <- p_model_drivers |> 
  mutate(
    params_siml = purrr::map(
      params_siml,
      ~mutate(
        ., 
        use_gs = TRUE,
        use_pml = TRUE,
        use_phydro = FALSE
        )
      ))

# run the model with the same model parameters
output <- rsofun::runread_pmodel_f(
  drivers_tmp,
  par = params_modl
)
```

Plot outputs. 

```{r}
# Create dataframe for plotting
df_plot <- output |>
  unnest(data) |>
  mutate(le =  le / (24 * 60 * 60)) |> 
  select(date, mod = le) |> 
  left_join(
    p_model_validation |>
      unnest(data) |> 
      select(date, obs = le),
    by = join_by(date)
  )

plot_eval(df_plot, is_gpp = FALSE, return_gg1 = TRUE)
ggsave("~/rsofun/fig/et_fr-pue.pdf", height = 4, width = 6)
```

## P-hydro run with diffusion

(For comparison)

For P-hydro, we must use the 3-hr daily max forcing as the acclimation forcing. So let's rename it in the data.

```{r}
phdro_model_drivers <- rsofun::p_model_drivers_format2024_08 |> 
  rename(forcing_acclim = forcing_3hrmax)
```

For P-hydro runs, we always set `use_phydro = T` and `use_gs = T`.

```{r}
# define model parameter values
params_modl <- list(
  kphio              = 0.0288,
  kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
  kphio_par_b        = 1.0,
  rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
  tau_acclim         = 30.0,
  kc_jmax            = 0.41,
  phydro_K_plant     = 5e-17,
  phydro_p50_plant   = -0.46,
  phydro_gamma       = 0.065,
  phydro_b_plant     = 1,
  phydro_alpha       = 0.08,
  bsoil              = 3,
  Ssoil              = 113
)

# run the model for these parameters
output <- rsofun::runread_pmodel_f(
  mutate(phdro_model_drivers, 
         params_siml = params_siml |> mutate(use_gs = TRUE, use_phydro = TRUE) |> list()),
  par = params_modl
)
```


```{r}
# gpp
df_plot <- output$data[[1]] |> 
  select(date, mod = gpp) |> 
  left_join(
    rsofun::p_model_validation_format2024_08$data[[1]] |> 
      select(date, obs = gpp),
    by = join_by(date)
  ) |> 
  as_tibble()

plot_eval(df_plot)

# le
df_plot <- output$data[[1]] |> 
  mutate(le =  le / (24 * 60 * 60)) |> 
  select(date, mod = le) |> 
  left_join(
    rsofun::p_model_validation_format2024_08$data[[1]] |> 
      mutate(le =  le / (24 * 60 * 60)) |> 
      select(date, obs = le),
    by = join_by(date)
  ) |> 
  as_tibble()

plot_eval(df_plot, is_gpp = FALSE)
```
