---
title: "Numerical solution of the P model"
output: html_notebook
---

```{r}
library(rsofun)
```

Get the analytical solution for one specific set of arguments.
```{r}
## Run for different temperatures and plot
tc <- 20
ppfd <- 800
vpd  <- 1000
co2  <- 400
elv  <- 0
fapar <- 1

out_analytical <- rpmodel( tc = tc, vpd = vpd, co2 = co2, elv = elv, kphio = 0.05, fapar = fapar, ppfd = ppfd, method_optci="prentice14", method_jmaxlim = "none", do_ftemp_kphio = FALSE )
```

Is it internally consistent?
$$
c_i = c_a - A / g_s
$$

```{r}
print( out_analytical$ci )
print( out_analytical$ca - (out_analytical$gpp / 12.0107) / out_analytical$gs )
```
Yes. 

And what about ...
$$
A = V_{\text{cmax}} \frac{c_i-\Gamma^{\ast}}{c_i + K}
$$

```{r}
print( out_analytical$gpp / 12.0107 )
print( out_analytical$vcmax * (out_analytical$ci - out_analytical$gammastar) / (out_analytical$ci + out_analytical$kmm ))
print( out_analytical$gs * (out_analytical$ca - out_analytical$ci) )
```
Yes.

Maximise the following term numerically:
$$
A - aE -bV_{\text{cmax}} = max.
$$
 
Define the function which is to be maximised.
```{r}
source("R/calc_net_assim.R")
```

Use the analytically optimal vcmax and ci as starting values and calculate net assimilation with these
```{r}
## starting values for par = [vcmax, gs]. Use the ones returned by rpmodel()
par <- c( out_analytical$vcmax, out_analytical$gs )
args <- c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, 146.0 )

out_numerical <- calc_net_assim( par, args )
```

Now, change Vcmax just a little bit to see whether the net assimilation declines.
```{r}
par <- c( out_analytical$vcmax * 1.01, out_analytical$gs )
out_numerical <- calc_net_assim( par, args )
```

Display this in 3D
```{r}
x <- seq(-10, 10, length = 30)  ## already in increasing order
y <- x  ## already in increasing order
f <- function(x, y) {r <- sqrt(x ^ 2 + y ^ 2); 10 * sin(r) / r}
z <- outer(x, y, f)  ## evaluation on grid; obtain a matrix `z`
persp(x, y, z)
```


Optimise Vcmax and gs simultaneously so that net assimilation is maximised. This should return the same values for Vcmax and gs as the starting values (which were taken from the analytical solution).
```{r}
out_optim <- optimr(
  par     = c( out_analytical$vcmax, out_analytical$gs ),
  fn      = calc_net_assim,
  args    = c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, 146.0 ),
  control = list( maxit = 1000, maximize = TRUE )
  )
```

