---
title: "P-model usage"
author: "Koen Hufkens, Pepa Aran"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{P-model usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(rsofun)
library(dplyr)
library(ggplot2)

# fake variable as optimization isn't used
pars <- list()
pars$par["kphio"] <- 0.05
```

The `rsofun` package and framework includes two main models. The `pmodel` and `biomee` (which in part relies on P-model components). Here we give a short example on how to run the `pmodel` on the included demo datasets to familiarize yourself with both the data structure and the outputs.

## Demo data

The package includes two demo datasets to run and validate pmodel output using GPP observations. These files can be directly loaded into your workspace by typing:

```{r}
library(rsofun)

p_model_drivers

p_model_validation
```

These are real data from the French FR-Pue fluxnet site. Information about data structure, variable names, and their meaning and units can be found in the reference pages of `p_model_drivers` and `p_model_validation`. We can use these data to run the model, together with observations of GPP we can also calibrate `pmodel` parameters. 

Another two datasets are provided as an example to validate the model against leaf traits data, rather than fluxes. Measurements of Vcmax25 (aggregated over species) for 72 sites from Atkin et al. 2015 are given in `p_model_validation_vcmax25` and the corresponding forcing for the P-model is given in `p_model_drivers_vcmax25`. Since leaf traits are only measured once per site, the forcing used is a single year of average climate (the average measurements between 2001 and 2015 on each day of the year).

```{r}
p_model_drivers_vcmax25

p_model_validation_vcmax25
```

For the remainder of this vignette, we will use the GPP flux datasets. The workflow is exactly the same for leaf traits data. 

To get your raw data into the structure used within `rsofun`, please see R packages [ingestr](https://github.com/geco-bern/ingestr) and [FluxDataKit](https://github.com/geco-bern/FluxDataKit).

## Running model

With all data prepared we can run the model using `runread_pmodel_f()`. This function takes the nested data structure and runs the model site by site, returning nested model output results matching the input drivers. 

```{r}

# optimized parameters from previous
# work
params_modl <- list(
    kphio              = 0.09423773,
    soilm_par_a        = 0.33349283,
    soilm_par_b        = 1.45602286,
    beta_unitcostratio = 146.000000,
    rd_to_vcmax        = 0.01400000,
    tau_acclim         = 30.0,
    kc_jmax            = 0.41
  )

# run the model for these parameters
output <- rsofun::runread_pmodel_f(
  p_model_drivers,
  par = params_modl
  )
```

### Plotting output

We can now visualize both the model output and the measured values together.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# we only have one site so we'll unnest
# the main model output
model_data <- output %>%
  filter(sitename == "FR-Pue") %>%
  tidyr::unnest(data)

validation_data <- p_model_validation %>%
  tidyr::unnest(data)

ggplot() +
    geom_line(
    data = model_data,
    aes(
      date,
      gpp
    ),
    colour = "red",
    alpha = 0.8
  ) +
  geom_line(
    data = validation_data,
    aes(
      date,
      gpp
    ),
    alpha = 0.8
  ) +
  labs(
    x = "Date",
    y = "GPP"
  )
```

## Calibrating model parameters

To optimize new parameters based upon driver data and a validation dataset we must first specify an optimization strategy and settings, as well as a cost function and parameter ranges.

```{r}
settings <- list(
  method              = "GenSA",
  metric              = cost_rmse_pmodel,
  control = list(
    maxit = 10),
  par = list(
    kphio = list(lower=0.04, upper=0.2, init = 0.05)
    )
)
```

By default `rsofun` supports both optimization using the `GenSA` and `BayesianTools` packages. The above statement provides settings for a `GenSA` optimization approach. For this example the maximum number of iterations is kept artificially low. In a real scenario you will have to increase this value orders of magnitude. Keep in mind that optimization routines rely on a cost function, which, depending on its structure influences parameter selection. A limited set of cost functions is provided but the model structure is transparent and custom cost functions can be easily written. More details can be found in the "Parameter calibration and cost functions" vignette.

In addition starting values and ranges are provided for the free parameters in the model. Free parameters include, `kphio`, `soilm_par_a` and `soilm_par_b`. Be mindful that with newer versions of `rsofun` additional parameters might be introduced, so re-check vignettes and model descriptions when updating existing code.

With all settings defined the optimization function `calib_sofun()` can be called with driver data and observations specified. Extra arguments for the cost function (like which parameters should be free, what variable should be used as target to compute the root mean squared error (RMSE) and previous values for the parameters that aren't calibrated, which are needed to run the P-model).

```{r eval=FALSE}
# calibrate the model and optimize free parameters
pars <- calib_sofun(
    drivers = p_model_drivers,  
    obs = p_model_validation,
    settings = settings,
    # extra arguments passed to the cost function:
    setup = "BRC",           # calibrate only kphio
    targets = "gpp",         # define target variable GPP
    par_fixed = params_modl  # fix non-calibrated parameters to previous values
  )
```

When successful the optimized parameters can be used to run subsequent modelling efforts, in this case slightly improving the model fit over a more global parameter set.

```{r}
# reformatting the parameter list
params_modl$kphio <- pars$par["kphio"]

# run the model for these parameters
output_new <- rsofun::runread_pmodel_f(
  p_model_drivers,
  par = params_modl
  )

# we only have one site so we'll unnest
# the main model output
model_data_new <- output_new %>%
  filter(sitename == "FR-Pue") %>%
  tidyr::unnest(data)

  ggplot() +
    geom_line(
    data = model_data,
    aes(
      date,
      gpp
    ),
    colour = "red",
    alpha = 0.8
  ) +
  geom_line(
    data = validation_data,
    aes(
      date,
      gpp
    ),
    alpha = 0.8
  ) +
    geom_line(
    data = model_data_new,
    aes(
      date,
      gpp
    ),
    colour = "blue",
    alpha = 0.7
  ) +
  labs(
    x = "Date",
    y = "GPP"
  )
```

For details on the optimization settings we refer to the manuals of [GenSA](https://cran.r-project.org/package=GenSA) and [BayesianTools](https://github.com/florianhartig/BayesianTools).
