---
title: "Prepare rsofun forcing data"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(rsofun)
library(ingestr)
library(dplyr)
library(readr)
library(tidyr)
```

The following describes how to use the [ingestr R package ](https://stineb.github.io/ingestr/) for collecting rsofun forcing data. This is to create the object `df_drivers`, required as an input (forcing) to rsofun. This vignette is specific for collecting forcing data where no site-specific measurements are available, but all data is extracted from global maps. Head over to the vignette `prepare_fluxnetinputs_rsofun.Rmd` for an example for collecting forcing data for a FLUXNET site, using site-specific meteorological data.

## Site selection and meta data

A small number of meta data variables have to be specified for each site specifically to define the simulation years and site location. This information is also used for input, calibration, and evaluation data ingestion. Required meta information is specified as a data frame with sites by rows and the following variables (by columns):

- `sitename`: any unique identifying of the site(s)
- `lat` for latitude (decimal degrees)
- `lon` for longitude (decimal degrees) - this is only used for data ingestion but not for the P-model simulation with `rsofun`.
- `elv` for elevation (m a.s.l.). If this is not known, *ingestr* may be used to extract information from a digital elevation model.
- `year_start` and `year_end` specifying years covered by the simulation. Data extraction and model simulations always cover all 365 days of each year.
- `whc` for the total rooting zone soil water holding capacity. This is not a temporary "fudge" solution. Set it to 170 mm if no better information is available.
- `koeppen_code` to group sites for evaluation by Koeppen-Geiger climate zones (may be ignored, only required for `eval_sofun()`).

The following is an example (`elv` is not specified) for a single-site run:
```{r}
siteinfo <- data.frame(
  sitename = "anywhere",
  lon = 100,
  lat = 50,
  year_start = 2003,
  year_end = 2005
)
```

## Simulation settings

Specify additional simulation parameters that are identical for all site-scale simulations in multi-site runs, 
```{r}
params_siml <- list(
  spinup             = TRUE,      # to bring soil moisture to steady state
  spinupyears        = 10,        # number of spinup years. 10 is enough for soil moisture.
  recycle            = 1,         # number of years recycled during spinup 
  soilmstress        = FALSE,     # boolean for whether soil moisture stress function is included
  tempstress         = FALSE,     # boolean for whether temperature stress function is included
  calc_aet_fapar_vpd = FALSE,     # set to FALSE - should be dropped again
  in_ppfd            = TRUE,      # if available from forcing files, set to TRUE
  in_netrad          = FALSE     # if available from forcing files, set to TRUE
	)
```

<!-- Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters), global simulation parameters are wrapped inside an additional column `params_siml`, added to the site meta info dataframe. -->
<!-- ```{r} -->
<!-- siteinfo <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml) -->
<!-- ``` -->

## Define model parameters

Specify model parameters based on prior model calibration. See `benchmarking/tag_v*/benchmark_gpp_FLUXNET2015_ensemble.html` for latest calibrations.
```{r}
params_modl <- list(
	kphio           = 0.09423773,
	soilm_par_a     = 0.33349283,
	soilm_par_b     = 1.45602286
	)
```

## Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific. Values entered here take no effect.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

## Get input

Input data, used as model forcing, is collected using the [ingestr](https://stineb.github.io/ingestr/) package. A brief description for how to use it for our present application is provided here. Data is collected by data source. 

## Meteo data

This extracts from original WATCH-WFDEI files, provided as NetCDF (global, 0.5 degree resolution), provided as monthly files containing all days in each month. The data directory specified here (dir = `"~/data/watch_wfdei/"`) contains sub-directories with names containing the variable names (corresponding to the ones specified by the argument `getvars = list(temp = "Tair"))`. 

Ingest can be used for a single site as follows:
```{r message=FALSE, warning=FALSE, eval=FALSE}
df_watch <- ingest_bysite(
  sitename  = "anywhere",  # can be any name
  source    = "watch_wfdei",
  getvars   = list(temp = "Tair"),  # additional variables required for rsofun: prec = "Rainf", vpd = "Qair", ppfd = "SWin"
  dir       = "~/data/watch_wfdei/",
  timescale = "d",
  year_start = 2001,
  year_end  = 2003,
  lon       = siteinfo$lon[1],
  lat       = siteinfo$lat[1],
  verbose   = FALSE
  )
```

For multi-site simulations with rsofun, it's preferable to use ingest for multiple sites at once, using the site meta information specified in `siteinfo`. Here, data object `df_watch` is organised as a nested table with rows for sites and time series nested inside the column `data`. See [here](https://tidyr.tidyverse.org/reference/nest.html) for how to handle nested dataframes. 
```{r message=FALSE, warning=FALSE}
df_watch <- ingest(
  siteinfo  = siteinfo,
  source    = "watch_wfdei",
  getvars   = list(temp = "Tair", prec = "Rainf", vpd = "Qair", ppfd = "SWin"),  # additional variables: 
  dir       = "~/data/watch_wfdei/"  # adjust this with your local path
  )
```


Some meteo data is not available from FLUXNET. Extract it from CRU global climate files instead.
```{r}
ddf_cru <- ingestr::ingest(
  siteinfo  = siteinfo,
  source    = "cru",
  getvars   = list(ccov = "cld"),
  dir       = "~/data/cru/ts_4.01/"
  )
```

Combine the two meteo data frames into one, containing `ccov` (cloud cover) from CRU and all other variables from FLUXNET.
```{r}
ddf_meteo <- df_watch %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru %>% 
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>% 
  group_by(sitename) %>% 
  tidyr::nest()
```

## fAPAR data

fAPAR data is prescribed in the P-model setup. The following extracts data MODIS FPAR data from the MODIS LP DAAC  (see a complete description [here](https://stineb.github.io/ingestr/articles/example.html#modis-lp-daac-1)). 

The following example is for downloading MODIS collection 6, MCD15A3H, band `Fpar_500m` data.
```{r}
settings_modis <- get_settings_modis(
  bundle            = "modis_fpar",
  data_path         = "~/data/modis_subsets/",
  method_interpol   = "loess",
  keep              = TRUE,
  overwrite_raw     = FALSE,
  overwrite_interpol= TRUE
  )
```

This is not yet implemented for multiple sites (using `ingest()`). Use the single-site ingest (`ingest_bysite()`) instead.
```{r}
df_modis_fpar <- ingest_bysite(
  sitename  = "anywhere",  # can be any name
  source    = "modis",
  year_start = 2003,
  year_end  = 2005,
  lon       = siteinfo$lon[1],
  lat       = siteinfo$lat[1],
  settings  = settings_modis,
  verbose   = FALSE
  )
```

Plot this for a quick check using the function `plot_fapar_ingestr_bysite()` from the ingestr package. **WARNING: FOR MODIS_FPAR, NO FILTERING IMPLEMENTED YET.**
```{r}
library(ggplot2)
plot_fapar_ingestr_bysite(
  df_modis_fpar,
  settings_modis)
```

To make this output consistent for use below, let's nest it and rename the desired column to `fapar` (this is a bit weird, sorry)
```{r}
df_modis_fpar <- df_modis_fpar %>% 
  rename(fapar = modisvar_filled) %>%
  group_by(sitename) %>% 
  nest()
```


## CO2

Ingesting CO2 data is particularly simple. We can safely assume it's well mixed in the atmosphere (independent of site location), and we can use a annual mean value for all days in respective years.  
```{r}
df_co2 <- ingestr::ingest(
  siteinfo,
  source  = "co2_mlo",
  verbose = FALSE
  )
```

## Collect all drivers

Finally, we can collect forcing data, simulation parameters, and site meta info into a single object that will be used to drive rsofun. All of the above steps can be customized. The function `collect_drivers_sofun()` can in general be used to process forcing data into the format required to run SOFUN. The arguments must have the following form:

- `siteinfo`: A data frame (tibble) with columns `sitename`, `lon`, `lat`, `elv` (elevation), `year_start`, `year_end`, `whc` (water holding capacity used for simulating the soil water balance). We have created this above.
```{r} 
siteinfo
```
- `meteo`: A nested data frame with columns `sitename` and `data`. The latter contains the nested meteorological forcing data frames, with columns `date`, `temp`, `prec`, `vpd`, `ppfd`, `patm`, and `ccov`. Like this:
```{r}
## nested data frame:
ddf_meteo

## inside .$data:
ddf_meteo$data[[1]] %>% dplyr::select(-ccov_int)
```


- `fapar`: A nested data frame with columns `sitename` and `data`. The latter contains the nested meteorological forcing data frames, with columns `date`, and `fapar`. Like this:
```{r}
## nested data frame:
df_modis_fpar

## inside .$data:
df_modis_fpar$data[[1]] %>% dplyr::select(date, fapar)
```
- `co2`  : A nested data frame with columns `sitename` and `data`. The latter contains the nested meteorological forcing data frames, with columns `date`, and `co2`. Like this:
```{r}
## nested data frame:
df_co2

## inside .$data:
df_co2$data[[1]] %>% dplyr::select(date, co2)
```
- `df_soiltexture`: See above ('Define soil parameters')

See [here](https://tidyr.tidyverse.org/reference/nest.html) for how to handle nested dataframes. 

Finally, all input data can be collected by:
```{r}
df_drivers <- collect_drivers_sofun( 
  siteinfo       = siteinfo,
  params_siml    = params_siml,
  meteo          = ddf_meteo, 
  fapar          = df_modis_fpar,
  co2            = df_co2,
  df_soiltexture = df_soiltexture
  )
df_drivers
```

