---
title: "Prepare forcing"
author: "Beni Stocker"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prepare forcing}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Libraries, check for Euler access ----
library(dplyr)
library(ggplot2)
library(rsofun)

p_model_fluxnet_drivers <- rsofun::p_model_drivers
```

Forcing data of `rsofun` adheres to a fixed format as described [here]. Any data
in this format will be valid `rsofun` input data. However to facilitate the
formatting of these datasets you may use the internal lab tool `ingestr`. This
toolset allows for the easy and consistent formatting of data.

However, this toolset is very dynamic in nature and therefore will not see a
formal CRAN release. Within this context we will also give very limited support.
Despite these caveats you might find this tool useful. The below routine shows
you how to format data using `ingestr`, assuming all 
[source data](https://github.com/geco-bern/ingestr) has been 
downloaded beforehand.

# Preparing driver data

To format data we need to load the required libraries first.

```{r eval = FALSE}
# Libraries
library(tidyverse)
library(rsofun)
library(ingestr)
```

To limit the computational times we'll use a subset of three
fluxnet site locations. We can also use arbitrary locations if desired.

```{r eval = FALSE}
fluxnet_sites <- ingestr::siteinfo_fluxnet2015 %>%
  dplyr::filter(sitename == "FR-Pue")
```
Using a data frame of site info we can subset and format fluxnet data into the
desired shape (assuming a local copy of the FLUXNET2015 dataset for the 
desired sites). We refer to the 
[FLUXNET2015 data website](https://fluxnet.org/data/fluxnet2015-dataset/)
for data downloads.

```{r eval = FALSE}
# grab fluxnet data ----
df_fluxnet <-
  suppressWarnings(
    suppressMessages(
      ingestr::ingest(
        siteinfo  = fluxnet_sites,
        source    = "fluxnet",
        getvars   = list(
          temp = "TA_F_DAY",
          prec = "P_F",
          vpd  = "VPD_F_DAY",
          ppfd = "SW_IN_F",
          patm = "PA_F"),
        dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
        settings  = list(
          dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", getswc = FALSE),
        timescale = "d"
      )
    )
  )
```

Not all required data is available within the fluxnet dataset. As such we amend
these data with additional data sources. In this case we use CRU data to 
complement the fluxnet data, and merged into a single 'meteo' data frame.

```{r eval = FALSE}
df_cru <- ingestr::ingest(
  siteinfo  = fluxnet_sites,
  source    = "cru",
  getvars   = "ccov",
  dir       = "~/data/cru/ts_4.01/"
  )

df_meteo <- df_fluxnet %>%
  tidyr::unnest(data) %>%
  left_join(
    df_cru %>%
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>%
  group_by(sitename) %>%
  tidyr::nest()
```

Additional fAPAR data is gathered from remote sensing products. We first specify
download settings, and initiate a download for our site locations.

```{r eval = FALSE}
settings_modis <- get_settings_modis(
  bundle            = "modis_fpar",
  data_path         = "~/data/modis_subsets/",
  method_interpol   = "loess",
  network = c("fluxnet","icos"),
  keep              = TRUE,
  overwrite_raw     = FALSE,
  overwrite_interpol= TRUE,
  n_focal           = 0
  )

df_modis_fpar <- ingest(
  fluxnet_sites,
  source = "modis",
  settings = settings_modis,
  parallel = FALSE,
  ncores = 1
  )

## renaming the variable
df_modis_fpar <- df_modis_fpar %>%
  mutate(
    data = purrr::map(data, ~rename(., fapar = modisvar_filled))
    )
```

We also download matching CO2 data for the dates covering the fluxnet sites.
This covers all data downloads!

```{r eval = FALSE}
# . grab CO2 data ----
df_co2 <- ingestr::ingest(
  fluxnet_sites,
  source  = "co2_mlo",
  verbose = FALSE
  )
```

A final set of required ancillary data covers the soil texture properties.

```{r eval = FALSE}
# . set soil parameters ----
df_soiltexture <- bind_rows(
  top    = tibble(
    layer = "top",
    fsand = 0.4,
    fclay = 0.3,
    forg = 0.1,
    fgravel = 0.1
    ),
  bottom = tibble(
    layer = "bottom",
    fsand = 0.4,
    fclay = 0.3,
    forg = 0.1,
    fgravel = 0.1)
)
```

## Simulation parameters

With all data downloads done we have to set some simulation parameters which
will be merged with all driver data. This data includes the number of spin-up
years, and how soil and temperature stress are dealt with.

```{r eval = FALSE}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)
```

## Merging data streams

With all pieces in place we can use the `rsofun` collect_drivers_sofun()
function to merge all datasets into one nested dataframe, which serves as
input to the `rsofun` p-model. This concludes the formatting of the driver data

```{r eval = FALSE}
p_model_fluxnet_drivers <- rsofun::collect_drivers_sofun(
  site_info      = fluxnet_sites,
  params_siml    = params_siml,
  meteo          = df_meteo,
  fapar          = df_modis_fpar,
  co2            = df_co2,
  params_soil    = df_soiltexture
  )
```

# Running a simulation

With the above formatted data we can now run the model, provided some additional
model parameters. For a model calibration routine see 
[this vignette](https://geco-bern.github.io/rsofun/articles/pmodel_use.html).

```{r}
# run the model for these parameters
# optimized parameters from previous work (Stocker et al., 2020 GMD)
params_modl <- list(
    kphio           = 0.09423773,
    soilm_par_a     = 0.33349283,
    soilm_par_b     = 1.45602286,
    tau_acclim_tempstress = 10,
    par_shape_tempstress  = 0.0
  )

output <- rsofun::runread_pmodel_f(
  p_model_fluxnet_drivers,
  par = params_modl
  )

output$data[[1]] %>% 
  ggplot(aes(date, gpp)) +
  geom_line()
```
