---
title: "Calibration of rsofun with Phydro"
author: "Jaideep Joshi"
date: "2023-09-07"
output: html_document
params: 
  output_dir: "."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(reshape2)
library(rsofun)
library(BayesianTools)
```

## Generate data if not already available

```{r}
site <- "GF-Guy"
```

```{r}
data_dir = "~/Downloads/fluxdatakit_oct3/Phydro_drivers/"
load(paste0(data_dir,"/",site,"_p_hydro_drivers.rda"))
load(paste0(data_dir,"/",site,"_p_hydro_validation.rda"))

p_hydro_validation$data[[1]] = p_hydro_validation$data[[1]] %>%
  rename(le=latenth) 
#   mutate(gpp = gpp*86400/1e6*12) %>% # convert to [gC m-2 day-1]
#   mutate(latenth = latenth*86400)    # convert [W m-2] to [J m-2 day-1]

p_hydro_drivers$forcing[[1]]$netrad[is.na(p_hydro_drivers$forcing[[1]]$netrad)] = 0
p_hydro_drivers$forcing[[1]]$patm[p_hydro_drivers$forcing[[1]]$patm < 0] = 1.0135e5

p_hydro_drivers$params_siml[[1]]$use_phydro = T
p_hydro_drivers$params_siml[[1]]$use_pml = F

```

Function to quickly run model and plot outputs

```{r}
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}

plot_pmodel = function(params_m, out_filename=""){
  output_p <- rsofun::runread_pmodel_f(
    p_hydro_drivers,
    par = params_m
  )

  if (out_filename != ""){
    save(output_p, file = paste0(data_dir, "/", site, "_", out_filename))
  }
  
  p1 = output_p$data[[1]] %>% select(date, gpp, le) %>% 
    melt("date") %>% 
    mutate(group="model") %>% 
    rbind(p_hydro_validation$data[[1]] %>% 
            select(date, gpp, le) %>%
            # mutate(gpp = gpp*86400/1e6*12) %>% # convert to gC m-2 day-1
            # mutate(latenth = latenth*86400) %>%  # convert W m-2 to J m-2 day-1 
            melt("date") %>% 
            mutate(group="obs")) %>% 
    ggplot(aes(y=value, x=as.Date(date))) +
    geom_line(aes(group=group, col=group), alpha=0.7) +
    theme_classic() +
    theme(strip.background = element_rect(color = "white", size = 1))+
    facet_wrap(~variable, scales = "free", nrow = 2)
  
  p2 = output_p$data[[1]] %>% select(date, gpp, le) %>% 
    drop_na %>% 
    melt("date", value.name = "pred") %>% 
    left_join(p_hydro_validation$data[[1]] %>% 
            select(date, gpp, le) %>%
            drop_na %>% 
            # mutate(gpp = gpp*86400/1e6*12) %>% # convert to gC m-2 day-1
            # mutate(le = le*86400) %>%  # convert W m-2 to J m-2 day-1 
            melt("date", value.name = "obs")) %>% 
    group_by(variable) %>% 
      reframe(obs = obs, pred = pred, density=scale(get_density(pred, obs,n=100))) %>%  
    ggplot(aes(y=obs, x=pred, colour=density)) +
    scale_color_viridis_c() +
    geom_point(alpha=0.7) +
    geom_abline(slope=1, intercept=0, col="red")+
    theme_classic() +
    theme(strip.background = element_rect(color = "white", size = 1))+
    facet_wrap(~variable, scales = "free", nrow = 1)+
    labs(colour="Density")
  
  print(p1)
  print(p2)
}

```

```{r}
params_modl <- list(
  kphio              = 7.013438e-02, # 0.089, # 0.11, #0.04998,    
  kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
  kphio_par_b        = 1.0,
  # soilm_thetastar    = 0.6 * 240,  # to recover old setup with soil moisture stress
  # soilm_betao        = 0.0,
  # beta_unitcostratio = 146.0,
  rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
  tau_acclim         = 30.0,
  kc_jmax            = 0.41,
  phydro_K_plant     = 1.421218e-16 , #0.3e-16,
  phydro_p50_plant   = -2.506519e+00, #-1,
  phydro_b_plant     = 1,
  phydro_alpha       = 1.334564e-01, #0.1,
  phydro_gamma       = 1.889438e+00, #1,
  bsoil              = 3.193891e+00, #3,
  whc                = 7.819509e+02 #200
)

plot_pmodel(params_modl)
output_p <- rsofun::runread_pmodel_f(
    p_hydro_drivers,
    par = params_modl
  )
```

## P-hydro full calibration (BayesianTools)

```{r}
p_hydro_drivers$params_siml[[1]]$use_phydro = T
p_hydro_drivers$params_siml[[1]]$use_pml = F

# Define calibration settings and parameter ranges from previous work
settings_bayes <- list(
  method = "BayesianTools",
  par = list(
    kphio = list(lower=0.04, upper=0.09, init=0.05),
    phydro_K_plant = list(lower=0.05e-16, upper=5e-16, init=0.3e-16),
    phydro_p50_plant = list(lower=-4, upper=-0.1, init=-1),
    phydro_alpha = list(lower=0.06, upper=0.15, init=0.1),
    phydro_gamma = list(lower=0.05, upper=2, init=1),
    bsoil = list(lower=1, upper=5, init=3),
    whc = list(lower=10, upper=1000, init=500),
    err_gpp = list(lower = 0.01, upper = 4, init = 2),
    err_le = list(lower = 0.01e7, upper = 4e7, init = 2e7)
  ),
  metric = rsofun::cost_likelihood_pmodel,
  control = list(
    sampler = "DEzs",
    settings = list(
      nrChains = 10,
      burnin = 5000,        
      iterations = 5000     # kept artificially low
    )
  )
)

# Calibrate the model and optimize the free parameters using
# demo datasets
pars_calib_bayes <- calib_sofun(
  # calib_sofun arguments:
  drivers = p_hydro_drivers,  
  obs = p_hydro_validation,
  settings = settings_bayes,
  # extra arguments passed to the cost function:
  par_fixed = list(         # fix all other parameters
    kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
    kphio_par_b        = 1.0,
    rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
    tau_acclim         = 30.0,
    kc_jmax            = 0.41,
    # phydro_K_plant     = 0.3e-16,
    # phydro_P50_plant     = -1,
    phydro_b_plant     = 1
    # phydro_alpha       = 0.1,
    # phydro_gamma       = 1
    # bsoil              = 3,
    # whc                = 90  
  ),
  targets = c("gpp", "le")           # define target variable GPP
)

plot(pars_calib_bayes$mod)
correlationPlot(pars_calib_bayes$mod)

print(pars_calib_bayes)

params_modl_opt = params_modl

params_modl_opt$kphio            = pars_calib_bayes$par[["kphio"]]
params_modl_opt$phydro_K_plant   = pars_calib_bayes$par[["phydro_K_plant"]]
params_modl_opt$phydro_p50_plant = pars_calib_bayes$par[["phydro_p50_plant"]]
params_modl_opt$phydro_alpha     = pars_calib_bayes$par[["phydro_alpha"]]
params_modl_opt$phydro_gamma     = pars_calib_bayes$par[["phydro_gamma"]]
params_modl_opt$bsoil            = pars_calib_bayes$par[["bsoil"]]
params_modl_opt$whc              = pars_calib_bayes$par[["whc"]]

plot_pmodel(params_modl_opt, "p_hydro_output.rda")

```


## P-model full calibration (BayesianTools)

```{r}
params_modl <- list(
  kphio              = 0.04998,    
  kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
  kphio_par_b        = 1.0,
  soilm_thetastar    = 0.6 * 240,  # to recover old setup with soil moisture stress
  soilm_betao        = 0.0,
  beta_unitcostratio = 146.0,
  rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
  tau_acclim         = 30.0,
  kc_jmax            = 0.41,
  # phydro_K_plant     = 0.3e-16,
  # phydro_p50_plant   = -1,
  # phydro_b_plant     = 1,
  # phydro_alpha       = 0.1,
  # phydro_gamma       = 1,
  # bsoil              = 3,
  whc                = 300
)
p_hydro_drivers$params_siml[[1]]$use_phydro = F
p_hydro_drivers$params_siml[[1]]$use_pml = F

plot_pmodel(params_modl)
```

```{r}
# Define calibration settings and parameter ranges from previous work
settings_bayes <- list(
  method = "BayesianTools",
  par = list(
    kphio = list(lower=0.04, upper=0.09, init=0.05),
    soilm_thetastar    = list(lower=0.1*240, upper=2*240, init=0.6*240),
    # soilm_betao        = list(lower=0, upper=10, init=0.1),
    beta_unitcostratio = list(lower=10, upper=1000, init=146),
    whc = list(lower=10, upper=1000, init=350),
    err_gpp = list(lower = 0.01, upper = 4, init = 2),
    err_le = list(lower = 0.01e7, upper = 4e7, init = 2e7)
  ),
  metric = rsofun::cost_likelihood_pmodel,
  control = list(
    sampler = "DEzs",
    settings = list(
      nrChains = 1,
      burnin = 1000,        
      iterations = 1300     # kept artificially low
    )
  )
)

# Calibrate the model and optimize the free parameters using
# demo datasets
pars_calib_bayes <- calib_sofun(
  # calib_sofun arguments:
  drivers = p_hydro_drivers,  
  obs = p_hydro_validation,
  settings = settings_bayes,
  # extra arguments passed to the cost function:
  par_fixed = list(         # fix all other parameters
    kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence 
    # of kphio, setup ORG
    kphio_par_b        = 1.0,
    soilm_betao        = 0.0,
    rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
    # beta_unitcostratio = 146,
    tau_acclim         = 30.0,
    kc_jmax            = 0.41
  ),
  targets = c("gpp", "le")           # define target variable GPP
)

print(pars_calib_bayes)

params_modl_opt = params_modl

params_modl_opt$kphio            = pars_calib_bayes$par[["kphio"]]
params_modl_opt$soilm_thetastar  = pars_calib_bayes$par[["soilm_thetastar"]]
# params_modl_opt$soilm_betao      = pars_calib_bayes$par[["soilm_betao"]]
params_modl_opt$beta_unitcostratio = pars_calib_bayes$par[["beta_unitcostratio"]]
params_modl_opt$whc              = pars_calib_bayes$par[["whc"]]

plot_pmodel(params_modl_opt)
