---
title: "Sensitivity analysis and calibration interpretation"
author: "Josefa Arán"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sensitivity analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)

library(rsofun)
library(dplyr)
library(ggplot2)
library(tidyr)
library(sensitivity)
library(BayesianTools)
```

Parameter calibration can have a big impact on our modeling effort and use
big computational resources. Hence, it is worth our time to explore which parameters
should actually be calibrated (the ones that impact the simulations greatly) 
and to examine if the calibration routines behave as expected. 
This vignette explains how to perform a simple
parameter sensitivity analysis for the P-model and how to interpret the outputs 
of the calibration using the `BayesianTools` package.

## Morris sensitivity analysis

The Morris method for global sensitivity analysis allows to explore which parameters have the
biggest influence on the model fit. In this example, we will quantify how different
values of the calibratable model parameters lead to more variability
in the match between GPP predicted by the P-model and GPP observations. It would
be wise to repeat this exercise for various targets because they may be simulated
by equations in the P-model that involve different model parameters.

If the P-model
has very low sensitivity to a certain parameter, calibrating it will
not improve the model substantially. But if it's very sensitive to another parameter,
calibrating this second parameter could improve the P-model fit greatly. We should
spend our computational resources on calibrating the parameters to which the
model is most sensitive.

First of all, let's define a function which measures the agreement between
GPP predictions from the P-model and GPP observations, for a set of values of
the calibratable parameters. It computes the normal log-likelihood of 
the GPP predictions, given the observed GPP and its uncertainty.
We want to see how sensitive this function is to changes in the parameter values.

```{r}
# Define log-likelihood function
ll_pmodel <- function(
    par_v                 # a vector of all calibratable parameters including errors
){
  rsofun::cost_likelihood_pmodel(        # reuse likelihood cost function
    par_v,
    obs = rsofun::p_model_validation,
    drivers = rsofun::p_model_drivers,
    targets = "gpp"
  )
}

# Compute log-likelihood for a given set of parameters
ll_pmodel( par_v = c(
  kphio              = 0.09423773, # setup ORG in Stocker et al. 2020 GMD
  kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
  kphio_par_b        = 1.0,
  soilm_thetastar    = 0.6 * 240,  # to recover old setup with soil moisture stress
  soilm_betao        = 0.0,
  beta_unitcostratio = 146.0,
  rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
  tau_acclim         = 30.0,
  kc_jmax            = 0.41,
  error_gpp          = 0.9         # value from previous simulations
))
```

Some parameters are constrained by their physical interpretation (e.g. `kphio > 0`)
and it's also necessary to provide a bounded parameter space for Morris'
method to sample the parameter space. We define the parameter space by their 
lower and upper bounds.
```{r}
# best parameter values (from previous literature)
par_cal_best <- c(
    kphio              = 0.09423773,
    kphio_par_a        = -0.0025,
    kphio_par_b        = 20,
    soilm_thetastar    = 0.6*240,
    soilm_betao        = 0.2,
    beta_unitcostratio = 146.0,
    rd_to_vcmax        = 0.014,
    tau_acclim         = 30.0,
    kc_jmax            = 0.41,
    error_gpp          = 1
  )

# lower bound
par_cal_min <- c(
    kphio              = 0.03,
    kphio_par_a        = -0.004,
    kphio_par_b        = 10,
    soilm_thetastar    = 0,
    soilm_betao        = 0,
    beta_unitcostratio = 50.0,
    rd_to_vcmax        = 0.01,
    tau_acclim         = 7.0,
    kc_jmax            = 0.2,
    error_gpp          = 0.01
  )

# upper bound
par_cal_max <- c(
    kphio              = 0.15,
    kphio_par_a        = -0.001,
    kphio_par_b        = 30,
    soilm_thetastar    = 240,
    soilm_betao        = 1,
    beta_unitcostratio = 200.0,
    rd_to_vcmax        = 0.1,
    tau_acclim         = 60.0,
    kc_jmax            = 0.8,
    error_gpp          = 4
  )
```

We use the `morris()` function from the `{sensitivity}` package to perform
the sensitivity analysis. As a target function, we will use the posterior
density (log-likelihood) of the parameters given the GPP data which we obtain
via the function `BayesianTools::createBayesianSetup()`. Note that, because of using
a uniform prior, the posterior distribution is proportional to the GPP 
log-likelihood (defined previously) 
wherever the parameter values are feasible and zero outside of the parameter ranges.
```{r eval = FALSE}
morris_setup <- BayesianTools::createBayesianSetup(
  likelihood = ll_pmodel,
  prior = BayesianTools::createUniformPrior(par_cal_min, par_cal_max, par_cal_best),
  names = names(par_cal_best)
)
```

In the following chunk, we run the Morris sensitivity analysis, using a grid with
`r=1000` values for each parameter and a one-at-a-time design. Running the
sensitivity analysis may take a few minutes, even for this small example dataset,
and is still computationally cheaper than running the parameter calibration.
```{r eval = FALSE, echo = TRUE}
set.seed(432)
morrisOut <- sensitivity::morris(
  model = morris_setup$posterior$density,
  factors = names(par_cal_best), 
  r = 1000, 
  design = list(type = "oat", levels = 20, grid.jump = 3), 
  binf = par_cal_min, 
  bsup = par_cal_max, 
  scale = TRUE)
```

```{r eval = FALSE, echo = FALSE}
# Save Morris sensitivity output because it takes very long to compute
save(morrisOut, file = "files/morrisOut.rda")
```

```{r eval = TRUE, echo = FALSE}
# Load Morris sensitivity output
load("files/morrisOut.rda")
```

The analysis evaluates the variability of the target function, i.e. the 
log-likelihood, for several points across the parameter space. It is an approximation
of the derivatives of the log-likelihood with respect to the model parameters.
Statistics $\mu *$ and $\sigma$ can be interpreted as the mean absolute derivative
and the standard deviation of the derivative,
respectively. The higher the value of these statistics for a given parameter, 
the more influential the parameter is.  
```{r eval = TRUE, fig.width=7, fig.height=4}
# Summarise the morris output
morrisOut.df <- data.frame(
  parameter = names(par_cal_best),
  mu.star = apply(abs(morrisOut$ee), 2, mean, na.rm = T),
  sigma = apply(morrisOut$ee, 2, sd, na.rm = T)
) %>%
  arrange( mu.star )

morrisOut.df |>
  tidyr::pivot_longer( -parameter, names_to = "variable", values_to = "value") |>
  ggplot(aes(
    reorder(parameter, value),
    value, 
    fill = variable),
    color = NA) +
  geom_bar(position = position_dodge(), stat = 'identity') +
  scale_fill_brewer("", labels = c('mu.star' = expression(mu * "*"),
                                   'sigma' = expression(sigma))) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 6),
    axis.title = element_blank(),
    legend.position = c(0.05, 0.95), legend.justification = c(0.05, 0.95)
  )

```

The outcome of the Morris sensitivity analysis depends strongly on the choice of
parameter ranges and how parameters interact with each other in the 
underlying model. In this example, we constrained the parameters based on 
their physical meaning (e.g. `soilm_betao` should be in `[0,1]`) and the site FR-Pue 
where the data is coming from (e.g. `kphio_par_b` around 25$^{o}$C).
When observing the figure above, we notice that parameters `kphio` and 
`kc_jmax` have a high impact on the model fit (big $\mu *$), but also the magnitude of this
dependence of GPP on the two parameters changes across the parameter space 
(big $\sigma$). This happens because parameters interact in the 
light use efficiency calculation
<!-- temperature dependence function for the quantum yield efficiency,  -->
and the calibration may be harder and require data from several sites.

The log-likelihood is most sensitive to `err_gpp` with a very large variation in
the magnitude of this dependence. This makes sense because for
higher values of the standard deviation the normal likelihood is flatter 
(and similar log-likelihood values are calculated, whether the model predictions
using the rest of parameters are good or bad) and for lower `err_gpp` values the
likelihood is pointy (hence good model fits achieve a big log-likelihood value and
and poor model fits, a very small value).

To help you interpret this sensitivity analysis and better understand the
parameter-model relationship, it may be wise to run it
several times for different parameter ranges and validation data. Note how `rd_to_vcmax`
does not affect GPP, but it would actually affect Vcmax25 predictions,
so trait data could also be added for validation.

## Interpretation of Bayesian calibration routine

It is always important to check the convergence of the MCMC algorithm used for the Bayesian calibration. Here we show some plots and statistics that may help you assess whether the parameter calibration has converged.

According to the previous sensitivity analysis, calibrating the error parameter for GPP and the quantum yield efficiency parameters will have a high impact on the model fit. Let's run the calibration:
```{r eval = FALSE, echo = TRUE}
set.seed(2023)

# Define calibration settings
settings_calib <- list(
  method = "BayesianTools",
  metric = rsofun::cost_likelihood_pmodel,
  control = list(
    sampler = "DEzs",
    settings = list(
      burnin = 3000,
      iterations = 9000,
      startValue = 3       # number of chains to be sampled
    )),
  par = list(
    kphio = list(lower = 0.03, upper = 0.15, init = 0.05),
    kphio_par_a = list(lower = -0.004, upper = -0.001, init = -0.0025),
    kphio_par_b = list(lower = 10, upper = 30, init =25),
    err_gpp = list(lower = 0.1, upper = 3, init = 0.8)
  )
)

# Calibrate kphio-related parameters and err_gpp 
par_calib <- calib_sofun(
  drivers = p_model_drivers,
  obs = p_model_validation,
  settings = settings_calib,
  par_fixed = list(
    soilm_thetastar    = 0.6*240,
    soilm_betao        = 0.2,
    beta_unitcostratio = 146.0,
    rd_to_vcmax        = 0.014,
    tau_acclim         = 30.0,
    kc_jmax            = 0.41),
  targets = "gpp"
)
```

```{r eval = FALSE, echo = FALSE}
# Calibrates kphio, betao, kc_jmax - top 3 model params
set.seed(2023)

# Define calibration settings
settings_calib <- list(
  method = "BayesianTools",
  metric = rsofun::cost_likelihood_pmodel,
  control = list(
    sampler = "DEzs",
    settings = list(
      burnin = 1500,
      iterations = 6000,
      startValue = 3       # number of chains to be sampled
    )),
  par = list(
    kphio = list(lower = 0.03, upper = 0.15, init = 0.05),
    soilm_betao = list(lower = 0, upper = 1, init = 0.2),
    kc_jmax = list(lower = 0.2, upper = 0.8, init = 0.41),
    err_gpp = list(lower = 0.1, upper = 3, init = 0.8)
  )
)

par_calib <- calib_sofun(
  drivers = p_model_drivers,
  obs = p_model_validation,
  settings = settings_calib,
  par_fixed = list(
    kphio_par_a = -0.0025,
    kphio_par_b = 20,
    soilm_thetastar    = 0.6*240,
    beta_unitcostratio = 146.0,
    rd_to_vcmax        = 0.014,
    tau_acclim         = 30.0),
  targets = "gpp"
)
```

```{r eval = FALSE, echo = FALSE}
# Save calibration output because it takes very long to compute
save(par_calib, file = "files/par_calib.rda")
```

```{r eval = TRUE, echo = FALSE}
# Load calibration output
load("files/par_calib.rda")
```

`BayesianTools` makes it easy to produce the trace plot of the MCMC chains and the posterior density plot for the parameters. Trace plots show the time series of the sampled chains, which should reach a stationary state. One can also choose a burnin visually, to discard the early iterations and keep only the samples from the stationary distribution to which they converge. We set \code{burnin = 3000} above from previous runs, and those iterations are not shown by the following trace plot. The samples after the burnin period should be used for inference.
```{r fig.height = 10, fig.width = 7}
plot(par_calib$mod)
```

The posterior density plots may be lumpy. In this case it's advisable to run the MCMC algorithm for more iterations, in order to get a better estimate of the parameters' posterior distributions. A good posterior should look more gaussian (although it can be skewed). A multimodal density indicates that the MCMC is still exploring the parameter space and hasn't converged yet. The posteriors can be plotted against the priors using `BayesianTools::marginalPlot()`.

When convergence has been reached, the oscillation of the time series should look like white noise. It's normal that consecutive MCMC samples are correlated because of the sampling algorithm's nature, but the presence of a more general trend indicates that convergence hasn't been reached.

Furthermore, trace plots can be deceiving and partial autocorrelation plots can throw some light. If autocorrelation is present, this can mean that the sampling is stuck in local maxima and the posterior parameter space may not be explored fully. Sometimes, thinning is used to deal with this autocorrelation (see [Link and Eaton 2012](https://besjournals.onlinelibrary.wiley.com/doi/pdf/10.1111/j.2041-210X.2011.00131.x)).
```{r fig.height = 10, fig.width = 7}
# Define function for plotting chains separately
plot_acf_mcmc <- function(chains, par_names){
  # chains: from the BayesianTools output
  n_chains <- length(chains)
  par(mfrow = c(length(par_names), n_chains))
  for(par_name in par_names){
    for(i in 1:n_chains){
      chains[[i]][, par_name] |>
        pacf(main = paste0("Series of ", par_name, " , chain ", i))
    }
  }
}

plot_acf_mcmc(par_calib$mod$chain, c("kphio", "kphio_par_a", "kphio_par_b", "err_gpp"))
```

Looking at the correlation between chains for different parameters is also helpful because parameter correlation may slow down convergence, or the chains may oscillate in the multivariate posterior space. In this calibration we expect parameter samples to be somewhat correlated, especially `kphio_par_a` and `kphio_par_b` because they specify the shape of the temperature dependence of the quantum yield efficiency, $\varphi_o(T)$. We can also see that `err_gpp` is correlated with `kphio` (to which the P-model is very sensitive), since the error represents how good the model fits the observed GPP.
```{r fig.width=5, fig.height=5}
correlationPlot(par_calib$mod, thin = 1)   # use all samples, no thinning
```

In addition to visualizations, it's helpful to compute some convergence diagnostics, like the Gelman-Brooks-Rubin (GBR) potential scale factors. This diagnostic compares the variance within chains to that across chains and should progressively get closer to 1. It is common in the literature to accept convergence with a GBR between 1.05 and 1.1.
```{r}
gelmanDiagnostics(par_calib$mod)
```

Finally, the parameter MAP estimates can be derived from the chains (that converged) after removing the burnin period. They can be seen, next to other statistics, using the `summary` function from the `BayesianTools` library.

```{r}
summary(par_calib$mod)
```

More details on diagnosing MCMC convergence can be found in [this vignette from BayesianTools](https://florianhartig.github.io/BayesianTools/articles/BayesianTools.html#running-mcmc-and-smc-functions) and [this blogpost](https://theoreticalecology.wordpress.com/2011/12/09/mcmc-chain-analysis-and-convergence-diagnostics-with-coda-in-r/).

### Plotting P-model output after calibration

After we have run and checked the calibration, let's see how the model performs. 
```{r echo = FALSE, eval = FALSE}
# Evaluation of the uncertainty coming from the model parameters' uncertainty
# without the `err_gpp` uncertainty.

# Sample kphio values from the posterior distribution
samples_kphio <- getSample(par_calib$mod, 
                             thin = 300,               # get every 10th sample
                             whichParameters = 1:3,      # ignore error_gpp
                             start = 100)              # remove burnin

# Run the P-model to get the posterior GPP predictions
pmodel_runs <- apply(samples_kphio, 1, function(kphio){
  out <- runread_pmodel_f(
    drivers = p_model_drivers,
    par =  list(                      # copied from par_fixed above
      kphio = kphio[1],
      kphio_par_a = kphio[2],
      kphio_par_b = kphio[3],
      soilm_thetastar    = 0.6*240,
      soilm_betao        = 0.2,
      beta_unitcostratio = 146.0,
      rd_to_vcmax        = 0.014,
      tau_acclim         = 30.0,
      kc_jmax            = 0.41)       # value from posterior
  )
  
  out$data[[1]][, c("date", "gpp")]                # return gpp prediction and date
})

# Load observations
validation_data <- p_model_validation$data[[1]]

# Compute standard deviation of predictions
gpp_runs_bound <- c()
for(i in 1:length(pmodel_runs)){
  
}


# Plot first year of P-model runs
posterior_plot <- ggplot() +
  geom_line(
    data = validation_data[1:365,],
    aes(date, gpp),
    alpha = 0.8
  ) +
  labs(
    x = "Date",
    y = "GPP"
  )

for(run in seq_along(pmodel_runs)){
  posterior_plot <- posterior_plot +
    geom_line(
      data = pmodel_runs[[run]][1:365,],
      aes(date, gpp),
      alpha = 0.1,
      color = 'blue'
    )
}

posterior_plot
```

Below we plot the first year of predicted GPP (in grey) against GPP observations (in black).
We use the MAP estimates
for all calibrated parameters to run the P-model and include a 90% credible interval based on the assumption that the
predicted GPP follows a normal distribution with standard deviation `err_gpp`, 
which we have also estimated. This gives a range that covers the observed GPP
in 90% of the cases.

```{r fig.width=7, fig.height=5}
# Plot observed and predicted GPP, with a 95% credible interval using err_gpp
plot_gpp_error <- ggplot(data = runread_pmodel_f(
    drivers = p_model_drivers,
    par =  list(                       
      kphio = par_calib$par[1],        # values from posterior
      kphio_par_a = par_calib$par[2],
      kphio_par_b = par_calib$par[3],
      soilm_thetastar    = 0.6*240,    # copied from par_fixed above
      soilm_betao        = 0.2,
      beta_unitcostratio = 146.0,
      rd_to_vcmax        = 0.014,
      tau_acclim         = 30.0,
      kc_jmax            = 0.41)       
  ) |>
    dplyr::select(sitename, data) |>
    tidyr::unnest(data) |>
    dplyr::slice(1:365)) +             # Plot only first year
  geom_ribbon(
    aes(ymin = gpp - 2*par_calib$par[4], 
        ymax = gpp + 2*par_calib$par[4],
        x = date),
    fill = 'grey40', alpha = 0.2) +
  geom_line(
    aes(
      date,
      gpp
    ),
    colour = "grey40",
    alpha = 0.8
  ) +
  theme_classic() +
  theme(panel.grid.major.y = element_line()) +
  labs(
    x = 'Date',
    y = expression(paste("GPP (g C m"^-2, "s"^-1, ")"))
  )

# Define GPP validation data (first year)
validation_data <- p_model_validation$data[[1]][1:365, ]

# Include observations in the plot
plot_gpp_error +  
  geom_line(
    data = validation_data,
    aes(
      date,
      gpp
    ),
    alpha = 0.8
  )
```
