---
title: "Steady-state estimation"
author: "Beni Stocker"
date: "2026-04-09"
output: html_document
---

This is to assess the accuracy of the steady-state estimation from transient model outputs, in support of Zhang-Zheng et al. (*in rev.*).

We consider two alternative models of the C cycle in a terrestrial ecosystem. The first represents all pool dynamics as a first-order decay process (constant turnover) and assumes constant parameters governing the downstream partition of C flows, determining the ratio of NPP/GPP and allocation fractions. Therefore, by design, it exhibits linear dynamics.

The second model, BiomeEP is a vegetation demography model that explicity simulates forest dynamics. Previous research has demonstrated that upon a change in biomass production (or NPP), steady-state biomass pools change in parallel, but to a lesser degree (in relative terms) (Marqués et al., 2023). The model thus exhibits non-linear dynamics, arising from the relationship between NPP and steady-state vegetation C pools.

This vignette evaluates estimated versus actual steady-state vegetation C pools ($C(t)$) based on the two alternative models. Steady-state vegetation C pools are estimated using an approximative function.

## Analytic approach

The steady-state vegetation C pool ($C^\ast$) can be estimated, assuming first-order decay (constant turnover $\tau$), as:
$$
C^\ast = \text{NPP}(t) / \tau
$$
Also following the first-order decay model, the temporal change in $C(t)$ is given by:
$$
dC/dt = \text{NPP}(t) - C(t)/\tau
$$
Combining these two equations, the turnover rate ($\tau$) can be estimated from NPP, the current (transiently changing) size ($C(t)$) and the temporal change of the vegetationa C pool ($dC/dt$):
$$
\tau = \frac{C(t)}{\text{NPP}(t) – dC/dt}
$$
This, in turn can be plugged into the first equation here. Thus, we get an expression for estimating of the steady-state vegetation C pool size from the transient dynamics as:
$$
C^\ast = \frac{C(t) \times \text{NPP}(t)}{\text{NPP}(t) – dC/dt}
$$
Eq. 12 is applied in the following for estimating the steady-state vegetation C pool size $C^\ast$ from the transient simulation outputs.

## Simulations

Load R libraries.
```{r}
library(dplyr)
library(rsofun)
library(ggplot2)
```

### Linear model

#### Define model

```{r}
#| code-fold: true
#| label: fig-ccascade-simulations
#| fig-cap: "Time series of the pool sizes in the C cascade model, forced with an increase of GPP from 100 to 120 PgC yr^-1^. The first 2000 simulation years are not shown as they are used for model spin-up. The years used for quantifying relative changes below are indicated by the vertical dotted lines."
#| out-width: 80%
ccascade <- function( 
  gpp, 
  state = initpools(),
  par = getpar(),
  fluxes = initfluxes()
  ){

  # determine integration length (number of time steps) from length of 'c_influx'
  len <- length(gpp)

  # output the state and fluxes in one data frame
  df <- dplyr::tibble()

  # integrate over each time step (this is an implementation of the differential equation)
  for (yr in seq(len)){

    # update states and fluxes
    out <- ccascade_onestep(gpp[yr], state, par, fluxes)
    state <- out$state
    fluxes <- out$fluxes

    # record for output
    df <- dplyr::bind_rows(
      df,
      dplyr::bind_cols( 
        dplyr::tibble(year = yr),
        dplyr::as_tibble(state), 
        dplyr::as_tibble(fluxes))
      )
  }
  return(df)
}

ccascade_onestep <- function(gpp, state, par, fluxes){

  # total biomass production
  bp <- gpp * par$bpe
  fluxes$ra <- gpp * (1 - par$bpe)

  # biomass turnover by pool, use exponential here to avoid numerical instability with k -> 1
  dleaf <- (1 - exp(-par$kleaf)) * state$cleaf
  dwood <- (1 - exp(-par$kwood)) * state$cwood
  droot <- (1 - exp(-par$kroot)) * state$croot

  # update biomass pools
  state$cleaf <- state$cleaf + bp * par$fleaf - dleaf
  state$cwood <- state$cwood + bp * par$fwood - dwood
  state$croot <- state$croot + bp * par$froot - droot

  # litter turnover
  dflitt <- (1 - exp(-par$kflitt)) * state$flitt
  dslitt <- (1 - exp(-par$kslitt)) * state$slitt

  # update litter: add biomass turnover and update litter pools
  state$flitt <- state$flitt + dleaf + droot - dflitt
  state$slitt <- state$slitt + dwood         - dslitt

  # heterotrophic respiration during litter decomposition
  rh_litt <- (1 - par$eff) * (dflitt + dslitt)

  # soil turnover
  dfsoil <- (1 - exp(-par$kfsoil)) * state$fsoil
  dssoil <- (1 - exp(-par$kssoil)) * state$ssoil

  # update soil: remainder of litter turnover goes to soil, subtract decomposing portion
  state$fsoil <- state$fsoil + par$eff * par$ffast       * (dflitt + dslitt) - dfsoil
  state$ssoil <- state$ssoil + par$eff * (1 - par$ffast) * (dflitt + dslitt) - dssoil

  # heterotrophic respiration from soil decomposition
  rh_soil <- dfsoil + dssoil
  fluxes$rh <- rh_litt + rh_soil

  return(list(state = state, fluxes = fluxes))
}

getpar <- function(){
  list(
    bpe = 0.4,
    fleaf = 0.3,
    froot = 3,
    fwood = 0.4,
    kleaf = 0.5,
    kwood = 0.02,
    kroot = 0.5,
    kflitt = 0.5,
    kslitt = 0.1,
    kfsoil = 0.1,
    kssoil = 0.003,
    eff = 0.6,
    ffast = 0.95
    )
}

initpools <- function(){
  list(
    cleaf = 0,
    cwood = 0,
    croot = 0,
    flitt = 0,
    slitt = 0,
    fsoil = 0,
    ssoil = 0
    )
}

initfluxes <- function(){
  list(
    ra = 0,
    rh = 0
    )
}
```

#### Create forcing

```{r}
#| code-fold: true
#| label: fig-ccascade-simulations
#| fig-cap: "Time series of the pool sizes in the C cascade model, forced with an increase of GPP from 100 to 120 PgC yr^-1^. The first 2000 simulation years are not shown as they are used for model spin-up. The years used for quantifying relative changes below are indicated by the vertical dotted lines."
#| out-width: 80%

# specify GPP for each simulation year as a vector
gpp <- c(rep(100, 3000), rep(120, 3000))
```

#### Run model

```{r}
# run the model, it returns a data frame
df <- ccascade(gpp)

df <- df |> 
  mutate(cveg = cleaf + cwood + croot)

df |> 
  ggplot() +
  geom_line(aes(year, cveg, color = "Vegetation C")) +
  geom_vline(xintercept = 2500, linetype = "dotted") +
  geom_vline(xintercept = 5500, linetype = "dotted") +
  xlim(2000, 6000) +
  khroma::scale_color_okabeito(name = "") +
  theme_classic() +
  labs(x = "Simulation year", y = "Pool size (PgC)")
```

### BiomeEP