---
title: "Compare P-model implementations"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

This Rmarkdown file is part of the rsofun package (`vignettes/versions_pmodel.Rmd`).

## Environment

Load the `rsofun` package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r message=FALSE, echo=FALSE}
library(rsofun)
library(captioner)
library(dplyr)
## load all rsofun dependencies
# load_dependencies_rsofun()

## for creating figure numbers
tab_nums <- captioner( prefix = "Table S", auto_space=FALSE, style = "i" )
fig_nums <- captioner( prefix = "Figure S", auto_space=FALSE, style = "i" )

## other crap
systr <- "''"    # for Mac
knitr::opts_knit$set( root.dir = rprojroot::find_rstudio_root_file() ) # does not work properly
```

## Run different implementations

```{r, eval=TRUE, message=FALSE, warning=FALSE}
## R, Vcmax based on Wang Han's formulation
pmodel_wanghan <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~rpmodel( tc = ., vpd = 1000, co2 = 400, elv = 0, kphio = 0.257, fapar = 1.0, ppfd = 800, method="wanghan", do_ftemp_kphio = FALSE ) 
  )

pmodel_smith <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~rpmodel( tc = ., vpd = 1000, co2 = 400, elv = 0, kphio = 0.257, fapar = 1.0, ppfd = 800, method="smith", do_ftemp_kphio = FALSE ) 
  )

## Vcmax based on Nick's code
source("~/optimal_vcmax_R/calc_optimal_vcmax.R")
pmodel_smithcode <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~calc_optimal_vcmax( tg_c = ., z = 0, vpdo = 1, cao = 400, paro = 800, q0 = 0.257, theta = 0.85 )
  )
```

## Evaluation

Below, I'm comparing different variables calculated in the different implementations. 'rsofun standard' refers to what is implemented in rsofun, based on Wang Han et al., 2017 and originally adopted from the GePiSaT code. 'Smith my implementation' is based on Beni's adoptation of Nick's code within rsofun, 'Smith original' is Nick Smith's code from [github](https://github.com/SmithEcophysLab/optimal_vcmax_R) (run with PAR and VPD not additionally adjusted for elevation-effects).

### Michaelis-Menten K

```{r, eval=TRUE, message=FALSE, warning=FALSE}
kmm_smithcode <- pmodel_smithcode %>% purrr::map_dbl("km")
kmm_wanghan   <- pmodel_wanghan   %>% purrr::map_dbl("kmm")
kmm_smith     <- pmodel_smith     %>% purrr::map_dbl("kmm")

plot(  seq( 0, 35, length.out = 100 ), kmm_wanghan, type = "l", xlab = "Temperature (deg C)", ylab = "Michaelis-Menten K", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), kmm_smith, col="blue", lwd=3 )
lines( seq( 0, 35, length.out = 100 ), kmm_smithcode, col="red" )
legend( "topright", c("rsofun standard", "Smith my implementation", "Smith original"), lty = 1, col = c("black", "blue", "red"), lwd=c(6,3,1), bty = "n")
```

### $c_i$

```{r, eval=TRUE, message=FALSE, warning=FALSE}
ci_smithcode <- pmodel_smithcode %>% purrr::map_dbl("ci")
ci_wanghan   <- pmodel_wanghan   %>% purrr::map_dbl("ci")
ci_smith     <- pmodel_smith     %>% purrr::map_dbl("ci")

plot(  seq( 0, 35, length.out = 100 ), ci_wanghan, type = "l", xlab = "Temperature (deg C)", ylab = "ci", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), ci_smith, col="blue", lwd=3 )
lines( seq( 0, 35, length.out = 100 ), ci_smithcode, col="red" )
legend( "topright", c("rsofun standard", "Smith my implementation", "Smith original"), lty = 1, col = c("black", "blue", "red"), lwd=c(6,3,1), bty = "n")
```

### $c_i:c_a$

```{r, eval=TRUE, message=FALSE, warning=FALSE}
chi_smithcode <- pmodel_smithcode %>% purrr::map_dbl("chi")
chi_wanghan   <- pmodel_wanghan   %>% purrr::map_dbl("chi")
chi_smith     <- pmodel_smith     %>% purrr::map_dbl("chi")

plot(  seq( 0, 35, length.out = 100 ), chi_wanghan, type = "l", xlab = "Temperature (deg C)", ylab = "ci:ca", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), chi_smith, col="blue", lwd=3 )
lines( seq( 0, 35, length.out = 100 ), chi_smithcode, col="red" )
legend( "topright", c("rsofun standard", "Smith my implementation", "Smith original"), lty = 1, col = c("black", "blue", "red"), lwd=c(6,3,1), bty = "n")
```

### m (=mj)

```{r, eval=TRUE, message=FALSE, warning=FALSE}
m_smith     <- pmodel_smith     %>% purrr::map_dbl("m")
m_smithcode <- pmodel_smithcode %>% purrr::map_dbl("m")

plot(  seq( 0, 35, length.out = 100 ), m_smith, type = "l", xlab = "Temperature (deg C)", ylab = "m", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), m_smithcode, col="red" )
legend( "topright", c("Smith my implementation", "Smith original"), lty = 1, col = c("black", "red"), lwd=c(6,1), bty = "n")
```

### mc

```{r, eval=TRUE, message=FALSE, warning=FALSE}
m_smith     <- pmodel_smith     %>% purrr::map_dbl("mc")
m_smithcode <- pmodel_smithcode %>% purrr::map_dbl("mc")

plot(  seq( 0, 35, length.out = 100 ), m_smith, type = "l", xlab = "Temperature (deg C)", ylab = "m", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), m_smithcode, col="red" )
legend( "topright", c("Smith my implementation", "Smith original"), lty = 1, col = c("black", "red"), lwd=c(6,1), bty = "n")
```

### $\omega$

```{r, eval=TRUE, message=FALSE, warning=FALSE}
omega_smith     <- pmodel_smith     %>% purrr::map_dbl("omega")
omega_smithcode <- pmodel_smithcode %>% purrr::map_dbl("omega")

plot(  seq( 0, 35, length.out = 100 ), omega_smith, type = "l", xlab = "Temperature (deg C)", ylab = "m", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), omega_smithcode, col="red" )
legend( "topright", c("Smith my implementation", "Smith original"), lty = 1, col = c("black", "red"), lwd=c(6,1), bty = "n")
```


### $\omega^\ast$

```{r, eval=TRUE, message=FALSE, warning=FALSE}
omega_star_smith     <- pmodel_smith     %>% purrr::map_dbl("omega_star")
omega_star_smithcode <- pmodel_smithcode %>% purrr::map_dbl("omega_star")

plot(  seq( 0, 35, length.out = 100 ), omega_star_smith, type = "l", xlab = "Temperature (deg C)", ylab = "m", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), omega_star_smithcode, col="red" )
legend( "topright", c("Smith my implementation", "Smith original"), lty = 1, col = c("black", "red"), lwd=c(6,1), bty = "n")
```

### f$_{temp}$

```{r, eval=TRUE, message=FALSE, warning=FALSE}
ftemp_inst_vcmax_smith     <- pmodel_smith     %>% purrr::map_dbl("ftemp_inst_vcmax")
ftemp_inst_vcmax_smithcode <- pmodel_smithcode %>% purrr::map_dbl("ftemp_vcmax")

plot(  seq( 0, 35, length.out = 100 ), ftemp_inst_vcmax_smith, type = "l", xlab = "Temperature (deg C)", ylab = "m", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), ftemp_inst_vcmax_smithcode, col="red" )
legend( "topleft", c("Smith my implementation", "Smith original"), lty = 1, col = c("black", "red"), lwd=c(6,1), bty = "n")
```


### Vcmax*

```{r, eval=TRUE, message=FALSE, warning=FALSE}
vcmax_star_smith     <- pmodel_smith     %>% purrr::map_dbl("vcmax_star")
vcmax_star_smithcode <- pmodel_smithcode %>% purrr::map_dbl("vcmax_star")

plot(  seq( 0, 35, length.out = 100 ), vcmax_star_smith, type = "l", xlab = "Temperature (deg C)", ylab = "Vcmax*", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), vcmax_star_smithcode, col="red" )
legend( "topleft", c("Smith my implementation", "Smith original"), lty = 1, col = c("black", "red"), lwd=c(6,1), bty = "n")
```



### Vcmax

```{r, eval=TRUE, message=FALSE, warning=FALSE}
vcmax_smithcode <- pmodel_smithcode %>% purrr::map_dbl("vcmax_prime")
vcmax_wanghan   <- pmodel_wanghan   %>% purrr::map_dbl("vcmax")
vcmax_smith     <- pmodel_smith     %>% purrr::map_dbl("vcmax")

plot(  seq( 0, 35, length.out = 100 ), vcmax_smith, type = "l", xlab = "Temperature (deg C)", ylab = "Vcmax", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), vcmax_wanghan, col="blue", lwd=3 )
lines( seq( 0, 35, length.out = 100 ), vcmax_smithcode, col="red" )
legend( "topleft", c("Smith my implementation", "rsofun standard", "Smith original"), lty = 1, col = c("black", "blue", "red"), lwd=c(6,3,1), bty = "n")

plot(  seq( 0, 35, length.out = 100 ), vcmax_wanghan, type = "l", xlab = "Temperature (deg C)", ylab = "Vcmax", lwd=6 )
legend( "topleft", c("rsofun standard (Wang Han)"), lty = 1, col = c("black"), lwd=6, bty = "n")
```


### Light use efficiency

```{r, eval=TRUE, message=FALSE, warning=FALSE}
lue_wanghan   <- pmodel_wanghan   %>% purrr::map_dbl("lue")
lue_smith     <- pmodel_smith     %>% purrr::map_dbl("lue")

plot(  seq( 0, 35, length.out = 100 ), lue_smith, type = "l", xlab = "Temperature (deg C)", ylab = "lue", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), lue_wanghan, col="blue", lwd=3 )
legend( "topright", c("Smith my implementation", "rsofun standard"), lty = 1, col = c("black", "blue", "red"), lwd=c(6,3,1), bty = "n")

plot(  seq( 0, 35, length.out = 100 ), lue_wanghan, type = "l", xlab = "Temperature (deg C)", ylab = "lue", lwd=6 )
legend( "topright", c("rsofun standard (Wang Han)"), lty = 1, col = c("black"), lwd=6, bty = "n")
```


