---
title: "Prepare FLUXNET 2015 ensemble forcing and evaluation data"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)
library(rsofun)
library(ingestr)
```

## Forcing data

Data is collected and available in file `~/data/stocker20gmd_outputs/ddf_fluxnet2015_pmodel_with_forcings_stocker19gmd.csv` (available on Euler). This here is to re-format it such that it corresponds to:
```{r}
rsofun::df_drivers
```

The final drivers data frame is saved as `"~/data/rsofun_benchmarking/df_drivers_fluxnet2015.Rdata"`.


### FLUXNET meteo data and MODIS FPAR

Reformat, but don't nest yet.
```{r}
ddf <- read_csv("~/data/stocker20gmd_outputs/ddf_fluxnet2015_pmodel_with_forcings_stocker19gmd.csv")
ddf_forcing <- ddf %>%
  dplyr::select(
    sitename, date, 
    temp = temp_day_fluxnet2015, 
    prec = prec_fluxnet2015, 
    vpd = vpd_day_fluxnet2015, 
    ppfd = ppfd_fluxnet2015, 
    patm = patm_fluxnet2015,
    fapar = fapar_spl
    )

usesites <- ddf_forcing %>% pull(sitename) %>% unique()

# ddf_obs <- ddf %>% 
#   dplyr::select(sitename, date, gpp_obs)
```


### Cloud cover from CRU

Some meteo data is not available from FLUXNET. Extract it from CRU global climate files instead.
```{r}
ddf_cru <- ingestr::ingest(
  siteinfo  = siteinfo_fluxnet2015 %>% 
    dplyr::filter(sitename %in% usesites),
  source    = "cru",
  getvars   = list(ccov = "cld"),
  dir       = "~/data/cru/ts_4.01/"
  )
```

```{r}
ddf_forcing <- ddf_forcing %>% 
  left_join(ddf_cru %>% unnest(data) %>% dplyr::select(sitename, date, ccov),
            by = c("sitename", "date"))
```

### CO2

Ingesting CO2 data is particularly simple. We can safely assume it's well mixed in the atmosphere (independent of site location), and we can use a annual mean value for all days in respective years.  
```{r}
df_co2 <- ingestr::ingest(
  siteinfo_fluxnet2015 %>% 
    dplyr::filter(sitename %in% usesites),
  source  = "co2_mlo",
  verbose = FALSE
  )

ddf_forcing <- ddf_forcing %>% 
  left_join(df_co2 %>% unnest(data),
            by = c("sitename", "date"))
```

### Save and nest

Save the flat file.
```{r}
write_csv(ddf_forcing, path = "~/data/stocker20gmd_outputs/ddf_forcings_stocker19gmd.csv")
```

## Collect all drivers

- `df_soiltexture`: dummy for all the same. Using WHC as a site parameter instead.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

Define simulation parameters common for all simulations:
```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)
```

Finally, all input data can be collected by:
```{r}
df_drivers_fluxnet2015 <- collect_drivers_sofun( 
  siteinfo       = siteinfo_fluxnet2015 %>% 
    dplyr::filter(sitename %in% usesites),
  params_siml    = params_siml,
  meteo          = ddf_forcing %>% dplyr::select(-fapar, -co2) %>% group_by(sitename) %>% nest(), 
  fapar          = ddf_forcing %>% dplyr::select(sitename, date, fapar) %>% group_by(sitename) %>% nest(),
  co2            = ddf_forcing %>% dplyr::select(sitename, date, co2) %>% group_by(sitename) %>% nest(),
  df_soiltexture = df_soiltexture
  )
df_drivers_fluxnet2015
```

Save the full nested object.
```{r}
save(df_drivers_fluxnet2015, file = "~/data/rsofun_benchmarking/df_drivers_fluxnet2015.Rdata")
```


## Appendix

### Site list

```{r}
siteinfo_fluxnet2015 %>% 
    dplyr::filter(sitename %in% usesites)%>% 
  kable()
```