---
title: "Prepare FLUXNET 2015 ensemble forcing and evaluation data"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)
library(rsofun)
library(ingestr)
```

## Forcing data

Collect data as described also [here](https://stineb.github.io/rsofun/articles/prepare_inputs_rsofun.html#network-site-ensemble-1).

### FLUXNET meteo data

```{r}
ddf_fluxnet <- ingestr::ingest(
  siteinfo  = ingestr::siteinfo_fluxnet2015,
  source    = "fluxnet",
  getvars   = list(temp = "TA_F_DAY", prec = "P_F", vpd  = "VPD_F_DAY", ppfd = "SW_IN_F", patm = "PA_F"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  settings  = list(dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", getswc = FALSE),
  timescale = "d"
  )
```

### Cloud cover from CRU

Some meteo data is not available from FLUXNET. Extract it from CRU global climate files instead.
```{r}
ddf_cru <- ingestr::ingest(
  siteinfo  = siteinfo_fluxnet2015,
  source    = "cru",
  getvars   = "ccov",
  dir       = "~/data/cru/ts_4.01/"
  )
```

```{r}
ddf_meteo <- ddf_fluxnet %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru %>% 
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>% 
  group_by(sitename) %>% 
  tidyr::nest() %>% 
  dplyr::filter(!(sitename %in% c("DE-Akm", "US-ORv", "DE-RuS")))
```


### MODIS fAPAR

```{r}
settings_modis <- get_settings_modis(
  bundle            = "modis_fpar",
  data_path         = "~/data/modis_subsets/",
  method_interpol   = "loess",
  keep              = TRUE,
  overwrite_raw     = FALSE,
  overwrite_interpol= TRUE,
  n_focal           = 0
  )

df_modis_fpar <- ingest(
  siteinfo_fluxnet2015 %>% dplyr::filter(!(sitename %in% c("DE-Akm", "US-ORv", "DE-RuS"))), 
  source = "modis",
  settings = settings_modis, 
  parallel = FALSE,
  ncores = 1  # allows parallelised download
  )

df_modis_fpar <- df_modis_fpar %>% 
  mutate(data = purrr::map(data, ~rename(., fapar = modisvar_filled)))
```

### CO2

Ingesting CO2 data is particularly simple. We can safely assume it's well mixed in the atmosphere (independent of site location), and we can use a annual mean value for all days in respective years.  
```{r}
df_co2 <- ingestr::ingest(
  siteinfo_fluxnet2015 %>% dplyr::filter(!(sitename %in% c("DE-Akm", "US-ORv", "DE-RuS"))),
  source  = "co2_mlo",
  verbose = FALSE
  )
```

### Save and nest

Save the flat file.
```{r}
write_csv(ddf_forcing, path = "~/data/stocker20gmd_outputs/ddf_forcings_fluxnet2015_allsites.csv")
```

## Collect all drivers

- `df_soiltexture`: dummy for all the same. Using WHC as a site parameter instead.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

Define simulation parameters common for all simulations:
```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)
```

Finally, all input data can be collected by:
```{r}
df_drivers_fluxnet2015_allsites <- collect_drivers_sofun( 
  siteinfo       = siteinfo_fluxnet2015 %>% dplyr::filter(!(sitename %in% c("DE-Akm", "US-ORv", "DE-RuS"))),
  params_siml    = params_siml,
  meteo          = ddf_meteo, 
  fapar          = df_modis_fpar,
  co2            = df_co2,
  df_soiltexture = df_soiltexture
  )
```

Save the full nested object.
```{r}
save(df_drivers_fluxnet2015_allsites, file = "~/data/rsofun_benchmarking/df_drivers_fluxnet2015_allsites.Rdata")
```


## Appendix

### Site list

```{r}
siteinfo_fluxnet2015 %>% 
    dplyr::filter(sitename %in% usesites)%>% 
  kable()
```