---
title: "prepare_drivers"
output: html_document
---

```{r setup, include=FALSE}
library(rsofun)
library(dplyr)
library(readr)
library(ingestr)
```

## Site selection and meta data

```{r}
mysites <- "CH-Lae"
```

```{r}
siteinfo <- data.frame(sitename="CH-Lae", lon = 8.365, lat = 47.47808, elv = 700, year_start = 2000, year_end = 2019, classid = NA, c4 = FALSE, whc = NA, koeppen_code = NA, igbp_land_use = "Mixed Forests", plant_functional_type = "Broadleaf trees")

siteinfo <- as_tibble(siteinfo)
siteinfo <- siteinfo %>% 
  dplyr::mutate(date_start = lubridate::ymd(paste0(year_start, "-01-01"))) %>%
  dplyr::mutate(date_end = lubridate::ymd(paste0(year_end, "-12-31")))
```

## Simulation settings

```{r}
params_siml <- list(
      spinup                = TRUE,
      spinupyears           = 1800, #1793,
      recycle               = 1, # 9 or 11 changed to 1 when aggregating forcing into 1 year
      firstyeartrend        = 1998,
      nyeartrend            = 1, #9 or 11 (longer transient years)
      outputhourly          = TRUE,
      outputdaily           = TRUE,
      do_U_shaped_mortality = FALSE,
      update_annualLAImax   = FALSE,
      do_closedN_run        = TRUE,
      method_photosynth     = "gs_leuning", # gs_leuning or pmodel
      method_mortality      = "dbh" # dbh or cstarvation or growthrate or const_selfthin
      )
```

```{r}
siteinfo <- tibble(siteinfo, params_siml = list(tibble(params_siml)))
```

## Define model parameters

#### Tile-level parameters

```{r}
params_tile <- list(
  soiltype     = 3,     # Sand = 1, LoamySand = 2, SandyLoam = 3, SiltLoam = 4, FrittedClay = 5, Loam = 6, Clay = 7
  FLDCAP       = 0.4,   # soil property: field capacity 
  WILTPT       = 0.05,  # soil property: wilting point
  K1           = 2.0,   # turnover rate of fast SOM per year
  K2           = 0.05,  # turnover rate of slow SOM per year
  K_nitrogen   = 8.0,   # mineral Nitrogen turnover rate
  MLmixRatio   = 0.8,   # the ratio of C and N returned to litters from microbes
  etaN         = 0.025, # loss rate with runoff
  LMAmin       = 0.02,  # minimum LMA, boundary condition
  fsc_fine     = 1.0,   # fraction of fast turnover carbon in fine biomass
  fsc_wood     = 0.0,   # fraction of fast turnover carbon in wood biomass
  GR_factor    = 0.33,  # growth respiration factor
  l_fract      = 0.0,   # fraction of the carbon retained after leaf drop
  retransN     = 0.0,   # retranslocation coefficient of Nitrogen
  f_initialBSW = 0.2,
  f_N_add      = 0.02   # re-fill of N for sapwood
  )

```

#### Species-level parameters

```{r}
params_species <- tibble(
  lifeform      = rep(1,16),                 # 0 for grasses; 1 for trees
  phenotype     = rep(0,16),                 # 0 for Deciduous; 1 for Evergreen
  pt            = rep(0,16),                 # 0 for C3; 1 for C4
  seedlingsize  = rep(0.05,16),              # initial size of seedlings
  LMA           = rep(0.035,16),             # Leaf mass per unit area: 0.035, 0.085, 0.135
  phiRL         = rep(3.5,16),               # Root:Shoot ratio:  4.0, 6.0, 8.0
  LNbase        = rep(0.8E-3,16),            # kgN m-2 leaf, Vmax = 0.03125*LNbase
  laimax        = rep(3.5,16),               # maximum crown LAI
  LAI_light     = rep(4.0,16),               # Light-limited crown LAI
  Nfixrate0     = rep(0,16),                 # 0.03 kgN kgRootC-1 yr-1
  NfixCost0     = rep(12,16),                # 12, 24 gC/gN
  phiCSA        = rep(0.25E-4,16),           # ratio of sapwood area to leaf area
  mortrate_d_c  = rep(0.01,16),              # canopy tree mortality rate, year-1
  mortrate_d_u  = rep(0.075,16),             # understory tree mortality rate, year-1
  maturalage    = rep(5,16),                 # the age that can reproduce
  fNSNmax       = rep(5,16)                  # multiplier for NSNmax as sum of potential bl and br
  ) 
```

## Define soil parameters

#### Soil parameters

By layers.
```{r}
# adopted from datatypes.mod.f90 l.538
params_soil <- tibble(
  type              = c("Coarse","Medium","Fine","CM","CF","MF","CMF","Peat","MCM"),
  GMD               = c(0.7, 0.4, 0.3, 0.1, 0.1, 0.07, 0.007, 0.3, 0.3),
  GSD               = c(5.0, 5.3, 7.4, 6.1, 6.1, 14.0, 15.0, 7.4, 7.4),
  vwc_sat           = c(0.380, 0.445, 0.448, 0.412, 0.414, 0.446, 0.424, 0.445, 0.445),
  chb               = c(3.5,6.4,11.0,4.8,6.3,8.4,6.3,6.4,6.4),
  psi_sat_ref       = c(-600, -790, -910, -1580, -1680, -1880, -5980, -790, -790), # Pa
  k_sat_ref         = c(130.8, 75.1, 53.2, 12.1, 11.1, 12.7, 1.69, 53.2, 53.2), # mol/(s MPa m)
  alphaSoil         = rep(1, 9),
  heat_capacity_dry = c(1.2e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.1e6, 1.4e6, 1.0)
  )
```

#### Initial cohort specification

```{r}
init_cohort <- tibble(
 init_cohort_species = rep(2, 10),
 init_cohort_nindivs = rep(1,10),    # initial individual density, individual/m2
 init_cohort_bsw     = rep(0.05,10), # initial biomass of sapwood, kg C/individual
 init_cohort_bHW     = rep(0.0, 10), # initial biomass of heartwood, kg C/tree
 init_cohort_nsc     = rep(0.05,10)  # initial non-structural biomass
)
```

#### Initial soil pools

high N input --> Deciduous, low  N input --> Evergreen
```{r}
init_soil <- list(
 init_fast_soil_C    = 0.0,    # initial fast soil C, kg C/m2
 init_slow_soil_C    = 0.0,    # initial slow soil C, kg C/m2
 init_Nmineral       = 0.015,  # Mineral nitrogen pool, (kg N/m2)
 N_input             = 0.0008  # annual N input to soil N pool, kgN m-2 yr-1
)
```

### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

## Get input (Forcing data)

## Meteo data

```{r message=FALSE, warning=FALSE}
library(tidyverse)
FLX_HH_LAE <- read.csv("/Users/lmarques/eulerdata/FLUXNET-2015_Tier1/20191024/HH/FLX_CH-Lae_FLUXNET2015_FULLSET_HH_2004-2014_1-3.csv")
names(FLX_HH_LAE)

FluxForcing <- FLX_HH_LAE %>% dplyr::select(TIMESTAMP_START,PPFD_IN,SW_IN_F,TA_F,TS_F_MDS_1,RH,P_F,WS_F,PA_F,SWC_F_MDS_1)
FluxForcing <- FluxForcing %>% dplyr::mutate_all(funs(ifelse(.==-9999, NA, .))) %>% mutate(YEAR=str_sub(TIMESTAMP_START, 1, 4)) %>% mutate(DOY=str_sub(TIMESTAMP_START, 7, 8)) %>% mutate(HOUR=str_sub(TIMESTAMP_START, 9, 10)) %>% relocate(YEAR, DOY, HOUR) %>% mutate(YEAR=as.integer(YEAR),DOY=as.numeric(DOY),HOUR=as.numeric(HOUR))
FluxForcing <- FluxForcing %>% dplyr::rename(PAR=PPFD_IN,Swdown=SW_IN_F,TEMP=TA_F,SoilT=TS_F_MDS_1,RH=RH,RAIN=P_F,WIND=WS_F,PRESSURE=PA_F,SWC=SWC_F_MDS_1) %>% dplyr::select(-TIMESTAMP_START) %>% 
  mutate(date = lubridate::ymd_hm( paste0( as.character(YEAR), "-01-01 00:00" ) ) + days(DOY-1) + hours(HOUR) ) %>% 
  mutate(sitename = "CH-Lae") %>% 
  dplyr::filter(!(mday(date)==29 & month(date)==2))
str(FluxForcing)

FluxForcing <- FluxForcing %>% 
    dplyr::group_by(lubridate::year(date),lubridate::month(date),lubridate::day(date),lubridate::hour(date)) %>% 
    summarise_at(vars(1:14), list(~mean(., na.rm = TRUE))) 
FluxForcing <- FluxForcing[,-c(1:4)]

```

Some meteo data is not available from FLUXNET. Extract it from CRU global climate files instead.
```{r}
ddf_cru <- ingestr::ingest(
  siteinfo  = siteinfo,
  source    = "cru",
  getvars   = list(ccov = "cld"),
  dir       = "~/data/cru/ts_4.01/"
  )
```

Combine the two meteo data frames into one, containing `ccov` (cloud cover) from CRU and all other variables from FLUXNET.
```{r}
ddf_meteo <- ddf_fluxnet %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru %>% 
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>% 
  group_by(sitename) %>% 
  tidyr::nest()
```

## fAPAR data

```{r}
fapar_LAE <- read.csv("fapar_MODIS_FPAR_MCD15A3H_gee_CH-Lae_subset.csv")
```


## CO2

Ingesting CO2 data is particularly simple. We can safely assume it's well mixed in the atmosphere (independent of site location), and we can use a annual mean value for all days in respective years.  
```{r}
df_co2 <- ingestr::ingest(
  siteinfo,
  source  = "co2_mlo",
  verbose = FALSE
  )
```

## Collect all drivers

Finally, we can collect forcing data, simulation parameters, and site meta info into a single object that will be used to drive rsofun. All of the above steps can be customized. The function `collect_drivers_sofun()` can in general be used to process forcing data into the format required to run SOFUN. The arguments must have the following form:

- `siteinfo`: A data frame (tibble) with columns `sitename`, `lon`, `lat`, `elv` (elevation), `year_start`, `year_end`, `whc` (water holding capacity used for simulating the soil water balance), `c4` (vegetation following C4 photosynthesis). An example for FLUXNET 2015 sites is given by 
```{r} 
ingestr::siteinfo_fluxnet2015 %>% dplyr::select(sitename, lon, lat, elv, year_start, year_end, whc)
```
- `meteo`: A nested data frame with columns `sitename` and `data`. The latter contains the nested meteorological forcing data frames, with columns `date`, `temp`, `prec`, `vpd`, `ppfd`, `patm`, and `ccov`. Like this:
```{r}
## nested data frame:
ddf_meteo

## inside .$data:
ddf_meteo$data[[1]] %>% dplyr::select(-ccov_int)
```


- `fapar`: A nested data frame with columns `sitename` and `data`. The latter contains the nested meteorological forcing data frames, with columns `date`, and `fapar`. Like this:
```{r}
## nested data frame:
df_gee_modis_fpar

## inside .$data:
df_gee_modis_fpar$data[[1]] %>% dplyr::select(date, fapar)
```
- `co2`  : A nested data frame with columns `sitename` and `data`. The latter contains the nested meteorological forcing data frames, with columns `date`, and `co2`. Like this:
```{r}
## nested data frame:
df_co2

## inside .$data:
df_co2$data[[1]] %>% dplyr::select(date, co2)
```
- `df_soiltexture`: See above ('Define soil parameters')

See [here](https://tidyr.tidyverse.org/reference/nest.html) for how to handle nested dataframes. 

Finally, all input data can be collected by:
```{r}
df_drivers <- collect_drivers_sofun( 
  siteinfo       = siteinfo,
  meteo          = ddf_meteo, 
  fapar          = df_gee_modis_fpar,
  co2            = df_co2,
  params_tile    = params_tile,
  params_species = params_species,
  params_soil    = params_soil,
  init_cohort    = init_cohort,
  init_soil      = init_soil,
  df_soiltexture = df_soiltexture
  )
df_drivers
```

