---
title: "P-model simulation at NEECF site"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(rsofun)
library(dplyr)
library(readr)
library(ingestr)
library(ggplot2)
```


## Site selection and meta data

```{r}
mysites <- "efg"
```

 Required meta information is specified for each site (in rows) and a number of variables:

- `lat` for latitude (decimal degrees)
- `lon` for longitude (decimal degrees) - this is only used for data ingestion but not for the P-model simulation with `rsofun`.
- `elv` for elevation (m a.s.l.)
- `year_start` and `year_end` specifying years covered by the simulation
- `whc` for the soil water holding capacity
- `koeppen_code` to group sites for evaluation by Koeppen-Geiger climate zones.

Let's create this info.
```{r}
siteinfo <- tibble(
  sitename = "efg",
  lon = 117.369258403,
  lat = 30.026229678,
  elv = 330,
  year_start = 2015,
  year_end = 2015,
  whc = 250
  )
```

## Simulation settings

Specify additional simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)
```

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters), global simulation parameters are wrapped inside an additional column `params_siml`, added to the site meta info dataframe.
```{r}
siteinfo <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```


## Define model parameters

First, let's do it by hand (calibration of parameters is shown later).
```{r}
params_modl <- list(
	kphio           = 0.09423773,
	soilm_par_a     = 0.33349283,
	soilm_par_b     = 1.45602286,
	vpdstress_par_a = 999,
	vpdstress_par_b = 999,
	vpdstress_par_m = 999
	)
```

## Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific. Values entered here take no effect.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

## Get input

Forcing data was collected as described in `ingestr/vignettes_add/get_sitedata_nonetwork.Rmd`.
```{r}
secs_per_tstep <- 60 * 60 * 24
forcing <- read_csv("~/data/neecf/climate_data/climate_data_collected.csv") %>% 
  mutate(prec = prec / secs_per_tstep, ppfd = ppfd / secs_per_tstep)
```

CO2 is separate.
```{r}
df_co2 <- ingest_bysite(
  sitename  = "efg",
  source  = "co2_mlo",
  year_start= 2015,
  year_end  = 2015,
  verbose = FALSE
  )
```



## Run the model

Run the model for all the sites specified in the first step.
```{r}
## run for a single site
mod <- run_pmodel_f_bysite( 
  "efg", 
  siteinfo$params_siml[[1]], 
  siteinfo, 
  forcing, 
  df_soiltexture, 
  params_modl = params_modl, 
  makecheck = TRUE 
  )
```

Check.
```{r}
forcing %>% 
  ggplot(aes(date, fapar)) + 
  geom_line()

forcing %>% 
  ggplot(aes(date, ppfd)) + 
  geom_line()

mod %>% 
  ggplot(aes(date, gpp)) + 
  geom_line()

mod %>% 
  ggplot(aes(date, vcmax25)) + 
  geom_line()
```
