---
title: "Example using P-model"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(rsofun)
library(dplyr)
library(readr)
library(ingestr)
library(ggplot2)
```

## Example using rsofun for P-model simulations

The following describes the use of `rsofun` for an ensemble of site-scale P-model simulations, including steps for model calibration and evaluation (benchmarking). `rsofun` is designed in a modular and hierarchical fashion. This enables multiple setups within the same modelling framework. The P-model implementation in `rsofun` is described in Stocker et al. (2020) *Geosci. Mod. Dev.*. 

In the P-model setup, the model requires time series of daily meteorological data as input and is calibrated against observational GPP data. This example describes simulations at a subset of FLUXNET 2015 Tier 1 sites, using GPP based on the night-time flux partitioning method as calibration target and benchmark (no worries, an out-of-sample calibration/evaluation function is available, too). 

### Site selection and meta data

We manually define a subset of sites that are part of the FLUXNET 2015 Tier 1 set of sites:
```{r}
mysites <- "efg"
```

A small number of meta data variables have to be specified for each site specifically to define the simulation years. This information is also used for input, calibration, and evaluation data ingestion. Required meta information is specified for each site (in rows) and a number of variables:

- `lat` for latitude (decimal degrees)
- `lon` for longitude (decimal degrees) - this is only used for data ingestion but not for the P-model simulation with `rsofun`.
- `elv` for elevation (m a.s.l.)
- `year_start` and `year_end` specifying years covered by the simulation
- `whc` for the soil water holding capacity
- `koeppen_code` to group sites for evaluation by Koeppen-Geiger climate zones.

Let's create this info.
```{r}
siteinfo <- tibble(
  sitename = "efg",
  lon = 117.369258403,
  lat = 30.026229678,
  elv = 330,
  year_start = 2015,
  year_end = 2015,
  whc = 250
  )
```

### Simulation settings

Specify additional simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)
```

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters), global simulation parameters are wrapped inside an additional column `params_siml`, added to the site meta info dataframe.
```{r}
siteinfo <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```


### Define model parameters

First, let's do it by hand (calibration of parameters is shown later).
```{r}
params_modl <- list(
	kphio           = 0.09423773,
	soilm_par_a     = 0.33349283,
	soilm_par_b     = 1.45602286,
	vpdstress_par_a = 999,
	vpdstress_par_b = 999,
	vpdstress_par_m = 999
	)
```

### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific. Values entered here take no effect.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Get input

Forcing data was collected as described in `ingestr/vignettes_add/get_sitedata_nonetwork.Rmd`.
```{r}
secs_per_tstep <- 60 * 60 * 24
forcing <- read_csv("~/data/neecf/climate_data/climate_data_collected.csv") %>% 
  mutate(prec = prec / secs_per_tstep, ppfd = ppfd / secs_per_tstep)
```

CO2 is separate.
```{r}
df_co2 <- ingest_bysite(
  sitename  = "efg",
  source  = "co2_mlo",
  year_start= 2015,
  year_end  = 2015,
  verbose = FALSE
  )
```



### Run the model

Run the model for all the sites specified in the first step.
```{r}
## run for a single site
mod <- run_pmodel_f_bysite( 
  "efg", 
  siteinfo$params_siml[[1]], 
  siteinfo, 
  forcing, 
  df_soiltexture, 
  params_modl = params_modl, 
  makecheck = TRUE 
  )
```

Check.
```{r}
forcing %>% 
  ggplot(aes(date, fapar)) + 
  geom_line()

forcing %>% 
  ggplot(aes(date, ppfd)) + 
  geom_line()

mod %>% 
  ggplot(aes(date, gpp)) + 
  geom_line()

mod %>% 
  ggplot(aes(date, vcmax25)) + 
  geom_line()
```
