---
title: "W-model site-scale simulations for Schenk & Jackson 2002 sites"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{splash}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Environment

Load the package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r message=FALSE}
library(rsofun)

## load all rsofun dependencies
load_dependencies_rsofun()

## other crap
systr <- "''"    # for Mac
knitr::opts_knit$set( root.dir = rprojroot::find_rstudio_root_file() ) # does not work properly
## for developer setup
if (!file.exists("bash"))    system("ln -s inst/bash bash")
if (!file.exists("extdata")) system("ln -s inst/extdata extdata")
```


## Simulation settings

Load meta info (longitude and latitude of sites).
```{r message=FALSE}
siteinfo <- read_csv("/alphadata01/bstocker/data/rootingdepth/root_profiles_schenkjackson02/data/root_profiles_D50D95.csv") %>%
  dplyr::filter(Wetland == "N" & Anthropogenic == "N" & Schenk_Jackson_2002 == "YES") %>% 
  dplyr::rename(sitename = ID, lat = Latitude, lon = Longitude) %>% 
  dplyr::mutate(elv = ifelse(elv==-999, NA, elv)) %>% 
  dplyr::filter(lon!=-999 & lat!=-999) %>% 
  dplyr::mutate(year_start = 1982, year_end = 2011) %>% 
  dplyr::select(sitename, lon, lat, elv, year_start, year_end)

write_csv(siteinfo, path = "/alphadata01/bstocker/rsofun/siteinfo_sj02zroot.csv")
```


Specify simulation settings.
```{r}
settings_sims <- list(
  siteinfo        = "/alphadata01/bstocker/rsofun/siteinfo_sj02zroot.csv",
  ensemble        = TRUE,
  setup           = "site",
  name            = "sj02zroot",
  dir_sofun       = options()$rsofun.dir.sofun,
  path_output     = "~/sofun/output_sj02zroot_sofun/",
  path_output_nc  = "~/sofun/output_nc_sj02zroot_sofun/",
  path_input      = "~/sofun/input_sj02zroot_sofun/",
  grid            = NA,
  implementation  = "fortran",
  in_ppfd         = TRUE,
  in_netrad       = FALSE,
  recycle         = 1,
  spinupyears     = 10,
  calibvars       = c(),  # needed later for calibration setup, any of "gpp", "fapar", and "transp"
  soilmstress     = TRUE,
  tempstress      = TRUE,
  loutdgpp        = FALSE,
  loutdwcont      = TRUE,
  loutdaet        = TRUE,
  loutdpet        = TRUE,
  loutdnetrad     = TRUE,
  loutdfapar      = TRUE,
  loutdwbal       = TRUE
  )
```


## Input settings

Specify input settings.
```{r}
settings_input <-  list(
  data                     = NA,
  temperature              = "watch_wfdei",
  precipitation            = "watch_wfdei",
  vpd                      = "watch_wfdei",
  ppfd                     = "watch_wfdei",
  netrad                   = NA,
  patm                     = NA,
  cloudcover               = "cru",
  fapar                    = "MODIS_FPAR_MCD15A3H", # "","MODIS_EVI_MOD13Q1", #  
  splined_fapar            = TRUE,
  path_co2                 = "/alphadata01/bstocker/data/co2/cCO2_rcp85_const850-1765.dat",
  path_watch_wfdei         = "~/data/watch_wfdei",
  path_fapar               = "/alphadata01/bstocker/data/fapar_MODIS_FPAR_MCD15A3H/sj02zroot/",
  path_MODIS_EVI_MOD13Q1   = "/alphadata01/bstocker/data/fapar_MODIS_EVI_MOD13Q1_gee_MOD13Q1_fluxnet2015_gee_subset/",
  get_from_remote          = FALSE,
  path_cru                 = "/alphadata01/bstocker/data/cru/ts_4.01/"
  )
```


## Model setup
Define model setup as a list.
```{r}
setup_sofun <- list(
  model      = "splash",
  dir        = options()$rsofun.dir.sofun,
  do_compile = FALSE,
  simsuite   = FALSE
  )
```

- `model`: For Fortran version: A character string specifying the compilation option. The name of the executable is derived from this as `"run<setup_sofun$model>"`.
- `dir`: A path (character) specifying the directory of where the executables are located (corresponds to the parent directory of the model git repository).
- `do_compile`: If `TRUE`, the model code is compiled as `make <setup_sofun$model>`. If `FALSE`, compiled executables are used (compiled with gfortran on a Mac 64-bit).
- `simsuite`: If `TRUE`, the SOFUN option for running an entire simulation suite (ensemble of simulations) with a single executable. (In the Fortran implementation, this this requires the main program file `sofun_simsuite.f90` to be compiled instead of `sofun.f90`). This is to be preferred over a set of individual runs submitted individually when doing calibration runs because the cost across the entire ensemble (mod-obs error) can thus be calculated online.


# Workflow

The example shown below is for a set of site-scale simulations.

## Prepare simulation setup

Create a run directory with all the simulation parameter files (defining simulation years, length, etc.). This returs the `settings_sims` list, complemented by additional information. Calibration settings are an optional argument. When passed on, simulation parameter files will contain information which target variables are to be written to special calibration output files (a single file written for an entire ensemble).

```{r, eval=TRUE, message=FALSE, warning=FALSE}
settings_sims <- prepare_setup_sofun(
  settings = settings_sims,
    setup = setup_sofun,
  write_paramfils = FALSE
  )
```

## Prepare inputs

Prepare SOFUN input (climate input, CO2, etc.). Complements `settings_input`. This will require inputs from the user through the prompt, entered in the console to specify whether data files should be downloaded from Imperial CX1. In case you chose to download, you must have access to CX1 and be connected to the Imperial VPN. Once asked (see console!), enter your user name on CX1. This also requires that no additional entering of the password is required. In order to set this up, you need to generate an SSH key pair beforehand (see [here](https://www.digitalocean.com/community/tutorials/how-to-set-up-ssh-keys--2).

```{r, eval=TRUE, message=FALSE, warning=FALSE, results="hide"}
inputdata <- prepare_input_sofun(
  settings_input = settings_input,
  settings_sims = settings_sims,
  return_data = TRUE,
  overwrite_climate = TRUE,
  overwrite_fapar = TRUE,
  verbose = TRUE,
  overwrite_csv_climate = TRUE,
  overwrite_csv_fapar = FALSE
  )
```

## Run the model

Run SOFUN with calibrated parameters.
```{r, message=FALSE, warning=FALSE}
settings_sims$in_netrad <- FALSE
nothing <- update_params( 
  params_opt, 
  settings_sims$dir_sofun, 
  setup = setup_sofun 
  )

## run and read at once
mod <- runread_sofun(
  settings = settings_sims,
  setup = setup_sofun
  )

## read only
mod <- read_sofun(
  settings = settings_sims,
  setup = setup_sofun
  )
save(mod, file = "~/mct/data/mod_v2_EVI.RData")
```

## Evaluate the model

Define model evaluation settings as a list.
```{r}
settings_eval <- list(
  sitenames = rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015$sitename,
  benchmark = list( aet = c("fluxnet2015"), netrad = c("fluxnet2015") ),
  sitenames_siteplots = "FR-Pue",
  agg = 5,
  path_fluxnet2015_d = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_fluxnet2015_w = "",
  path_fluxnet2015_m = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1m/original/unpacked/",
  path_fluxnet2015_y = "/alphadata01/bstocker/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1y/original/unpacked/",
  path_gepisat_d     = "",
  dir_figs           = "~/mct/fig/",
  remove_premodis    = FALSE
  )
```

- `benchmark`: Named list of character strings with data source identifiers for each calibration target variable. The list is named corresponding to variable names defined by 'targetvars'. The identifier triggers certain functions to be used for reading and processing observational data. Use, e.g., `benchmark = list( gpp = "fluxnet2015_NT" )` to specify that observational data for the target variable `"gpp"` comes from FLUXNET 2015 dataset with GPP data based on the night-time flux decomposition method (`"NT"`). Alternatively, use GPP data based on the daytime method (`"fluxnet2015_NT"`) or Tyler Davis' new method (unpublished) (`"fluxnet2015_Ty"`). If multiple data sources are selected (e.g., `benchmark = list( gpp = c("fluxnet2015_NT", "fluxnet2015_DT") )`), their mean is used for calibration.

Get evaluation results.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
## Get observational data for evaluation
obs_eval <- get_obs_eval( 
  settings_eval = settings_eval, 
  settings_sims = settings_sims, 
  overwrite = TRUE 
  )

## Evaluate the whole shabang
out_eval <- eval_sofun(
  mod, 
  settings_eval, 
  settings_sims, 
  obs_eval = obs_eval, 
  overwrite = TRUE 
  )
save(obs_eval, file = "~/mct/data/obs_eval_v2_EVI.RData")
save(out_eval, file = "~/mct/data/out_eval_v2_EVI.RData")
```

Net radiation
```{r}
out_plot <- out_eval$netrad$fluxnet2015$data$xdf %>% 
  mutate(mod=mod/(60*60*24), obs=obs/(60*60*24)) %>% 
  analyse_modobs2("mod", "obs", type="heat")
out_plot$gg + 
  labs(x=expression(paste("Modelled net radiation (W m"^-2, ")")), 
       y=expression(paste("Simulated net radiation (W m"^-2, ")")))
ggsave("fig/modobs_netrad.pdf")
```

Actual evapotranspiration. Modelled vs. observed.
```{r}
out_plot <- out_eval$aet$fluxnet2015$data$xdf %>% 
  analyse_modobs2("mod", "obs", type="heat")
out_plot$gg
```

```{r}
out_eval$aet$fluxnet2015$plot$by_doy_allsites()
```

The object returned by `eval_sofun()` includes a set of standard functions that can be applied to the evaluation data itself. Here is one example for observed versus modelled values.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
modobs_daily   <- out_eval$aet$fluxnet2015$plot$modobs_daily()
modobs_xdf     <- out_eval$aet$fluxnet2015$plot$modobs_xdaily()
modobs_mdf     <- out_eval$aet$fluxnet2015$plot$modobs_monthly()
modobs_meandoy <- out_eval$aet$fluxnet2015$plot$modobs_meandoy()
modobs_meandoy <- out_eval$aet$fluxnet2015$plot$modobs_annual()
out_eval$aet$fluxnet2015$plot$by_doy_allsites()
out_eval$aet$fluxnet2015$plot$by_xoy_allsites()
modobs_anomalies_daily  <- out_eval$aet$fluxnet2015$plot$modobs_anomalies_daily()
modobs_anomalies_xdaily <- out_eval$aet$fluxnet2015$plot$modobs_anomalies_xdaily()
modobs_anomalies_annual <- out_eval$aet$fluxnet2015$plot$modobs_anomalies_annual()
modobs_meanxoy <- out_eval$aet$fluxnet2015$plot$modobs_meanxoy()
out_eval$aet$fluxnet2015$plot$by_doy_allzones()
#out_eval$aet$fluxnet2015$plot$modobs_spatial_annual()
modobs_spatial <- out_eval$aet$fluxnet2015$plot$modobs_spatial()
```



