---
title: "W-model site-scale simulations"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{splash}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
# output:
#   html_document:
#     toc: true
#     toc_float: true
#     toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
---

# Environment

Load the package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r message=FALSE}
library(rsofun)

## load all rsofun dependencies
load_dependencies_rsofun()

## other crap
systr <- "''"    # for Mac
knitr::opts_knit$set( root.dir = rprojroot::find_rstudio_root_file() ) # does not work properly
## for developer setup
if (!file.exists("bash"))    system("ln -s inst/bash bash")
if (!file.exists("extdata")) system("ln -s inst/extdata extdata")

# if (is.null(options()$rsofun.dir.sofun)) rlang::abort( "Option rsofun.dir.sofun not set. Do so by `options( list( rsofun.dir.sofun=string_path_where_sofun_is ))`" )  
#options( list( rsofun.dir.sofun="~/sofun/trunk/" ) )
options( list( rsofun.dir.sofun="~/sofun/trunk/" ) )
```


## Simulation settings

Meta information for the FLUXNET 2015 sites is provided along with the rsofun package. Load this and write to file
```{r message=FALSE}
siteinfo <- rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015 %>% 
  write_csv(path = "./siteinfo_pet_fluxnet2015.csv")
```

To conduct P-model simulations with the Fortran implementation, respective model code needs to be downloaded separately from the Github [sofun](https://github.com/stineb/sofun) repository and placed in a separate directory (`dir_sofun` below). Builds of this Fortran code are available for public download at XXX in order to avoid having to compile from source.


```{r}
settings_sims <- list(
  siteinfo        = "./siteinfo_pet_fluxnet2015.csv",
  ensemble        = TRUE,
  setup           = "site",
  name            = "pet_fluxnet2015",
  dir_sofun       = "/alphadata01/bstocker/sofun/trunk/",
  path_output     = "~/sofun/output_pet_fluxnet2015_sofun/",
  path_output_nc  = "~/sofun/output_nc_pet_fluxnet2015_sofun/",
  path_input      = "~/sofun/input_pet_fluxnet2015_sofun/",
  grid            = NA,
  implementation  = "fortran",
  in_ppfd         = TRUE,
  in_netrad       = FALSE,
  recycle         = 1,
  spinupyears     = 10,
  calibvars       = c(),  # needed later for calibration setup, any of "gpp", "fapar", and "transp"
  soilmstress     = TRUE,
  tempstress      = TRUE,
  loutdgpp        = FALSE,
  loutdwcont      = TRUE,
  loutdaet        = TRUE,
  loutdpet        = TRUE,
  loutdnetrad     = TRUE,
  loutdfapar      = TRUE,
  loutdwbal       = TRUE
  )
```

- `siteinfo`: Path (character string) of a CSV file that contains all the information needed for defining simulations. One row for each simulation in an ensemble, typically sites. Specifying this element is required if `ensemble = TRUE`. Columns are as follows:
    - site name, must be: column number 1
    - longitude of site, column must be named 'lon'
    - latitude of site, column must be named 'lat'
    - elevation of site, column must be named 'elv'
    - years for which simulation is to be done (corresponding to data availability from site),
      requires two columns named 'year_start' and 'year_end'.
- `ensemble`: TRUE if an ensemble of site-level simulations are to be run (This may become obsolte, this information is given already if the number of rows in the CSV file `siteinfo` is larger than one.)
- `setup`: String. One of `"simple"` (SOFUN used as a function, single time step, single location), `"site"` (site-scale simulation), or `"lonlat"` (spatial simulation on a longitude-latitude grid).
- `name`: a character string specifying the name of the ensemble (e.g. 'fluxnet2015') or of a single simulation.
- `dir_sofun`: Path (character string) where the model sits (corresponding to the parent directory of the respective git repository).
- `path_output`: Path (character string) where model output (ascii text files) is written to.
- `path_output_nc`: Path (character string) where NetCDF model output is written to.
- `path_input`: Path (character string) where model input is located.
- `grid`: Character string defining the type of grid used, e.g. `halfdeg` for half-degree resolution in lon. and lat. (only used in `lonlat` setup).
- `implementation`: Character string specifying whether Fortran (`implementation= "fortran"`) or Python (`implementation= "python"`) version is to be used.
- `in_ppfd`: Switch (`TRUE` of `FALSE`) whether PPFD should be read from data (prescribed) or simulated online using SPLASH and prescribed fractional cloud cover data.
- `recycle`: Periodicity (integer, number of years) of repeating forcing years during spinup (e.g. if the simulation start year is 1982 and `recycle=3`, then forcing years 1979-1981 are repeated during the duration of the model spinup, so that the last year of the spinup is forcing year 1981. For `recycle=1` and simulation start year 1982, forcing year 1981 is used for all years of the model spinup. Here, 'forcing year' refers to the year AD in the climate, CO2, fAPAR, etc. data used as model forcing.
- `spinupyears`: Integer, number of model spinup years before the transient simulation. Use `spinupyears > 0` if the model contains pool variables that that are simulated by dynamics that depend on their current state (typically soil water storage, or plant and soil carbon pools). Typically `spinupyears = 10` is sufficient to bring soil water pools to equilibrium across the globe (unless you chose a large soil water holding capacity).
- `calibvars`: A vector of strings specifying which variables are to be calibrated. Use `c()` (used as default) for a standard model setup. See vignette `calib_sofun.pdf` for an example of a model calibration setup.
- `soilmstress`: Switch (`TRUE` of `FALSE`) defining whether soil moisture stress function is to be applied to GPP.
- `tempstress`: Switch (`TRUE` of `FALSE`) defining whether temperature stress function is to be applied to GPP.
- `lout<var>`: Switch (`TRUE` of `FALSE`) whether variable `<var>` is to be written to output (ascii text file for time series). To be anbandoned, only NetCDF output should be maintained.
- `lout<var>`: Switch (`TRUE` of `FALSE`) whether variable `<var>` is to be written to output (NetCDF file).


## Input settings

Input settings define data sources used for SOFUN input. Two modes of specifying inputs are available. Either providing the input data directly (element `data`, not available for `"lonlat"` simulations), or by specifying a keyword for the source dataset as strings. E.g. `"fluxnet2015"` triggers specific functions that read input data from specifically formatted files, here all the site-level meteo data from FLUXNET 2015. Define the input settings as a list.

Use this for 
```{r}
settings_input <-  list(
  data                     = NA,
  temperature              = "fluxnet2015",
  precipitation            = "fluxnet2015",
  vpd                      = "fluxnet2015",
  ppfd                     = "fluxnet2015",
  netrad                   = "fluxnet2015",  #  c("fluxnet2015", "watch_wfdei"),
  patm                     = "fluxnet2015",
  cloudcover               = "cru",
  fapar                    = "MODIS_EVI_MOD13Q1", #  "MODIS_FPAR_MCD15A3H", # "",
  splined_fapar            = TRUE,
  path_co2                 = "~/data/co2/cCO2_rcp85_const850-1765.dat",
  path_fluxnet2015         = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_fluxnet2015_hh      = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_0.5h/original/unpacked/",
  path_MODIS_FPAR_MCD15A3H = "~/data/fapar_MODIS_FPAR_MCD15A3H_gee_MCD15A3H_fluxnet2015_gee_subset/",
  path_MODIS_EVI_MOD13Q1   = "~/data/fapar_MODIS_EVI_MOD13Q1_gee_MOD13Q1_fluxnet2015_gee_subset/",
  get_from_remote          = FALSE,
  path_cru                 = "~/data/cru/ts_4.01/",
  settings_gee             = get_settings_gee( 
    bundle = "fpar", 
    python_path = "/Users/benjaminstocker/Library/Enthought/Canopy_64bit/User/bin/python",
    gee_path = "/alphadata01/bstocker/gee_subset/gee_subset/"
    )
  )
```

- `data`: A named list of data frames, each containing the input data for one site with a column `"date"` specifying the date, and columns `fapar`, `tempvarnam`, `precvarnam`, `vpdvarnam`, `ppfdvarnam`, and `netradvarnam` specifying the fraction of absorbed photosynthetically active radiation, temperature, precipitation, vapour pressure deficit, photosynthetic photon flux density, and net radiation, respictively. Elements of the list `data` must be named according to site names. If `data = NA`, data will be read from files. If `data` is provided, all other elements of the input settings will be ignored.
- `temperature`: A character string specifying the source for temperature data. Any of `"fluxnet2015"`, `"watch_wfdei"`, and/or `"cru"`. This can also be a vector of strings (e.g. `c("fluxnet2015", "watch_wfdei")`, to specify priorities: first take FLUXNET 2015 data for periods where data is available. For remaining years (given by `date_start` and `date_end`), use WATCH-WFDEI data. If `"fluxnet2015"` is specified for any of `temperature`, `precipitation`, `vpd`, `ppfd`, or `netrad`, then `path_fluxnet2015` must be specified as well in the settings.
- `precipitation`: See `temperature`.
- `vpd`: See `temperature`.
- `ppfd`: See `temperature`.
- `netrad`: See `temperature`.
- `fapar`: A character string specifying the type of fAPAR data used as input. Implemented for use of data from CX1 are `"MODIS_FPAR_MCD15A3H"` and `"MODIS_EVI_MOD13Q1"`. Use `NA` in case no fAPAR data is used as forcing (internally simulated fAPAR).
- `splined_fapar`: Logical defining whether splined fAPAR data is to be used. If `FALSE`, linearly interpolated fAPAR data is used.
- `co2`: A character string specifying which CO$_2$ file should be used (globally uniform and identical for each site in an ensemble). All available CO$_2$ forcing files are located on CX1 at `/work/bstocker/labprentice/data/co2/`. `co2="cmip"` specifies that the CMIP-standard CO2 file `cCO2_rcp85_const850-1765.dat` should be used.
- `path_fluxnet2015`: A character string specifying the path where standard FLUXNET 2015 CSV files are located.
- `path_watch_wfdei`: A character string specifying the path where standard WATCH-WFDEI NetCDF files are located.
- `path_cru_ts4_01`: A character string specifying the path where standard CRU NetCDF files are located.
- `path<fapar>`: A character string specifying the path where site-specific fapar files are located. This element is named according to the `fapar` setting (element `fapar`). E.g., an element named `path_MODIS_FPAR_MCD15A3H` is required if `fapar = MODIS_FPAR_MCD15A3H`.


## Model setup
Define model setup as a list.
```{r}
setup_sofun <- list(
  model      = "splash",
  dir        = "/alphadata01/bstocker/sofun/trunk/",
  do_compile = FALSE,
  simsuite   = FALSE
  )
```

- `model`: For Fortran version: A character string specifying the compilation option. The name of the executable is derived from this as `"run<setup_sofun$model>"`.
- `dir`: A path (character) specifying the directory of where the executables are located (corresponds to the parent directory of the model git repository).
- `do_compile`: If `TRUE`, the model code is compiled as `make <setup_sofun$model>`. If `FALSE`, compiled executables are used (compiled with gfortran on a Mac 64-bit).
- `simsuite`: If `TRUE`, the SOFUN option for running an entire simulation suite (ensemble of simulations) with a single executable. (In the Fortran implementation, this this requires the main program file `sofun_simsuite.f90` to be compiled instead of `sofun.f90`). This is to be preferred over a set of individual runs submitted individually when doing calibration runs because the cost across the entire ensemble (mod-obs error) can thus be calculated online.


# Workflow

The example shown below is for a set of site-scale simulations.

## Prepare simulation setup

Create a run directory with all the simulation parameter files (defining simulation years, length, etc.). This returs the `settings_sims` list, complemented by additional information. Calibration settings are an optional argument. When passed on, simulation parameter files will contain information which target variables are to be written to special calibration output files (a single file written for an entire ensemble).

```{r, eval=TRUE, message=FALSE, warning=FALSE}
settings_sims <- prepare_setup_sofun(
  settings = settings_sims,
    setup = setup_sofun,
  write_paramfils = FALSE
  )
```

## Prepare inputs

Prepare SOFUN input (climate input, CO2, etc.). Complements `settings_input`. This will require inputs from the user through the prompt, entered in the console to specify whether data files should be downloaded from Imperial CX1. In case you chose to download, you must have access to CX1 and be connected to the Imperial VPN. Once asked (see console!), enter your user name on CX1. This also requires that no additional entering of the password is required. In order to set this up, you need to generate an SSH key pair beforehand (see [here](https://www.digitalocean.com/community/tutorials/how-to-set-up-ssh-keys--2).

```{r, eval=TRUE, message=FALSE, warning=TRUE, results="hide"}
inputdata <- prepare_input_sofun(
  settings_input = settings_input,
  settings_sims = settings_sims,
  return_data = FALSE,
  overwrite_climate = TRUE,
  overwrite_fapar = TRUE,
  verbose = TRUE,
  overwrite_csv_climate = TRUE,
  overwrite_csv_fapar = TRUE
  )
```

## Run the model

Run SOFUN with calibrated parameters.
```{r, message=FALSE, warning=FALSE}
settings_sims$in_netrad <- FALSE
nothing <- update_params( 
  params_opt, 
  settings_sims$dir_sofun, 
  setup = setup_sofun 
  )

## run and read at once
mod <- runread_sofun(
  settings = settings_sims,
  setup = setup_sofun
  )

## read only
mod <- read_sofun(
  settings = settings_sims,
  setup = setup_sofun
  )
save(mod, file = "~/mct/data/mod_v2_EVI.RData")
```

## Evaluate the model

Define model evaluation settings as a list.
```{r}
settings_eval <- list(
  sitenames = rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015$sitename,
  benchmark = list( aet = c("fluxnet2015"), netrad = c("fluxnet2015") ),
  sitenames_siteplots = "FR-Pue",
  agg = 5,
  path_fluxnet2015_d = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_fluxnet2015_w = "",
  path_fluxnet2015_m = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1m/original/unpacked/",
  path_fluxnet2015_y = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1y/original/unpacked/",
  path_gepisat_d     = "",
  dir_figs           = "~/mct/fig/",
  remove_premodis    = FALSE
  )
```

- `benchmark`: Named list of character strings with data source identifiers for each calibration target variable. The list is named corresponding to variable names defined by 'targetvars'. The identifier triggers certain functions to be used for reading and processing observational data. Use, e.g., `benchmark = list( gpp = "fluxnet2015_NT" )` to specify that observational data for the target variable `"gpp"` comes from FLUXNET 2015 dataset with GPP data based on the night-time flux decomposition method (`"NT"`). Alternatively, use GPP data based on the daytime method (`"fluxnet2015_NT"`) or Tyler Davis' new method (unpublished) (`"fluxnet2015_Ty"`). If multiple data sources are selected (e.g., `benchmark = list( gpp = c("fluxnet2015_NT", "fluxnet2015_DT") )`), their mean is used for calibration.

Get evaluation results.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
## Get observational data for evaluation
obs_eval <- get_obs_eval( 
  settings_eval = settings_eval, 
  settings_sims = settings_sims, 
  overwrite = TRUE 
  )

## Evaluate the whole shabang
out_eval <- eval_sofun(
  mod, 
  settings_eval, 
  settings_sims, 
  obs_eval = obs_eval, 
  overwrite = TRUE 
  )
save(obs_eval, file = "~/mct/data/obs_eval_v2_EVI.RData")
save(out_eval, file = "~/mct/data/out_eval_v2_EVI.RData")
```

Net radiation
```{r}
out_plot <- out_eval$netrad$fluxnet2015$data$xdf %>% 
  mutate(mod=mod/(60*60*24), obs=obs/(60*60*24)) %>% 
  analyse_modobs2("mod", "obs", type="heat")
out_plot$gg + 
  labs(x=expression(paste("Modelled net radiation (W m"^-2, ")")), 
       y=expression(paste("Simulated net radiation (W m"^-2, ")")))
ggsave("fig/modobs_netrad.pdf")
```

Actual evapotranspiration. Modelled vs. observed.
```{r}
out_plot <- out_eval$aet$fluxnet2015$data$xdf %>% 
  analyse_modobs2("mod", "obs", type="heat")
out_plot$gg
```

```{r}
out_eval$aet$fluxnet2015$plot$by_doy_allsites()
```

The object returned by `eval_sofun()` includes a set of standard functions that can be applied to the evaluation data itself. Here is one example for observed versus modelled values.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
modobs_daily   <- out_eval$aet$fluxnet2015$plot$modobs_daily()
modobs_xdf     <- out_eval$aet$fluxnet2015$plot$modobs_xdaily()
modobs_mdf     <- out_eval$aet$fluxnet2015$plot$modobs_monthly()
modobs_meandoy <- out_eval$aet$fluxnet2015$plot$modobs_meandoy()
modobs_meandoy <- out_eval$aet$fluxnet2015$plot$modobs_annual()
out_eval$aet$fluxnet2015$plot$by_doy_allsites()
out_eval$aet$fluxnet2015$plot$by_xoy_allsites()
modobs_anomalies_daily  <- out_eval$aet$fluxnet2015$plot$modobs_anomalies_daily()
modobs_anomalies_xdaily <- out_eval$aet$fluxnet2015$plot$modobs_anomalies_xdaily()
modobs_anomalies_annual <- out_eval$aet$fluxnet2015$plot$modobs_anomalies_annual()
modobs_meanxoy <- out_eval$aet$fluxnet2015$plot$modobs_meanxoy()
out_eval$aet$fluxnet2015$plot$by_doy_allzones()
#out_eval$aet$fluxnet2015$plot$modobs_spatial_annual()
modobs_spatial <- out_eval$aet$fluxnet2015$plot$modobs_spatial()
```



