---
title: "Compare P-model implementations"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

## Environment

Load the `rsofun` package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r message=FALSE, echo=FALSE, warning=FALSE}
library(rsofun)

## load all rsofun dependencies
load_dependencies_rsofun()
```

## Simulation settings

```{r}
path_siteinfo <- "~/.siteinfo_example_fortran.csv"
siteinfo <- rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015 %>% 
  # dplyr::filter(sitename %in% c("FR-Pue", "FR-LBr", "IT-Noe")) %>%
  # dplyr::slice(61) %>%
  write_csv(path = path_siteinfo)
```


```{r}
params_siml <- list(
  spinup         = TRUE,
	spinupyears    = 10,
	recycle        = 1,
	soilmstress               = FALSE,
	tempstress                = FALSE,
	in_ppfd                   = TRUE,
	in_netrad                 = FALSE,
	const_clim_year           = -9999,
	const_lu_year             = -9999,
	const_co2_year            = -9999,
	const_ndep_year           = -9999,
	const_nfert_year          = -9999,
	daily_out_startyr         = 1982,
	daily_out_endyr           = 1982,
	outdt                     = 1,
	ltre                      = FALSE,
	ltne                      = FALSE,
	ltrd                      = FALSE,
	ltnd                      = FALSE,
	lgr3                      = TRUE,
	lgn3                      = FALSE,
	lgr4                      = FALSE,
	loutplant                 = FALSE,
	loutgpp                   = FALSE,
	loutwaterbal              = FALSE,
	loutforcing               = FALSE,
	loutdgpp                  = FALSE,
	loutdrd                   = FALSE,
	loutdtransp               = FALSE,
	loutdwcont                = FALSE,
	loutdaet                  = FALSE,
	loutdpet                  = FALSE,
	loutdnetrad               = FALSE,
	loutdwbal                 = FALSE,
	loutdtemp                 = FALSE,
	loutdfapar                = FALSE,
	loutdtemp_soil            = FALSE,
	lcalibgpp                 = FALSE,
	lcalibfapar               = FALSE,
	lcalibtransp              = FALSE,
	lcaliblatenth             = FALSE
	)

settings_sims <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```


### Define model parameters
```{r}
params_modl <- list(
	kphio           = 0.04997714009213085,
	soilm_par_a     = 1.0,
	soilm_par_b     = 0.0,
	vpdstress_par_a = 0.2,
	vpdstress_par_b = 0.2,
	vpdstress_par_m = 5
	)
```

### Define soil parameters

```{r}
list_soiltexture <- list(
  top = list(fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = list(fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

## Get input

First, define input settings.
```{r}
settings_input <-  list(
    data                     = NA,
    temperature              = "fluxnet2015",
    precipitation            = "fluxnet2015",
    vpd                      = "fluxnet2015",
    ppfd                     = "fluxnet2015",
    netrad                   = "fluxnet2015",  #  c("fluxnet2015", "watch_wfdei"),
    patm                     = "fluxnet2015",
    netrad                   = NA,
    cloudcover               = "cru",
    path_input               = "~/sofun_inputs/example/",
    path_watch_wfdei         = "~/data/watch_wfdei/",
    path_cru                 = "~/data/cru/ts_4.01/",
    path_MODIS_FPAR_MCD15A3H = "~/data/fluxnet_subsets/fapar_MODIS_FPAR_MCD15A3H_gee_MCD15A3H_fluxnet2015_gee_subset/",
    path_MODIS_EVI_MOD13Q1   = "~/data/fluxnet_subsets/fapar_MODIS_EVI_MOD13Q1_gee_MOD13Q1_fluxnet2015_gee_subset/",
    path_co2                 = "~/data/co2/cCO2_rcp85_const850-1765.csv",
    path_fluxnet2015         = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
    path_fluxnet2015_hh      = "~/data/FLUXNET-2015_Tier1/20191024/HH/",
    get_from_remote          = FALSE,
    settings_gee             = get_settings_gee( 
      bundle = "fpar", 
      python_path = "/Users/benjaminstocker/Library/Enthought/Canopy_64bit/User/bin/python",
      gee_path = "~/gee_subset/gee_subset/"
      ),
  fapar = "MODIS_FPAR_MCD15A3H",
  splined_fapar = TRUE
  )
```


Then, get the input data.
```{r message=FALSE, warning=FALSE, echo=FALSE}
ddf_input <- prepare_input_sofun(
  settings_input = settings_input,
  settings_sims = settings_sims,
  overwrite_csv_climate_lev1 = FALSE,
  overwrite_csv_climate_lev2 = FALSE,
  overwrite_csv_climate_lev3 = FALSE,
  overwrite_rdata_climate = FALSE,
  overwrite_csv_fapar = FALSE,
  verbose = FALSE
  )
```

## Run and compare implementations

### Compare input data

First, compare input data with the dataset used in Stocker et al. (2019) GMD. 
```{r}
ddf_stocker19gmd <- read_csv("~/data/stocker19gmd_outputs/ddf_fluxnet2015_pmodel_with_forcings_stocker19gmd.csv")

## temperature: ok
ddf_input %>% 
  left_join(ddf_stocker19gmd, by = c("sitename", "date")) %>% 
  ggplot(aes(x = temp, y = temp_day_fluxnet2015)) + 
  geom_point(size = 0.1)

## VPD
ddf_input %>% 
  left_join(ddf_stocker19gmd, by = c("sitename", "date")) %>% 
  ggplot(aes(x = vpd, y = vpd_day_fluxnet2015)) + 
  geom_point(size = 0.1)

## VPD data is different for CZ-BK1, no idea why
ddf_test <- ddf_input %>% 
  left_join(ddf_stocker19gmd, by = c("sitename", "date")) %>% 
  mutate(vpd_res = vpd - vpd_day_fluxnet2015) %>% 
  filter(vpd_res != 0) %>% 
  group_by(sitename) %>% 
  summarise(tot = n()) %>% 
  print()

## VPD without data from CZ-BK1
ddf_input %>% 
  filter(sitename!="CZ-BK1") %>% 
  left_join(ddf_stocker19gmd, by = c("sitename", "date")) %>% 
  ggplot(aes(x = vpd, y = vpd_day_fluxnet2015)) + 
  geom_point(size = 0.1)

## PPFD: ok
ddf_input %>% 
  left_join(ddf_stocker19gmd, by = c("sitename", "date")) %>% 
  ggplot(aes(x = ppfd, y = ppfd_fluxnet2015)) + 
  geom_point(size = 0.1)

## prec: ok
ddf_input %>% 
  left_join(ddf_stocker19gmd, by = c("sitename", "date")) %>% 
  ggplot(aes(x = prec, y = prec_fluxnet2015)) + 
  geom_point(size = 0.1)

## fapar: looks quite different :-/
ddf_test <- ddf_input %>% 
  left_join(ddf_stocker19gmd, by = c("sitename", "date"))

ddf_test %>% 
  ggplot(aes(x = fapar, y = fapar_spl)) + 
  geom_point(size = 0.1)

## seems to be related to a stiffer (or just different?) spline applied now
## not gonna resolve this for now,
ddf_test %>%
  filter(sitename == "CH-Lae" & year(date)==2013) %>% 
  rename(fapar_new = fapar, fapar_stocker19gmd = fapar_spl) %>% 
  pivot_longer(cols = c(fapar_new, fapar_stocker19gmd), names_to = "set", values_to = "fapar") %>% 
  ggplot(aes(x = date, y = fapar, color = set)) +
  geom_line()
```

As shown above, there is a slight difference in VPD values (exclusively from one site) and some difference in the splined fapar data.

### SOFUN results

Re-create input data from the file `ddf_fluxnet2015_pmodel_with_forcings_stocker19gmd.csv`, run rsofun with these forcings and evaluate against outputs from Stocker et al. (2019) GMD.
```{r}
ddf_input_test <- select(ddf_input, sitename, date, ccov, year, co2) %>% 
  left_join(
    select(ddf_stocker19gmd, sitename, date, temp = temp_day_fluxnet2015, prec = prec_fluxnet2015, vpd = vpd_day_fluxnet2015, ppfd = ppfd_fluxnet2015, fapar = fapar_spl),
    by = c("sitename", "date")
  )
  
test_bysite <- function(usesitename){
  
  mod_new <- try( run_sofun_f_bysite( 
    settings         = dplyr::filter(settings_sims, sitename == usesitename), 
    params_modl      = params_modl, 
    list_soiltexture = list_soiltexture,
    forcing          = dplyr::filter(ddf_input_test, sitename == usesitename)
  ) )
  
  if (nrow(mod_new) != 1){
    
    mod_new <- mod_new %>%
      left_join(select(ddf_stocker19gmd, sitename, date, gpp_mod_ORG), by = c("sitename", "date"))
    
    modobs <- mod_new %>% 
      rbeni::analyse_modobs2("gpp", "gpp_mod_ORG", type = "heat")
    gg_modobs <- modobs$gg + 
      labs(title = usesitename)
    
    ## There are some small differences for a substantial number of points (this is just one site)
    ddf_test <- mod_new %>% 
      mutate(gpp_res = gpp - gpp_mod_ORG) %>% 
      filter(abs(gpp_res) > 1e-6) %>% 
      arrange(-gpp_res) %>% 
      left_join(select(ddf_input_test, -fapar), by = c("sitename", "date")) %>% 
      left_join(select(settings_sims, sitename, elv), by = "sitename")
    
    gg_hist <- ddf_test %>% 
      ggplot() +
      geom_histogram(aes(x = gpp_res / gpp, y = ..density..), position = "identity", color = "black", fill = "black", alpha = 0.3) + 
      labs(title = usesitename)
    
    return(list(gg_modobs, gg_hist))
    
  }
  
}

purrr::map(as.list(settings_sims$sitename), ~test_bysite(.))
```

The histograms are for a subset of data, where the residual between rsofun-calculated GPP and Stocker et al. (2019) GMD results differ by more than 1e-6 g C m-2 d-1. 

This suggests that there are some differences in the exact values of GPP simulated by the `rsofun` implementation, but these look like numerical inaccuracies.

